### 2018-11-21 ggseg_paper/ggseg-paper.tex
started paper

% !TeX root = RJwrapper.tex
\title{Visualisation of Brain Statistics with R-package ggseg}
\author{by Athanasia M. Mowinckel, Didac Vidal Piñeiro}

\maketitle

\abstract{%
(150 words) An abstract of less than 150 words.
}

% Any extra LaTeX you need in the preamble

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Neuroscientific analysis usually requires the use of multiple softwares
to analyse, visualise, and summarise data. This often makes the process
of preparing results for publication laborious, as results must be
exported and imported in various formats. In neuroimaging, images of
probands brain as collected and merged together to provide three
dimensional representations. Much of the neuroimaging analyses done are
done not on the individal voxel (i.e.~3-dimentional pixels of brain
images) level, but rather on pre-defined brain segmentations, called
brain atlases. These brain atlases are plentiful and are different ways
of segmenting the brain into functionally or structurally similar
regions, and the use of these is wide-spread, as these atlases provide
larger meaningful divisions of the brain. While neuroimaging analyses on
the voxel-level are usually computed by special software for such
analyses, analyses of brain atlas data is usually done in standard
statistical software, like R \citep{R}.

A key part of understanding and disseminating analysis results, is the
visualiasion of these in a meaningful way. With regards to results from
brain atlas analyses, it is most meaningfully represented if projected
onto a representation of the brain, rather than other standard types of
charts. Each atlas has its own labelling depending on what is meaningful
for the type of segmentation it is based on, and as such for the reader
to fully understand a bar chart with atlas labels, they need to be very
familiar with the location of each label to create a clear comprehension
of the results. A projection directly onto a brain shape, eases the
readability of the results for the reader, and provides clear references
even if the used atlas is unfamiliar.

While there are several tools that aid R-users in plotting neuroimaging
data directly through R using the grammar of graphics as implemented in
ggplot2(REF), such as ggBrain(REF) and ggneuro(REF)(see
\href{https://neuroconductor.org/}{neuroconductor} for compiled
neuroimaging packages for R), these are based on plotting imaging files,
not results from analyses of brain atlases. We here introduce the
ggseg-package for visualising results from brain atlas analyses. The
ggseg-package was developed to create templates that others might use to
project their brain atlas results on to. It is based on the grammar of
graphics of ggplot2 using polygons, and while its plotting functions are
the what users are drawn to, it is the pre-compiled number of datasets
for different brain atlases that provides the real functionality needed
for visualisation of brain atlas results.

\hypertarget{rationale}{%
\subsection{Rationale}\label{rationale}}

\hypertarget{brain-atlas-selection}{%
\subsection{Brain atlas selection}\label{brain-atlas-selection}}

dkt, jhu, glasser, yeo7, yeo17, aseg, mid sagittal.

\hypertarget{ggseg3d---the-plotly-surface-plot}{%
\subsection{ggseg3d - the plotly surface
plot}\label{ggseg3d---the-plotly-surface-plot}}

\hypertarget{just-keep-this-in-for-now-so-we-remember-to-check}{%
\subsection{Just keep this in for now, so we remember to
check}\label{just-keep-this-in-for-now-so-we-remember-to-check}}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
for Authors}.

\bibliography{RJreferences}

\address{%
Athanasia M. Mowinckel\\
Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo\\
Postboks 1094 Blindern\\ 0317 Oslo\\ Norway\\
}
\href{mailto:a.m.mowinckel@psykologi.uio.bo}{\nolinkurl{a.m.mowinckel@psykologi.uio.bo}}

\address{%
Didac Vidal Piñeiro\\
Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo\\
Postboks 1094 Blindern\\ 0317 Oslo\\ Norway\\
}
\href{mailto:d.v.pineiro@psykologi.uio.no}{\nolinkurl{d.v.pineiro@psykologi.uio.no}}

### 2018-11-21 ggseg_paper/ggseg_paper.Rmd
started paper

--- 2018-11-21 ggseg_paper/ggseg-paper.tex
@@
-% !TeX root = RJwrapper.tex
-\title{Visualisation of Brain Statistics with R-package ggseg}
-\author{by Athanasia M. Mowinckel, Didac Vidal Piñeiro}
+---
+title: Visualisation of Brain Statistics with R-package ggseg
+author:
+  - name: Athanasia M. Mowinckel
+    affiliation: Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo
+    address:
+    - Postboks 1094 Blindern
+    - 0317 Oslo
+    - Norway
+    email:  a.m.mowinckel@psykologi.uio.bo
+  - name: Didac Vidal Piñeiro
+    affiliation: Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo
+    address:
+    - Postboks 1094 Blindern
+    - 0317 Oslo
+    - Norway
+    email:  d.v.pineiro@psykologi.uio.no
+abstract: >
+  (150 words) An abstract of less than 150 words.
+output:
+  rticles::rjournal_article:
+    includes:
+      in_header: preamble.tex
+---
 
-\maketitle
+## Introduction
 
-\abstract{%
-(150 words) An abstract of less than 150 words.
-}
+Neuroscientific analysis usually requires the use of multiple softwares to analyse, visualise, and summarise data. 
+This often makes the process of preparing results for publication laborious, as results must be exported and imported in various formats. 
+In neuroimaging, images of probands brain as collected and merged together to provide three dimensional representations.
+Much of the neuroimaging analyses done are done not on the individal voxel (i.e. 3-dimentional pixels of brain images) level, but rather on pre-defined brain segmentations, called brain atlases.
+These brain atlases are plentiful and are different ways of segmenting the brain into functionally or structurally similar regions, and the use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain.
+While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R \citep{R}.
 
-% Any extra LaTeX you need in the preamble
+A key part of understanding and disseminating analysis results, is the visualiasion of these in a meaningful way. 
+With regards to results from brain atlas analyses, it is most meaningfully represented if projected onto a representation of the brain, rather than other standard types of charts. 
+Each atlas has its own labelling depending on what is meaningful for the type of segmentation it is based on, and as such for the reader to fully understand a bar chart with atlas labels, they need to be very familiar with the location of each label to create a clear comprehension of the results.
+A projection directly onto a brain shape, eases the readability of the results for the reader, and provides clear references even if the used atlas is unfamiliar.
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+While there are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2(REF), such as ggBrain(REF) and ggneuro(REF)(see [neuroconductor](https://neuroconductor.org/) for compiled neuroimaging packages for R), these are based on plotting imaging files, not results from analyses of brain atlases.
+We here introduce the ggseg-package for visualising results from brain atlas analyses.
+The ggseg-package was developed to create templates that others might use to project their brain atlas results on to. 
+It is based on the grammar of graphics of ggplot2 using polygons, and while its plotting functions are the what users are drawn to, it is the pre-compiled number of datasets for different brain atlases that provides the real functionality needed for visualisation of brain atlas results.
 
-Neuroscientific analysis usually requires the use of multiple softwares
-to analyse, visualise, and summarise data. This often makes the process
-of preparing results for publication laborious, as results must be
-exported and imported in various formats. In neuroimaging, images of
-probands brain as collected and merged together to provide three
-dimensional representations. Much of the neuroimaging analyses done are
-done not on the individal voxel (i.e.~3-dimentional pixels of brain
-images) level, but rather on pre-defined brain segmentations, called
-brain atlases. These brain atlases are plentiful and are different ways
-of segmenting the brain into functionally or structurally similar
-regions, and the use of these is wide-spread, as these atlases provide
-larger meaningful divisions of the brain. While neuroimaging analyses on
-the voxel-level are usually computed by special software for such
-analyses, analyses of brain atlas data is usually done in standard
-statistical software, like R \citep{R}.
+## Rationale
 
-A key part of understanding and disseminating analysis results, is the
-visualiasion of these in a meaningful way. With regards to results from
-brain atlas analyses, it is most meaningfully represented if projected
-onto a representation of the brain, rather than other standard types of
-charts. Each atlas has its own labelling depending on what is meaningful
-for the type of segmentation it is based on, and as such for the reader
-to fully understand a bar chart with atlas labels, they need to be very
-familiar with the location of each label to create a clear comprehension
-of the results. A projection directly onto a brain shape, eases the
-readability of the results for the reader, and provides clear references
-even if the used atlas is unfamiliar.
-
-While there are several tools that aid R-users in plotting neuroimaging
-data directly through R using the grammar of graphics as implemented in
-ggplot2(REF), such as ggBrain(REF) and ggneuro(REF)(see
-\href{https://neuroconductor.org/}{neuroconductor} for compiled
-neuroimaging packages for R), these are based on plotting imaging files,
-not results from analyses of brain atlases. We here introduce the
-ggseg-package for visualising results from brain atlas analyses. The
-ggseg-package was developed to create templates that others might use to
-project their brain atlas results on to. It is based on the grammar of
-graphics of ggplot2 using polygons, and while its plotting functions are
-the what users are drawn to, it is the pre-compiled number of datasets
-for different brain atlases that provides the real functionality needed
-for visualisation of brain atlas results.
-
-\hypertarget{rationale}{%
-\subsection{Rationale}\label{rationale}}
-
-\hypertarget{brain-atlas-selection}{%
-\subsection{Brain atlas selection}\label{brain-atlas-selection}}
-
+## Brain atlas selection
 dkt, jhu, glasser, yeo7, yeo17, aseg, mid sagittal.
 
-\hypertarget{ggseg3d---the-plotly-surface-plot}{%
-\subsection{ggseg3d - the plotly surface
-plot}\label{ggseg3d---the-plotly-surface-plot}}
+## ggseg3d - the plotly surface plot
 
-\hypertarget{just-keep-this-in-for-now-so-we-remember-to-check}{%
-\subsection{Just keep this in for now, so we remember to
-check}\label{just-keep-this-in-for-now-so-we-remember-to-check}}
-
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
-for Authors}.
-
+## Just keep this in for now, so we remember to check
+This file is only a basic article template.
+For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
 \bibliography{RJreferences}
-
-\address{%
-Athanasia M. Mowinckel\\
-Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo\\
-Postboks 1094 Blindern\\ 0317 Oslo\\ Norway\\
-}
-\href{mailto:a.m.mowinckel@psykologi.uio.bo}{\nolinkurl{a.m.mowinckel@psykologi.uio.bo}}
-
-\address{%
-Didac Vidal Piñeiro\\
-Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo\\
-Postboks 1094 Blindern\\ 0317 Oslo\\ Norway\\
-}
-\href{mailto:d.v.pineiro@psykologi.uio.no}{\nolinkurl{d.v.pineiro@psykologi.uio.no}}

### 2019-04-10 draft_1.Rmd
first draft

---
title: Visualisation of Brain Statistics with R-package `ggseg`
preprint: true
author: 
  - name: Athanasia M. Mowinckel
    affiliation: 1
    corresponding: true
    email: a.m.mowinckel@psykologi.uio.no
  - name: Didac Vidal Piñeiro
    affiliation: 1
affiliation:
  - code: 1
    address: Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway
abstract: >
  The abstract of the article.
  It can also be on *multiple* lines.
header-includes: 
- \usepackage{lipsum} 
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
bibliography: sample.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
    fig_caption: true
  rticles::peerj_article: default
---
```{r setup, include=F}
knitr::opts_chunk$set(eval=TRUE, echo=TRUE)
library(kableExtra)
library(tidyverse)
library(plotly) 
library(ggplot2)
library(knitr)
options(tinytex.verbose = TRUE)

add_plotly <- function(plotly, file){
  orca(p, file = file, format = "png")
  knitr::include_graphics(file)
}

# Hook for setting latex text sizes
def.chunk.hook  <- knitr::knit_hooks$get("chunk") 
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

# Hook for setting code chunk wrapping
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

dt <- data(package='ggseg')
dt <- as.data.frame(dt$results, stringsAsFactors = FALSE)

dtE <- data(package='ggsegExtra')
dtE <- as.data.frame(dtE$results, stringsAsFactors = FALSE)

dd <- bind_rows(dt, dtE) %>% 
  filter(!grepl("pal", Item)) %>% 
  mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
         Item = gsub("_3d", "", Item),
         val = "Yes") %>% 
  select(Package, Title, Type, Item, val) %>% 
  spread(Type, val, fill = "No") %>% 
  mutate(Title = linebreak(stringr::str_wrap(Title, width = 40)))

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```

# Introduction

Neuroscientific analysis usually requires the use of multiple tools to analyse, visualise, and summarise data. 
This often makes the process of preparing results for publication laborious, as results must be exported and imported in various formats. 
In neuroimaging, images of probands' brains are collected and merged together to provide three dimensional representations.
Much of the neuroimaging analyses are not necessarily done on the individual voxel level (3-dimensional pixels), but rather on pre-defined brain segmentations, called brain atlases.
These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (REF). 
The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain (REF). 
While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (REF), python (REF), or Matlab (REF).

A key part of understanding and disseminating analysis results, is the visualization of these in a meaningful way. 
With regards to results from brain atlas analyses, it is most meaningfully represented if projected onto a representation of the brain, rather than other common types of charts (like bar-charts) accompanied by a diagram of the brain label positions. 
Each atlas has its own labelling depending on what is meaningful for the type of segmentation it is based on. 
As such, for the reader to fully understand a bar chart with atlas labels, they need to be very familiar with the location of each label to have a clear comprehension of the results.
A projection directly onto a brain shape, eases the readability of the results for the reader, and provides clear point of reference even if the used atlas is unfamiliar.

There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (REF), such as ggBrain(REF) and ggneuro(REF)(see [neuroconductor](https://neuroconductor.org/) for compiled neuroimaging packages for R). 
These are based on plotting imaging files, not results from analyses of brain atlases. 
Here we introduce the ggseg-package for visualizing results from brain atlas analyses.
The ggseg-package was developed to create templates that others might use to project their brain atlas results on to. 
While its plotting functions are what the users are drawn to, it is the pre-compiled number of data sets for different brain atlases that provides the real functionality needed for this visualization.
The package also includes functions to plot atlases in tri-surface mesh plots using plotly (REF).

## Brain atlas selection
There is a multitude of brain atlases to chose from, all with different purposes and segmented based on different criteria. 
We have initially focused on commonly used and well-established atlases with manageable numbers of segmentations. 
The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer (REF); dkt and aseg. 
The Desikan-Killany cortical atlas (dkt) (REF) and the Automatic Segmentation of Subcortical Structures (aseg) (REF) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.

While functions and base data can be found in the main `ggseg`-package, a companion package called `ggsegExtra` has numerous other atlases, and is continuously expanded. 
We encourage the community to help build and expand the atlases available, and to create a repository of established and often used atlases. 
A summary of all available atlases may be found in Table \@ref(tab:atlasTab), where in addition to the two default atlases, ggsegExtra has another `r dd %>% filter(Package == "ggsegExtra") %>% nrow()` atlases.
The [ggseg wiki](https://github.com/LCBC-UiO/ggseg/wiki) currently has two different ways you may create and supply atlases to the package as 2D polygons.
Work is being done to include instructions on 3D-mesh plot contributions also. 
The package includes functions like `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to aid users in testing and making custom atlases.

```{r "atlasTab", echo=F, warning=FALSE, message=FALSE}
dd %>% 
  kable(format="latex", escape = FALSE, booktabs = TRUE,
               caption = "Table of currently available atlases in either ggseg or ggsegExtra R-package. Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful. ") %>% 
  collapse_rows(1)
```

# Tutorial
The aim of this tutorial is to familiarize the reader with the main package functions and the general use of the package.  
It will cover the two main functions: `ggseg()` which plots 2D polygons,  and `ggseg3d()` which plots tri-surface mesh plots.

## Plotting polygon data

The function to plot two-dimensional polygon data is the `ggseg`-function. 
This function will automatically plot the dkt-atlas by default (see Figure \@ref(fig:init)). 
As this function is a wrapper for `geom_polygon` from `ggplot2` it can be built upon like all ggplot objects. 
The plot is kept as simple as possible, with as little extra information to ggplot as possible. 
This is done for it to be easy to extend the plot with the many ggplot options available.

```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in shaded polygons."}
library(ggseg)
library(tidyverse)
ggseg()
```

Because of the way `ggseg` is implemented, you should be able to safely use the function together with other ggplot functions, like themes, scales, and such (see Figure \@ref(fig:void)). 
There are many extra options to ggseg, to alter the plot as you wish.

```{r void, fig.width=5, fig.cap="ggseg plots can be used with most standard scales, themes and such that work with ggplot. The special ggseg options for hemispheres, view etc. depend on the atlas used, and some options are only available for certain atlases. There is no 'lateral' or 'medial' views of subcortical atlases. "}
p1 <- ggseg(position = "stacked") + 
  theme_dark() + 
  labs(title="dkt", subtitle = "dark theme")
p2 <- ggseg(position = "stacked") + 
  theme_classic() + 
  labs(title="dkt", subtitle = "classic theme")

# Combine plots
pp <- cowplot::plot_grid(p1, p2, nrow = 2)

p3 <- ggseg(atlas=aseg) + 
  labs(title="aseg", subtitle = "default theme")

# Create a subplot grid
cowplot::plot_grid(pp, p3, nrow = 1)
```

You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these (see Figure \@ref(fig:stacked)).
The possibilities of altering the plot depends on the atlas used. 
There are generally more options available for cortical atlases than subcortical ones, as the atlases have more meaningful ways of viewing the polygons.
Subcortical atlases also tend to have structures that span the hemispheres or other typically used referece points of the brain, and as such it is more difficult to label all segments as either, e.g., left or right.

```{r stacked, fig.width=5, fig.cap="There are several ggseg-special options that may be supplied to control how the plots looks and is organised. Top left: plot using medial view. Bottom left: plot using left hemisphere. Right: plot combining both the options. "}
lat <- ggseg(view = "medial") + 
  labs(title = "Medial view")
left <- ggseg(hemisphere = "left") + 
  labs(title = "Left hemi.")
pp <- cowplot::plot_grid(lat, left, nrow=2)

combo <- ggseg(view = "medial",
      hemisphere = "left") + 
  labs(title = "Combination")

cowplot::plot_grid(pp, combo, nrow=1)
```

### Using own data with fill and colour
`ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality. 
Standard arguments like `fill` to flood the segments with a colour, or `colour` to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.

```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas."}
ggseg(mapping=aes(fill = area), colour="black") + 
  scale_fill_brain("dkt") +
  theme(legend.justification=c(1,0),
        legend.position="bottom",
        legend.text = element_text(size = 5)) +
  guides(fill = guide_legend(ncol = 3))
```

Typically you would want each area to have a different colour, corresponding to some descriptive or inference statistic that has been computed.
The bulk of the package originates not in its functions, but in the datasets that accompany the package. 
There is one dataset per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas. 
Before setting up the data from the statistics to project onto the segments, looking at the atlas datasets might help understand what the data needs to look like, or avoid looking like.

```{r, fig.cap="All ggseg polygon atlases are required to have at minimum 6 columns with the names: 'long', 'lat', 'id', 'hemi', 'area', 'side'. These columns are necessary for the ggseg function to work. The additional columns in the dkt were added for various practical reasons.", size='small'}
dkt
```

The column 'label' was added to the atlas for easy matching with the output from Freesurfer statistics tables. 
The data is in long-format, meaning each segment has its own row, meaning the data you wish to project onto the segments also need to be in long-format. 
As data directly from Freesurfer comes in wide-format, this requires some extra wranling.

```{r}
freesurfer_stats <- data.frame(
  id = c(10:12),
  lh_superiortemporal = c(3.32, 4.1, 3.5),
  lh_precentral = c(2.3, 2.5, 2.1),
  lh_rostralmiddlefrontal = c(3.3, 3.2, 3.1)
  )
freesurfer_stats

freesurfer_long <- freesurfer_stats %>% 
  gather(label, thickness, -id)
freesurfer_long
```

This data can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
The data **must** include a column that has the same name at atleast _some_ data matching the values in the corresponding column in the atlas data.
In the next example we create some data with 4 rows, and an 'area' and 'p' column. 
The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins. 
We use the `p` column as the column flooding the segment with colour.

```{r datasupp, fig.width=5}
someData = data.frame(
  area = c("transverse temporal", "insula",
           "pre central","superior parietal"), 
  p = sample(seq(0,.5,.001), 4),
  stringsAsFactors = FALSE)

p <- ggseg(.data=someData, mapping=aes(fill=p))
p
```

We can also change how this looks, by various ggplot functions, like scales, labs, themes and such. 
```{r datasupp2, fig.width=5}
p +
  theme_void() +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  labs(title="A nice plot title", fill="p-value")
```

If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column `hemi` with either "right" or "left" for this to happen.
```{r datasuppX, fig.width=5}
someData$hemi = "right"

ggseg(.data=someData, colour="white",mapping=aes(fill=p)) +
  theme_void() +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  labs(title="Another nice plot title")
```

### Creating subplots
If your data has different statistics for different groups, and you want to plot them all, we can use ggplot's `facet_wrap` or `facet_grid` for that.
The data needs to be in long format for this to work, as needed by ggplot, meaning the data from the groups should appear in seperate rows in the data, not in separate columns. Secondly, the data needs to be grouped using `dplyr`'s `group_by` function _before_ providing the data to the `ggseg` function. 
The ggseg-function will detect if the data is grouped, and perform the necessary steps for the `facet`'s to work when they are applied later.

```{r datasupp3, fig.width=5}
someData = data.frame(
  area = rep(c("transverse temporal", "insula",
               "pre central","superior parietal"),2), 
  p = sample(seq(0,.5,.001), 8),
  AgeG = c(rep("Young",4), rep("Old",4)),
  stringsAsFactors = FALSE) %>% 
  group_by(AgeG)

ggseg(.data=someData, colour="white", mapping=aes(fill=p)) +
  facet_wrap(~AgeG, ncol=1) +
  theme(legend.position = "bottom")
```

Depending on what type of column is used for fill (continuous or discrete),
you can also use the differenct `scales` to alter the colours.

```{r datasupp5, fig.width=5}
ggseg(.data = someData, atlas=dkt, 
      colour="white", position="stacked",
      mapping=aes(fill=p)) +
  facet_wrap(~AgeG, ncol=2) +
  theme_dark() +
  theme(legend.position = "bottom",
        axis.text = element_blank(),
        axis.title = element_blank()
        ) +
  scale_fill_gradient2(high="goldenrod",
                       low="firebrick",
                       na.value="transparent")
```

All the consepts described above also work with the `aseg` atlas for subcrotical structures, except arguments to see only one hemisphere or lateral or medial surfaces, as these arguments do not make sense for this atlas. 
Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas
```{r , size = 'small', out.width='90%'}
aseg
```

```{r atlases, fig.width=5, fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
ggseg(atlas="aseg", mapping=aes(fill=area))
```

## Plotting 3D mesh data

Creating plots of the brain as 2-dimensional can at times make it difficult to properly understand where an atlas segment is located in the true 3-dimensional brain. 
This is particularly true when looking within the brain as subcortical structures or white matter tracts.
To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using [plotly](https://plotly-book.cpsievert.me/).
The data is somewhat more complex than the 2D ggplot polygon version `ggseg`, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
A lot of credit goes to A.M.Winkler and his [Brainder](https://brainder.org/) work, which supplied us with the first examples of going from Freesurfer `.srf`-files to `.ply`-files, and whose scripts massively aided us in making this work.

The function `ggseg3d()`, is based in the plotly, it is recommended to get a little familiarized with with [plotly](https://plotly-book.cpsievert.me/).

Out-of-the-box, `ggseg3d()` works without supplying any extra information, and plots the `dkt_3d` atlas in 'LCBC' surface.
The 'LCBC' surface is the white matter surface, inflated in 10 steps, to increase visibility within folds of the brain, while still maintining curvature to indicate where folds have been.
All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the segments.

```{r ggseg3d_1_out, message=FALSE, echo=F, eval=TRUE, out.width='30%'}
p <- ggseg3d(surface = "white")  %>% 
  layout(title="\nggseg3d(surface='white')",
         font =  list(size = 30))
add_plotly(p, "png/ggseg3d_white.png")

p <- ggseg3d()  %>% 
  layout(title="\nggseg3d(surface='LCBC')",
         font =  list(size = 30))
add_plotly(p, "png/ggseg3d.png")

p <- ggseg3d(surface = "inflated") %>%
  layout(title="\nggseg3d(surface='inflated')",
         font =  list(size = 30))
add_plotly(p, "png/ggseg3d_inflated.png")
```

The cortical atlas data is stored in nested tibbles, which have datasets for three different surfaces and the two hemispheres.
The 'data'-column within includes all necessary information to `ggseg3d()` to create a mesh-plot. 
```{r, out.width='90%', size='small'}
dkt_3d
```

To grab all the data for a surface and hemisphere, you should reduce the data to one line, and then `unnest()`
```{r out.width='40%', linewidth=70}
dkt_3d %>% 
  filter(surf == "inflated" & hemi == "right") %>% 
  unnest(ggseg_3d) %>% 
  select(-lobe, -acronym)
```

### External data supply
Particularly notice the `mesh` column, which is a list column of lists. 
In there is all the 6 vectors needed to create the mesh of the tri-surface plot. 
You'll also need to notice the `label`, `annot` and `area` columns, which are likely the columns you will be matching on when proviging with your own data for colours. 
You need to be meticulous when fixing your data, be sure it matches.
The function _should_ give you a warning if it's struggling to match something.

The column you want to use for colour, needs to be supplied to the `colour` option, and you'll likely want to supply it to the `text` option, as this will add another line to the plotly hover information.

```{r, eval=FALSE, out.width='40%'}
someData = dkt_3d %>% 
  filter(surf == "inflated" & 
         hemi == "right") %>% 
  unnest(ggseg_3d) %>% 
  select(area) %>% 
  na.omit() %>% 
  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>% 
           round(2))

  ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
```

```{r, out.width='40%', echo=FALSE}
someData = dkt_3d %>% 
  filter(surf == "inflated" & hemi == "right") %>% 
  unnest(ggseg_3d) %>% 
  select(area) %>% 
  na.omit() %>% 
  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>% 
           round(2))

p <- ggseg3d(.data = someData, atlas = dkt_3d, 
             colour = "p", text = "p")
add_plotly(p, "png/ggseg3d_data.png")
```

### Colours

You can provide custom colour palettes either in hex or R-names.
Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.

```{r,fig.width=5, eval=FALSE, out.width='45%'}
  ggseg3d(.data = someData, atlas = dkt_3d, 
          colour = "p", text = "p", 
          palette = c("#ff0000", "#00ff00", "#0000ff"))
```
```{r, out.width='40%', echo=F, eval=T}
p <- ggseg3d(.data = someData, atlas = dkt_3d, 
          colour = "p", text = "p", 
          palette = c("#ff0000", "#00ff00", "#0000ff"))
add_plotly(p, "png/ggseg3d_colour.png")
```

If you want to alter the colour of `NA` regions, supply `na.colour`, either as HEX colour or colour name.
This option only takes a single colour.
```{r, out.width='40%', eval=FALSE,  out.width='45%'}
  ggseg3d(.data = someData, atlas = dkt_3d, 
          colour = "p", text = "p", 
          na.colour = "black", camera="medial")
```
```{r, out.width='40%', echo=F, eval=T}
p <-   ggseg3d(.data = someData, atlas = dkt_3d, 
          colour = "p", text = "p", 
          na.colour = "black", camera="medial")
add_plotly(p, "png/ggseg3d_nacolour.png")
```

If you are plotting the sub-cortical structures, you might want to reduce the opacity of the `NA` structures, so that you can see the more medial structures. 
You may also want to add the `glassbrain` option, any value above 0 and up to 1 increases the opacity of a glass brain overlay.
The glassbrain option provides a frame of reference for the subcortical structures, and as such it is recommended to add this when viewing these structures.
There are three glassbrain options, controlling the opacity, which hemispheres to add, and the colour.

```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
somData_aseg = aseg_3d %>% 
  unnest() %>% 
  select(label) %>% 
  filter(!grepl("Ventricle|Putamen", label)) %>% 
  mutate(p = seq(1, nrow(.)))

  ggseg3d(.data = somData_aseg, atlas = aseg_3d, 
          colour = "p", text = "p", 
          na.alpha= .5, glassbrain = .5, 
          glassbrain_hemisphere = "left")
```
```{r, out.width='40%', echo=F, eval=T}
somData_aseg = aseg_3d %>% 
  unnest() %>% 
  select(label) %>% 
  filter(!grepl("Ventricle|Putamen", label)) %>% 
  mutate(p = seq(1, nrow(.)))

p <- ggseg3d(.data = somData_aseg, atlas = aseg_3d, 
          colour = "p", text = "p", 
          na.alpha= .5, glassbrain = .5, 
          glassbrain_hemisphere = "left")
add_plotly(p, "png/ggseg3d_glass.png")
```

There are many plotly options that may help you adapt the plot to look the way you wish. 
In addition to Carson Sievert's [book on plotly in R](https://plotly-book.cpsievert.me/), we recommend resources for [modifying axes in 3D plots](https://plot.ly/r/axes/#modifying-axes-for-3D-plots), the [basic introduction to tri-surface plots](https://plot.ly/r/trisurf/), and this tutorial on [tri-surface plots with plotly in R](https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/).

# Summary
The aim of the `ggseg`-package is to aid researchers easily disseminate analysis results in an easier and more stream-lined fashion. 
This tutorial should provide instructions on how to use the main package functions, and cover the most commonly used options.
The two main functions in the package, `ggseg()` and `ggseg3d()`, create polygon and mesh-plots of brain atlases, using the accompanying package data.
In addition to the two default atlases available in `ggseg`, the `ggsegExtra` package includes `r dd %>% filter(Package == "ggsegExtra") %>% nrow()` more atlases.
We encourage users of this package to contribute more atlases they find uses for, and help build an extensive repository of brain atlases.

### 2019-04-10 draft_1.tex
first draft

%% Submissions for peer-review must enable line-numbering
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

% ZNK -- Adding headers for pandoc

\setlength{\emergencystretch}{3em}
\providecommand{\tightlist}{
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{lipsum}
\usepackage[unicode=true]{hyperref}
\usepackage{longtable}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}

\author[1]{Athanasia M. Mowinckel}

\corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
\author[1]{Didac Vidal Piñeiro}

\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}

%
% \author[1]{First Author}
% \author[2]{Second Author}
% \affil[1]{Address of first author}
% \affil[2]{Address of second author}
% \corrauthor[1]{First Author}{f.author@email.com}

%

\begin{abstract}
The abstract of the article. It can also be on \emph{multiple} lines.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Neuroscientific analysis usually requires the use of multiple tools to analyse, visualise, and summarise data.
This often makes the process of preparing results for publication laborious, as results must be exported and imported in various formats.
In neuroimaging, images of probands' brains are collected and merged together to provide three dimensional representations.
Much of the neuroimaging analyses are not necessarily done on the individual voxel level (3-dimensional pixels), but rather on pre-defined brain segmentations, called brain atlases.
These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (REF).
The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain (REF).
While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (REF), python (REF), or Matlab (REF).

A key part of understanding and disseminating analysis results, is the visualization of these in a meaningful way.
With regards to results from brain atlas analyses, it is most meaningfully represented if projected onto a representation of the brain, rather than other common types of charts (like bar-charts) accompanied by a diagram of the brain label positions.
Each atlas has its own labelling depending on what is meaningful for the type of segmentation it is based on.
As such, for the reader to fully understand a bar chart with atlas labels, they need to be very familiar with the location of each label to have a clear comprehension of the results.
A projection directly onto a brain shape, eases the readability of the results for the reader, and provides clear point of reference even if the used atlas is unfamiliar.

There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (REF), such as ggBrain(REF) and ggneuro(REF)(see \href{https://neuroconductor.org/}{neuroconductor} for compiled neuroimaging packages for R).
These are based on plotting imaging files, not results from analyses of brain atlases.
Here we introduce the ggseg-package for visualizing results from brain atlas analyses.
The ggseg-package was developed to create templates that others might use to project their brain atlas results on to.
While its plotting functions are what the users are drawn to, it is the pre-compiled number of data sets for different brain atlases that provides the real functionality needed for this visualization.
The package also includes functions to plot atlases in tri-surface mesh plots using plotly (REF).

\hypertarget{brain-atlas-selection}{%
\subsection{Brain atlas selection}\label{brain-atlas-selection}}

There is a multitude of brain atlases to chose from, all with different purposes and segmented based on different criteria.
We have initially focused on commonly used and well-established atlases with manageable numbers of segmentations.
The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer (REF); dkt and aseg.
The Desikan-Killany cortical atlas (dkt) (REF) and the Automatic Segmentation of Subcortical Structures (aseg) (REF) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.

While functions and base data can be found in the main \texttt{ggseg}-package, a companion package called \texttt{ggsegExtra} has numerous other atlases, and is continuously expanded.
We encourage the community to help build and expand the atlases available, and to create a repository of established and often used atlases.
A summary of all available atlases may be found in Table \ref{tab:atlasTab}, where in addition to the two default atlases, ggsegExtra has another 12 atlases.
The \href{https://github.com/LCBC-UiO/ggseg/wiki}{ggseg wiki} currently has two different ways you may create and supply atlases to the package as 2D polygons.
Work is being done to include instructions on 3D-mesh plot contributions also.
The package includes functions like \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to aid users in testing and making custom atlases.

\begin{table}[t]

\caption{\label{tab:atlasTab}Table of currently available atlases in either ggseg or ggsegExtra R-package. Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful. }
\centering
\begin{tabular}{lllll}
\toprule
Package & Title & Item & Mesh & Polygon\\
\midrule
 & Desikan-Killiany Cortical Atlas & dkt & No & Yes\\
\cmidrule{2-5}
 & \makecell[l]{Desikan-Killiany Cortical Atlas - tri-\\surf mesh Mesh data for the Desikan-\\Killany Cortical atlas, with 40 regions\\in on the cortical surface of the brain.} & dkt & Yes & No\\
\cmidrule{2-5}
\multirow{-3}{*}{\raggedright\arraybackslash ggseg} & \makecell[l]{Freesurfer automatic subcortical\\segmentation of a brain volume} & aseg & Yes & Yes\\
\cmidrule{1-5}
 & Desterieux cortical parcellations & desterieux & Yes & No\\
\cmidrule{2-5}
 & Harvard-Oxford Cortical atlas & hoCort & No & Yes\\
\cmidrule{2-5}
 & Parcellation from a midsagittal slice & midsagittal & No & Yes\\
\cmidrule{2-5}
 & Parcellation from JHU & jhu & No & Yes\\
\cmidrule{2-5}
 & Parcellation from of JHU & jhu & Yes & No\\
\cmidrule{2-5}
 & Parcellation from of white matter & icbm & Yes & No\\
\cmidrule{2-5}
 & Parcellation from of white matter & tracula & Yes & Yes\\
\cmidrule{2-5}
 & \makecell[l]{Parcellation from the Human Connectome\\Project} & glasser & Yes & Yes\\
\cmidrule{2-5}
 & \makecell[l]{Schaefer 17 Resting-state Cortical\\Parcellations} & schaefer17 & Yes & No\\
\cmidrule{2-5}
 & \makecell[l]{Schaefer 7 Resting-state Cortical\\Parcellations} & schaefer7 & Yes & No\\
\cmidrule{2-5}
 & \makecell[l]{Yeo 17 Resting-state Cortical\\Parcellations} & yeo17 & Yes & Yes\\
\cmidrule{2-5}
\multirow{-12}{*}{\raggedright\arraybackslash ggsegExtra} & \makecell[l]{Yeo 7 Resting-state Cortical\\Parcellations} & yeo7 & Yes & Yes\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{tutorial}{%
\section{Tutorial}\label{tutorial}}

The aim of this tutorial is to familiarize the reader with the main package functions and the general use of the package.\\
It will cover the two main functions: \texttt{ggseg()} which plots 2D polygons, and \texttt{ggseg3d()} which plots tri-surface mesh plots.

\hypertarget{plotting-polygon-data}{%
\subsection{Plotting polygon data}\label{plotting-polygon-data}}

The function to plot two-dimensional polygon data is the \texttt{ggseg}-function.
This function will automatically plot the dkt-atlas by default (see Figure \ref{fig:init}).
As this function is a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2} it can be built upon like all ggplot objects.
The plot is kept as simple as possible, with as little extra information to ggplot as possible.
This is done for it to be easy to extend the plot with the many ggplot options available.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(ggseg)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{ggseg}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{draft_1_files/figure-latex/init-1.pdf}
\caption{\label{fig:init}By default \texttt{ggseg} will plot the dkt atlas in shaded polygons.}
\end{figure}

Because of the way \texttt{ggseg} is implemented, you should be able to safely use the function together with other ggplot functions, like themes, scales, and such (see Figure \ref{fig:void}).
There are many extra options to ggseg, to alter the plot as you wish.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"dark theme"}\NormalTok{)}
\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"classic theme"}\NormalTok{)}

\CommentTok{# Combine plots}
\NormalTok{pp <-}\StringTok{ }\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(p1, p2, }\DataTypeTok{nrow =} \DecValTok{2}\NormalTok{)}

\NormalTok{p3 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"default theme"}\NormalTok{)}

\CommentTok{# Create a subplot grid}
\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(pp, p3, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{draft_1_files/figure-latex/void-1.pdf}
\caption{\label{fig:void}ggseg plots can be used with most standard scales, themes and such that work with ggplot. The special ggseg options for hemispheres, view etc. depend on the atlas used, and some options are only available for certain atlases. There is no `lateral' or `medial' views of subcortical atlases.}
\end{figure}

You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these (see Figure \ref{fig:stacked}).
The possibilities of altering the plot depends on the atlas used.
There are generally more options available for cortical atlases than subcortical ones, as the atlases have more meaningful ways of viewing the polygons.
Subcortical atlases also tend to have structures that span the hemispheres or other typically used referece points of the brain, and as such it is more difficult to label all segments as either, e.g., left or right.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lat <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Medial view"}\NormalTok{)}
\NormalTok{left <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Left hemi."}\NormalTok{)}
\NormalTok{pp <-}\StringTok{ }\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(lat, left, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{)}

\NormalTok{combo <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
      \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Combination"}\NormalTok{)}

\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(pp, combo, }\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{draft_1_files/figure-latex/stacked-1.pdf}
\caption{\label{fig:stacked}There are several ggseg-special options that may be supplied to control how the plots looks and is organised. Top left: plot using medial view. Bottom left: plot using left hemisphere. Right: plot combining both the options.}
\end{figure}

\hypertarget{using-own-data-with-fill-and-colour}{%
\subsubsection{Using own data with fill and colour}\label{using-own-data-with-fill-and-colour}}

\texttt{ggseg} accepts any argument you can supply to \texttt{geom\_polygon} and therefore is easy to work with for those familiar with ggplot functionality.
Standard arguments like \texttt{fill} to flood the segments with a colour, or \texttt{colour} to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option \texttt{aes}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
        \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{draft_1_files/figure-latex/fill-1.pdf}
\caption{\label{fig:fill}Supplying `area' to the fill option in ggseg, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.}
\end{figure}

Typically you would want each area to have a different colour, corresponding to some descriptive or inference statistic that has been computed.
The bulk of the package originates not in its functions, but in the datasets that accompany the package.
There is one dataset per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
Before setting up the data from the statistics to project onto the segments, looking at the atlas datasets might help understand what the data needs to look like, or avoid looking like.

\small

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dkt}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 80 x 6
##    atlas area               hemi  side    label             ggseg          
##    <chr> <chr>              <chr> <chr>   <chr>             <list>         
##  1 dkt   superior temporal  left  lateral lh_superiortempo~ <tibble [1,494~
##  2 dkt   pre central        left  lateral lh_precentral     <tibble [1,314~
##  3 dkt   post central       left  lateral lh_postcentral    <tibble [1,164~
##  4 dkt   rostral middle fr~ left  lateral lh_rostralmiddle~ <tibble [1,194~
##  5 dkt   insula             left  lateral lh_insula         <tibble [870 x~
##  6 dkt   superior parietal  left  lateral lh_superiorparie~ <tibble [480 x~
##  7 dkt   inferior temporal  left  lateral lh_inferiortempo~ <tibble [906 x~
##  8 dkt   lateral occipital  left  lateral lh_lateraloccipi~ <tibble [738 x~
##  9 dkt   lateral orbitofro~ left  lateral lh_lateralorbito~ <tibble [528 x~
## 10 dkt   superior frontal   left  lateral lh_superiorfront~ <tibble [420 x~
## # ... with 70 more rows
\end{verbatim}

\normalsize

The column `label' was added to the atlas for easy matching with the output from Freesurfer statistics tables.
The data is in long-format, meaning each segment has its own row, meaning the data you wish to project onto the segments also need to be in long-format.
As data directly from Freesurfer comes in wide-format, this requires some extra wranling.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{freesurfer_stats <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{id =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\OperatorTok{:}\DecValTok{12}\NormalTok{),}
  \DataTypeTok{lh_superiortemporal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.32}\NormalTok{, }\FloatTok{4.1}\NormalTok{, }\FloatTok{3.5}\NormalTok{),}
  \DataTypeTok{lh_precentral =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.3}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.1}\NormalTok{),}
  \DataTypeTok{lh_rostralmiddlefrontal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.3}\NormalTok{, }\FloatTok{3.2}\NormalTok{, }\FloatTok{3.1}\NormalTok{)}
\NormalTok{  )}
\NormalTok{freesurfer_stats}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   id lh_superiortemporal lh_precentral lh_rostralmiddlefrontal
## 1 10                3.32           2.3                     3.3
## 2 11                4.10           2.5                     3.2
## 3 12                3.50           2.1                     3.1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{-}\NormalTok{id)}
\NormalTok{freesurfer_long}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   id                   label thickness
## 1 10     lh_superiortemporal      3.32
## 2 11     lh_superiortemporal      4.10
## 3 12     lh_superiortemporal      3.50
## 4 10           lh_precentral      2.30
## 5 11           lh_precentral      2.50
## 6 12           lh_precentral      2.10
## 7 10 lh_rostralmiddlefrontal      3.30
## 8 11 lh_rostralmiddlefrontal      3.20
## 9 12 lh_rostralmiddlefrontal      3.10
\end{verbatim}

This data can now be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
The data \textbf{must} include a column that has the same name at atleast \emph{some} data matching the values in the corresponding column in the atlas data.
In the next example we create some data with 4 rows, and an `area' and `p' column.
The \texttt{ggseg} function will recognise the matching column `area', and merge the supplied data into the atlas using \texttt{dplyr} joins.
We use the \texttt{p} column as the column flooding the segment with colour.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{area =} \KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
           \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{), }
  \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{4}\NormalTok{),}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{p <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))}
\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft_1_files/figure-latex/datasupp-1.pdf}

We can also change how this looks, by various ggplot functions, like scales, labs, themes and such.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_void}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p-value"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft_1_files/figure-latex/datasupp2-1.pdf}

If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column \texttt{hemi} with either ``right'' or ``left'' for this to happen.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{someData}\OperatorTok{$}\NormalTok{hemi =}\StringTok{ "right"}

\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{,}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_void}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"Another nice plot title"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft_1_files/figure-latex/datasuppX-1.pdf}

\hypertarget{creating-subplots}{%
\subsubsection{Creating subplots}\label{creating-subplots}}

If your data has different statistics for different groups, and you want to plot them all, we can use ggplot's \texttt{facet\_wrap} or \texttt{facet\_grid} for that.
The data needs to be in long format for this to work, as needed by ggplot, meaning the data from the groups should appear in seperate rows in the data, not in separate columns. Secondly, the data needs to be grouped using \texttt{dplyr}'s \texttt{group\_by} function \emph{before} providing the data to the \texttt{ggseg} function.
The ggseg-function will detect if the data is grouped, and perform the necessary steps for the \texttt{facet}'s to work when they are applied later.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{area =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
               \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}\DecValTok{2}\NormalTok{), }
  \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{8}\NormalTok{),}
  \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{)),}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(AgeG)}

\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft_1_files/figure-latex/datasupp3-1.pdf}

Depending on what type of column is used for fill (continuous or discrete),
you can also use the differenct \texttt{scales} to alter the colours.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas=}\NormalTok{dkt, }
      \DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"stacked"}\NormalTok{,}
      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{,}
        \DataTypeTok{axis.text =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.title =} \KeywordTok{element_blank}\NormalTok{()}
\NormalTok{        ) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient2}\NormalTok{(}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{,}
                       \DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}
                       \DataTypeTok{na.value=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft_1_files/figure-latex/datasupp5-1.pdf}

All the consepts described above also work with the \texttt{aseg} atlas for subcrotical structures, except arguments to see only one hemisphere or lateral or medial surfaces, as these arguments do not make sense for this atlas.
Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas
\small

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aseg}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 34 x 6
##    atlas area            hemi  side  label                ggseg            
##    <chr> <chr>           <chr> <chr> <fct>                <list>           
##  1 aseg  <NA>            right axial <NA>                 <tibble [808 x 5~
##  2 aseg  <NA>            left  axial <NA>                 <tibble [1,014 x~
##  3 aseg  thalamus proper left  axial Left-Thalamus-Proper <tibble [56 x 5]>
##  4 aseg  thalamus proper left  axial Right-Thalamus-Prop~ <tibble [56 x 5]>
##  5 aseg  thalamus proper right axial Left-Thalamus-Proper <tibble [61 x 5]>
##  6 aseg  thalamus proper right axial Right-Thalamus-Prop~ <tibble [61 x 5]>
##  7 aseg  lateral ventri~ right axial Left-Lateral-Ventri~ <tibble [66 x 5]>
##  8 aseg  lateral ventri~ right axial Right-Lateral-Ventr~ <tibble [66 x 5]>
##  9 aseg  hippocampus     left  axial Left-Hippocampus     <tibble [47 x 5]>
## 10 aseg  hippocampus     left  axial Right-Hippocampus    <tibble [47 x 5]>
## # ... with 24 more rows
\end{verbatim}

\normalsize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area))}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{draft_1_files/figure-latex/atlases-1.pdf}
\caption{\label{fig:atlases}The first 10 rows of the \texttt{aseg} dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way.}
\end{figure}

\hypertarget{plotting-3d-mesh-data}{%
\subsection{Plotting 3D mesh data}\label{plotting-3d-mesh-data}}

Creating plots of the brain as 2-dimensional can at times make it difficult to properly understand where an atlas segment is located in the true 3-dimensional brain.
This is particularly true when looking within the brain as subcortical structures or white matter tracts.
To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using \href{https://plotly-book.cpsievert.me/}{plotly}.
The data is somewhat more complex than the 2D ggplot polygon version \texttt{ggseg}, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
A lot of credit goes to A.M.Winkler and his \href{https://brainder.org/}{Brainder} work, which supplied us with the first examples of going from Freesurfer \texttt{.srf}-files to \texttt{.ply}-files, and whose scripts massively aided us in making this work.

The function \texttt{ggseg3d()}, is based in the plotly, it is recommended to get a little familiarized with with \href{https://plotly-book.cpsievert.me/}{plotly}.

Out-of-the-box, \texttt{ggseg3d()} works without supplying any extra information, and plots the \texttt{dkt\_3d} atlas in `LCBC' surface.
The `LCBC' surface is the white matter surface, inflated in 10 steps, to increase visibility within folds of the brain, while still maintining curvature to indicate where folds have been.
All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the segments.

\includegraphics[width=0.3\linewidth]{png/ggseg3d_white}
\includegraphics[width=0.3\linewidth]{png/ggseg3d}
\includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated}

The cortical atlas data is stored in nested tibbles, which have datasets for three different surfaces and the two hemispheres.
The `data'-column within includes all necessary information to \texttt{ggseg3d()} to create a mesh-plot.
\small

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dkt_3d}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 6 x 4
##   atlas  surf     hemi  ggseg_3d         
##   <chr>  <chr>    <chr> <list>           
## 1 dkt_3d inflated left  <tibble [36 x 8]>
## 2 dkt_3d inflated right <tibble [36 x 8]>
## 3 dkt_3d LCBC     left  <tibble [36 x 8]>
## 4 dkt_3d LCBC     right <tibble [36 x 8]>
## 5 dkt_3d white    left  <tibble [36 x 8]>
## 6 dkt_3d white    right <tibble [36 x 8]>
\end{verbatim}

\normalsize

To grab all the data for a surface and hemisphere, you should reduce the data to one line, and then \texttt{unnest()}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dkt_3d }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{lobe, }\OperatorTok{-}\NormalTok{acronym)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 36 x 9
## atlas surf hemi area colour mesh label roi annot
## <chr> <chr> <chr> <chr> <chr> <list> <chr> <chr> <chr>
## 1 dkt_3d infla~ right <NA> <NA> <list~ rh_medial~ 0001 medialwa~
## 2 dkt_3d infla~ right banks supe~ #1964~ <list~ rh_bankss~ 0002
bankssts
## 3 dkt_3d infla~ right caudal ant~ #7D64~ <list~ rh_caudal~ 0003
caudalan~
## 4 dkt_3d infla~ right caudal mid~ #6419~ <list~ rh_caudal~ 0004
caudalmi~
## 5 dkt_3d infla~ right corpus cal~ <NA> <list~ rh_corpus~ 0005
corpusca~
## 6 dkt_3d infla~ right cuneus #DC14~ <list~ rh_cuneus 0006 cuneus
## 7 dkt_3d infla~ right entorhinal #DC14~ <list~ rh_entorh~ 0007
entorhin~
## 8 dkt_3d infla~ right fusiform #B4DC~ <list~ rh_fusifo~ 0008
fusiform
## 9 dkt_3d infla~ right inferior p~ #DC3C~ <list~ rh_inferi~ 0009
inferior~
## 10 dkt_3d infla~ right inferior t~ #B428~ <list~ rh_inferi~ 0010
inferior~
## # ... with 26 more rows
\end{verbatim}

\hypertarget{external-data-supply}{%
\subsubsection{External data supply}\label{external-data-supply}}

Particularly notice the \texttt{mesh} column, which is a list column of lists.
In there is all the 6 vectors needed to create the mesh of the tri-surface plot.
You'll also need to notice the \texttt{label}, \texttt{annot} and \texttt{area} columns, which are likely the columns you will be matching on when proviging with your own data for colours.
You need to be meticulous when fixing your data, be sure it matches.
The function \emph{should} give you a warning if it's struggling to match something.

The column you want to use for colour, needs to be supplied to the \texttt{colour} option, and you'll likely want to supply it to the \texttt{text} option, as this will add another line to the plotly hover information.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{someData =}\StringTok{ }\NormalTok{dkt_3d }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }
\StringTok{         }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(area) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{na.omit}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{, }\DataTypeTok{length.out =} \DecValTok{100}\NormalTok{ ), }\KeywordTok{nrow}\NormalTok{(.)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{           }\KeywordTok{round}\NormalTok{(}\DecValTok{2}\NormalTok{)) }

  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }\DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=0.4\linewidth]{png/ggseg3d_data}

\hypertarget{colours}{%
\subsubsection{Colours}\label{colours}}

You can provide custom colour palettes either in hex or R-names.
Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.

\begin{Shaded}
\begin{Highlighting}[]
  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }
          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
          \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"#ff0000"}\NormalTok{, }\StringTok{"#00ff00"}\NormalTok{, }\StringTok{"#0000ff"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=0.4\linewidth]{png/ggseg3d_colour}

If you want to alter the colour of \texttt{NA} regions, supply \texttt{na.colour}, either as HEX colour or colour name.
This option only takes a single colour.

\begin{Shaded}
\begin{Highlighting}[]
  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }
          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
          \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{camera=}\StringTok{"medial"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=0.4\linewidth]{png/ggseg3d_nacolour}

If you are plotting the sub-cortical structures, you might want to reduce the opacity of the \texttt{NA} structures, so that you can see the more medial structures.
You may also want to add the \texttt{glassbrain} option, any value above 0 and up to 1 increases the opacity of a glass brain overlay.
The glassbrain option provides a frame of reference for the subcortical structures, and as such it is recommended to add this when viewing these structures.
There are three glassbrain options, controlling the opacity, which hemispheres to add, and the colour.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{somData_aseg =}\StringTok{ }\NormalTok{aseg_3d }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(label) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"Ventricle|Putamen"}\NormalTok{, label)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{nrow}\NormalTok{(.))) }

  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ somData_aseg, }\DataTypeTok{atlas =}\NormalTok{ aseg_3d, }
          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{, }\DataTypeTok{glassbrain =} \FloatTok{.5}\NormalTok{, }
          \DataTypeTok{glassbrain_hemisphere =} \StringTok{"left"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=0.4\linewidth]{png/ggseg3d_glass}

There are many plotly options that may help you adapt the plot to look the way you wish.
In addition to Carson Sievert's \href{https://plotly-book.cpsievert.me/}{book on plotly in R}, we recommend resources for \href{https://plot.ly/r/axes/\#modifying-axes-for-3D-plots}{modifying axes in 3D plots}, the \href{https://plot.ly/r/trisurf/}{basic introduction to tri-surface plots}, and this tutorial on \href{https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/}{tri-surface plots with plotly in R}.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

The aim of the \texttt{ggseg}-package is to aid researchers easily disseminate analysis results in an easier and more stream-lined fashion.
This tutorial should provide instructions on how to use the main package functions, and cover the most commonly used options.
The two main functions in the package, \texttt{ggseg()} and \texttt{ggseg3d()}, create polygon and mesh-plots of brain atlases, using the accompanying package data.
In addition to the two default atlases available in \texttt{ggseg}, the \texttt{ggsegExtra} package includes 12 more atlases.
We encourage users of this package to contribute more atlases they find uses for, and help build an extensive repository of brain atlases.

\end{document}

### 2019-04-12 draft_1_DVP.Rmd
Update and rename draft_1.Rmd to draft_1_DVP.Rmd

--- 2019-04-10 draft_1.Rmd
@@
 # Tutorial
-The aim of this tutorial is to familiarize the reader with the main package functions and the general use of the package.  
-It will cover the two main functions: `ggseg()` which plots 2D polygons,  and `ggseg3d()` which plots tri-surface mesh plots.
+This tutorial will introduce the `ggseg` and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the package. 
+The tutorial will cover the following functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
 
-## Plotting polygon data
+## Plotting 2D polygon data with `ggseg`
 
-The function to plot two-dimensional polygon data is the `ggseg`-function. 
-This function will automatically plot the dkt-atlas by default (see Figure \@ref(fig:init)). 
-As this function is a wrapper for `geom_polygon` from `ggplot2` it can be built upon like all ggplot objects. 
-The plot is kept as simple as possible, with as little extra information to ggplot as possible. 
-This is done for it to be easy to extend the plot with the many ggplot options available.
+`ggseg` is the main function for plotting 2D data.  
+By default, the function automatically plots the DKT atlas (see Figure \@ref(fig:init)). 
+`ggseg` function is a wrapper for `geom_polygon` from `ggplot2` and it can be built upon and combined like any ggplot object. 
+The image plot consists of a simple brain representation containing no extra information. 
+Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options.
 
@@
 
-Because of the way `ggseg` is implemented, you should be able to safely use the function together with other ggplot functions, like themes, scales, and such (see Figure \@ref(fig:void)). 
-There are many extra options to ggseg, to alter the plot as you wish.
+[comment]: <> (Figure 2 and 3 can be merged together. We can then use the figure legend to specify a little bit more. Not modified depending on your opinion)
+`ggseg` is implemented as a fully-functional part of ggplot. 
+Thus one should be able to use the function together with any other ggplot function such as themes or scales. 
+See Figure \@ref(fig:void) for an example.
 
@@
 
-You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these (see Figure \@ref(fig:stacked)).
-The possibilities of altering the plot depends on the atlas used. 
-There are generally more options available for cortical atlases than subcortical ones, as the atlases have more meaningful ways of viewing the polygons.
-Subcortical atlases also tend to have structures that span the hemispheres or other typically used referece points of the brain, and as such it is more difficult to label all segments as either, e.g., left or right.
+One have several options for plotting the main brain representations.
+These options are atlas-specific. 
+For cortical atlases, such as the DKT, one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \@ref(fig:stacked)).
+For subcortical atlases, such as the aseg, the options are more limited but one can often choose between axial, sagital, and coronal views.
 

### 2019-04-12 draft_1_DVP.Rmd
Update draft_1_DVP.Rmd

@@
 
-```{r void, fig.width=5, fig.cap="ggseg plots can be used with most standard scales, themes and such that work with ggplot.
-The special ggseg options for hemispheres, view etc.
-depend on the atlas used, and some options are only available for certain atlases.
-There is no 'lateral' or 'medial' views of subcortical atlases.
+```{r void, fig.width=5, fig.cap="ggseg plots can be used together with most standard scales, themes and such that work with ggplot.
+The additional ggseg options for hemispheres, view etc.
+are atlas-specific.
+For example 'lateral' or 'medial' views are only available for cortical atlases such as the DKT.
 "}
@@
 
-One have several options for plotting the main brain representations.
+One has several options for plotting the main brain representations.
 These options are atlas-specific. 
@@
 
-```{r stacked, fig.width=5, fig.cap="There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
+```{r stacked, fig.width=5, fig.cap="One can control the organization and the global visualization of a ggseg plot with gggseg-specific options such as hemisphere or view.
 Top left: plot using medial view.
@@
 Standard arguments like `fill` to flood the segments with a colour, or `colour` to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales. See an example in Figure \@ref(fig:fill).
 
-```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas."}
+```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The scale_fill_brain("dkt") option will colour the plot according the original colors used in FreeSurfer."}
 ggseg(mapping=aes(fill = area), colour="black") + 
@@
 
-Typically you would want each area to have a different colour, corresponding to some descriptive or inference statistic that has been computed.
+[comment]: <> (Starting here I do not comment, as maybe re-structuring is better; waiting for input)
+Most users will use ggseg to display through a color scale some descriptive or inferential statistics such as mean cortical thickness or thickness – cognition relationship across the different brain regions.
 The bulk of the package originates not in its functions, but in the datasets that accompany the package. 
@@
 
-The column 'label' was added to the atlas for easy matching with the output from Freesurfer statistics tables. 
-The data is in long-format, meaning each segment has its own row, meaning the data you wish to project onto the segments also need to be in long-format. 
-As data directly from Freesurfer comes in wide-format, this requires some extra wranling.
+[comment]: <> (End of no comments, as maybe re-structuring is better; waiting for input)
+In any atlas, the column ‘label’ is particularly useful for combining the data of interest with the ‘ggseg’ polygons. 
+The column ‘label’ contains the label (region) names as in the original neuroimaging software. 
+For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
+Yet the data in ‘ggseg’ is in a long format - that is each region has its own row - thus data of interest needs to be converted into this format. 
+Note that most datasets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column. 
+See below an example of wide-to-long conversion.
 
@@
 
-This data can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
-The data **must** include a column that has the same name at atleast _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an 'area' and 'p' column. 
+This data - in long format - can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
+The data **must** include a column that has the same name at least _some_ data matching the values in the corresponding column in the atlas data.
+In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of an analysis. 
 The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins. 
@@
 
-We can also change how this looks, by various ggplot functions, like scales, labs, themes and such. 
+We can then modify the appearance similarly to any other ggplot graph using ggplot functions such as scales, labs, themes, etc. 
 ```{r datasupp2, fig.width=5}
@@
 ### Creating subplots
-If your data has different statistics for different groups, and you want to plot them all, we can use ggplot's `facet_wrap` or `facet_grid` for that.
-The data needs to be in long format for this to work, as needed by ggplot, meaning the data from the groups should appear in seperate rows in the data, not in separate columns.
+One can plot a statistic of interest in different groups (i.e. thickness or significance in young or older adults) using ggplot's `facet_wrap` or `facet_grid`.
+As before, data needs to be in long format (group data should appear in seperate rows, not in separate columns). 
 Secondly, the data needs to be grouped using `dplyr`'s `group_by` function _before_ providing the data to the `ggseg` function. 
-The ggseg-function will detect if the data is grouped, and perform the necessary steps for the `facet`'s to work when they are applied later.
+The ggseg function will detect grouped data, and adapt it to `facet`'s requirements.
 
@@
   p = sample(seq(0,.5,.001), 8),
-  AgeG = c(rep("Young",4), rep("Old",4)),
+  AgeG = c(rep("Young",4), rep("Older",4)),
   stringsAsFactors = FALSE) %>% 
@@
 
-Depending on what type of column is used for fill (continuous or discrete),
-you can also use the differenct `scales` to alter the colours.
+As before, one can apply the different ggplot and facet options, such as `scales` to modify the plot's appearance.
 
@@
 
-All the consepts described above also work with the `aseg` atlas for subcrotical structures, except arguments to see only one hemisphere or lateral or medial surfaces, as these arguments do not make sense for this atlas.
+All the concepts described above also work with the `aseg` atlas for subcortical structures, except hemisphere and view arguments that do not apply in subcortical atlases such as aseg. 
 Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas
@@
 
-Creating plots of the brain as 2-dimensional can at times make it difficult to properly understand where an atlas segment is located in the true 3-dimensional brain. 
-This is particularly true when looking within the brain as subcortical structures or white matter tracts.
-To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using [plotly](https://plotly-book.cpsievert.me/).
-The data is somewhat more complex than the 2D ggplot polygon version `ggseg`, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
-Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
-A lot of credit goes to A.M.Winkler and his [Brainder](https://brainder.org/) work, which supplied us with the first examples of going from Freesurfer `.srf`-files to `.ply`-files, and whose scripts massively aided us in making this work.
-
-The function `ggseg3d()`, is based in the plotly, it is recommended to get a little familiarized with with [plotly](https://plotly-book.cpsievert.me/).
+Representing brains as 2D polygons is a good solution for fast, efficient and flexible plotting, based on ggplot (a powerful tool for visualization) and that can be easily combined with interactive apps such as in shiny.
+Yet brains are intrinsically 3D and it can be challenging to recognize the location of a region.
+This problem exacerbated in atlases representing subcortical structures – after all, cortical surfaces are 2D – such as grey matter structures or white matter tracts.
+Hence, here we also provide the ggseg3D function to plot, view and print 3D atlases in R. 
+ggseg3D is based on tri-surface mesh plots using plotly. 
+The data structure is somewhat more complex than the 2D ggplot polygon version and includes additional options for brain inflation, glass brains, camera locations, etc.  
+As ggseg3D is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination. 
+We recommend users to get a little familiarized with with [plotly](https://plotly-book.cpsievert.me/).
 

### 2019-04-12 draft_1_DVP.Rmd
Update draft_1_DVP.Rmd

@@
 The image plot consists of a simple brain representation containing no extra information. 
-Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options.
+Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options. We recommend users to get familiarized with [ggplot2](https://ggplot2.tidyverse.org/) grammar.
 
@@
 Yet brains are intrinsically 3D and it can be challenging to recognize the location of a region.
-This problem exacerbated in atlases representing subcortical structures – after all, cortical surfaces are 2D – such as grey matter structures or white matter tracts.
+This problem exacerbates in atlases that represent subcortical features – after all, cortical surfaces are 2D – such as grey matter structures or white matter tracts.
 Hence, here we also provide the ggseg3D function to plot, view and print 3D atlases in R. 
@@
 As ggseg3D is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination. 
-We recommend users to get a little familiarized with with [plotly](https://plotly-book.cpsievert.me/).
+We recommend users to familiarize with [plotly](https://plotly-book.cpsievert.me/).
 

### 2019-04-15 draft_1_DVP.Rmd
Update draft_1_DVP.Rmd

@@
 The image plot consists of a simple brain representation containing no extra information. 
-Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options. We recommend users to get familiarized with [ggplot2](https://ggplot2.tidyverse.org/) grammar.
+Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options. 
+We recommend users to get familiarized with [ggplot2](https://ggplot2.tidyverse.org/) grammar.
 
@@
 
-Out-of-the-box, `ggseg3d()` works without supplying any extra information, and plots the `dkt_3d` atlas in 'LCBC' surface.
-The 'LCBC' surface is the white matter surface, inflated in 10 steps, to increase visibility within folds of the brain, while still maintining curvature to indicate where folds have been.
-All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the segments.
+Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface.
+The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage_ template subject. See *SCRIPT* for reproducibility.
+All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the colorlut used in the neuroimaging sofware.
 
@@
 
-The cortical atlas data is stored in nested tibbles, which have datasets for three different surfaces and the two hemispheres.
-The 'data'-column within includes all necessary information to `ggseg3d()` to create a mesh-plot. 
+The cortical atlas data is stored in nested tibbles. Each atlas have datasets for three different surfaces (see Figure X) and the two hemispheres.
+The 'ggseg3d' column includes all necessary information to `ggseg3d()` to create a mesh-plot. 
+The additional `[..._3d]` atlases in the ggsegExtra contain identical structures except for the subcortical atlases in which surface and hemisphere do not apply.
+
 ```{r, out.width='90%', size='small'}
@@
 
-To grab all the data for a surface and hemisphere, you should reduce the data to one line, and then `unnest()`
+### External data supply
+Similarly as in the 2D-atlas, the user will use ggseg3D to display through a color scale some descriptive or inferential statistics.
+First, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest()` it.
+
 ```{r out.width='40%', linewidth=70}
@@
 
-### External data supply
-Particularly notice the `mesh` column, which is a list column of lists. 
-In there is all the 6 vectors needed to create the mesh of the tri-surface plot. 
-You'll also need to notice the `label`, `annot` and `area` columns, which are likely the columns you will be matching on when proviging with your own data for colours. 
-You need to be meticulous when fixing your data, be sure it matches.
-The function _should_ give you a warning if it's struggling to match something.
+Note the `mesh` column, which contains lists.
+Each list corresponds to a regio and contains 6 vectors required to create the mesh of the tri-surface plot.
+
+Note als the `label`, `annot` and `area` columns which provide matching values for your own data. 
+Similarly to ggseg function, the label column should match the region names used in the original neuroimaging software while area and annot provide alternative/ secundary names. 
+It is thus important to match your identifiers for region with those used in the atlas. 
+A warning will be issued in case of any mismatch.
 
@@
 
-If you are plotting the sub-cortical structures, you might want to reduce the opacity of the `NA` structures, so that you can see the more medial structures. 
-You may also want to add the `glassbrain` option, any value above 0 and up to 1 increases the opacity of a glass brain overlay.
-The glassbrain option provides a frame of reference for the subcortical structures, and as such it is recommended to add this when viewing these structures.
-There are three glassbrain options, controlling the opacity, which hemispheres to add, and the colour.
+Subcortical atlases include cortical surfaces and other landmark structures to visualization purposes only. 
+One can control the opacity of the these `NA` structures, to improve visualization. 
+Additionaly, one can add the `glassbrain` option, which provides a frame of reference for the subcortical structures.
+Glassbrain can be controlled with thre different arguments: opacity (0-1), hemisphere (left, right), and colour.
 
@@
 
-There are many plotly options that may help you adapt the plot to look the way you wish. 
+ggsed3d() are based on plotly and thus can be additional plotly functionalities will can be used to modify and improve the 3D atlas representations. 
 In addition to Carson Sievert's [book on plotly in R](https://plotly-book.cpsievert.me/), we recommend resources for [modifying axes in 3D plots](https://plot.ly/r/axes/#modifying-axes-for-3D-plots), the [basic introduction to tri-surface plots](https://plot.ly/r/trisurf/), and this tutorial on [tri-surface plots with plotly in R](https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/).
+Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line to save ggseg3d atlas snapshots.
 

### 2019-04-15 draft_1.tex
moved some files

@@
 % 
+\usepackage{natbib}
+\bibliographystyle{plainnat}
 
@@
 Much of the neuroimaging analyses are not necessarily done on the individual voxel level (3-dimensional pixels), but rather on pre-defined brain segmentations, called brain atlases.
-These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (REF).
-The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain (REF).
-While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (REF), python (REF), or Matlab (REF).
+These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (\citet{dkt}).
+The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain.
+While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (\citet{R}), python (\citet{python}), or MATLAB (\citet{MATLAB}).
 
@@
 
-There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (REF), such as ggBrain(REF) and ggneuro(REF)(see \href{https://neuroconductor.org/}{neuroconductor} for compiled neuroimaging packages for R).
+There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (\citet{ggplot}), such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
 These are based on plotting imaging files, not results from analyses of brain atlases.
@@
 While its plotting functions are what the users are drawn to, it is the pre-compiled number of data sets for different brain atlases that provides the real functionality needed for this visualization.
-The package also includes functions to plot atlases in tri-surface mesh plots using plotly (REF).
+The package also includes functions to plot atlases in tri-surface mesh plots using plotly (\citet{plotly}).
 
@@
 We have initially focused on commonly used and well-established atlases with manageable numbers of segmentations.
-The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer (REF); dkt and aseg.
-The Desikan-Killany cortical atlas (dkt) (REF) and the Automatic Segmentation of Subcortical Structures (aseg) (REF) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.
+The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer \citep{fischl:99}; dkt and aseg.
+The Desikan-Killany cortical atlas (dkt; \citet{dkt}) and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.
 
@@
 This is particularly true when looking within the brain as subcortical structures or white matter tracts.
-To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using \href{https://plotly-book.cpsievert.me/}{plotly}.
+To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using plotly \citep{plotly}.
 The data is somewhat more complex than the 2D ggplot polygon version \texttt{ggseg}, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
 Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
-A lot of credit goes to A.M.Winkler and his \href{https://brainder.org/}{Brainder} work, which supplied us with the first examples of going from Freesurfer \texttt{.srf}-files to \texttt{.ply}-files, and whose scripts massively aided us in making this work.
+A lot of credit goes to A.M.Winkler and his Brainder (\citet{brainder}) work, which supplied us with the first examples of going from Freesurfer \texttt{.srf}-files to \texttt{.ply}-files, and whose scripts massively aided us in making this work.
 
-The function \texttt{ggseg3d()}, is based in the plotly, it is recommended to get a little familiarized with with \href{https://plotly-book.cpsievert.me/}{plotly}.
+The function \texttt{ggseg3d()}, is based in the plotly, it is recommended to get a little familiarized with with plotly (\citet{plotly}).
 
@@
 There are many plotly options that may help you adapt the plot to look the way you wish.
-In addition to Carson Sievert's \href{https://plotly-book.cpsievert.me/}{book on plotly in R}, we recommend resources for \href{https://plot.ly/r/axes/\#modifying-axes-for-3D-plots}{modifying axes in 3D plots}, the \href{https://plot.ly/r/trisurf/}{basic introduction to tri-surface plots}, and this tutorial on \href{https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/}{tri-surface plots with plotly in R}.
+In addition to Carson Sievert's book on plotly in R (\citeyearpar{plotly}) we recommend resources for modifying axes in 3D plots (\citet{plotly-ax}), the basic introduction to tri-surface plots (\citet{plotly-tri}), and this tutorial on tri-surface plots with plotly in R (\citet{plotly-trisurf}).
 
@@
 
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+Didac Vidal-Piñeiro generated the idea for the plotting tool, and the initial scripts that made the plots.
+He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
+Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and making the functions more user-friendly.
+She also was responsible for adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible.
+A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+We also thanks John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (\citeyearpar{neuroconductor}), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
+
+\hypertarget{funding}{%
+\section{Funding}\label{funding}}
+
+This work is funded by EU Horizon 2020 Grant `Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+
+\hypertarget{prior-versions}{%
+\section{Prior versions}\label{prior-versions}}
+
+Some of the content presented here also apprears in the package vignette of \texttt{ggseg}, which may be accessed through R or in the package website (\citet{ggseg}).
+Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creationg and functionality (\citet{ggsegAnim}, \citet{ggsegIntro}).
+
+\renewcommand\refname{References}
+\bibliography{references.bib}
+
 \end{document}

### 2019-04-15 draft_1_AMM.rmd
moved some files

--- 2019-04-10 draft_1.Rmd
@@
 - \usepackage{xcolor}
-bibliography: sample.bib
+bibliography: references.bib
 output:
@@
     base_format: rticles::peerj_article # for using bookdown features like \@ref()
+    citation_package: natbib
     fig_caption: true
@@
 Much of the neuroimaging analyses are not necessarily done on the individual voxel level (3-dimensional pixels), but rather on pre-defined brain segmentations, called brain atlases.
-These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (REF). 
-The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain (REF). 
-While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (REF), python (REF), or Matlab (REF).
+These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (@dkt). 
+The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain. 
+While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (@R), python (@python), or MATLAB (@MATLAB).
 
@@
 
-There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (REF), such as ggBrain(REF) and ggneuro(REF)(see [neuroconductor](https://neuroconductor.org/) for compiled neuroimaging packages for R).
+There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (@ggplot), such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
 These are based on plotting imaging files, not results from analyses of brain atlases. 
@@
 While its plotting functions are what the users are drawn to, it is the pre-compiled number of data sets for different brain atlases that provides the real functionality needed for this visualization.
-The package also includes functions to plot atlases in tri-surface mesh plots using plotly (REF).
+The package also includes functions to plot atlases in tri-surface mesh plots using plotly (@plotly).
 
 ## Brain atlas selection
-There is a multitude of brain atlases to chose from, all with different purposes and segmented based on different criteria. 
+There are multitudes of brain atlases to chose from, all with different purposes and segmented based on different criteria. 
 We have initially focused on commonly used and well-established atlases with manageable numbers of segmentations. 
-The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer (REF); dkt and aseg. 
-The Desikan-Killany cortical atlas (dkt) (REF) and the Automatic Segmentation of Subcortical Structures (aseg) (REF) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.
+The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer [@fischl:99]; dkt and aseg. 
+The Desikan-Killany cortical atlas (dkt; @dkt) and the Automatic Segmentation of Subcortical Structures (aseg; @aseg) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.
 
@@
 This is particularly true when looking within the brain as subcortical structures or white matter tracts.
-To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using [plotly](https://plotly-book.cpsievert.me/).
+To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using plotly [@plotly].
 The data is somewhat more complex than the 2D ggplot polygon version `ggseg`, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
 Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
-A lot of credit goes to A.M.Winkler and his [Brainder](https://brainder.org/) work, which supplied us with the first examples of going from Freesurfer `.srf`-files to `.ply`-files, and whose scripts massively aided us in making this work.
+A lot of credit goes to A.M.Winkler and his Brainder (@brainder) work, which supplied us with the first examples of going from Freesurfer `.srf`-files to `.ply`-files, and whose scripts massively aided us in making this work.
 
-The function `ggseg3d()`, is based in the plotly, it is recommended to get a little familiarized with with [plotly](https://plotly-book.cpsievert.me/).
+The function `ggseg3d()`, is based in the plotly, it is recommended to get a little familiarized with with plotly (@plotly).
 
@@
 There are many plotly options that may help you adapt the plot to look the way you wish. 
-In addition to Carson Sievert's [book on plotly in R](https://plotly-book.cpsievert.me/), we recommend resources for [modifying axes in 3D plots](https://plot.ly/r/axes/#modifying-axes-for-3D-plots), the [basic introduction to tri-surface plots](https://plot.ly/r/trisurf/), and this tutorial on [tri-surface plots with plotly in R](https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/).
+In addition to Carson Sievert's book on plotly in R (-@plotly) we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with plotly in R (@plotly-trisurf).
 
@@
 We encourage users of this package to contribute more atlases they find uses for, and help build an extensive repository of brain atlases.
+
+# Author Contributions
+Didac Vidal-Piñeiro generated the idea for the plotting tool, and the initial scripts that made the plots. 
+He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R. 
+Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and making the functions more user-friendly.
+She also was responsible for adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible. 
+A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.
+
+# Conflicts of Interest
+The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+# Acknowledgements
+We also thanks John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
+
+# Funding 
+This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+
+# Prior versions
+Some of the content presented here also apprears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg). 
+Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creationg and functionality (@ggsegAnim, @ggsegIntro).
+
+# References

### 2019-04-15 draft_1_AMM.rmd
Update draft_1_AMM.rmd

@@
 # Author Contributions
-Didac Vidal-Piñeiro generated the idea for the plotting tool, and the initial scripts that made the plots. 
-He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R. 
+DVP designed the earliest 2D polygon plotting functionality together with the initial scripts, and contributed to the image-to-polygon and the image-to-mesh pipelines. 
 Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and making the functions more user-friendly.
@@
 # Acknowledgements
-We also thanks John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
+We particularly thank Anders M Fjell and Kristine B Walhovd for insightful discussion and funding support and Ingle K Amlien for inspiration and code comments. 
+We also thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
 
 # Funding 
-This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592 and ERC Starting Grant schemes (grant agreements 283634, 725025 and 313440).
 
 # Prior versions
-Some of the content presented here also apprears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg). 
+Some of the content presented here also appears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg). 
 Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creationg and functionality (@ggsegAnim, @ggsegIntro).

### 2019-04-15 draft_1_AMM.rmd
Update draft_1_AMM.rmd

@@
 # Author Contributions
-DVP designed the earliest 2D polygon plotting functionality together with the initial scripts, and contributed to the image-to-polygon and the image-to-mesh pipelines. 
+DVP designed the earliest 2D polygon plotting functionality together with the initial scripts and contributed to the image-to-polygon and the image-to-mesh pipelines. 
 Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and making the functions more user-friendly.
@@
 We particularly thank Anders M Fjell and Kristine B Walhovd for insightful discussion and funding support and Ingle K Amlien for inspiration and code comments. 
-We also thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
+We also thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor) and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
 
 # Funding 
-This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592 and ERC Starting Grant schemes (grant agreements 283634, 725025 and 313440).
+This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimising the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592 and ERC Starting Grant schemes (grant agreements 283634, 725025, and 313440).
 

### 2019-04-29 draft_2.Rmd
updated ms from comments and added Didacs discussion

---
title: Visualisation of Brain Statistics with R-package `ggseg`
preprint: true
author: 
  - name: Athanasia M. Mowinckel
    affiliation: 1
    corresponding: true
    email: a.m.mowinckel@psykologi.uio.no
  - name: Didac Vidal Piñeiro
    affiliation: 1
affiliation:
  - code: 1
    address: Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway
abstract: >
  The abstract of the article.
  It can also be on *multiple* lines.
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
    citation_package: natbib
    fig_caption: true
    includes:
      in_header: preamble.tex
  rticles::peerj_article: preprint
---
```{r setup, include=F}
knitr::opts_chunk$set(eval=TRUE, echo=TRUE)
options(tinytex.verbose = TRUE)

library(kableExtra)
library(tidyverse)
library(plotly) 
library(ggplot2)
library(knitr)
library(ggseg)

include_plotly <- function(plot, file){
  orca(plot, file = file, format = "png")
  knitr::include_graphics(file)
}

# Hook for setting latex text sizes
def.chunk.hook  <- knitr::knit_hooks$get("chunk") 
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

# Hook for setting code chunk wrapping
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

dt <- data(package='ggseg')
dt <- as.data.frame(dt$results, stringsAsFactors = FALSE)

dtE <- data(package='ggsegExtra')
dtE <- as.data.frame(dtE$results, stringsAsFactors = FALSE)

dd <- bind_rows(dt, dtE) %>% 
  filter(!grepl("pal", Item)) %>% 
  mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
         Item = gsub("_3d", "", Item),
         val = "Yes") %>% 
  select(Package, Title, Type, Item, val) %>% 
  spread(Type, val, fill = "No") %>% 
  mutate(Title = linebreak(stringr::str_wrap(Title, width = 40))) 
rm(dt, dtE)
options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```

# Introduction

Visualization is increasingly important for accurate guidance and interpretation of neuroimaging results, as current research is able to generate a high amount of data and outcomes.
For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (>100.000).
Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or connectivity properties (@eickhoff_2018).
The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses.
Complex analyses and post-processing is often performed in specialized software environments such as R (@R).

MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts).
The projection of data into brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and informs on the spatial patterns of the data.
Adopting the grammar of graphics implemented in ggplot2 (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
Yet these tools display whole-brain image files and are not suited for representing brain atlas data.

In this tutorial, we introduce the ggseg-package for visualizing brain atlas data in R.
The ggseg – and the complimentary ggsegExtra – package include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
Two-dimensional functionality is based on polygons and ggplot-based grammar of graphics (@ggplot), while the 3D functionality is based on tri-surface mesh plots and plotly (@plotly).

The ggseg-package presents both compiled data sets, tailored scripts to allow brain data integration and plotting and, other minor features such as custom color palettes.
The data featured are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
The ggsegExtra-package currently includes >`r dd %>% filter(Package == "ggsegExtra") %>% nrow()` additional precompiled atlases and it is continuously expanding.
See Table \@ref(tab:atlasTab) for a summary of the available atlases.

We encourage users to contribute to the ggsegExtra brain atlas repository.
In SI text and [github wiki](https://github.com/LCBC-UiO/ggseg/wiki), we offer a pipeline to create and supply atlases for 2D plotting.
See SI text and SI scripts for instruction to include atlases for 3D plotting.

```{r "atlasTab", echo=F, warning=FALSE, message=FALSE}
dd %>%
  kable(format="latex", escape = FALSE, booktabs = TRUE,
               caption = "Table of currently available atlases in either ggseg or ggsegExtra R-package. Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.") %>%
  collapse_rows(1)
```

# Tutorial
This tutorial will introduce the `ggseg` and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
The tutorial will cover the following functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.

## Plotting polygon data (ggplot2)

`ggseg` is the main function for plotting 2D data.
By default, the function automatically plots the DKT atlas (see Figure \@ref(fig:init)).
The `ggseg` function is a wrapper for `geom_polygon` from `ggplot2`, and it can be built upon and combined like any ggplot object.
The image plot consists of a simple brain representation containing no extra information.
Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options.
We recommend users to get familiarized with ggplot2 grammar (@ggplot).

```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in shaded polygons."}
library(ggseg)
library(tidyverse)
ggseg()
```

`ggseg` is implemented as a fully-functional part of ggplot.
Thus one should be able to use the function together with any other ggplot function such as themes or scales.
You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these.
One has several options for plotting the main brain representations.
These options are atlas-specific.
For cortical atlases, such as the 'dkt', one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \@ref(fig:collection)).
For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.

```{r collection, fig.width=5, fig.cap="ggseg plots can be used with most standard scales, themes and such that work with ggplot. The special ggseg options for hemispheres, view etc. depend on the atlas used, and some options are only available for certain atlases. There is no 'lateral' or 'medial' views of subcortical atlases. There are several ggseg-special options that may be supplied to control how the plots looks and is organised. **A:** dkt atlas, stacked with dark theme ; **B:** dkt with medial view only; **C:** dkt atlas with only left medial display; **D:** dkt atlas, stacked, with classic theme; **E:** dkt atlas with left hemisphere only; **F:** complete aseg atlas"}
# dkt dark theme
p1 <- ggseg(position = "stacked") +
  theme_dark() +
  labs(title=" ")

# dkt classic theme
p2 <- ggseg(position = "stacked") +
  theme_classic() +
  labs(title = " ")

# dkt medial view
med <- ggseg(view = "medial") +
  labs(title = " ")

# dkt left hemisphere
left <- ggseg(hemisphere = "left") +
  labs(title = " ")

# aseg default theme
p3 <- ggseg(atlas=aseg) +
  labs(title = " ")

# dkt left medial alone
combo <- ggseg(view = "medial",
      hemisphere = "left") +
  labs(title=" ")

# Combine plots
cowplot::plot_grid(p1, med, combo,  p2, left, p3,
                   labels = c("A: dkt - dark", "B: dkt - medial",
                              "C: dkt - combo", "D: dkt - classic",
                              "E: dkt - left", "F: aseg"),
                   hjust = -.05)

```

### Using own data with fill and colour
`ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality.
Standard arguments like `fill` to flood the segments with a colour, or `colour` to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales.
See an example in Figure \@ref(fig:fill).

```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas."}
ggseg(mapping=aes(fill = area), colour="black") +
  scale_fill_brain("dkt") +
  theme(legend.justification=c(1,0),
        legend.position="bottom",
        legend.text = element_text(size = 5)) +
  guides(fill = guide_legend(ncol = 3))
```

Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or cognition relationships etc.,  across the different brain regions.
The bulk of the package originates not in its functions, but in the data sets that accompany the package.
There is one data set per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
Before setting up the data from the statistics to project onto the segments, looking at the atlas data sets might help understand what the data needs to look like, or avoid looking like.

```{r, fig.cap="All ggseg polygon atlases are required to have at minimum 6 columns with the names: 'long', 'lat', 'id', 'hemi', 'area', 'side'. These columns are necessary for the ggseg function to work. The additional columns in the dkt were added for various practical reasons.", size='small'}
dkt
```

In any atlas, the column 'label' is particularly useful for combining the data of interest with the `ggseg`-polygons.
The column 'label' contains the label (region) names as in the original neuroimaging software.
For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
Yet the data in `ggseg` is in a long format - that is each region has its own row - and data of interest needs to be in this same format.
Most data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
See below an example of wide-to-long conversion.

```{r }
freesurfer_stats <- data.frame(
  id = c(10:12),
  lh_superiortemporal = c(3.32, 4.1, 3.5),
  lh_precentral = c(2.3, 2.5, 2.1),
  lh_rostralmiddlefrontal = c(3.3, 3.2, 3.1)
  )
freesurfer_stats

freesurfer_long <- freesurfer_stats %>%
  gather(label, thickness, -id)
freesurfer_long
```

This data - in long format - can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
The data **must** include a column that has the same name and at least _some_ data matching the values in the corresponding column in the atlas data.
In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of an analysis.
The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins.
We use the `p` column as the column flooding the segment with colour.

```{r datasupp, fig.width=5}
someData = data.frame(
  area = c("transverse temporal", "insula",
           "pre central","superior parietal"),
  p = sample(seq(0,.5,.001), 4),
  stringsAsFactors = FALSE)

p <- ggseg(.data=someData, mapping=aes(fill=p))
p
```

The appearance of this plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.

```{r datasupp2, fig.width=5}
p +
  theme_void() +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  labs(title="A nice plot title", fill="p-value")
```

If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column `hemi` with either 'right' or 'left' for this to happen.

```{r datasuppX, fig.width=5}
someData$hemi = "right"

ggseg(.data=someData, colour="white",mapping=aes(fill=p)) +
  theme_void() +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  labs(title="Another nice plot title")
```

### Creating subplots
There is often the need to plot a statistic of interest in different groups (i.e. thickness or significance in young or older adults).
This may be obtained also with ggseg, using ggplot's `facet_wrap` or `facet_grid`, using two guiding rules:
**1)** as before, data needs to be in long format (group data should appear in seperate rows, not in separate columns).
**2)** The data needs to be grouped using `dplyr`'s `group_by` function _before_ providing the data to the `ggseg` function.
The ggseg function will detect grouped data, and adapt it to `facet`'s requirements.

```{r datasupp3, fig.width=5}
someData = data.frame(
  area = rep(c("transverse temporal", "insula",
               "pre central","superior parietal"),2),
  p = sample(seq(0,.5,.001), 8),
  AgeG = c(rep("Young",4), rep("Old",4)),
  stringsAsFactors = FALSE) %>%
  group_by(AgeG)

ggseg(.data=someData, colour="white", mapping=aes(fill=p)) +
  facet_wrap(~AgeG, ncol=1) +
  theme(legend.position = "bottom")
```

As before, one can apply the different ggplot and facet options, such as `scales` to modify the plot's appearance.

```{r datasupp5, fig.width=5}
ggseg(.data = someData, atlas=dkt,
      colour="white", position="stacked",
      mapping=aes(fill=p)) +
  facet_wrap(~AgeG, ncol=2) +
  theme_dark() +
  theme(legend.position = "bottom",
        axis.text = element_blank(),
        axis.title = element_blank()
        ) +
  scale_fill_gradient2(high="goldenrod",
                       low="firebrick",
                       na.value="transparent")
```

All the concepts described above also work with the 'aseg' atlas for subcortical structures, except hemisphere and view arguments that do not apply in the same way in subcortical atlases such as aseg.
Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.

```{r , size = 'small', out.width='90%'}
aseg
```

```{r atlases, fig.width=5, fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
ggseg(atlas="aseg", mapping=aes(fill=area))
```

## Plotting 3D mesh data

Representing brains as 2D polygons is a good solution for fast, efficient, and flexible plotting, and can be easily combined with interactive apps such as Shiny (@shiny).
Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
This problem is exacerbated in atlases that represent subcortical features – after all, cortical surfaces are 2-dimensional – such as grey matter structures or white matter tracts.
Hence, here we also provide the `ggseg3d` function to plot, view, and print 3D-atlases in R.
`ggseg3d` is based on tri-surface mesh plots using plotly (@plotly).
The data structure is somewhat more complex than the ggplot polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
As `ggseg3d` is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
We recommend users to familiarize themselves with plotly (@plotly) when using this function.

Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface.
The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage_ template subject.
See *SI SCRIPT* for reproducibility.
All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the colorlut used in the neuroimaging software.

```{r ggseg3d-1-out, message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3 , fig.cap="The three surface options provided in ggseg 3d atlases. From left: the 'white' surface is the white matter surface, 'LCBC' surface is the white matter surface inflated in 10 steps, and the 'inflated' surface is a fully inflated sortical surface."}
p <- ggseg3d(surface = "white")  %>%
  layout(title="\nggseg3d(surface='white')",
         font =  list(size = 30))
orca(p, file = "png/ggseg3d_white.png", format = "png")

p <- ggseg3d()  %>%
  layout(title="\nggseg3d(surface='LCBC')",
         font =  list(size = 30))
orca(p, file = "png/ggseg3d.png", format = "png")

p <- ggseg3d(surface = "inflated") %>%
  layout(title="\nggseg3d(surface='inflated')",
         font =  list(size = 30))
orca(p, file = "png/ggseg3d_inflated.png", format = "png")

include_graphics(c("png/ggseg3d_white.png","png/ggseg3d.png", "png/ggseg3d_inflated.png"))
```

The 3D-atlas data is stored in nested tibbles.
Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d_1_out)) and the two hemispheres.
As there are no meaningful other surfaces for subcortical atlases, these only have a single surface.
The 'ggseg_3d' column includes all necessary information to `ggseg3d()` to create a mesh-plot.
The additional 3D-atlases in ggsegExtra have the same data structure.

```{r, out.width='90%', size='small'}
dkt_3d
```

### External data supply

Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a color scale some descriptive or inferential statistics.
If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas is expecting.

```{r out.width='40%', linewidth=70}
dkt_3d %>%
  filter(surf == "inflated" & hemi == "right") %>%
  unnest(ggseg_3d) %>%
  select(-lobe, -acronym)
```

Note the `mesh` column, which contains lists.
Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
It should also be noted that the 'label', 'annot' and 'area' columns could provide matching values for your own data.
Similarly to the `ggseg`-function, the 'label' column should match the region names used in the original neuroimaging software while 'area' and 'annot' provide alternative/secondary names.
It is thus important to match your regional identifiers with those used in the atlas.
A warning will be issued in case of any mismatch.
The column you want to use for colour, needs to be supplied to the `colour` option, and you'll likely want to supply it to the `text` option, as this will add another line to the plotly hover information.

```{r, eval=FALSE, out.width='40%'}
someData = dkt_3d %>%
  filter(surf == "inflated" &
         hemi == "right") %>%
  unnest(ggseg_3d) %>%
  select(area) %>%
  na.omit() %>%
  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
           round(2))

  ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
```

```{r, out.width='40%', echo=FALSE}
someData = dkt_3d %>%
  filter(surf == "inflated" & hemi == "right") %>%
  unnest(ggseg_3d) %>%
  select(area) %>%
  na.omit() %>%
  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
           round(2))

p <- ggseg3d(.data = someData, atlas = dkt_3d,
             colour = "p", text = "p")
include_plotly(p, "png/ggseg3d_data.png")
```

### Colours

You can provide custom colour palettes either in hex or R-names.
Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.

```{r,fig.width=5, eval=FALSE, out.width='45%'}
  ggseg3d(.data = someData, atlas = dkt_3d,
          colour = "p", text = "p",
          palette = c("#ff0000", "#00ff00", "#0000ff"))
```
```{r, out.width='40%', echo=F, eval=T}
p <- ggseg3d(.data = someData, atlas = dkt_3d,
          colour = "p", text = "p",
          palette = c("#ff0000", "#00ff00", "#0000ff"))
include_plotly(p, "png/ggseg3d_colour.png")
```

If you want to alter the colour of `NA` regions, supply `na.colour`, either as HEX colour or colour name.
This option only takes a single colour.

```{r, out.width='40%', eval=FALSE,  out.width='45%'}
  ggseg3d(.data = someData, atlas = dkt_3d,
          colour = "p", text = "p",
          na.colour = "black", camera="medial")
```
```{r, out.width='40%', echo=F, eval=T}
p <-   ggseg3d(.data = someData, atlas = dkt_3d,
          colour = "p", text = "p",
          na.colour = "black", camera="medial")
include_plotly(p, "png/ggseg3d_nacolour.png")
```

Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
One can control the opacity of the these `NA` structures, to improve visualization.
Additionally, one can add the `glassbrain` option, which provides a frame of reference for the subcortical structures.
Glassbrains can be controlled with three different arguments: opacity (0-1), hemisphere (left, right), and colour.

```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
somData_aseg = aseg_3d %>%
  unnest() %>%
  select(label) %>%
  filter(!grepl("Ventricle|Putamen", label)) %>%
  mutate(p = seq(1, nrow(.)))

  ggseg3d(.data = somData_aseg, atlas = aseg_3d,
          colour = "p", text = "p",
          na.alpha= .5, glassbrain = .5,
          glassbrain_hemisphere = "left")
```
```{r, out.width='40%', echo=F, eval=T}
somData_aseg = aseg_3d %>%
  unnest() %>%
  select(label) %>%
  filter(!grepl("Ventricle|Putamen", label)) %>%
  mutate(p = seq(1, nrow(.)))

p <- ggseg3d(.data = somData_aseg, atlas = aseg_3d,
          colour = "p", text = "p",
          na.alpha= .5, glassbrain = .5,
          glassbrain_hemisphere = "left")
include_plotly(p, "png/ggseg3d_glass.png")
```

`ggsed3d()` is based on plotly and thus additional plotly functionalities can be used to modify and improve the 3D atlas representations.
In addition to Carson Sievert's book on plotly in R (-@plotly), we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with plotly in R (@plotly-trisurf).
Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line tool to save ggseg3d atlas snapshots.

# Discussion

The main aim of the ggseg – and ggsegExtra - package is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
In this tutorial, we introduce the package to the readers by presenting some use examples and highlighting the main functions and options that are available.
As a visualization tool, this package adds up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of ggseg as a visualization tool.
After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
On the contrary, we believe the ggseg niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.

The ggseg and ggsegExtra packages contain three main features:
**1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases. The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make i easy to use for ones own purposes.
**2)** `ggseg()` and `ggseg3d()` functions for visualization.
`ggseg()` as a wrapper for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
`ggseg3d()` as a plotly wrapper for tri-surface mesh plots which prints 3D atlases.
Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
**3)** Complimentary features – e.g. color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format. These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository.

The foundations of the ggseg-package trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al. (-@vidal_2019)) .
That is, ggseg appears with the need to inspect and display 4D information - including a spatial dimension - overcoming the constrains of printed journals and classical 2D plots (e.g. bar plots).
The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public [ggseg GitHub repository](https://github.com/LCBC-UiO/ggseg) and [ggsegExtra GitHub repository](https://github.com/LCBC-UiO/ggsegExtra).
See X and Y for detailed guidelines on how to create new atlases.
Finally - while the ggseg-package is circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (@gganatogram).

# Conclusion
Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation and, communicate with colleagues and general audience.
In this tutorial, we have introduced the ggseg-package, a tool for visualizing brain statistics through brain parcellation atlases in R.
This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines. We hope this tool and tutorial proves useful to neuroscientists and inspires others to apply the functions in a wide variety of fields and structures.

# Author Contributions
Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts that made the plots.
He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
She is also responsible for conceiving and adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible.
A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.

# Conflicts of Interest
The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.

# Acknowledgements
We wish to thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
A. M. Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into `.ply` files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
We are indebted to those users that actively contributed to the package and all the scientists and engineers that worked on the framework on which the package is sustained (e.g. R-project, neuroimaging software).

# Funding
This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.

# Prior versions
Some of the content presented here also appears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg).
Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (@ggsegAnim, @ggsegIntro).

# References

### 2019-04-29 draft_2.tex
updated ms from comments and added Didacs discussion

--- 2019-04-10 draft_1.tex
@@
 \usepackage{longtable}
+
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
+
+\author[1]{Athanasia M. Mowinckel}
+
+\corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
+\author[1]{Didac Vidal Piñeiro}
+
+\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
+
+%
+% \author[1]{First Author}
+% \author[2]{Second Author}
+% \affil[1]{Address of first author}
+% \affil[2]{Address of second author}
+% \corrauthor[1]{First Author}{f.author@email.com}
+
+% 
+\usepackage{natbib}
+\bibliographystyle{plainnat}
+
+\begin{abstract}
+The abstract of the article. It can also be on \emph{multiple} lines.
+% Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+\end{abstract}
 
@@
 
-\usepackage{lipsum}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
-
-\author[1]{Athanasia M. Mowinckel}
-
-\corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
-\author[1]{Didac Vidal Piñeiro}
-
-\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
-
-%
-% \author[1]{First Author}
-% \author[2]{Second Author}
-% \affil[1]{Address of first author}
-% \affil[2]{Address of second author}
-% \corrauthor[1]{First Author}{f.author@email.com}
-
-%
-
-\begin{abstract}
-The abstract of the article. It can also be on \emph{multiple} lines.
-% Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-\end{abstract}
-
 \begin{document}
@@
 
-Neuroscientific analysis usually requires the use of multiple tools to analyse, visualise, and summarise data.
-This often makes the process of preparing results for publication laborious, as results must be exported and imported in various formats.
-In neuroimaging, images of probands' brains are collected and merged together to provide three dimensional representations.
-Much of the neuroimaging analyses are not necessarily done on the individual voxel level (3-dimensional pixels), but rather on pre-defined brain segmentations, called brain atlases.
-These brain atlases are plentiful and represent different ways of segmenting the brain into functionally or structurally similar regions (REF).
-The use of these is wide-spread, as these atlases provide larger meaningful divisions of the brain (REF).
-While neuroimaging analyses on the voxel-level are usually computed by special software for such analyses, analyses of brain atlas data is usually done in standard statistical software, like R (REF), python (REF), or Matlab (REF).
+Visualization is increasingly important for accurate guidance and interpretation of neuroimaging results, as current research is able to generate a high amount of data and outcomes.
+For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (\textgreater{}100.000).
+Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or connectivity properties (\citet{eickhoff_2018}).
+The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses.
+Complex analyses and post-processing is often performed in specialized software environments such as R (\citet{R}).
 
-A key part of understanding and disseminating analysis results, is the visualization of these in a meaningful way.
-With regards to results from brain atlas analyses, it is most meaningfully represented if projected onto a representation of the brain, rather than other common types of charts (like bar-charts) accompanied by a diagram of the brain label positions.
-Each atlas has its own labelling depending on what is meaningful for the type of segmentation it is based on.
-As such, for the reader to fully understand a bar chart with atlas labels, they need to be very familiar with the location of each label to have a clear comprehension of the results.
-A projection directly onto a brain shape, eases the readability of the results for the reader, and provides clear point of reference even if the used atlas is unfamiliar.
+MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g.~bar charts).
+The projection of data into brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and informs on the spatial patterns of the data.
+Adopting the grammar of graphics implemented in ggplot2 (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
+Yet these tools display whole-brain image files and are not suited for representing brain atlas data.
 
-There are several tools that aid R-users in plotting neuroimaging data directly through R using the grammar of graphics as implemented in ggplot2 (REF), such as ggBrain(REF) and ggneuro(REF)(see \href{https://neuroconductor.org/}{neuroconductor} for compiled neuroimaging packages for R).
-These are based on plotting imaging files, not results from analyses of brain atlases.
-Here we introduce the ggseg-package for visualizing results from brain atlas analyses.
-The ggseg-package was developed to create templates that others might use to project their brain atlas results on to.
-While its plotting functions are what the users are drawn to, it is the pre-compiled number of data sets for different brain atlases that provides the real functionality needed for this visualization.
-The package also includes functions to plot atlases in tri-surface mesh plots using plotly (REF).
+In this tutorial, we introduce the ggseg-package for visualizing brain atlas data in R.
+The ggseg -- and the complimentary ggsegExtra -- package include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+Two-dimensional functionality is based on polygons and ggplot-based grammar of graphics (\citet{ggplot}), while the 3D functionality is based on tri-surface mesh plots and plotly (\citet{plotly}).
 
-\hypertarget{brain-atlas-selection}{%
-\subsection{Brain atlas selection}\label{brain-atlas-selection}}
+The ggseg-package presents both compiled data sets, tailored scripts to allow brain data integration and plotting and, other minor features such as custom color palettes.
+The data featured are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
+Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
+The ggsegExtra-package currently includes \textgreater{}13 additional precompiled atlases and it is continuously expanding.
+See Table \ref{tab:atlasTab} for a summary of the available atlases.
 
-There is a multitude of brain atlases to chose from, all with different purposes and segmented based on different criteria.
-We have initially focused on commonly used and well-established atlases with manageable numbers of segmentations.
-The ggseg-package thus features data for two well-established parcellations, as implemented in the neuroimaging software Freesurfer (REF); dkt and aseg.
-The Desikan-Killany cortical atlas (dkt) (REF) and the Automatic Segmentation of Subcortical Structures (aseg) (REF) together cover all grey matter of the brain, and are commonly used in analyses elucidating the associations between thickness, volume, or area of these segments and mental health and cognitive function.
-
-While functions and base data can be found in the main \texttt{ggseg}-package, a companion package called \texttt{ggsegExtra} has numerous other atlases, and is continuously expanded.
-We encourage the community to help build and expand the atlases available, and to create a repository of established and often used atlases.
-A summary of all available atlases may be found in Table \ref{tab:atlasTab}, where in addition to the two default atlases, ggsegExtra has another 12 atlases.
-The \href{https://github.com/LCBC-UiO/ggseg/wiki}{ggseg wiki} currently has two different ways you may create and supply atlases to the package as 2D polygons.
-Work is being done to include instructions on 3D-mesh plot contributions also.
-The package includes functions like \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to aid users in testing and making custom atlases.
+We encourage users to contribute to the ggsegExtra brain atlas repository.
+In SI text and \href{https://github.com/LCBC-UiO/ggseg/wiki}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
+See SI text and SI scripts for instruction to include atlases for 3D plotting.
 
@@
 \caption{\label{tab:atlasTab}Table of currently available atlases in either ggseg or ggsegExtra R-package.
-Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.
-}
+Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.}
 \centering
@@
 \midrule
- & Desikan-Killiany Cortical Atlas & dkt & No & Yes\\
+ & Desikan-Killiany Cortical Atlas & dkt & Yes & Yes\\
 \cmidrule{2-5}
- & \makecell[l]{Desikan-Killiany Cortical Atlas - tri-\\surf mesh Mesh data for the Desikan-\\Killany Cortical atlas, with 40 regions\\in on the cortical surface of the brain.} & dkt & Yes & No\\
-\cmidrule{2-5}
-\multirow{-3}{*}{\raggedright\arraybackslash ggseg} & \makecell[l]{Freesurfer automatic subcortical\\segmentation of a brain volume} & aseg & Yes & Yes\\
+\multirow{-2}{*}{\raggedright\arraybackslash ggseg} & \makecell[l]{Freesurfer automatic subcortical\\segmentation of a brain volume} & aseg & Yes & Yes\\
 \cmidrule{1-5}
  & Desterieux cortical parcellations & desterieux & Yes & No\\
+\cmidrule{2-5}
+ & \makecell[l]{Genetic topography of brain area\\morphology} & chenAr & No & Yes\\
+\cmidrule{2-5}
+ & \makecell[l]{Genetic topography of brain thickness\\morphology} & chenTh & No & Yes\\
 \cmidrule{2-5}
@@
 \cmidrule{2-5}
- & Parcellation from JHU & jhu & No & Yes\\
-\cmidrule{2-5}
- & Parcellation from of JHU & jhu & Yes & No\\
+ & Parcellation from JHU & jhu & Yes & Yes\\
 \cmidrule{2-5}
  & Parcellation from of white matter & icbm & Yes & No\\
-\cmidrule{2-5}
- & Parcellation from of white matter & tracula & Yes & Yes\\
 \cmidrule{2-5}
@@
 \cmidrule{2-5}
+ & White matter tract parcellations & tracula & Yes & Yes\\
+\cmidrule{2-5}
  & \makecell[l]{Yeo 17 Resting-state Cortical\\Parcellations} & yeo17 & Yes & Yes\\
 \cmidrule{2-5}
-\multirow{-12}{*}{\raggedright\arraybackslash ggsegExtra} & \makecell[l]{Yeo 7 Resting-state Cortical\\Parcellations} & yeo7 & Yes & Yes\\
+\multirow{-13}{*}{\raggedright\arraybackslash ggsegExtra} & \makecell[l]{Yeo 7 Resting-state Cortical\\Parcellations} & yeo7 & Yes & Yes\\
 \bottomrule
@@
 
-The aim of this tutorial is to familiarize the reader with the main package functions and the general use of the package.\\
-It will cover the two main functions: \texttt{ggseg()} which plots 2D polygons, and \texttt{ggseg3d()} which plots tri-surface mesh plots.
+This tutorial will introduce the \texttt{ggseg} and \texttt{ggsegExtra} packages and familiarize the reader with the main functions and the general use of the packages.
+The tutorial will cover the following functions: \texttt{ggseg()} for plotting 2D polygons and \texttt{ggseg3d()} for plotting 3D brains based on tri-surface mesh plots.
 
-\hypertarget{plotting-polygon-data}{%
-\subsection{Plotting polygon data}\label{plotting-polygon-data}}
+\hypertarget{plotting-polygon-data-ggplot2}{%
+\subsection{Plotting polygon data (ggplot2)}\label{plotting-polygon-data-ggplot2}}
 
-The function to plot two-dimensional polygon data is the \texttt{ggseg}-function.
-This function will automatically plot the dkt-atlas by default (see Figure \ref{fig:init}).
-As this function is a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2} it can be built upon like all ggplot objects.
-The plot is kept as simple as possible, with as little extra information to ggplot as possible.
-This is done for it to be easy to extend the plot with the many ggplot options available.
+\texttt{ggseg} is the main function for plotting 2D data.
+By default, the function automatically plots the DKT atlas (see Figure \ref{fig:init}).
+The \texttt{ggseg} function is a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2}, and it can be built upon and combined like any ggplot object.
+The image plot consists of a simple brain representation containing no extra information.
+Hence, \texttt{ggseg} plots can be easily complemented with any of the available ggplot features and options.
+We recommend users to get familiarized with ggplot2 grammar (\citet{ggplot}).
 
@@
 \centering
-\includegraphics{draft_1_files/figure-latex/init-1.pdf}
+\includegraphics{draft_2_files/figure-latex/init-1.pdf}
 \caption{\label{fig:init}By default \texttt{ggseg} will plot the dkt atlas in shaded polygons.}
@@
 
-Because of the way \texttt{ggseg} is implemented, you should be able to safely use the function together with other ggplot functions, like themes, scales, and such (see Figure \ref{fig:void}).
-There are many extra options to ggseg, to alter the plot as you wish.
+\texttt{ggseg} is implemented as a fully-functional part of ggplot.
+Thus one should be able to use the function together with any other ggplot function such as themes or scales.
+You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these.
+One has several options for plotting the main brain representations.
+These options are atlas-specific.
+For cortical atlases, such as the `dkt', one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \ref{fig:collection}).
+For subcortical atlases, such as the `aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.
 
@@
 \begin{Highlighting}[]
-\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"dark theme"}\NormalTok{)}
-\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"classic theme"}\NormalTok{)}
+\CommentTok{# dkt dark theme}
+\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
+
+\CommentTok{# dkt classic theme}
+\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+
+\CommentTok{# dkt medial view}
+\NormalTok{med <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+
+\CommentTok{# dkt left hemisphere}
+\NormalTok{left <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+
+\CommentTok{# aseg default theme}
+\NormalTok{p3 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+
+\CommentTok{# dkt left medial alone}
+\NormalTok{combo <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
+      \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
 
 \CommentTok{# Combine plots}
-\NormalTok{pp <-}\StringTok{ }\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(p1, p2, }\DataTypeTok{nrow =} \DecValTok{2}\NormalTok{)}
-
-\NormalTok{p3 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"default theme"}\NormalTok{)}
-
-\CommentTok{# Create a subplot grid}
-\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(pp, p3, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{)}
+\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
+                   \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A: dkt - dark"}\NormalTok{, }\StringTok{"B: dkt - medial"}\NormalTok{,}
+                              \StringTok{"C: dkt - combo"}\NormalTok{, }\StringTok{"D: dkt - classic"}\NormalTok{,}
+                              \StringTok{"E: dkt - left"}\NormalTok{, }\StringTok{"F: aseg"}\NormalTok{),}
+                   \DataTypeTok{hjust =} \FloatTok{-.05}\NormalTok{)}
 \end{Highlighting}
@@
 \centering
-\includegraphics{draft_1_files/figure-latex/void-1.pdf}
-\caption{\label{fig:void}ggseg plots can be used with most standard scales, themes and such that work with ggplot.
+\includegraphics{draft_2_files/figure-latex/collection-1.pdf}
+\caption{\label{fig:collection}ggseg plots can be used with most standard scales, themes and such that work with ggplot.
 The special ggseg options for hemispheres, view etc.
 depend on the atlas used, and some options are only available for certain atlases.
-There is no `lateral' or `medial' views of subcortical atlases.}
-\end{figure}
-
-You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these (see Figure \ref{fig:stacked}).
-The possibilities of altering the plot depends on the atlas used.
-There are generally more options available for cortical atlases than subcortical ones, as the atlases have more meaningful ways of viewing the polygons.
-Subcortical atlases also tend to have structures that span the hemispheres or other typically used referece points of the brain, and as such it is more difficult to label all segments as either, e.g., left or right.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{lat <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Medial view"}\NormalTok{)}
-\NormalTok{left <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Left hemi."}\NormalTok{)}
-\NormalTok{pp <-}\StringTok{ }\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(lat, left, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{)}
-
-\NormalTok{combo <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
-      \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Combination"}\NormalTok{)}
-
-\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(pp, combo, }\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\centering
-\includegraphics{draft_1_files/figure-latex/stacked-1.pdf}
-\caption{\label{fig:stacked}There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
-Top left: plot using medial view.
-Bottom left: plot using left hemisphere.
-Right: plot combining both the options.}
+There is no `lateral' or `medial' views of subcortical atlases.
+There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
+\textbf{A:} dkt atlas, stacked with dark theme ; \textbf{B:} dkt with medial view only; \textbf{C:} dkt atlas with only left medial display; \textbf{D:} dkt atlas, stacked, with classic theme; \textbf{E:} dkt atlas with left hemisphere only; \textbf{F:} complete aseg atlas}
 \end{figure}
@@
 Standard arguments like \texttt{fill} to flood the segments with a colour, or \texttt{colour} to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option \texttt{aes}.
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales.
+See an example in Figure \ref{fig:fill}.
 
@@
 \begin{Highlighting}[]
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
@@
 \centering
-\includegraphics{draft_1_files/figure-latex/fill-1.pdf}
+\includegraphics{draft_2_files/figure-latex/fill-1.pdf}
 \caption{\label{fig:fill}Supplying `area' to the fill option in ggseg, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.}
@@
 
-Typically you would want each area to have a different colour, corresponding to some descriptive or inference statistic that has been computed.
-The bulk of the package originates not in its functions, but in the datasets that accompany the package.
-There is one dataset per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
-Before setting up the data from the statistics to project onto the segments, looking at the atlas datasets might help understand what the data needs to look like, or avoid looking like.
+Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or cognition relationships etc., across the different brain regions.
+The bulk of the package originates not in its functions, but in the data sets that accompany the package.
+There is one data set per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
+Before setting up the data from the statistics to project onto the segments, looking at the atlas data sets might help understand what the data needs to look like, or avoid looking like.
 
@@
 
-The column `label' was added to the atlas for easy matching with the output from Freesurfer statistics tables.
-The data is in long-format, meaning each segment has its own row, meaning the data you wish to project onto the segments also need to be in long-format.
-As data directly from Freesurfer comes in wide-format, this requires some extra wranling.
+In any atlas, the column `label' is particularly useful for combining the data of interest with the \texttt{ggseg}-polygons.
+The column `label' contains the label (region) names as in the original neuroimaging software.
+For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
+Yet the data in \texttt{ggseg} is in a long format - that is each region has its own row - and data of interest needs to be in this same format.
+Most data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+See below an example of wide-to-long conversion.
 
@@
 \begin{Highlighting}[]
-\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{-}\NormalTok{id)}
@@
 
-This data can now be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
-The data \textbf{must} include a column that has the same name at atleast \emph{some} data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an `area' and `p' column.
+This data - in long format - can now be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
+The data \textbf{must} include a column that has the same name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
+In the next example we create some data with 4 rows, and an `area' and `p' column, representing the results of an analysis.
 The \texttt{ggseg} function will recognise the matching column `area', and merge the supplied data into the atlas using \texttt{dplyr} joins.
@@
   \DataTypeTok{area =} \KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
-           \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{), }
+           \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}
   \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{4}\NormalTok{),}
@@
 
-\includegraphics{draft_1_files/figure-latex/datasupp-1.pdf}
+\includegraphics{draft_2_files/figure-latex/datasupp-1.pdf}
 
-We can also change how this looks, by various ggplot functions, like scales, labs, themes and such.
+The appearance of this plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
 
@@
 
-\includegraphics{draft_1_files/figure-latex/datasupp2-1.pdf}
+\includegraphics{draft_2_files/figure-latex/datasupp2-1.pdf}
 
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column \texttt{hemi} with either ``right'' or ``left'' for this to happen.
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column \texttt{hemi} with either `right' or `left' for this to happen.
 
@@
 
-\includegraphics{draft_1_files/figure-latex/datasuppX-1.pdf}
+\includegraphics{draft_2_files/figure-latex/datasuppX-1.pdf}
 
@@
 
-If your data has different statistics for different groups, and you want to plot them all, we can use ggplot's \texttt{facet\_wrap} or \texttt{facet\_grid} for that.
-The data needs to be in long format for this to work, as needed by ggplot, meaning the data from the groups should appear in seperate rows in the data, not in separate columns.
-Secondly, the data needs to be grouped using \texttt{dplyr}'s \texttt{group\_by} function \emph{before} providing the data to the \texttt{ggseg} function.
-The ggseg-function will detect if the data is grouped, and perform the necessary steps for the \texttt{facet}'s to work when they are applied later.
+There is often the need to plot a statistic of interest in different groups (i.e.~thickness or significance in young or older adults).
+This may be obtained also with ggseg, using ggplot's \texttt{facet\_wrap} or \texttt{facet\_grid}, using two guiding rules:
+\textbf{1)} as before, data needs to be in long format (group data should appear in seperate rows, not in separate columns).
+\textbf{2)} The data needs to be grouped using \texttt{dplyr}'s \texttt{group\_by} function \emph{before} providing the data to the \texttt{ggseg} function.
+The ggseg function will detect grouped data, and adapt it to \texttt{facet}'s requirements.
 
@@
   \DataTypeTok{area =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
-               \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}\DecValTok{2}\NormalTok{), }
+               \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}\DecValTok{2}\NormalTok{),}
   \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{8}\NormalTok{),}
   \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{)),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{group_by}\NormalTok{(AgeG)}
@@
 
-\includegraphics{draft_1_files/figure-latex/datasupp3-1.pdf}
+\includegraphics{draft_2_files/figure-latex/datasupp3-1.pdf}
 
-Depending on what type of column is used for fill (continuous or discrete),
-you can also use the differenct \texttt{scales} to alter the colours.
+As before, one can apply the different ggplot and facet options, such as \texttt{scales} to modify the plot's appearance.
 
@@
 \begin{Highlighting}[]
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas=}\NormalTok{dkt, }
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas=}\NormalTok{dkt,}
       \DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"stacked"}\NormalTok{,}
@@
 
-\includegraphics{draft_1_files/figure-latex/datasupp5-1.pdf}
+\includegraphics{draft_2_files/figure-latex/datasupp5-1.pdf}
 
-All the consepts described above also work with the \texttt{aseg} atlas for subcrotical structures, except arguments to see only one hemisphere or lateral or medial surfaces, as these arguments do not make sense for this atlas.
-Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas
+All the concepts described above also work with the `aseg' atlas for subcortical structures, except hemisphere and view arguments that do not apply in the same way in subcortical atlases such as aseg.
+Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
+
 \small
@@
 \centering
-\includegraphics{draft_1_files/figure-latex/atlases-1.pdf}
+\includegraphics{draft_2_files/figure-latex/atlases-1.pdf}
 \caption{\label{fig:atlases}The first 10 rows of the \texttt{aseg} dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way.}
@@
 
-Creating plots of the brain as 2-dimensional can at times make it difficult to properly understand where an atlas segment is located in the true 3-dimensional brain.
-This is particularly true when looking within the brain as subcortical structures or white matter tracts.
-To provide the option of viewing and printing the plots in 3-dimensional space, another function was created using \href{https://plotly-book.cpsievert.me/}{plotly}.
-The data is somewhat more complex than the 2D ggplot polygon version \texttt{ggseg}, as it includes options for brain inflation, glass brains and much more, to ease viewing and understanding.
-Another positive effect of basing this function on plotly, is the interactivity of the plot, which also makes it nice for public dissemination.
-A lot of credit goes to A.M.Winkler and his \href{https://brainder.org/}{Brainder} work, which supplied us with the first examples of going from Freesurfer \texttt{.srf}-files to \texttt{.ply}-files, and whose scripts massively aided us in making this work.
+Representing brains as 2D polygons is a good solution for fast, efficient, and flexible plotting, and can be easily combined with interactive apps such as Shiny (\citet{shiny}).
+Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
+This problem is exacerbated in atlases that represent subcortical features -- after all, cortical surfaces are 2-dimensional -- such as grey matter structures or white matter tracts.
+Hence, here we also provide the \texttt{ggseg3d} function to plot, view, and print 3D-atlases in R.
+\texttt{ggseg3d} is based on tri-surface mesh plots using plotly (\citet{plotly}).
+The data structure is somewhat more complex than the ggplot polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
+As \texttt{ggseg3d} is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
+We recommend users to familiarize themselves with plotly (\citet{plotly}) when using this function.
 
-The function \texttt{ggseg3d()}, is based in the plotly, it is recommended to get a little familiarized with with \href{https://plotly-book.cpsievert.me/}{plotly}.
+Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dkt\_3d} atlas in `LCBC' surface.
+The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage} template subject.
+See \emph{SI SCRIPT} for reproducibility.
+All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the regions, based on the colorlut used in the neuroimaging software.
 
-Out-of-the-box, \texttt{ggseg3d()} works without supplying any extra information, and plots the \texttt{dkt\_3d} atlas in `LCBC' surface.
-The `LCBC' surface is the white matter surface, inflated in 10 steps, to increase visibility within folds of the brain, while still maintining curvature to indicate where folds have been.
-All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the segments.
+\begin{figure}
+\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in ggseg 3d atlases.
+From left: the 'white' surface is the white matter surface, 'LCBC' surface is the white matter surface inflated in 10 steps, and the 'inflated' surface is a fully inflated sortical surface.}\label{fig:ggseg3d-1-out}
+\end{figure}
 
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_white}
-\includegraphics[width=0.3\linewidth]{png/ggseg3d}
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated}
+The 3D-atlas data is stored in nested tibbles.
+Each cortical atlas has data sets for three different surfaces (see Figure @ref(fig:ggseg3d\_1\_out)) and the two hemispheres.
+As there are no meaningful other surfaces for subcortical atlases, these only have a single surface.
+The `ggseg\_3d' column includes all necessary information to \texttt{ggseg3d()} to create a mesh-plot.
+The additional 3D-atlases in ggsegExtra have the same data structure.
 
-The cortical atlas data is stored in nested tibbles, which have datasets for three different surfaces and the two hemispheres.
-The `data'-column within includes all necessary information to \texttt{ggseg3d()} to create a mesh-plot.
 \small
@@
 
-To grab all the data for a surface and hemisphere, you should reduce the data to one line, and then \texttt{unnest()}
+\hypertarget{external-data-supply}{%
+\subsubsection{External data supply}\label{external-data-supply}}
+
+Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d} to display through a color scale some descriptive or inferential statistics.
+If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see what the atlas is expecting.
 
@@
 \begin{Highlighting}[]
-\NormalTok{dkt_3d }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{dkt_3d }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{lobe, }\OperatorTok{-}\NormalTok{acronym)}
@@
 
-\hypertarget{external-data-supply}{%
-\subsubsection{External data supply}\label{external-data-supply}}
-
-Particularly notice the \texttt{mesh} column, which is a list column of lists.
-In there is all the 6 vectors needed to create the mesh of the tri-surface plot.
-You'll also need to notice the \texttt{label}, \texttt{annot} and \texttt{area} columns, which are likely the columns you will be matching on when proviging with your own data for colours.
-You need to be meticulous when fixing your data, be sure it matches.
-The function \emph{should} give you a warning if it's struggling to match something.
-
+Note the \texttt{mesh} column, which contains lists.
+Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
+It should also be noted that the `label', `annot' and `area' columns could provide matching values for your own data.
+Similarly to the \texttt{ggseg}-function, the `label' column should match the region names used in the original neuroimaging software while `area' and `annot' provide alternative/secondary names.
+It is thus important to match your regional identifiers with those used in the atlas.
+A warning will be issued in case of any mismatch.
 The column you want to use for colour, needs to be supplied to the \texttt{colour} option, and you'll likely want to supply it to the \texttt{text} option, as this will add another line to the plotly hover information.
@@
 \begin{Highlighting}[]
-\NormalTok{someData =}\StringTok{ }\NormalTok{dkt_3d }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }
-\StringTok{         }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(area) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{na.omit}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{, }\DataTypeTok{length.out =} \DecValTok{100}\NormalTok{ ), }\KeywordTok{nrow}\NormalTok{(.)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{           }\KeywordTok{round}\NormalTok{(}\DecValTok{2}\NormalTok{)) }
+\NormalTok{someData =}\StringTok{ }\NormalTok{dkt_3d }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}
+\StringTok{         }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(area) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{na.omit}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{, }\DataTypeTok{length.out =} \DecValTok{100}\NormalTok{ ), }\KeywordTok{nrow}\NormalTok{(.)) }\OperatorTok{%>%}
+\StringTok{           }\KeywordTok{round}\NormalTok{(}\DecValTok{2}\NormalTok{))}
 
@@
 \begin{Highlighting}[]
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
+  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
+          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
           \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"#ff0000"}\NormalTok{, }\StringTok{"#00ff00"}\NormalTok{, }\StringTok{"#0000ff"}\NormalTok{))}
@@
 \begin{Highlighting}[]
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
+  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
+          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
           \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{camera=}\StringTok{"medial"}\NormalTok{)}
@@
 
-If you are plotting the sub-cortical structures, you might want to reduce the opacity of the \texttt{NA} structures, so that you can see the more medial structures.
-You may also want to add the \texttt{glassbrain} option, any value above 0 and up to 1 increases the opacity of a glass brain overlay.
-The glassbrain option provides a frame of reference for the subcortical structures, and as such it is recommended to add this when viewing these structures.
-There are three glassbrain options, controlling the opacity, which hemispheres to add, and the colour.
+Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
+One can control the opacity of the these \texttt{NA} structures, to improve visualization.
+Additionally, one can add the \texttt{glassbrain} option, which provides a frame of reference for the subcortical structures.
+Glassbrains can be controlled with three different arguments: opacity (0-1), hemisphere (left, right), and colour.
 
@@
 \begin{Highlighting}[]
-\NormalTok{somData_aseg =}\StringTok{ }\NormalTok{aseg_3d }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{unnest}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(label) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"Ventricle|Putamen"}\NormalTok{, label)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{nrow}\NormalTok{(.))) }
+\NormalTok{somData_aseg =}\StringTok{ }\NormalTok{aseg_3d }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(label) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"Ventricle|Putamen"}\NormalTok{, label)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{nrow}\NormalTok{(.)))}
 
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ somData_aseg, }\DataTypeTok{atlas =}\NormalTok{ aseg_3d, }
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{, }
-          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{, }\DataTypeTok{glassbrain =} \FloatTok{.5}\NormalTok{, }
+  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ somData_aseg, }\DataTypeTok{atlas =}\NormalTok{ aseg_3d,}
+          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
+          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{, }\DataTypeTok{glassbrain =} \FloatTok{.5}\NormalTok{,}
           \DataTypeTok{glassbrain_hemisphere =} \StringTok{"left"}\NormalTok{)}
@@
 
-There are many plotly options that may help you adapt the plot to look the way you wish.
-In addition to Carson Sievert's \href{https://plotly-book.cpsievert.me/}{book on plotly in R}, we recommend resources for \href{https://plot.ly/r/axes/\#modifying-axes-for-3D-plots}{modifying axes in 3D plots}, the \href{https://plot.ly/r/trisurf/}{basic introduction to tri-surface plots}, and this tutorial on \href{https://moderndata.plot.ly/trisurf-plots-in-r-using-plotly/}{tri-surface plots with plotly in R}.
+\texttt{ggsed3d()} is based on plotly and thus additional plotly functionalities can be used to modify and improve the 3D atlas representations.
+In addition to Carson Sievert's book on plotly in R (\citeyearpar{plotly}), we recommend resources for modifying axes in 3D plots (\citet{plotly-ax}), the basic introduction to tri-surface plots (\citet{plotly-tri}), and this tutorial on tri-surface plots with plotly in R (\citet{plotly-trisurf}).
+Finally, we recommend \href{https://github.com/plotly/orca\#installation}{orca} command line tool to save ggseg3d atlas snapshots.
 
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-The aim of the \texttt{ggseg}-package is to aid researchers easily disseminate analysis results in an easier and more stream-lined fashion.
-This tutorial should provide instructions on how to use the main package functions, and cover the most commonly used options.
-The two main functions in the package, \texttt{ggseg()} and \texttt{ggseg3d()}, create polygon and mesh-plots of brain atlases, using the accompanying package data.
-In addition to the two default atlases available in \texttt{ggseg}, the \texttt{ggsegExtra} package includes 12 more atlases.
-We encourage users of this package to contribute more atlases they find uses for, and help build an extensive repository of brain atlases.
+The main aim of the ggseg -- and ggsegExtra - package is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
+In this tutorial, we introduce the package to the readers by presenting some use examples and highlighting the main functions and options that are available.
+As a visualization tool, this package adds up to manifold functionalities such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}) in R, and software-specific image viewers such as FSLeyes (\citet{fsleyes}) and Freeview (\citet{dale_99}).
+In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of ggseg as a visualization tool.
+After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
+On the contrary, we believe the ggseg niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
+The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (\citet{binder}) or Shiny (\citet{shiny}), is an added benefit.
+
+The ggseg and ggsegExtra packages contain three main features:
+\textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
+The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make i easy to use for ones own purposes.
+\textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
+\texttt{ggseg()} as a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2} and it can be built upon like any ggplot object.
+\texttt{ggseg3d()} as a plotly wrapper for tri-surface mesh plots which prints 3D atlases.
+Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
+\textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
+These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository.
+
+The foundations of the ggseg-package trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
+(\citeyearpar{vidal_2019})) .
+That is, ggseg appears with the need to inspect and display 4D information - including a spatial dimension - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
+The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
+Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public \href{https://github.com/LCBC-UiO/ggseg}{ggseg GitHub repository} and \href{https://github.com/LCBC-UiO/ggsegExtra}{ggsegExtra GitHub repository}.
+See X and Y for detailed guidelines on how to create new atlases.
+Finally - while the ggseg-package is circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
+We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (\citet{gganatogram}).
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation and, communicate with colleagues and general audience.
+In this tutorial, we have introduced the ggseg-package, a tool for visualizing brain statistics through brain parcellation atlases in R.
+This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines.
+We hope this tool and tutorial proves useful to neuroscientists and inspires others to apply the functions in a wide variety of fields and structures.
+
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts that made the plots.
+He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
+Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
+She is also responsible for conceiving and adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible.
+A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+We wish to thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (\citeyearpar{neuroconductor}), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
+A.
+M.
+Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into \texttt{.ply} files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
+We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
+We are indebted to those users that actively contributed to the package and all the scientists and engineers that worked on the framework on which the package is sustained (e.g.~R-project, neuroimaging software).
+
+\hypertarget{funding}{%
+\section{Funding}\label{funding}}
+
+This work is funded by EU Horizon 2020 Grant `Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+
+\hypertarget{prior-versions}{%
+\section{Prior versions}\label{prior-versions}}
+
+Some of the content presented here also appears in the package vignette of \texttt{ggseg}, which may be accessed through R or in the package website (\citet{ggseg}).
+Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (\citet{ggsegAnim}, \citet{ggsegIntro}).
+
+\renewcommand\refname{References}
+\bibliography{references.bib}
 

### 2019-05-02 draft_2_DVP.Rmd

--- 2019-04-29 draft_2.Rmd
@@
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
-Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or connectivity properties (@eickhoff_2018).
-The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses.
-Complex analyses and post-processing is often performed in specialized software environments such as R (@R).
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or on connectivity properties (@eickhoff_2018).
+The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).
 
@@
 Adopting the grammar of graphics implemented in ggplot2 (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
-Yet these tools display whole-brain image files and are not suited for representing brain atlas data.
+Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
 
@@
 
-The ggseg-package presents both compiled data sets, tailored scripts to allow brain data integration and plotting and, other minor features such as custom color palettes.
-The data featured are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
+The ggseg package presents both compiled data sets, tailored scripts that allow brain data integration and plotting and, other minor features such as custom color palettes.
+The data featured in the ggseg package are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The ggsegExtra-package currently includes >`r dd %>% filter(Package == "ggsegExtra") %>% nrow()` additional precompiled atlases and it is continuously expanding.
-See Table \@ref(tab:atlasTab) for a summary of the available atlases.
-
-We encourage users to contribute to the ggsegExtra brain atlas repository.
+The ggsegExtra package includes a library of precompiled atlases (currently including `r dd %>% filter(Package == "ggsegExtra") %>% nrow()` additional atlases) and it is updated continuously.
+See Table \@ref(tab:atlasTab) for a summary of the available atlases (as for May 2019).
+We encourage users to contribute to the ggsegExtra brain atlas repository by including additional brain atlases.
 In SI text and [github wiki](https://github.com/LCBC-UiO/ggseg/wiki), we offer a pipeline to create and supply atlases for 2D plotting.
@@
   kable(format="latex", escape = FALSE, booktabs = TRUE,
-               caption = "Table of currently available atlases in either ggseg or ggsegExtra R-package.
-Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.") %>%
+               caption = "Table of currently available atlases in the ggseg and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlase representations.") %>%
   collapse_rows(1)
@@
 This tutorial will introduce the `ggseg` and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
-The tutorial will cover the following functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
+The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
 
@@
 
-`ggseg` is implemented as a fully-functional part of ggplot.
-Thus one should be able to use the function together with any other ggplot function such as themes or scales.
-You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these.
-One has several options for plotting the main brain representations.
+As `ggseg` is implemented as a fully-functional part of ggplot, one can combine the function together with any other ggplot function such as themes or scales.
+Also, one has several options for plotting the main brain representations.
 These options are atlas-specific.
-For cortical atlases, such as the 'dkt', one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \@ref(fig:collection)).
+For cortical atlases, such as the 'dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
+See Figure \@ref(fig:collection)) for an example.
 For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.
 
-```{r collection, fig.width=5, fig.cap="ggseg plots can be used with most standard scales, themes and such that work with ggplot.
-The special ggseg options for hemispheres, view etc.
-depend on the atlas used, and some options are only available for certain atlases.
-There is no 'lateral' or 'medial' views of subcortical atlases.
-There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
+```{r collection, fig.width=5, fig.cap="ggseg plots can be used with any ggplot feasture such as standard scales and themes.
+For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
 **A:** dkt atlas, stacked with dark theme ; **B:** dkt with medial view only; **C:** dkt atlas with only left medial display; **D:** dkt atlas, stacked, with classic theme; **E:** dkt atlas with left hemisphere only; **F:** complete aseg atlas"}
@@
 `ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like `fill` to flood the segments with a colour, or `colour` to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
+Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
 To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales.
@@
 
-```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas."}
+```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The 'dkt' atlas default palette corresponds to the FreeSurferColorLut scheme."}
 ggseg(mapping=aes(fill = area), colour="black") +
@@
 
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or cognition relationships etc.,  across the different brain regions.
-The bulk of the package originates not in its functions, but in the data sets that accompany the package.
-There is one data set per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
-Before setting up the data from the statistics to project onto the segments, looking at the atlas data sets might help understand what the data needs to look like, or avoid looking like.
+Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
+Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
+The atlas dataset structure will help us understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data set.
+All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
@@
 The column 'label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
-Yet the data in `ggseg` is in a long format - that is each region has its own row - and data of interest needs to be in this same format.
-Most data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+Yet the data in `ggseg` is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
+Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
 See below an example of wide-to-long conversion.
@@
 
-This data - in long format - can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
+Data in long format can then be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
 The data **must** include a column that has the same name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of an analysis.
+In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of a hypothetical analysis.
 The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins.
@@
 If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column `hemi` with either 'right' or 'left' for this to happen.
+In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
 
@@
 ### Creating subplots
-There is often the need to plot a statistic of interest in different groups (i.e. thickness or significance in young or older adults).
+There is often the need to plot a statistic of interest in different groups (e.g. thickness or brain - cognition relationships in young or older adults).
 This may be obtained also with ggseg, using ggplot's `facet_wrap` or `facet_grid`, using two guiding rules:
@@
 
-All the concepts described above also work with the 'aseg' atlas for subcortical structures, except hemisphere and view arguments that do not apply in the same way in subcortical atlases such as aseg.
+All the concepts described above also work with the 'aseg' atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases such as aseg.
 Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas. 
+Note that both the equivalent structures between 'aseg' and the 'dkt' atlases.
 
@@
 See *SI SCRIPT* for reproducibility.
-All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the colorlut used in the neuroimaging software.
+All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
 
 ```{r ggseg3d-1-out, message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3 , fig.cap="The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the white matter surface inflated in 10 steps, and the 'inflated' surface is a fully inflated sortical surface."}
+From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (10 iterations), and the 'inflated' surface is a fully inflated white matter cortical surface."}
 p <- ggseg3d(surface = "white")  %>%
@@
 Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d_1_out)) and the two hemispheres.
-As there are no meaningful other surfaces for subcortical atlases, these only have a single surface.
-The 'ggseg_3d' column includes all necessary information to `ggseg3d()` to create a mesh-plot.
+Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
+The 'ggseg_3d' column includes all necessary information to `ggseg3d()` to create a 3D mesh-plot.
 The additional 3D-atlases in ggsegExtra have the same data structure.
@@
 Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a color scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas is expecting.
+If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas expects.
 
@@
 **1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make i easy to use for ones own purposes.
+The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
 **2)** `ggseg()` and `ggseg3d()` functions for visualization.
-`ggseg()` as a wrapper for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
-`ggseg3d()` as a plotly wrapper for tri-surface mesh plots which prints 3D atlases.
+`ggseg()` is a wrapper function for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
+`ggseg3d()` is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
 Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
@@
 (-@vidal_2019)) .
-That is, ggseg appears with the need to inspect and display 4D information - including a spatial dimension - overcoming the constrains of printed journals and classical 2D plots (e.g. bar plots).
+That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial diemension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
+bar plots).
 The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.

### 2019-11-28 draft_2.Rmd
revised based on didacs comments

@@
 abstract: >
-  The abstract of the article.
-  It can also be on *multiple* lines.
+  There is an increasing emphasis on visualizing results from statistical analysis in more intuitive ways.
+Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualization of these data should reflect these same dimensional properties.
+The ggseg package makes it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
+Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.
+In this tutorial, we present the data and functions in the ggseg and ggseg3d R-packages for brain atlas visualization.
+The tutorial highlights the main functions to create brain segmentation plots in R, and also introduces the accompanying package data that makes these visualizations possible.
 bibliography: references.bib
@@
 ```{r setup, include=F}
-knitr::opts_chunk$set(eval=TRUE, echo=TRUE)
+knitr::opts_chunk$set(eval=TRUE, echo=TRUE, collapse = TRUE )#, fig.pos = 'H')
 options(tinytex.verbose = TRUE)
@@
 library(ggseg)
+library(ggseg3d)
 
@@
 
-dt <- data(package='ggseg')
-dt <- as.data.frame(dt$results, stringsAsFactors = FALSE)
+dt <- as_tibble(data(package='ggseg')$results)
+dt3 <- as_tibble(data(package='ggseg3d')$results)
+dtE <- as_tibble(data(package='ggsegExtra')$results)
 
-dtE <- data(package='ggsegExtra')
-dtE <- as.data.frame(dtE$results, stringsAsFactors = FALSE)
-
-dd <- bind_rows(dt, dtE) %>% 
+dd <-  bind_rows(dt, dt3) %>% 
+  bind_rows(dtE) %>% 
   filter(!grepl("pal", Item)) %>% 
   mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
-         Item = gsub("_3d", "", Item),
-         val = "Yes") %>% 
-  select(Package, Title, Type, Item, val) %>% 
-  spread(Type, val, fill = "No") %>% 
-  mutate(Title = linebreak(stringr::str_wrap(Title, width = 40))) 
-rm(dt, dtE)
+         Item = gsub("_3d", "", Item)) %>% 
+  select(Package, Title, Type, Item) %>% 
+  spread(Type, Package, fill = "None")  %>% 
+  mutate(citation = case_when(
+    grepl("dkt", Item) ~ "\\citet{dkt}",
+    grepl("yeo", Item) ~ "\\citet{yeo2011}",
+    grepl("aseg", Item) ~ "\\citet{aseg}",
+    grepl("desterieux", Item) ~ "\\citet{desterieux}",
+    grepl("^chen", Item) ~ "\\citet{chen}",
+    grepl("glasser", Item) ~ "\\citet{glasser}",
+    grepl("icbm", Item) ~ "\\citet{icbm}",
+    grepl("jhu", Item) ~ "\\citet{jhu}",
+    grepl("schaefer", Item) ~ "\\citet{schaefer}",
+    grepl("tracula", Item) ~ "\\citet{tracula}",
+    grepl("^ho", Item) ~ "\\citet{ho}",
+    TRUE ~ ""
+    ))
+rm(dt, dt3, dtE)
 options(warnPartialMatchAttr = FALSE,
@@
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
-Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or connectivity properties (@eickhoff_2018).
-The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses.
-Complex analyses and post-processing is often performed in specialized software environments such as R (@R).
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or on connectivity properties (@eickhoff_2018).
+The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons. 
+This facilitates replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).
 
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts).
-The projection of data into brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and informs on the spatial patterns of the data.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts) alone.
+The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
 Adopting the grammar of graphics implemented in ggplot2 (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
-Yet these tools display whole-brain image files and are not suited for representing brain atlas data.
+Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
 
-In this tutorial, we introduce the ggseg-package for visualizing brain atlas data in R.
-The ggseg – and the complimentary ggsegExtra – package include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
-Two-dimensional functionality is based on polygons and ggplot-based grammar of graphics (@ggplot), while the 3D functionality is based on tri-surface mesh plots and plotly (@plotly).
+In this tutorial, we introduce two packages for visualizing brain atlas data in R.
+The ggseg and ggseg3d – pluss the complimentary ggsegExtra – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+Two-dimensional functionality in ggseg is based on polygons and ggplot2-based grammar of graphics (@ggplot), while the 3D functionality in ggseg3d is based on tri-surface mesh plots and plotly (@plotly).
 
-The ggseg-package presents both compiled data sets, tailored scripts to allow brain data integration and plotting and, other minor features such as custom color palettes.
-The data featured are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
+Both packages present compiled data sets, tailored functions that allow brain data integration, and plotting, as well as other minor features such as custom colour palettes.
+The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The ggsegExtra-package currently includes >`r dd %>% filter(Package == "ggsegExtra") %>% nrow()` additional precompiled atlases and it is continuously expanding.
-See Table \@ref(tab:atlasTab) for a summary of the available atlases.
-
-We encourage users to contribute to the ggsegExtra brain atlas repository.
-In SI text and [github wiki](https://github.com/LCBC-UiO/ggseg/wiki), we offer a pipeline to create and supply atlases for 2D plotting.
-See SI text and SI scripts for instruction to include atlases for 3D plotting.
+The ggsegExtra package includes a library of precompiled atlases (currently including `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is updated continuously.
+See Table \@ref(tab:atlasTab) for a summary of the available atlases (as of `r format(Sys.Date(), "%B %Y")`).
 
@@
   kable(format="latex", escape = FALSE, booktabs = TRUE,
-               caption = "Table of currently available atlases in either ggseg or ggsegExtra R-package.
-Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.") %>%
-  collapse_rows(1)
+               caption = "Table of currently available atlases in the ggseg, ggseg3d, and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlas representations, respectively") %>%
+  #collapse_rows(1) %>% 
+  #column_spec(1, width="20em") %>% 
+  kable_styling(latex_options = c("scale_down"))
 ```
@@
 # Tutorial
-This tutorial will introduce the `ggseg` and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
-The tutorial will cover the following functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
+This tutorial will introduce the `ggseg`, `ggseg3d`, and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
+The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
 
@@
 We recommend users to get familiarized with ggplot2 grammar (@ggplot).
+The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (@cran) some time during 2020.
 
-```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in shaded polygons."}
+```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in grey shaded polygons."}
+# remotes::install_github("LCBC-UiO/ggseg")
 library(ggseg)
 library(tidyverse)
+
+# Figure 1
 ggseg()
@@
 
-`ggseg` is implemented as a fully-functional part of ggplot.
-Thus one should be able to use the function together with any other ggplot function such as themes or scales.
-You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these.
-One has several options for plotting the main brain representations.
+In addition to the standard options for ggplot polygon geoms, the function also has several options for plotting the main brain representations.
 These options are atlas-specific.
-For cortical atlases, such as the 'dkt', one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \@ref(fig:collection)).
-For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.
+For cortical atlases, such as the 'dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
+See Figure \@ref(fig:collection)) for examples.
+For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagittal, and coronal views.
 
-```{r collection, fig.width=5, fig.cap="ggseg plots can be used with most standard scales, themes and such that work with ggplot.
-The special ggseg options for hemispheres, view etc.
-depend on the atlas used, and some options are only available for certain atlases.
-There is no 'lateral' or 'medial' views of subcortical atlases.
-There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
+```{r collection, fig.width=5, fig.cap="ggseg plots can be used with any ggplot feasture such as standard scales and themes.
+For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
 **A:** dkt atlas, stacked with dark theme ; **B:** dkt with medial view only; **C:** dkt atlas with only left medial display; **D:** dkt atlas, stacked, with classic theme; **E:** dkt atlas with left hemisphere only; **F:** complete aseg atlas"}
@@
 
-# Combine plots
+# Combine plots to Figure 2
 cowplot::plot_grid(p1, med, combo,  p2, left, p3,
@@
 `ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like `fill` to flood the segments with a colour, or `colour` to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
+Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the ggplot mapping function `aes`.
 To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales.
@@
 
-```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas."}
+```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The 'dkt' atlas default palette corresponds to the FreeSurferColorLut scheme."}
+# Figure 3
 ggseg(mapping=aes(fill = area), colour="black") +
@@
 
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or cognition relationships etc.,  across the different brain regions.
-The bulk of the package originates not in its functions, but in the data sets that accompany the package.
-There is one data set per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
-Before setting up the data from the statistics to project onto the segments, looking at the atlas data sets might help understand what the data needs to look like, or avoid looking like.
+Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
+Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
+The atlas data set structure will help us understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data set.
+All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
@@
 The additional columns in the dkt were added for various practical reasons.", size='small'}
-dkt
+head(dkt, 5)
 ```
@@
 The column 'label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
-Yet the data in `ggseg` is in a long format - that is each region has its own row - and data of interest needs to be in this same format.
-Most data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+Yet the data in `ggseg` is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
+Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
 See below an example of wide-to-long conversion.
 
-```{r }
+```{r, fig.cap="Transforming wide data from Freesurfer to comply with the long data format of ggplot"}
 freesurfer_stats <- data.frame(
@@
 
-This data - in long format - can now be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
+Data in long format can then be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
 The data **must** include a column that has the same name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of an analysis.
+In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of a hypothetical analysis.
 The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins.
 We use the `p` column as the column flooding the segment with colour.
+The appearance of the plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
 
-```{r datasupp, fig.width=5}
+```{r datasupp, fig.width=5, fig.cap="Supplying data through the .data option in ggseg will make it possible to use columns in the supplied data to aestheticals (such as fill) in ggseg.
+The ggseg object can be used with any other polygon compatible function from ggplot or ggplot extentions, like here adding title, changing the legend name and the colour scheme with standard ggplot2 functions."}
 someData = data.frame(
@@
 
-p <- ggseg(.data=someData, mapping=aes(fill=p))
-p
+# Figure 4
+ggseg(.data=someData, mapping=aes(fill=p))  +
+  labs(title="A nice plot title", fill="p-value") +
+  scale_fill_gradient(low="firebrick",high="goldenrod")
 ```
 
-The appearance of this plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
-
-```{r datasupp2, fig.width=5}
-p +
-  theme_void() +
-  scale_fill_gradient(low="firebrick",high="goldenrod") +
-  labs(title="A nice plot title", fill="p-value")
-```
-
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column `hemi` with either 'right' or 'left' for this to happen.
-
-```{r datasuppX, fig.width=5}
-someData$hemi = "right"
-
-ggseg(.data=someData, colour="white",mapping=aes(fill=p)) +
-  theme_void() +
-  scale_fill_gradient(low="firebrick",high="goldenrod") +
-  labs(title="Another nice plot title")
-```
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi` with either 'right' or 'left' for this to happen.
+In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
 
 ### Creating subplots
-There is often the need to plot a statistic of interest in different groups (i.e. thickness or significance in young or older adults).
+There is often the need to plot a statistic of interest in different groups (e.g. thickness or brain - cognition relationships in young or older adults).
 This may be obtained also with ggseg, using ggplot's `facet_wrap` or `facet_grid`, using two guiding rules:
@@
 
-```{r datasupp3, fig.width=5}
+```{r datasupp3, fig.width=5, fig.cap="Data with different groups need the data to be grouped_by before providing the data to ggseg"}
 someData = data.frame(
@@
 
-ggseg(.data=someData, colour="white", mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=1) +
+# Figure 5
+ggseg(.data=someData, colour="white", position = "stacked",
+      mapping=aes(fill=p)) +
+  facet_wrap(~AgeG, ncol=2) +
   theme(legend.position = "bottom")
@@
 
-As before, one can apply the different ggplot and facet options, such as `scales` to modify the plot's appearance.
+All the concepts described above also work with the 'aseg' atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases such as aseg.
+Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
+Note the equivalent structures between 'aseg' and the 'dkt' atlases.
 
-```{r datasupp5, fig.width=5}
-ggseg(.data = someData, atlas=dkt,
-      colour="white", position="stacked",
-      mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=2) +
-  theme_dark() +
-  theme(legend.position = "bottom",
-        axis.text = element_blank(),
-        axis.title = element_blank()
-        ) +
-  scale_fill_gradient2(high="goldenrod",
-                       low="firebrick",
-                       na.value="transparent")
+```{r , size = 'small', out.width='90%', fig.cap="The aseg atlashas no different sides to view."}
+head(aseg, 5)
 ```
 
-All the concepts described above also work with the 'aseg' atlas for subcortical structures, except hemisphere and view arguments that do not apply in the same way in subcortical atlases such as aseg.
-Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
-
-```{r , size = 'small', out.width='90%'}
-aseg
-```
-
-```{r atlases, fig.width=5, fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
+```{r atlases, fig.width=5,out.width='60%', fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
+# Figure 6
 ggseg(atlas="aseg", mapping=aes(fill=area))
@@
 This problem is exacerbated in atlases that represent subcortical features – after all, cortical surfaces are 2-dimensional – such as grey matter structures or white matter tracts.
-Hence, here we also provide the `ggseg3d` function to plot, view, and print 3D-atlases in R.
+Hence, here we also provide the `ggseg3d` package to plot, view, and print 3D-atlases in R.
 `ggseg3d` is based on tri-surface mesh plots using plotly (@plotly).
@@
 See *SI SCRIPT* for reproducibility.
-All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the colorlut used in the neuroimaging software.
+All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
 
 ```{r ggseg3d-1-out, message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3 , fig.cap="The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the white matter surface inflated in 10 steps, and the 'inflated' surface is a fully inflated sortical surface."}
+From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface."}
 p <- ggseg3d(surface = "white")  %>%
+  pan_camera("lateral") %>% 
+  remove_axes() %>% 
   layout(title="\nggseg3d(surface='white')",
@@
 p <- ggseg3d()  %>%
+  pan_camera("lateral") %>% 
+  remove_axes() %>% 
   layout(title="\nggseg3d(surface='LCBC')",
@@
 p <- ggseg3d(surface = "inflated") %>%
+  pan_camera("lateral") %>% 
+  remove_axes() %>% 
   layout(title="\nggseg3d(surface='inflated')",
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d_1_out)) and the two hemispheres.
-As there are no meaningful other surfaces for subcortical atlases, these only have a single surface.
-The 'ggseg_3d' column includes all necessary information to `ggseg3d()` to create a mesh-plot.
+Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
+Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
+The 'ggseg_3d' column includes all necessary information for `ggseg3d()` to create a 3D mesh-plot, but that the user should not attempt to alter.
 The additional 3D-atlases in ggsegExtra have the same data structure.
+It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological  coordinate system, but rather Cartesian plot coordinates without further meaning.
 
 ```{r, out.width='90%', size='small'}
+# remotes::install_github("LCBC-UiO/ggseg")
+library(ggseg3d)
 dkt_3d
@@
 
-Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a color scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas is expecting.
+Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a colour scale some descriptive or inferential statistics.
+If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas expects.
 
-```{r out.width='40%', linewidth=70}
+```{r out.width='40%', linewidth=70, size='small'}
 dkt_3d %>%
@@
   unnest(ggseg_3d) %>%
-  select(-lobe, -acronym)
+  select(-lobe, -acronym, -roi, -annot) %>% 
+  head(5)
 ```
@@
            round(2))
-
+  # Figure 8
   ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
@@
 
-```{r, out.width='40%', echo=FALSE}
+```{r ggseg3d-data1, out.width='60%', echo=FALSE, fig.cap="Supplying data so ggseg3d works in much the same way as for ggseg.
+Since ggseg3d is based on plotly rather than ggplot, much adaptation of the plots aesthetics must be done immediately in the main function.
+Here we set the parcellation colours with `colour` and accompanying mouse-hover text with `text`."}
 someData = dkt_3d %>%
@@
 p <- ggseg3d(.data = someData, atlas = dkt_3d,
-             colour = "p", text = "p")
+             colour = "p", text = "p") %>%   
+  pan_camera("lateral") %>% 
+  remove_axes()
 include_plotly(p, "png/ggseg3d_data.png")
@@
 Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.
-
-```{r,fig.width=5, eval=FALSE, out.width='45%'}
-  ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          palette = c("#ff0000", "#00ff00", "#0000ff"))
-```
-```{r, out.width='40%', echo=F, eval=T}
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          palette = c("#ff0000", "#00ff00", "#0000ff"))
-include_plotly(p, "png/ggseg3d_colour.png")
-```
-
 If you want to alter the colour of `NA` regions, supply `na.colour`, either as HEX colour or colour name.
@@
 
-```{r, out.width='40%', eval=FALSE,  out.width='45%'}
+```{r,fig.width=5, eval=FALSE, out.width='45%'}
+# Figure 9 left
 ggseg3d(.data = someData, atlas = dkt_3d,
           colour = "p", text = "p",
-          na.colour = "black", camera="medial")
+          palette = c("#ff0000", "#00ff00", "#0000ff")) %>%   
+  pan_camera("lateral") %>% 
+  remove_axes()
+
+# Figure 9 right
+ggseg3d(.data = someData, atlas = dkt_3d,
+          colour = "p", text = "p",
+          na.colour = "black") %>%   
+  pan_camera("lateral") %>% 
+  remove_axes()
 ```
-```{r, out.width='40%', echo=F, eval=T}
+
+```{r ggseg3d-nacol, out.width='40%', echo=F, eval=T,fig.show='hold', fig.ncol = 2, fig.cap="Left: Specifying the palette for ggseg3d must be done directly in the main call to the plot, rather than a separate function later like in ggseg.
+The `palette` option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
+Right: The colour of the `NA` values can also be changed through the option `na.colour`.
+Here the medial wall (which is `NA`), has been coloured black using this option."}
 p <- ggseg3d(.data = someData, atlas = dkt_3d,
           colour = "p", text = "p",
-          na.colour = "black", camera="medial")
-include_plotly(p, "png/ggseg3d_nacolour.png")
+          palette = c("#ff0000", "#00ff00", "#0000ff")) %>%   
+  pan_camera("lateral") %>% 
+  remove_axes()
+orca(p, file = "png/ggseg3d_colour.png", format = "png")
+
+p <-   ggseg3d(.data = someData, atlas = dkt_3d,
+          colour = "p", text = "p",
+          na.colour = "black") %>%   
+  pan_camera("medial") %>% 
+  remove_axes()
+orca(p, file = "png/ggseg3d_nacolour.png", format = "png")
+
+include_graphics(c("png/ggseg3d_colour.png","png/ggseg3d_nacolour.png"))
 ```
 
+## Glass brain
 Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
@@
 Additionally, one can add the `glassbrain` option, which provides a frame of reference for the subcortical structures.
-Glassbrains can be controlled with three different arguments: opacity (0-1), hemisphere (left, right), and colour.
+Glass brains can be added with the function `add_glassbrain()` which takes three extra arguments: hemisphere, colour, and opacity.
 
 ```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
-somData_aseg = aseg_3d %>%
-  unnest() %>%
-  select(label) %>%
-  filter(!grepl("Ventricle|Putamen", label)) %>%
-  mutate(p = seq(1, nrow(.)))
-
-  ggseg3d(.data = somData_aseg, atlas = aseg_3d,
-          colour = "p", text = "p",
-          na.alpha= .5, glassbrain = .5,
-          glassbrain_hemisphere = "left")
+# Figure 10
+ggseg3d(atlas = aseg_3d,
+          na.alpha= .5) %>%   
+  add_glassbrain("left") %>% 
+  pan_camera("lateral") %>% 
+  remove_axes()
 ```
-```{r, out.width='40%', echo=F, eval=T}
-somData_aseg = aseg_3d %>%
-  unnest() %>%
-  select(label) %>%
-  filter(!grepl("Ventricle|Putamen", label)) %>%
-  mutate(p = seq(1, nrow(.)))
-
-p <- ggseg3d(.data = somData_aseg, atlas = aseg_3d,
-          colour = "p", text = "p",
-          na.alpha= .5, glassbrain = .5,
-          glassbrain_hemisphere = "left")
+```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="When visualising subcortical structures, it might be an idea to add a glassbrain to the plot.
+This will help with understanding structure locations in the brain and make the plot easier to interpret.
+The glassbrain is controlled by three options: opacity:0-1, hemisphere left, right or both, and colour."}
+p <- ggseg3d(atlas = aseg_3d,
+          na.alpha= .5) %>%   
+  add_glassbrain("left") %>% 
+  pan_camera("lateral") %>% 
+  remove_axes()
 include_plotly(p, "png/ggseg3d_glass.png")
@@
 
-The main aim of the ggseg – and ggsegExtra - package is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
-In this tutorial, we introduce the package to the readers by presenting some use examples and highlighting the main functions and options that are available.
+The main aim of the ggseg, ggseg3d, and ggsegExtra packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
+In this tutorial, we introduce the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
 As a visualization tool, this package adds up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of ggseg as a visualization tool.
+In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the ggseg-packages as visualization tools.
 After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
+An example of this is in the online supplementary information of Vidal-Piñeiro et al. (-@vidal_2019).
 
-The ggseg and ggsegExtra packages contain three main features:
+The ggseg, ggseg3d, and ggsegExtra packages contain three main features:
 **1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make i easy to use for ones own purposes.
+The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
 **2)** `ggseg()` and `ggseg3d()` functions for visualization.
-`ggseg()` as a wrapper for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
-`ggseg3d()` as a plotly wrapper for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
+`ggseg()` is a wrapper function for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
+`ggseg3d()` is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from ggplot and plotly, respectively.
 **3)** Complimentary features – e.g.
 color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format.
-These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository.
+These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository in ggsegExtra.
 
-The foundations of the ggseg-package trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
+The foundations of the ggseg-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
 (-@vidal_2019)) .
-That is, ggseg appears with the need to inspect and display 4D information - including a spatial dimension - overcoming the constrains of printed journals and classical 2D plots (e.g. bar plots).
+That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
+bar plots).
 The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
-Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public [ggseg GitHub repository](https://github.com/LCBC-UiO/ggseg) and [ggsegExtra GitHub repository](https://github.com/LCBC-UiO/ggsegExtra).
-See X and Y for detailed guidelines on how to create new atlases.
-Finally - while the ggseg-package is circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
+Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public [ggseg GitHub repository](https://github.com/LCBC-UiO/ggseg), [ggseg3d GitHub repository](https://github.com/LCBC-UiO/ggseg3d), and [ggsegExtra GitHub repository](https://github.com/LCBC-UiO/ggsegExtra).
+
+Finally - while the ggseg-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
 We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (@gganatogram).
+
+# Planned package improvements
+In the package [github wiki](https://github.com/LCBC-UiO/ggseg/wiki), we offer a pipeline to create and supply atlases for 2D plotting.
+However, the current way to do this is convoluted and difficult, and we are putting efforts into improving the way atlases can be made more easily.
+The aim is to create a set of functions that will call specialized tools like FSL(@fsl), Freesurfer(@fischl_99, @dale_99, @Fischl2000) and Imagemagick(@magick) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid ggseg atlases.
+We encourage users to contribute to the ggsegExtra brain atlas repository by including additional brain atlases.
 
@@
 Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into `.ply` files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
+Additionally, Inge Amlien has provided helpful pointers to plot Freesurfer images in R, which resulted in us finally managing to plot mesh segmented brains in R.
+He also provided the 'LCBC' surface files, which can be found in the ggsegExtra github repository (@ggsegExt12_lcbc).
 We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
@@
 # Funding
+
 This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+The project has also received funding from the European Research Council’s Starting Grant scheme under grant agreements 283634, 725025 (to center PI A.M.
+Fjell), and 313440 (to center PI K.B.
+Walhovd).
 
 # Prior versions
-Some of the content presented here also appears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg).
+Some of the content presented here also appears in the package vignette of `ggseg` and `ggseg3d`, which may be accessed through R or in the package websites (@ggseg, @ggseg3d).
 Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (@ggsegAnim, @ggsegIntro).

### 2019-11-28 draft_2.tex
revised based on didacs comments

@@
 
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
-
-\author[1]{Athanasia M. Mowinckel}
-
-\corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
-\author[1]{Didac Vidal Piñeiro}
-
-\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
-
-%
-% \author[1]{First Author}
-% \author[2]{Second Author}
-% \affil[1]{Address of first author}
-% \affil[2]{Address of second author}
-% \corrauthor[1]{First Author}{f.author@email.com}
-
-% 
-\usepackage{natbib}
-\bibliographystyle{plainnat}
-
-\begin{abstract}
-The abstract of the article. It can also be on \emph{multiple} lines.
-% Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-\end{abstract}
-
+% Pandoc syntax highlighting
+% See https://github.com/rstudio/rticles/issues/182
 \usepackage{color}
@@
 
+% Pandoc Header
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
+
+\author[1]{Athanasia M. Mowinckel}
+
+\corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
+\author[1]{Didac Vidal Piñeiro}
+
+\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
+
+%
+% \author[1]{First Author}
+% \author[2]{Second Author}
+% \affil[1]{Address of first author}
+% \affil[2]{Address of second author}
+% \corrauthor[1]{First Author}{f.author@email.com}
+
+% 
+\usepackage{natbib}
+\bibliographystyle{plainnat}
+
+\begin{abstract}
+There is an increasing emphasis on visualizing results from statistical analysis in more intuitive ways.
+Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualization of these data should reflect these same dimensional properties.
+The ggseg package makes it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
+Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.
+In this tutorial, we present the data and functions in the ggseg and ggseg3d R-packages for brain atlas visualization.
+The tutorial highlights the main functions to create brain segmentation plots in R, and also introduces the accompanying package data that makes these visualizations possible.
+% Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+\end{abstract}
+
 \begin{document}
@@
 Visualization is increasingly important for accurate guidance and interpretation of neuroimaging results, as current research is able to generate a high amount of data and outcomes.
-For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (\textgreater{}100.000).
+For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (\textgreater100.000).
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
-Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or connectivity properties (\citet{eickhoff_2018}).
-The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons and facilitating replicability and data sharing in otherwise computationally expensive analyses.
-Complex analyses and post-processing is often performed in specialized software environments such as R (\citet{R}).
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or on connectivity properties (\citet{eickhoff_2018}).
+The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons.
+This facilitates replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (\citet{R}).
 
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g.~bar charts).
-The projection of data into brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and informs on the spatial patterns of the data.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g.~bar charts) alone.
+The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
 Adopting the grammar of graphics implemented in ggplot2 (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
-Yet these tools display whole-brain image files and are not suited for representing brain atlas data.
+Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
 
-In this tutorial, we introduce the ggseg-package for visualizing brain atlas data in R.
-The ggseg -- and the complimentary ggsegExtra -- package include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
-Two-dimensional functionality is based on polygons and ggplot-based grammar of graphics (\citet{ggplot}), while the 3D functionality is based on tri-surface mesh plots and plotly (\citet{plotly}).
+In this tutorial, we introduce two packages for visualizing brain atlas data in R.
+The ggseg and ggseg3d -- pluss the complimentary ggsegExtra -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+Two-dimensional functionality in ggseg is based on polygons and ggplot2-based grammar of graphics (\citet{ggplot}), while the 3D functionality in ggseg3d is based on tri-surface mesh plots and plotly (\citet{plotly}).
 
-The ggseg-package presents both compiled data sets, tailored scripts to allow brain data integration and plotting and, other minor features such as custom color palettes.
-The data featured are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
+Both packages present compiled data sets, tailored functions that allow brain data integration, and plotting, as well as other minor features such as custom colour palettes.
+The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
-The ggsegExtra-package currently includes \textgreater{}13 additional precompiled atlases and it is continuously expanding.
-See Table \ref{tab:atlasTab} for a summary of the available atlases.
+The ggsegExtra package includes a library of precompiled atlases (currently including 13 additional atlases) and it is updated continuously.
+See Table \ref{tab:atlasTab} for a summary of the available atlases (as of November 2019).
 
-We encourage users to contribute to the ggsegExtra brain atlas repository.
-In SI text and \href{https://github.com/LCBC-UiO/ggseg/wiki}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
-See SI text and SI scripts for instruction to include atlases for 3D plotting.
+\begin{table}
 
-\begin{table}[t]
-
-\caption{\label{tab:atlasTab}Table of currently available atlases in either ggseg or ggsegExtra R-package.
-Most atlases have both polygon and mesh atlases, but the mesh atlases are somewhat easier to create and are thus more plentiful.}
+\caption{\label{tab:atlasTab}Table of currently available atlases in the ggseg, ggseg3d, and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlas representations, respectively}
 \centering
-\begin{tabular}{lllll}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllll}
 \toprule
-Package & Title & Item & Mesh & Polygon\\
+Title & Item & Mesh & Polygon & citation\\
 \midrule
- & Desikan-Killiany Cortical Atlas & dkt & Yes & Yes\\
-\cmidrule{2-5}
-\multirow{-2}{*}{\raggedright\arraybackslash ggseg} & \makecell[l]{Freesurfer automatic subcortical\\segmentation of a brain volume} & aseg & Yes & Yes\\
-\cmidrule{1-5}
- & Desterieux cortical parcellations & desterieux & Yes & No\\
-\cmidrule{2-5}
- & \makecell[l]{Genetic topography of brain area\\morphology} & chenAr & No & Yes\\
-\cmidrule{2-5}
- & \makecell[l]{Genetic topography of brain thickness\\morphology} & chenTh & No & Yes\\
-\cmidrule{2-5}
- & Harvard-Oxford Cortical atlas & hoCort & No & Yes\\
-\cmidrule{2-5}
- & Parcellation from a midsagittal slice & midsagittal & No & Yes\\
-\cmidrule{2-5}
- & Parcellation from JHU & jhu & Yes & Yes\\
-\cmidrule{2-5}
- & Parcellation from of white matter & icbm & Yes & No\\
-\cmidrule{2-5}
- & \makecell[l]{Parcellation from the Human Connectome\\Project} & glasser & Yes & Yes\\
-\cmidrule{2-5}
- & \makecell[l]{Schaefer 17 Resting-state Cortical\\Parcellations} & schaefer17 & Yes & No\\
-\cmidrule{2-5}
- & \makecell[l]{Schaefer 7 Resting-state Cortical\\Parcellations} & schaefer7 & Yes & No\\
-\cmidrule{2-5}
- & White matter tract parcellations & tracula & Yes & Yes\\
-\cmidrule{2-5}
- & \makecell[l]{Yeo 17 Resting-state Cortical\\Parcellations} & yeo17 & Yes & Yes\\
-\cmidrule{2-5}
-\multirow{-13}{*}{\raggedright\arraybackslash ggsegExtra} & \makecell[l]{Yeo 7 Resting-state Cortical\\Parcellations} & yeo7 & Yes & Yes\\
+Desikan-Killiany Cortical Atlas & dkt & ggseg3d & ggseg & \citet{dkt}\\
+Desterieux cortical parcellations & desterieux & ggsegExtra & None & \citet{desterieux}\\
+Freesurfer automatic subcortical segmentation of a brain volume & aseg & ggseg3d & ggseg & \citet{aseg}\\
+Genetic topography of brain area morphology & chenAr & None & ggsegExtra & \citet{chen}\\
+Genetic topography of brain thickness morphology & chenTh & None & ggsegExtra & \citet{chen}\\
+\addlinespace
+Harvard-Oxford Cortical atlas & hoCort & None & ggsegExtra & \citet{ho}\\
+Parcellation from a midsagittal slice & midsagittal & None & ggsegExtra & \\
+Parcellation from JHU & jhu & ggsegExtra & ggsegExtra & \citet{jhu}\\
+Parcellation from of white matter & icbm & ggsegExtra & None & \citet{icbm}\\
+Parcellation from the Human Connectome Project & glasser & ggsegExtra & ggsegExtra & \citet{glasser}\\
+\addlinespace
+Schaefer 17 Resting-state Cortical Parcellations & schaefer17 & ggsegExtra & None & \citet{schaefer}\\
+Schaefer 7 Resting-state Cortical Parcellations & schaefer7 & ggsegExtra & None & \citet{schaefer}\\
+White matter tract parcellations & tracula & ggsegExtra & ggsegExtra & \citet{tracula}\\
+Yeo 17 Resting-state Cortical Parcellations & yeo17 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
+Yeo 7 Resting-state Cortical Parcellations & yeo7 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 
-This tutorial will introduce the \texttt{ggseg} and \texttt{ggsegExtra} packages and familiarize the reader with the main functions and the general use of the packages.
-The tutorial will cover the following functions: \texttt{ggseg()} for plotting 2D polygons and \texttt{ggseg3d()} for plotting 3D brains based on tri-surface mesh plots.
+This tutorial will introduce the \texttt{ggseg}, \texttt{ggseg3d}, and \texttt{ggsegExtra} packages and familiarize the reader with the main functions and the general use of the packages.
+The tutorial will focus on the two main functions: \texttt{ggseg()} for plotting 2D polygons and \texttt{ggseg3d()} for plotting 3D brains based on tri-surface mesh plots.
 
@@
 We recommend users to get familiarized with ggplot2 grammar (\citet{ggplot}).
+The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (\citet{cran}) some time during 2020.
 
@@
 \begin{Highlighting}[]
+\CommentTok{# remotes::install_github("LCBC-UiO/ggseg")}
 \KeywordTok{library}\NormalTok{(ggseg)}
 \KeywordTok{library}\NormalTok{(tidyverse)}
+
+\CommentTok{# Figure 1}
 \KeywordTok{ggseg}\NormalTok{()}
@@
 \includegraphics{draft_2_files/figure-latex/init-1.pdf}
-\caption{\label{fig:init}By default \texttt{ggseg} will plot the dkt atlas in shaded polygons.}
+\caption{\label{fig:init}By default \texttt{ggseg} will plot the dkt atlas in grey shaded polygons.}
 \end{figure}
 
-\texttt{ggseg} is implemented as a fully-functional part of ggplot.
-Thus one should be able to use the function together with any other ggplot function such as themes or scales.
-You can stack the hemispheres, view only the medial or lateral side, choose either or both of the hemispheres, or a combination of any of these.
-One has several options for plotting the main brain representations.
+In addition to the standard options for ggplot polygon geoms, the function also has several options for plotting the main brain representations.
 These options are atlas-specific.
-For cortical atlases, such as the `dkt', one can stack the hemispheres, plot one or both hemisphere, or view only the medial or lateral side (or any combination of these; see Figure \ref{fig:collection}).
-For subcortical atlases, such as the `aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.
+For cortical atlases, such as the `dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
+See Figure \ref{fig:collection}) for examples.
+For subcortical atlases, such as the `aseg', the options are more limited but one can often choose between axial, sagittal, and coronal views.
 
@@
 
-\CommentTok{# Combine plots}
+\CommentTok{# Combine plots to Figure 2}
 \NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
@@
 \includegraphics{draft_2_files/figure-latex/collection-1.pdf}
-\caption{\label{fig:collection}ggseg plots can be used with most standard scales, themes and such that work with ggplot.
-The special ggseg options for hemispheres, view etc.
-depend on the atlas used, and some options are only available for certain atlases.
-There is no `lateral' or `medial' views of subcortical atlases.
-There are several ggseg-special options that may be supplied to control how the plots looks and is organised.
+\caption{\label{fig:collection}ggseg plots can be used with any ggplot feasture such as standard scales and themes.
+For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
 \textbf{A:} dkt atlas, stacked with dark theme ; \textbf{B:} dkt with medial view only; \textbf{C:} dkt atlas with only left medial display; \textbf{D:} dkt atlas, stacked, with classic theme; \textbf{E:} dkt atlas with left hemisphere only; \textbf{F:} complete aseg atlas}
@@
 \texttt{ggseg} accepts any argument you can supply to \texttt{geom\_polygon} and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like \texttt{fill} to flood the segments with a colour, or \texttt{colour} to colour the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option \texttt{aes}.
+Standard arguments like \texttt{fill} that floods the segments with a colour, or \texttt{colour} that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the ggplot mapping function \texttt{aes}.
 To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales.
@@
 \begin{Highlighting}[]
+\CommentTok{# Figure 3}
 \KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
@@
 \includegraphics{draft_2_files/figure-latex/fill-1.pdf}
-\caption{\label{fig:fill}Supplying `area' to the fill option in ggseg, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.}
+\caption{\label{fig:fill}Supplying `area' to the fill option in ggseg, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The `dkt' atlas default palette corresponds to the FreeSurferColorLut scheme.}
 \end{figure}
 
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or cognition relationships etc., across the different brain regions.
-The bulk of the package originates not in its functions, but in the data sets that accompany the package.
-There is one data set per atlas, which contains key information regarding the atlas, and coordinates for the segment polygons of the atlas.
-Before setting up the data from the statistics to project onto the segments, looking at the atlas data sets might help understand what the data needs to look like, or avoid looking like.
+Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
+Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
+The atlas data set structure will help us understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data set.
+All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
@@
 \begin{Highlighting}[]
-\NormalTok{dkt}
+\KeywordTok{head}\NormalTok{(dkt, }\DecValTok{5}\NormalTok{)}
+\CommentTok{## # A tibble: 5 x 6}
+\CommentTok{##   atlas area               hemi  side    label              ggseg          }
+\CommentTok{##   <chr> <chr>              <chr> <chr>   <chr>              <list>         }
+\CommentTok{## 1 dkt   superior temporal  left  lateral lh_superiortempor~ <tibble [1,494~}
+\CommentTok{## 2 dkt   pre central        left  lateral lh_precentral      <tibble [1,314~}
+\CommentTok{## 3 dkt   post central       left  lateral lh_postcentral     <tibble [1,164~}
+\CommentTok{## 4 dkt   rostral middle fr~ left  lateral lh_rostralmiddlef~ <tibble [1,194~}
+\CommentTok{## 5 dkt   insula             left  lateral lh_insula          <tibble [870 x~}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 80 x 6
-##    atlas area               hemi  side    label             ggseg          
-##    <chr> <chr>              <chr> <chr>   <chr>             <list>         
-##  1 dkt   superior temporal  left  lateral lh_superiortempo~ <tibble [1,494~
-##  2 dkt   pre central        left  lateral lh_precentral     <tibble [1,314~
-##  3 dkt   post central       left  lateral lh_postcentral    <tibble [1,164~
-##  4 dkt   rostral middle fr~ left  lateral lh_rostralmiddle~ <tibble [1,194~
-##  5 dkt   insula             left  lateral lh_insula         <tibble [870 x~
-##  6 dkt   superior parietal  left  lateral lh_superiorparie~ <tibble [480 x~
-##  7 dkt   inferior temporal  left  lateral lh_inferiortempo~ <tibble [906 x~
-##  8 dkt   lateral occipital  left  lateral lh_lateraloccipi~ <tibble [738 x~
-##  9 dkt   lateral orbitofro~ left  lateral lh_lateralorbito~ <tibble [528 x~
-## 10 dkt   superior frontal   left  lateral lh_superiorfront~ <tibble [420 x~
-## # ... with 70 more rows
-\end{verbatim}
 
@@
 The column `label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics tables.
-Yet the data in \texttt{ggseg} is in a long format - that is each region has its own row - and data of interest needs to be in this same format.
-Most data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+Yet the data in \texttt{ggseg} is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
+Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
 See below an example of wide-to-long conversion.
@@
 \NormalTok{freesurfer_stats}
+\CommentTok{##   id lh_superiortemporal lh_precentral lh_rostralmiddlefrontal}
+\CommentTok{## 1 10                3.32           2.3                     3.3}
+\CommentTok{## 2 11                4.10           2.5                     3.2}
+\CommentTok{## 3 12                3.50           2.1                     3.1}
+
+\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{-}\NormalTok{id)}
+\NormalTok{freesurfer_long}
+\CommentTok{##   id                   label thickness}
+\CommentTok{## 1 10     lh_superiortemporal      3.32}
+\CommentTok{## 2 11     lh_superiortemporal      4.10}
+\CommentTok{## 3 12     lh_superiortemporal      3.50}
+\CommentTok{## 4 10           lh_precentral      2.30}
+\CommentTok{## 5 11           lh_precentral      2.50}
+\CommentTok{## 6 12           lh_precentral      2.10}
+\CommentTok{## 7 10 lh_rostralmiddlefrontal      3.30}
+\CommentTok{## 8 11 lh_rostralmiddlefrontal      3.20}
+\CommentTok{## 9 12 lh_rostralmiddlefrontal      3.10}
 \end{Highlighting}
@@
 
-\begin{verbatim}
-##   id lh_superiortemporal lh_precentral lh_rostralmiddlefrontal
-## 1 10                3.32           2.3                     3.3
-## 2 11                4.10           2.5                     3.2
-## 3 12                3.50           2.1                     3.1
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{-}\NormalTok{id)}
-\NormalTok{freesurfer_long}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##   id                   label thickness
-## 1 10     lh_superiortemporal      3.32
-## 2 11     lh_superiortemporal      4.10
-## 3 12     lh_superiortemporal      3.50
-## 4 10           lh_precentral      2.30
-## 5 11           lh_precentral      2.50
-## 6 12           lh_precentral      2.10
-## 7 10 lh_rostralmiddlefrontal      3.30
-## 8 11 lh_rostralmiddlefrontal      3.20
-## 9 12 lh_rostralmiddlefrontal      3.10
-\end{verbatim}
-
-This data - in long format - can now be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
+Data in long format can then be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
 The data \textbf{must} include a column that has the same name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an `area' and `p' column, representing the results of an analysis.
+In the next example we create some data with 4 rows, and an `area' and `p' column, representing the results of a hypothetical analysis.
 The \texttt{ggseg} function will recognise the matching column `area', and merge the supplied data into the atlas using \texttt{dplyr} joins.
 We use the \texttt{p} column as the column flooding the segment with colour.
+The appearance of the plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
 
@@
 
-\NormalTok{p <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))}
-\NormalTok{p}
+\CommentTok{# Figure 4}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))  }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p-value"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}
 \end{Highlighting}
@@
 
+\begin{figure}
+\centering
 \includegraphics{draft_2_files/figure-latex/datasupp-1.pdf}
+\caption{\label{fig:datasupp}Supplying data through the .data option in ggseg will make it possible to use columns in the supplied data to aestheticals (such as fill) in ggseg.
+The ggseg object can be used with any other polygon compatible function from ggplot or ggplot extentions, like here adding title, changing the legend name and the colour scheme with standard ggplot2 functions.}
+\end{figure}
 
-The appearance of this plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{p }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_void}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p-value"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{draft_2_files/figure-latex/datasupp2-1.pdf}
-
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column \texttt{hemi} with either `right' or `left' for this to happen.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{someData}\OperatorTok{$}\NormalTok{hemi =}\StringTok{ "right"}
-
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{,}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_void}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"Another nice plot title"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{draft_2_files/figure-latex/datasuppX-1.pdf}
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column \texttt{hemi} with either `right' or `left' for this to happen.
+In this case, data will be merged into the atlas both by `area' and by `hemi'.
 
@@
 
-There is often the need to plot a statistic of interest in different groups (i.e.~thickness or significance in young or older adults).
+There is often the need to plot a statistic of interest in different groups (e.g.~thickness or brain - cognition relationships in young or older adults).
 This may be obtained also with ggseg, using ggplot's \texttt{facet\_wrap} or \texttt{facet\_grid}, using two guiding rules:
@@
 
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{) }\OperatorTok{+}
+\CommentTok{# Figure 5}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
+      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
@@
 
+\begin{figure}
+\centering
 \includegraphics{draft_2_files/figure-latex/datasupp3-1.pdf}
+\caption{\label{fig:datasupp3}Data with different groups need the data to be grouped\_by before providing the data to ggseg}
+\end{figure}
 
-As before, one can apply the different ggplot and facet options, such as \texttt{scales} to modify the plot's appearance.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas=}\NormalTok{dkt,}
-      \DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"stacked"}\NormalTok{,}
-      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{,}
-        \DataTypeTok{axis.text =} \KeywordTok{element_blank}\NormalTok{(),}
-        \DataTypeTok{axis.title =} \KeywordTok{element_blank}\NormalTok{()}
-\NormalTok{        ) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_gradient2}\NormalTok{(}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{,}
-                       \DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}
-                       \DataTypeTok{na.value=}\StringTok{"transparent"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{draft_2_files/figure-latex/datasupp5-1.pdf}
-
-All the concepts described above also work with the `aseg' atlas for subcortical structures, except hemisphere and view arguments that do not apply in the same way in subcortical atlases such as aseg.
+All the concepts described above also work with the `aseg' atlas for subcortical structures, with the exception of `hemisphere' and `view' arguments that are superfluous in subcortical atlases such as aseg.
 Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
+Note the equivalent structures between `aseg' and the `dkt' atlases.
 
@@
 \begin{Highlighting}[]
-\NormalTok{aseg}
+\KeywordTok{head}\NormalTok{(aseg, }\DecValTok{5}\NormalTok{)}
+\CommentTok{## # A tibble: 5 x 6}
+\CommentTok{##   atlas area             hemi  side  label                ggseg            }
+\CommentTok{##   <chr> <chr>            <chr> <chr> <fct>                <list>           }
+\CommentTok{## 1 aseg  <NA>             right axial <NA>                 <tibble [808 x 5~}
+\CommentTok{## 2 aseg  <NA>             left  axial <NA>                 <tibble [1,014 x~}
+\CommentTok{## 3 aseg  thalamus proper  left  axial Left-Thalamus-Proper <tibble [56 x 5]>}
+\CommentTok{## 4 aseg  thalamus proper  right axial Right-Thalamus-Prop~ <tibble [61 x 5]>}
+\CommentTok{## 5 aseg  lateral ventric~ right axial Right-Lateral-Ventr~ <tibble [66 x 5]>}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 34 x 6
-##    atlas area            hemi  side  label                ggseg            
-##    <chr> <chr>           <chr> <chr> <fct>                <list>           
-##  1 aseg  <NA>            right axial <NA>                 <tibble [808 x 5~
-##  2 aseg  <NA>            left  axial <NA>                 <tibble [1,014 x~
-##  3 aseg  thalamus proper left  axial Left-Thalamus-Proper <tibble [56 x 5]>
-##  4 aseg  thalamus proper left  axial Right-Thalamus-Prop~ <tibble [56 x 5]>
-##  5 aseg  thalamus proper right axial Left-Thalamus-Proper <tibble [61 x 5]>
-##  6 aseg  thalamus proper right axial Right-Thalamus-Prop~ <tibble [61 x 5]>
-##  7 aseg  lateral ventri~ right axial Left-Lateral-Ventri~ <tibble [66 x 5]>
-##  8 aseg  lateral ventri~ right axial Right-Lateral-Ventr~ <tibble [66 x 5]>
-##  9 aseg  hippocampus     left  axial Left-Hippocampus     <tibble [47 x 5]>
-## 10 aseg  hippocampus     left  axial Right-Hippocampus    <tibble [47 x 5]>
-## # ... with 24 more rows
-\end{verbatim}
 
@@
 \begin{Highlighting}[]
+\CommentTok{# Figure 6}
 \KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area))}
@@
 \begin{figure}
-\centering
-\includegraphics{draft_2_files/figure-latex/atlases-1.pdf}
-\caption{\label{fig:atlases}The first 10 rows of the \texttt{aseg} dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way.}
+\includegraphics[width=0.6\linewidth]{draft_2_files/figure-latex/atlases-1} \caption{The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way.}\label{fig:atlases}
 \end{figure}
@@
 This problem is exacerbated in atlases that represent subcortical features -- after all, cortical surfaces are 2-dimensional -- such as grey matter structures or white matter tracts.
-Hence, here we also provide the \texttt{ggseg3d} function to plot, view, and print 3D-atlases in R.
+Hence, here we also provide the \texttt{ggseg3d} package to plot, view, and print 3D-atlases in R.
 \texttt{ggseg3d} is based on tri-surface mesh plots using plotly (\citet{plotly}).
@@
 See \emph{SI SCRIPT} for reproducibility.
-All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the regions, based on the colorlut used in the neuroimaging software.
+All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
 
@@
 \includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the white matter surface inflated in 10 steps, and the 'inflated' surface is a fully inflated sortical surface.}\label{fig:ggseg3d-1-out}
+From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface.}\label{fig:ggseg3d-1-out}
 \end{figure}
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure @ref(fig:ggseg3d\_1\_out)) and the two hemispheres.
-As there are no meaningful other surfaces for subcortical atlases, these only have a single surface.
-The `ggseg\_3d' column includes all necessary information to \texttt{ggseg3d()} to create a mesh-plot.
+Each cortical atlas has data sets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
+Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
+The `ggseg\_3d' column includes all necessary information for \texttt{ggseg3d()} to create a 3D mesh-plot, but that the user should not attempt to alter.
 The additional 3D-atlases in ggsegExtra have the same data structure.
+It is important to note that the coordinates in the plot (X, Y, Z) are \textbf{not} any type of radiological coordinate system, but rather Cartesian plot coordinates without further meaning.
 
@@
 \begin{Highlighting}[]
+\CommentTok{# remotes::install_github("LCBC-UiO/ggseg")}
+\KeywordTok{library}\NormalTok{(ggseg3d)}
 \NormalTok{dkt_3d}
+\CommentTok{## # A tibble: 6 x 4}
+\CommentTok{##   atlas  surf     hemi  ggseg_3d         }
+\CommentTok{##   <chr>  <chr>    <chr> <list>           }
+\CommentTok{## 1 dkt_3d inflated left  <tibble [36 x 8]>}
+\CommentTok{## 2 dkt_3d inflated right <tibble [36 x 8]>}
+\CommentTok{## 3 dkt_3d LCBC     left  <tibble [36 x 8]>}
+\CommentTok{## 4 dkt_3d LCBC     right <tibble [36 x 8]>}
+\CommentTok{## 5 dkt_3d white    left  <tibble [36 x 8]>}
+\CommentTok{## 6 dkt_3d white    right <tibble [36 x 8]>}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 6 x 4
-##   atlas  surf     hemi  ggseg_3d         
-##   <chr>  <chr>    <chr> <list>           
-## 1 dkt_3d inflated left  <tibble [36 x 8]>
-## 2 dkt_3d inflated right <tibble [36 x 8]>
-## 3 dkt_3d LCBC     left  <tibble [36 x 8]>
-## 4 dkt_3d LCBC     right <tibble [36 x 8]>
-## 5 dkt_3d white    left  <tibble [36 x 8]>
-## 6 dkt_3d white    right <tibble [36 x 8]>
-\end{verbatim}
 
@@
 
-Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d} to display through a color scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see what the atlas is expecting.
+Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d} to display through a colour scale some descriptive or inferential statistics.
+If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see what the atlas expects.
+
+\small
 
@@
 \StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{lobe, }\OperatorTok{-}\NormalTok{acronym)}
+\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{lobe, }\OperatorTok{-}\NormalTok{acronym, }\OperatorTok{-}\NormalTok{roi, }\OperatorTok{-}\NormalTok{annot) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\CommentTok{## # A tibble: 5 x 7}
+\CommentTok{## atlas surf hemi area colour mesh label}
+\CommentTok{## <chr> <chr> <chr> <chr> <chr> <list> <chr>}
+\CommentTok{## 1 dkt_3d inflat~ right <NA> <NA> <named li~ rh_medialwall}
+\CommentTok{## 2 dkt_3d inflat~ right banks superior ~ #196428 <named li~}
+\NormalTok{rh_bankssts}
+\CommentTok{## 3 dkt_3d inflat~ right caudal anterior~ #7D64A0 <named li~}
+\NormalTok{rh_caudalanteri}\OperatorTok{~}
+\CommentTok{## 4 dkt_3d inflat~ right caudal middle f~ #641900 <named li~}
+\NormalTok{rh_caudalmiddle}\OperatorTok{~}
+\CommentTok{## 5 dkt_3d inflat~ right corpus callosum <NA> <named li~}
+\NormalTok{rh_corpuscallos}\OperatorTok{~}
 \end{Highlighting}
@@
 
-\begin{verbatim}
-## # A tibble: 36 x 9
-## atlas surf hemi area colour mesh label roi annot
-## <chr> <chr> <chr> <chr> <chr> <list> <chr> <chr> <chr>
-## 1 dkt_3d infla~ right <NA> <NA> <list~ rh_medial~ 0001 medialwa~
-## 2 dkt_3d infla~ right banks supe~ #1964~ <list~ rh_bankss~ 0002
-bankssts
-## 3 dkt_3d infla~ right caudal ant~ #7D64~ <list~ rh_caudal~ 0003
-caudalan~
-## 4 dkt_3d infla~ right caudal mid~ #6419~ <list~ rh_caudal~ 0004
-caudalmi~
-## 5 dkt_3d infla~ right corpus cal~ <NA> <list~ rh_corpus~ 0005
-corpusca~
-## 6 dkt_3d infla~ right cuneus #DC14~ <list~ rh_cuneus 0006 cuneus
-## 7 dkt_3d infla~ right entorhinal #DC14~ <list~ rh_entorh~ 0007
-entorhin~
-## 8 dkt_3d infla~ right fusiform #B4DC~ <list~ rh_fusifo~ 0008
-fusiform
-## 9 dkt_3d infla~ right inferior p~ #DC3C~ <list~ rh_inferi~ 0009
-inferior~
-## 10 dkt_3d infla~ right inferior t~ #B428~ <list~ rh_inferi~ 0010
-inferior~
-## # ... with 26 more rows
-\end{verbatim}
+\normalsize
 
@@
 \StringTok{           }\KeywordTok{round}\NormalTok{(}\DecValTok{2}\NormalTok{))}
-
+  \CommentTok{# Figure 8}
   \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }\DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{)}
@@
 
-\includegraphics[width=0.4\linewidth]{png/ggseg3d_data}
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{png/ggseg3d_data} \caption{Supplying data so ggseg3d works in much the same way as for ggseg.
+Since ggseg3d is based on plotly rather than ggplot, much adaptation of the plots aesthetics must be done immediately in the main function.
+Here we set the parcellation colours with `colour` and accompanying mouse-hover text with `text`.}\label{fig:ggseg3d-data1}
+\end{figure}
 
@@
 Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-          \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"#ff0000"}\NormalTok{, }\StringTok{"#00ff00"}\NormalTok{, }\StringTok{"#0000ff"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics[width=0.4\linewidth]{png/ggseg3d_colour}
-
 If you want to alter the colour of \texttt{NA} regions, supply \texttt{na.colour}, either as HEX colour or colour name.
@@
 \begin{Highlighting}[]
+\CommentTok{# Figure 9 left}
 \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
           \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-          \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{camera=}\StringTok{"medial"}\NormalTok{)}
+          \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"#ff0000"}\NormalTok{, }\StringTok{"#00ff00"}\NormalTok{, }\StringTok{"#0000ff"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{   }
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
+
+\CommentTok{# Figure 9 right}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
+          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
+          \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{%>%}\StringTok{   }
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
 \end{Highlighting}
@@
 
-\includegraphics[width=0.4\linewidth]{png/ggseg3d_nacolour}
+\begin{figure}
+\includegraphics[width=0.4\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.4\linewidth]{png/ggseg3d_nacolour} \caption{Left: Specifying the palette for ggseg3d must be done directly in the main call to the plot, rather than a separate function later like in ggseg.
+The `palette` option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
+Right: The colour of the `NA` values can also be changed through the option `na.colour`.
+Here the medial wall (which is `NA`), has been coloured black using this option.}\label{fig:ggseg3d-nacol}
+\end{figure}
+
+\hypertarget{glass-brain}{%
+\subsection{Glass brain}\label{glass-brain}}
 
@@
 Additionally, one can add the \texttt{glassbrain} option, which provides a frame of reference for the subcortical structures.
-Glassbrains can be controlled with three different arguments: opacity (0-1), hemisphere (left, right), and colour.
+Glass brains can be added with the function \texttt{add\_glassbrain()} which takes three extra arguments: hemisphere, colour, and opacity.
 
@@
 \begin{Highlighting}[]
-\NormalTok{somData_aseg =}\StringTok{ }\NormalTok{aseg_3d }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(label) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"Ventricle|Putamen"}\NormalTok{, label)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{nrow}\NormalTok{(.)))}
-
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ somData_aseg, }\DataTypeTok{atlas =}\NormalTok{ aseg_3d,}
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{, }\DataTypeTok{glassbrain =} \FloatTok{.5}\NormalTok{,}
-          \DataTypeTok{glassbrain_hemisphere =} \StringTok{"left"}\NormalTok{)}
+\CommentTok{# Figure 10}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg_3d,}
+          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{%>%}\StringTok{   }
+\StringTok{  }\KeywordTok{add_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
 \end{Highlighting}
@@
 
-\includegraphics[width=0.4\linewidth]{png/ggseg3d_glass}
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{When visualising subcortical structures, it might be an idea to add a glassbrain to the plot.
+This will help with understanding structure locations in the brain and make the plot easier to interpret.
+The glassbrain is controlled by three options: opacity:0-1, hemisphere left, right or both, and colour.}\label{fig:glassbrain1}
+\end{figure}
 
@@
 
-The main aim of the ggseg -- and ggsegExtra - package is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
-In this tutorial, we introduce the package to the readers by presenting some use examples and highlighting the main functions and options that are available.
+The main aim of the ggseg, ggseg3d, and ggsegExtra packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
+In this tutorial, we introduce the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
 As a visualization tool, this package adds up to manifold functionalities such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}) in R, and software-specific image viewers such as FSLeyes (\citet{fsleyes}) and Freeview (\citet{dale_99}).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of ggseg as a visualization tool.
+In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the ggseg-packages as visualization tools.
 After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (\citet{binder}) or Shiny (\citet{shiny}), is an added benefit.
+An example of this is in the online supplementary information of Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}).
 
-The ggseg and ggsegExtra packages contain three main features:
+The ggseg, ggseg3d, and ggsegExtra packages contain three main features:
 \textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make i easy to use for ones own purposes.
+The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
 \textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
-\texttt{ggseg()} as a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2} and it can be built upon like any ggplot object.
-\texttt{ggseg3d()} as a plotly wrapper for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
+\texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \texttt{ggplot2} and it can be built upon like any ggplot object.
+\texttt{ggseg3d()} is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from ggplot and plotly, respectively.
 \textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
-These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository.
+These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository in ggsegExtra.
 
-The foundations of the ggseg-package trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
-(\citeyearpar{vidal_2019})) .
-That is, ggseg appears with the need to inspect and display 4D information - including a spatial dimension - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
+The foundations of the ggseg-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019})) .
+That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
 The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
-Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public \href{https://github.com/LCBC-UiO/ggseg}{ggseg GitHub repository} and \href{https://github.com/LCBC-UiO/ggsegExtra}{ggsegExtra GitHub repository}.
-See X and Y for detailed guidelines on how to create new atlases.
-Finally - while the ggseg-package is circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
+Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public \href{https://github.com/LCBC-UiO/ggseg}{ggseg GitHub repository}, \href{https://github.com/LCBC-UiO/ggseg3d}{ggseg3d GitHub repository}, and \href{https://github.com/LCBC-UiO/ggsegExtra}{ggsegExtra GitHub repository}.
+
+Finally - while the ggseg-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
 We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (\citet{gganatogram}).
+
+\hypertarget{planned-package-improvements}{%
+\section{Planned package improvements}\label{planned-package-improvements}}
+
+In the package \href{https://github.com/LCBC-UiO/ggseg/wiki}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
+However, the current way to do this is convoluted and difficult, and we are putting efforts into improving the way atlases can be made more easily.
+The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid ggseg atlases.
+We encourage users to contribute to the ggsegExtra brain atlas repository by including additional brain atlases.
 
@@
 Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into \texttt{.ply} files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
+Additionally, Inge Amlien has provided helpful pointers to plot Freesurfer images in R, which resulted in us finally managing to plot mesh segmented brains in R.
+He also provided the `LCBC' surface files, which can be found in the ggsegExtra github repository (\citet{ggsegExt12_lcbc}).
 We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
@@
 This work is funded by EU Horizon 2020 Grant `Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
+The project has also received funding from the European Research Council's Starting Grant scheme under grant agreements 283634, 725025 (to center PI A.M.
+Fjell), and 313440 (to center PI K.B.
+Walhovd).
 
@@
 
-Some of the content presented here also appears in the package vignette of \texttt{ggseg}, which may be accessed through R or in the package website (\citet{ggseg}).
+Some of the content presented here also appears in the package vignette of \texttt{ggseg} and \texttt{ggseg3d}, which may be accessed through R or in the package websites (\citet{ggseg}, \citet{ggseg3d}).
 Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (\citet{ggsegAnim}, \citet{ggsegIntro}).

### 2019-11-28 test.Rmd
revised based on didacs comments

--- 2019-05-02 draft_2_DVP.Rmd
@@
 abstract: >
-  The abstract of the article.
-  It can also be on *multiple* lines.
+  There is an increasing emphasis on visualising results from statistical analysis in more intuitive ways.
+Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualisation of these data should reflect these same dimensional properties.
+The ggseg package makes it possible to visualise pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
+Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.
+In this tutorial, we present the data and functions in the ggseg R-package for brain atlas visualisation.
+The tutorial highlights the main functions to create brain segmentation plots in R, and also introduces the accompanying package data that makes these visualisations possible.
 bibliography: references.bib
@@
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts).
-The projection of data into brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and informs on the spatial patterns of the data.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts) alone.
+The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
 Adopting the grammar of graphics implemented in ggplot2 (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
@@
 
-The ggseg package presents both compiled data sets, tailored scripts that allow brain data integration and plotting and, other minor features such as custom color palettes.
+The ggseg package presents both compiled data sets, tailored scripts that allow brain data integration and plotting, and other minor features such as custom color palettes.
 The data featured in the ggseg package are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
@@
 dd %>%
-  kable(format="latex", escape = FALSE, booktabs = TRUE,
+  mutate(citation = case_when(grepl("dkt", Item) ~ "@dkt")) %>%
+  kable(format="latex", escape = TRUE, booktabs = TRUE,
         caption = "Table of currently available atlases in the ggseg and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlase representations.") %>%
-  collapse_rows(1)
+  #collapse_rows(1) %>% 
+  kable_styling(latex_options = c("scale_down"))
 ```
-
-# Tutorial
-This tutorial will introduce the `ggseg` and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
-The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
-
-## Plotting polygon data (ggplot2)
-
-`ggseg` is the main function for plotting 2D data.
-By default, the function automatically plots the DKT atlas (see Figure \@ref(fig:init)).
-The `ggseg` function is a wrapper for `geom_polygon` from `ggplot2`, and it can be built upon and combined like any ggplot object.
-The image plot consists of a simple brain representation containing no extra information.
-Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options.
-We recommend users to get familiarized with ggplot2 grammar (@ggplot).
-
-```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in shaded polygons."}
-library(ggseg)
-library(tidyverse)
-ggseg()
-```
-
-As `ggseg` is implemented as a fully-functional part of ggplot, one can combine the function together with any other ggplot function such as themes or scales.
-Also, one has several options for plotting the main brain representations.
-These options are atlas-specific.
-For cortical atlases, such as the 'dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
-See Figure \@ref(fig:collection)) for an example.
-For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagital, and coronal views.
-
-```{r collection, fig.width=5, fig.cap="ggseg plots can be used with any ggplot feasture such as standard scales and themes.
-For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
-**A:** dkt atlas, stacked with dark theme ; **B:** dkt with medial view only; **C:** dkt atlas with only left medial display; **D:** dkt atlas, stacked, with classic theme; **E:** dkt atlas with left hemisphere only; **F:** complete aseg atlas"}
-# dkt dark theme
-p1 <- ggseg(position = "stacked") +
-  theme_dark() +
-  labs(title=" ")
-
-# dkt classic theme
-p2 <- ggseg(position = "stacked") +
-  theme_classic() +
-  labs(title = " ")
-
-# dkt medial view
-med <- ggseg(view = "medial") +
-  labs(title = " ")
-
-# dkt left hemisphere
-left <- ggseg(hemisphere = "left") +
-  labs(title = " ")
-
-# aseg default theme
-p3 <- ggseg(atlas=aseg) +
-  labs(title = " ")
-
-# dkt left medial alone
-combo <- ggseg(view = "medial",
-      hemisphere = "left") +
-  labs(title=" ")
-
-# Combine plots
-cowplot::plot_grid(p1, med, combo,  p2, left, p3,
-                   labels = c("A: dkt - dark", "B: dkt - medial",
-                              "C: dkt - combo", "D: dkt - classic",
-                              "E: dkt - left", "F: aseg"),
-                   hjust = -.05)
-
-```
-
-### Using own data with fill and colour
-`ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single value or within the ggplot mapping option `aes`.
-To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales.
-See an example in Figure \@ref(fig:fill).
-
-```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
-The 'dkt' atlas default palette corresponds to the FreeSurferColorLut scheme."}
-ggseg(mapping=aes(fill = area), colour="black") +
-  scale_fill_brain("dkt") +
-  theme(legend.justification=c(1,0),
-        legend.position="bottom",
-        legend.text = element_text(size = 5)) +
-  guides(fill = guide_legend(ncol = 3))
-```
-
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
-The atlas dataset structure will help us understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique data set.
-All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
-
-```{r, fig.cap="All ggseg polygon atlases are required to have at minimum 6 columns with the names: 'long', 'lat', 'id', 'hemi', 'area', 'side'.
-These columns are necessary for the ggseg function to work.
-The additional columns in the dkt were added for various practical reasons.", size='small'}
-dkt
-```
-
-In any atlas, the column 'label' is particularly useful for combining the data of interest with the `ggseg`-polygons.
-The column 'label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
-Yet the data in `ggseg` is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
-Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
-See below an example of wide-to-long conversion.
-
-```{r }
-freesurfer_stats <- data.frame(
-  id = c(10:12),
-  lh_superiortemporal = c(3.32, 4.1, 3.5),
-  lh_precentral = c(2.3, 2.5, 2.1),
-  lh_rostralmiddlefrontal = c(3.3, 3.2, 3.1)
-  )
-freesurfer_stats
-
-freesurfer_long <- freesurfer_stats %>%
-  gather(label, thickness, -id)
-freesurfer_long
-```
-
-Data in long format can then be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
-The data **must** include a column that has the same name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of a hypothetical analysis.
-The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins.
-We use the `p` column as the column flooding the segment with colour.
-
-```{r datasupp, fig.width=5}
-someData = data.frame(
-  area = c("transverse temporal", "insula",
-           "pre central","superior parietal"),
-  p = sample(seq(0,.5,.001), 4),
-  stringsAsFactors = FALSE)
-
-p <- ggseg(.data=someData, mapping=aes(fill=p))
-p
-```
-
-The appearance of this plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
-
-```{r datasupp2, fig.width=5}
-p +
-  theme_void() +
-  scale_fill_gradient(low="firebrick",high="goldenrod") +
-  labs(title="A nice plot title", fill="p-value")
-```
-
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.fame includes the column `hemi` with either 'right' or 'left' for this to happen.
-In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
-
-```{r datasuppX, fig.width=5}
-someData$hemi = "right"
-
-ggseg(.data=someData, colour="white",mapping=aes(fill=p)) +
-  theme_void() +
-  scale_fill_gradient(low="firebrick",high="goldenrod") +
-  labs(title="Another nice plot title")
-```
-
-### Creating subplots
-There is often the need to plot a statistic of interest in different groups (e.g. thickness or brain - cognition relationships in young or older adults).
-This may be obtained also with ggseg, using ggplot's `facet_wrap` or `facet_grid`, using two guiding rules:
-**1)** as before, data needs to be in long format (group data should appear in seperate rows, not in separate columns).
-**2)** The data needs to be grouped using `dplyr`'s `group_by` function _before_ providing the data to the `ggseg` function.
-The ggseg function will detect grouped data, and adapt it to `facet`'s requirements.
-
-```{r datasupp3, fig.width=5}
-someData = data.frame(
-  area = rep(c("transverse temporal", "insula",
-               "pre central","superior parietal"),2),
-  p = sample(seq(0,.5,.001), 8),
-  AgeG = c(rep("Young",4), rep("Old",4)),
-  stringsAsFactors = FALSE) %>%
-  group_by(AgeG)
-
-ggseg(.data=someData, colour="white", mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=1) +
-  theme(legend.position = "bottom")
-```
-
-As before, one can apply the different ggplot and facet options, such as `scales` to modify the plot's appearance.
-
-```{r datasupp5, fig.width=5}
-ggseg(.data = someData, atlas=dkt,
-      colour="white", position="stacked",
-      mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=2) +
-  theme_dark() +
-  theme(legend.position = "bottom",
-        axis.text = element_blank(),
-        axis.title = element_blank()
-        ) +
-  scale_fill_gradient2(high="goldenrod",
-                       low="firebrick",
-                       na.value="transparent")
-```
-
-All the concepts described above also work with the 'aseg' atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases such as aseg.
-Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas. 
-Note that both the equivalent structures between 'aseg' and the 'dkt' atlases.
-
-```{r , size = 'small', out.width='90%'}
-aseg
-```
-
-```{r atlases, fig.width=5, fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
-ggseg(atlas="aseg", mapping=aes(fill=area))
-```
-
-## Plotting 3D mesh data
-
-Representing brains as 2D polygons is a good solution for fast, efficient, and flexible plotting, and can be easily combined with interactive apps such as Shiny (@shiny).
-Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
-This problem is exacerbated in atlases that represent subcortical features – after all, cortical surfaces are 2-dimensional – such as grey matter structures or white matter tracts.
-Hence, here we also provide the `ggseg3d` function to plot, view, and print 3D-atlases in R.
-`ggseg3d` is based on tri-surface mesh plots using plotly (@plotly).
-The data structure is somewhat more complex than the ggplot polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
-As `ggseg3d` is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
-We recommend users to familiarize themselves with plotly (@plotly) when using this function.
-
-Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface.
-The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage_ template subject.
-See *SI SCRIPT* for reproducibility.
-All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
-
-```{r ggseg3d-1-out, message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3 , fig.cap="The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (10 iterations), and the 'inflated' surface is a fully inflated white matter cortical surface."}
-p <- ggseg3d(surface = "white")  %>%
-  layout(title="\nggseg3d(surface='white')",
-         font =  list(size = 30))
-orca(p, file = "png/ggseg3d_white.png", format = "png")
-
-p <- ggseg3d()  %>%
-  layout(title="\nggseg3d(surface='LCBC')",
-         font =  list(size = 30))
-orca(p, file = "png/ggseg3d.png", format = "png")
-
-p <- ggseg3d(surface = "inflated") %>%
-  layout(title="\nggseg3d(surface='inflated')",
-         font =  list(size = 30))
-orca(p, file = "png/ggseg3d_inflated.png", format = "png")
-
-include_graphics(c("png/ggseg3d_white.png","png/ggseg3d.png", "png/ggseg3d_inflated.png"))
-```
-
-The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d_1_out)) and the two hemispheres.
-Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
-The 'ggseg_3d' column includes all necessary information to `ggseg3d()` to create a 3D mesh-plot.
-The additional 3D-atlases in ggsegExtra have the same data structure.
-
-```{r, out.width='90%', size='small'}
-dkt_3d
-```
-
-### External data supply
-
-Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a color scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas expects.
-
-```{r out.width='40%', linewidth=70}
-dkt_3d %>%
-  filter(surf == "inflated" & hemi == "right") %>%
-  unnest(ggseg_3d) %>%
-  select(-lobe, -acronym)
-```
-
-Note the `mesh` column, which contains lists.
-Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
-It should also be noted that the 'label', 'annot' and 'area' columns could provide matching values for your own data.
-Similarly to the `ggseg`-function, the 'label' column should match the region names used in the original neuroimaging software while 'area' and 'annot' provide alternative/secondary names.
-It is thus important to match your regional identifiers with those used in the atlas.
-A warning will be issued in case of any mismatch.
-The column you want to use for colour, needs to be supplied to the `colour` option, and you'll likely want to supply it to the `text` option, as this will add another line to the plotly hover information.
-
-```{r, eval=FALSE, out.width='40%'}
-someData = dkt_3d %>%
-  filter(surf == "inflated" &
-         hemi == "right") %>%
-  unnest(ggseg_3d) %>%
-  select(area) %>%
-  na.omit() %>%
-  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
-           round(2))
-
-  ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
-```
-
-```{r, out.width='40%', echo=FALSE}
-someData = dkt_3d %>%
-  filter(surf == "inflated" & hemi == "right") %>%
-  unnest(ggseg_3d) %>%
-  select(area) %>%
-  na.omit() %>%
-  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
-           round(2))
-
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-             colour = "p", text = "p")
-include_plotly(p, "png/ggseg3d_data.png")
-```
-
-### Colours
-
-You can provide custom colour palettes either in hex or R-names.
-Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.
-
-```{r,fig.width=5, eval=FALSE, out.width='45%'}
-  ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          palette = c("#ff0000", "#00ff00", "#0000ff"))
-```
-```{r, out.width='40%', echo=F, eval=T}
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          palette = c("#ff0000", "#00ff00", "#0000ff"))
-include_plotly(p, "png/ggseg3d_colour.png")
-```
-
-If you want to alter the colour of `NA` regions, supply `na.colour`, either as HEX colour or colour name.
-This option only takes a single colour.
-
-```{r, out.width='40%', eval=FALSE,  out.width='45%'}
-  ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          na.colour = "black", camera="medial")
-```
-```{r, out.width='40%', echo=F, eval=T}
-p <-   ggseg3d(.data = someData, atlas = dkt_3d,
-          colour = "p", text = "p",
-          na.colour = "black", camera="medial")
-include_plotly(p, "png/ggseg3d_nacolour.png")
-```
-
-Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these `NA` structures, to improve visualization.
-Additionally, one can add the `glassbrain` option, which provides a frame of reference for the subcortical structures.
-Glassbrains can be controlled with three different arguments: opacity (0-1), hemisphere (left, right), and colour.
-
-```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
-somData_aseg = aseg_3d %>%
-  unnest() %>%
-  select(label) %>%
-  filter(!grepl("Ventricle|Putamen", label)) %>%
-  mutate(p = seq(1, nrow(.)))
-
-  ggseg3d(.data = somData_aseg, atlas = aseg_3d,
-          colour = "p", text = "p",
-          na.alpha= .5, glassbrain = .5,
-          glassbrain_hemisphere = "left")
-```
-```{r, out.width='40%', echo=F, eval=T}
-somData_aseg = aseg_3d %>%
-  unnest() %>%
-  select(label) %>%
-  filter(!grepl("Ventricle|Putamen", label)) %>%
-  mutate(p = seq(1, nrow(.)))
-
-p <- ggseg3d(.data = somData_aseg, atlas = aseg_3d,
-          colour = "p", text = "p",
-          na.alpha= .5, glassbrain = .5,
-          glassbrain_hemisphere = "left")
-include_plotly(p, "png/ggseg3d_glass.png")
-```
-
-`ggsed3d()` is based on plotly and thus additional plotly functionalities can be used to modify and improve the 3D atlas representations.
-In addition to Carson Sievert's book on plotly in R (-@plotly), we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with plotly in R (@plotly-trisurf).
-Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line tool to save ggseg3d atlas snapshots.
-
-# Discussion
-
-The main aim of the ggseg – and ggsegExtra - package is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
-In this tutorial, we introduce the package to the readers by presenting some use examples and highlighting the main functions and options that are available.
-As a visualization tool, this package adds up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of ggseg as a visualization tool.
-After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
-On the contrary, we believe the ggseg niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
-The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
-
-The ggseg and ggsegExtra packages contain three main features:
-**1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
-**2)** `ggseg()` and `ggseg3d()` functions for visualization.
-`ggseg()` is a wrapper function for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
-`ggseg3d()` is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to its environment as can be combined with any additional argument from ggplot and plotly, respectively.
-**3)** Complimentary features – e.g.
-color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format.
-These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository.
-
-The foundations of the ggseg-package trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
-(-@vidal_2019)) .
-That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial diemension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
-bar plots).
-The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
-Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public [ggseg GitHub repository](https://github.com/LCBC-UiO/ggseg) and [ggsegExtra GitHub repository](https://github.com/LCBC-UiO/ggsegExtra).
-See X and Y for detailed guidelines on how to create new atlases.
-Finally - while the ggseg-package is circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
-We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (@gganatogram).
-
-# Conclusion
-Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation and, communicate with colleagues and general audience.
-In this tutorial, we have introduced the ggseg-package, a tool for visualizing brain statistics through brain parcellation atlases in R.
-This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines.
-We hope this tool and tutorial proves useful to neuroscientists and inspires others to apply the functions in a wide variety of fields and structures.
-
-# Author Contributions
-Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts that made the plots.
-He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
-Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
-She is also responsible for conceiving and adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible.
-A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.
-
-# Conflicts of Interest
-The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
-
-# Acknowledgements
-We wish to thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
-A.
-M.
-Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into `.ply` files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
-We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
-We are indebted to those users that actively contributed to the package and all the scientists and engineers that worked on the framework on which the package is sustained (e.g.
-R-project, neuroimaging software).
-
-# Funding
-This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
-
-# Prior versions
-Some of the content presented here also appears in the package vignette of `ggseg`, which may be accessed through R or in the package website (@ggseg).
-Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (@ggsegAnim, @ggsegIntro).
-
-# References

### 2019-11-28 draft_2.Rmd
title and abstract

@@
 ---
-title: Visualisation of Brain Statistics with R-package `ggseg`
+title: Visualisation of Brain Statistics with R-packages `ggseg` and `ggseg3d`
 preprint: true
@@
 Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualization of these data should reflect these same dimensional properties.
-The ggseg package makes it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
+The ggseg-packages make it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
 Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.

### 2019-11-28 draft_2.tex
title and abstract

@@
 
-\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
+\title{Visualisation of Brain Statistics with R-packages \texttt{ggseg} and \texttt{ggseg3d}}
 

### 2019-12-11 draft_2.Rmd
revised as discussed

--- 2019-11-28 draft_2.Rmd
@@
 ---
-title: Visualisation of Brain Statistics with R-package `ggseg`
+title: Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_
 preprint: true
@@
     email: a.m.mowinckel@psykologi.uio.no
-  - name: Didac Vidal Piñeiro
+  - name: Didac Vidal-Piñeiro
     affiliation: 1
@@
 abstract: >
-  There is an increasing emphasis on visualizing results from statistical analysis in more intuitive ways.
-Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualization of these data should reflect these same dimensional properties.
-The ggseg package makes it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
-Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.
-In this tutorial, we present the data and functions in the ggseg and ggseg3d R-packages for brain atlas visualization.
-The tutorial highlights the main functions to create brain segmentation plots in R, and also introduces the accompanying package data that makes these visualizations possible.
+  There is increased emphasis on visualizing results from statistical analysis in more intuitive ways.
+Results from imaging data inherently have a spatial component.
+Commonly used tools to disseminate results (i.e.
+bar-charts or similar) ignore this spatial component.
+A tool that easily integrates the spatial aspect of the data could greatly improve dissemination of results, and also increase efficiency of workflows.
+The ggseg-packages make it possible to visualize pre-defined brain segmentations as both 2D polygons or 3D meshes.
+Presenting both the relative position and size of segments in addition to the statistical summary information eases interpretation and dissemination of results.
+Here we present the main data and functions in the _ggseg_ and _ggseg3d_ R-packages for brain atlas visualization.
+The main highlighted functions create brain segmentation plots in R, and datasets included in the packages, as well as in the _ggsegExtra_ companion package, fascilitate these visualizations.
 bibliography: references.bib
@@
 ---
+
 ```{r setup, include=F}
-knitr::opts_chunk$set(eval=TRUE, echo=TRUE, collapse = TRUE )#, fig.pos = 'H')
+knitr::opts_chunk$set(eval=TRUE, echo=TRUE, collapse = TRUE, results = 'hold', cache=TRUE, fig.pos = 'H') 
 options(tinytex.verbose = TRUE)
@@
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
-Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or on connectivity properties (@eickhoff_2018).
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on structural or connectivity properties (@eickhoff_2018).
 The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons. 
-This facilitates replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).
+This enables replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).
 
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g. bar charts) alone.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, thus it is desirable that any visual representation respects this relation.
 The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
-Adopting the grammar of graphics implemented in ggplot2 (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
+Adopting the grammar of graphics implemented in _ggplot2_ (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
 Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The ggseg and ggseg3d – pluss the complimentary ggsegExtra – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
-Two-dimensional functionality in ggseg is based on polygons and ggplot2-based grammar of graphics (@ggplot), while the 3D functionality in ggseg3d is based on tri-surface mesh plots and plotly (@plotly).
+The _ggseg_ and _ggseg3d_ – pluss the complimentary _ggsegExtra_ – packages include pre-compiled datasets for different brain atlases that allow for 2D and 3D visualization.
+The two-dimensional functionality in _ggseg_ is based on polygons and _ggplot2_-based grammar of graphics (@ggplot), while the 3D functionality in _ggseg3d_ is based on tri-surface mesh plots and _plotly_ (@plotly).
 
-Both packages present compiled data sets, tailored functions that allow brain data integration, and plotting, as well as other minor features such as custom colour palettes.
+Both packages present compiled datasets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
 The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The ggsegExtra package includes a library of precompiled atlases (currently including `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is updated continuously.
-See Table \@ref(tab:atlasTab) for a summary of the available atlases (as of `r format(Sys.Date(), "%B %Y")`).
+The _ggsegExtra_ package contains a collection of precompiled atlases (currently `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is updated frequently.
+A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in \@ref(tab:atlasTab).
 
-```{r "atlasTab", echo=F, warning=FALSE, message=FALSE}
+```{r 'atlasTab', echo=F, warning=FALSE, message=FALSE}
 dd %>%
@@
         caption = "Table of currently available atlases in the ggseg, ggseg3d, and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlas representations, respectively") %>%
-  #collapse_rows(1) %>% 
-  #column_spec(1, width="20em") %>% 
   kable_styling(latex_options = c("scale_down"))
@@
 # Tutorial
-This tutorial will introduce the `ggseg`, `ggseg3d`, and `ggsegExtra` packages and familiarize the reader with the main functions and the general use of the packages.
+This tutorial will introduce the _ggseg_, _ggseg3d_, and _ggsegExtra_ packages and familiarize the reader with the main functions and the general use of the packages.
 The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
@@
 
-`ggseg` is the main function for plotting 2D data.
+_ggseg_ is the main function for plotting 2D data.
 By default, the function automatically plots the DKT atlas (see Figure \@ref(fig:init)).
-The `ggseg` function is a wrapper for `geom_polygon` from `ggplot2`, and it can be built upon and combined like any ggplot object.
+The `ggseg()`-function is a wrapper for `geom_polygon()` from _ggplot2_, and it can be built upon and combined like any _ggplot_-object.
 The image plot consists of a simple brain representation containing no extra information.
-Hence, `ggseg` plots can be easily complemented with any of the available ggplot features and options.
-We recommend users to get familiarized with ggplot2 grammar (@ggplot).
+Hence, _ggseg_ plots can be easily complemented with any of the available _ggplot2_ features and options.
+We recommend users to get familiarized with _ggplot2_ (@ggplot).
 The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (@cran) some time during 2020.
 
-```{r "init", fig.width=3, message=FALSE, fig.cap="By default `ggseg` will plot the dkt atlas in grey shaded polygons."}
+```{r 'init', fig.width=3, message=FALSE, fig.cap='By default `ggseg()` will plot the dkt atlas in grey shaded polygons.'}
 # remotes::install_github("LCBC-UiO/ggseg")
 library(ggseg)
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 
@@
 
-In addition to the standard options for ggplot polygon geoms, the function also has several options for plotting the main brain representations.
+In addition to the standard options for _ggplot2_ polygon geoms, the function also has several options for plotting the main brain representations.
 These options are atlas-specific.
-For cortical atlases, such as the 'dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
-See Figure \@ref(fig:collection)) for examples.
-For subcortical atlases, such as the 'aseg', the options are more limited but one can often choose between axial, sagittal, and coronal views.
+For cortical atlases, such as the `dkt`, one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:collection) for examples).
+For subcortical atlases, such as the `aseg` atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
 
-```{r collection, fig.width=5, fig.cap="ggseg plots can be used with any ggplot feasture such as standard scales and themes.
-For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
-**A:** dkt atlas, stacked with dark theme ; **B:** dkt with medial view only; **C:** dkt atlas with only left medial display; **D:** dkt atlas, stacked, with classic theme; **E:** dkt atlas with left hemisphere only; **F:** complete aseg atlas"}
+```{r 'collection', fig.width=5, fig.cap='*ggseg* plots can be used with any *ggplot2* feature such as standard scales and themes.
+For cortical atlases, one can supply special *ggseg* options to determine hemisphere or view, and control how the plots looks like and is organised.
+**A:** `dkt` atlas, stacked with dark theme ; **B:** `dkt` with medial view only; **C:** `dkt` atlas with only left medial display; **D:** `dkt` atlas, stacked, with classic theme; **E:** `dkt` atlas with left hemisphere only; **F:** complete `aseg` atlas'}
 # dkt dark theme
@@
                    hjust = -.05)
-
 ```
@@
 ### Using own data with fill and colour
-`ggseg` accepts any argument you can supply to `geom_polygon` and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the ggplot mapping function `aes`.
-To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales.
-See an example in Figure \@ref(fig:fill).
+`ggseg()` accepts any argument you can supply to `geom_polygon()`, and therefore is easy to work with for those familiar with _ggplot2_ functionality.
+Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the _ggplot2_ mapping function `aes`.
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales (Figure \@ref(fig:fill)).
 
-```{r fill, fig.width=5, fig.height=8, fig.cap="Supplying 'area' to the fill option in ggseg, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
-The 'dkt' atlas default palette corresponds to the FreeSurferColorLut scheme."}
+```{r 'fill', fig.width=5, fig.height=5, fig.cap="Supplying 'area' to the fill option in `ggseg()`, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The `dkt` atlas default palette corresponds to the FreeSurferColorLut scheme."}
 # Figure 3
@@
 
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
-The atlas data set structure will help us understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique data set.
-All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Most users will use `ggseg()` to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas datasets.
+The atlas dataset structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique dataset.
+All datasets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
-```{r, fig.cap="All ggseg polygon atlases are required to have at minimum 6 columns with the names: 'long', 'lat', 'id', 'hemi', 'area', 'side'.
-These columns are necessary for the ggseg function to work.
-The additional columns in the dkt were added for various practical reasons.", size='small'}
+```{r, size='small', eval=FALSE}
+# Look at the top 5 rows of the dkt dataset
 head(dkt, 5)
@@
 
-In any atlas, the column 'label' is particularly useful for combining the data of interest with the `ggseg`-polygons.
+In any atlas, the column 'label' is particularly useful for combining the data of interest with the _ggseg_-polygons.
 The column 'label' contains the label (region) names as in the original neuroimaging software.
 For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
-Yet the data in `ggseg` is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
-Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+Yet, the data in _ggseg_ is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
+Often datasets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
 See below an example of wide-to-long conversion.
 
-```{r, fig.cap="Transforming wide data from Freesurfer to comply with the long data format of ggplot"}
+```{r, size='small'}
+# Create a mock freesurfer output
 freesurfer_stats <- data.frame(
@@
 
+# Wrangle wide format data into long format
 freesurfer_long <- freesurfer_stats %>%
@@
 
-Data in long format can then be used directly with the `ggseg` function, as the `label` column corresponds in name and content with the `label` column in the atlas data of dkt.
-The data **must** include a column that has the same name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an 'area' and 'p' column, representing the results of a hypothetical analysis.
-The `ggseg` function will recognise the matching column 'area', and merge the supplied data into the atlas using `dplyr` joins.
-We use the `p` column as the column flooding the segment with colour.
-The appearance of the plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
+Data in long format can then be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dkt`.
+The data **must** include a column that has the same column name and at least _some_ data matching the values in the corresponding column in the atlas data.
+In the next example we create some data with 4 rows, with an 'area' and 'p' column, representing the results of a hypothetical analysis.
+The `ggseg()`-function will recognise the matching column 'area', and merge the supplied data into the atlas using _dplyr_-joins.
+We use the 'p' column as the column flooding the segment with colour.
+The appearance of the plot can then be modified similarly to any other _ggplot2_ graph using functions such as scales, labs, themes, etc., as seen in Figure \@ref(fig:datasupp)
 
-```{r datasupp, fig.width=5, fig.cap="Supplying data through the .data option in ggseg will make it possible to use columns in the supplied data to aestheticals (such as fill) in ggseg.
-The ggseg object can be used with any other polygon compatible function from ggplot or ggplot extentions, like here adding title, changing the legend name and the colour scheme with standard ggplot2 functions."}
+```{r 'datasupp', fig.width=5, fig.cap="Supplying data through the '.data' option in `ggseg()` will make it possible to use columns in the supplied data to aesthetical arguments (such as 'fill').
+The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions."}
+# Make some mock data
 someData = data.frame(
@@
 
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi` with either 'right' or 'left' for this to happen.
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column 'hemi' with either 'right' or 'left' for this to happen.
 In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
+For more information about adapting data and vieweing only one hemisphere or side, the [package vignettes](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.
 
@@
 There is often the need to plot a statistic of interest in different groups (e.g. thickness or brain - cognition relationships in young or older adults).
-This may be obtained also with ggseg, using ggplot's `facet_wrap` or `facet_grid`, using two guiding rules:
-**1)** as before, data needs to be in long format (group data should appear in seperate rows, not in separate columns).
-**2)** The data needs to be grouped using `dplyr`'s `group_by` function _before_ providing the data to the `ggseg` function.
-The ggseg function will detect grouped data, and adapt it to `facet`'s requirements.
+This may be obtained also with `ggseg()`, using _ggplot2_'s `facet_wrap` or `facet_grid`, with three guiding rules: 
+**1)** as before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in seperate rows, not in separate columns).
+**2)** The data needs to be grouped using _dplyr_'s `group_by()` function _before_ providing the data to the `ggseg()`-function.
+The `ggseg()`-function will detect grouped data, and adapt it to `facet`'s requirements. 
+**3)** Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules. 
+An example of this can be seen in Figure \@ref(fig:datasupp3), where a mock dataset including summary statistics for two groups ('Young' and 'Old') is used when facetting a _ggseg_-plot.
 
-```{r datasupp3, fig.width=5, fig.cap="Data with different groups need the data to be grouped_by before providing the data to ggseg"}
+All the concepts described above also work with the `aseg` atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:atlases)).
+
+```{r 'datasupp3', fig.width=5, fig.cap='Data including different groups need to be grouped using `group_by()` **before** plottinh with `ggseg()` for facetting to work.'}
+# Make some mock data
 someData = data.frame(
@@
 
-All the concepts described above also work with the 'aseg' atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases such as aseg.
-Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
-Note the equivalent structures between 'aseg' and the 'dkt' atlases.
-
-```{r , size = 'small', out.width='90%', fig.cap="The aseg atlashas no different sides to view."}
-head(aseg, 5)
-```
-
-```{r atlases, fig.width=5,out.width='60%', fig.cap="The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way."}
+```{r 'atlases', fig.width=5, fig.cap='The `aseg` atlas, showing subcortical structures, has some distinct differences from the `dkt`.
+For instance, there is no option to show only a single hemisphere, as the resulting plot would be very difficult to understand.
+Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.'}
 # Figure 6
-ggseg(atlas="aseg", mapping=aes(fill=area))
+ggseg(atlas="aseg", mapping=aes(fill=area)) + 
+  theme(legend.justification=c(1,0),
+        legend.position="bottom",
+        legend.text = element_text(size = 5)) +
+  guides(fill = guide_legend(ncol = 3))
 ```
@@
 Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
-This problem is exacerbated in atlases that represent subcortical features – after all, cortical surfaces are 2-dimensional – such as grey matter structures or white matter tracts.
-Hence, here we also provide the `ggseg3d` package to plot, view, and print 3D-atlases in R.
-`ggseg3d` is based on tri-surface mesh plots using plotly (@plotly).
-The data structure is somewhat more complex than the ggplot polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
-As `ggseg3d` is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
-We recommend users to familiarize themselves with plotly (@plotly) when using this function.
+This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical strucctures, such as grey matter structures or white matter tracts, can be flattened to 2-dimensions.
+Hence, here we also provide the _ggseg3d_ package to plot, view, and print 3D-atlases in R.
+_ggseg3d_ is based on tri-surface mesh plots using _plotly_ (@plotly).
+The data structure is more complex than the _ggplot2_ polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
+As _ggseg3d_ is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
+We recommend users to familiarize themselves with _plotly_ (@plotly) when using this function.
 
-Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface.
+Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \@ref(fig:ggseg3d-1-out))
 The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage_ template subject.
-See *SI SCRIPT* for reproducibility.
-All `[...]_3d` atlases have a built in `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
+All `[...]_3d` atlases include a `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
 
-```{r ggseg3d-1-out, message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3 , fig.cap="The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface."}
+```{r "ggseg3d-1-out", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap="The three surface options provided in *ggseg3d* atlases.
+**From left:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface."}
 p <- ggseg3d(surface = "white")  %>%
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes() %>% 
@@
 p <- ggseg3d()  %>%
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes() %>% 
@@
 p <- ggseg3d(surface = "inflated") %>%
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes() %>% 
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
+Each cortical atlas has datasets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
 Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
-The 'ggseg_3d' column includes all necessary information for `ggseg3d()` to create a 3D mesh-plot, but that the user should not attempt to alter.
-The additional 3D-atlases in ggsegExtra have the same data structure.
-It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological  coordinate system, but rather Cartesian plot coordinates without further meaning.
+The 'ggseg_3d' column includes all necessary information for `ggseg3d()` to create a 3D mesh-plot, and should not be modified by the user.
+The additional 3D-atlases in _ggsegExtra_ have the same data structure.
+It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
 
 ```{r, out.width='90%', size='small'}
-# remotes::install_github("LCBC-UiO/ggseg")
+# remotes::install_github("LCBC-UiO/ggseg3d")
 library(ggseg3d)
-dkt_3d
 ```
@@
 
-Similarly as in the 2D-atlas, the user will use `ggseg3d` to display through a colour scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see what the atlas expects.
+Similarly as in the 2D-atlas, the user will use `ggseg3d()` to display through a colour scale some descriptive or inferential statistic.
+If the data is not already in the correct long format, or uses similar naming as the atlas, The users should inspect the atlas data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see how the atlas data is organised.
 
-```{r out.width='40%', linewidth=70, size='small'}
+```{r out.width='80%', linewidth=100, size='small', eval=FALSE}
+# Select surface and hemisphere, and then unnest to inspect the atlas data
 dkt_3d %>%
@@
   unnest(ggseg_3d) %>%
-  select(-lobe, -acronym, -roi, -annot) %>% 
   head(5)
@@
 It should also be noted that the 'label', 'annot' and 'area' columns could provide matching values for your own data.
-Similarly to the `ggseg`-function, the 'label' column should match the region names used in the original neuroimaging software while 'area' and 'annot' provide alternative/secondary names.
+Similarly to the `ggseg()`-function, the 'label' column should match the region names used in the original neuroimaging software while 'area' and 'annot' provide alternative/secondary names.
 It is thus important to match your regional identifiers with those used in the atlas.
-A warning will be issued in case of any mismatch.
-The column you want to use for colour, needs to be supplied to the `colour` option, and you'll likely want to supply it to the `text` option, as this will add another line to the plotly hover information.
+To colour the segments using a column from the data, that column name needs to be supplied to the `colour` option, and providing it to the `text` option will add another line to the _plotly_ hover information.
 
 ```{r, eval=FALSE, out.width='40%'}
-someData = dkt_3d %>%
-  filter(surf == "inflated" &
-         hemi == "right") %>%
-  unnest(ggseg_3d) %>%
-  select(area) %>%
-  na.omit() %>%
-  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
-           round(2))
-  # Figure 8
+# Figure 8 left
 ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
-```
 
-```{r ggseg3d-data1, out.width='60%', echo=FALSE, fig.cap="Supplying data so ggseg3d works in much the same way as for ggseg.
-Since ggseg3d is based on plotly rather than ggplot, much adaptation of the plots aesthetics must be done immediately in the main function.
-Here we set the parcellation colours with `colour` and accompanying mouse-hover text with `text`."}
-someData = dkt_3d %>%
-  filter(surf == "inflated" & hemi == "right") %>%
-  unnest(ggseg_3d) %>%
-  select(area) %>%
-  na.omit() %>%
-  mutate(p = sample(seq(0,.5, length.out = 100 ), nrow(.)) %>%
-           round(2))
-
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-             colour = "p", text = "p") %>%   
-  pan_camera("lateral") %>% 
-  remove_axes()
-include_plotly(p, "png/ggseg3d_data.png")
-```
-
-### Colours
-
-You can provide custom colour palettes either in hex or R-names.
-Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.
-If you want to alter the colour of `NA` regions, supply `na.colour`, either as HEX colour or colour name.
-This option only takes a single colour.
-
-```{r,fig.width=5, eval=FALSE, out.width='45%'}
-# Figure 9 left
-ggseg3d(.data = someData, atlas = dkt_3d,
+# Figure 8 middle
+ggseg3d(.data = someData, 
+        atlas = dkt_3d,
         colour = "p", text = "p",
         palette = c("#ff0000", "#00ff00", "#0000ff")) %>%   
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes()
 
-# Figure 9 right
-ggseg3d(.data = someData, atlas = dkt_3d,
+# Figure 8 right
+ggseg3d(.data = someData, 
+        atlas = dkt_3d,
         colour = "p", text = "p",
         na.colour = "black") %>%   
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes()
@@
 
-```{r ggseg3d-nacol, out.width='40%', echo=F, eval=T,fig.show='hold', fig.ncol = 2, fig.cap="Left: Specifying the palette for ggseg3d must be done directly in the main call to the plot, rather than a separate function later like in ggseg.
-The `palette` option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
-Right: The colour of the `NA` values can also be changed through the option `na.colour`.
-Here the medial wall (which is `NA`), has been coloured black using this option."}
+```{r ggseg3d-data1, out.width='30%', echo=FALSE, fig.cap="**Left:** Supplying data to `ggseg3d()` works in much the same way as for `ggseg()`.
+Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function.
+Here we set the parcellation colours with the 'colour' column and accompanying mouse-hover text with the 'text' column.
+**Middle:** Specifying the palette for `ggseg3d()` must be done directly in the main call to the plot.
+The 'palette' option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
+**Right:** The colour of the `NA` values can also be changed through the option 'na.colour', which is here set to be 'black'", fig.show='hold', fig.ncol = 3}
+p <- ggseg3d(.data = someData, atlas = dkt_3d,
+             colour = "p", text = "p") %>%   
+  pan_camera("right lateral") %>% 
+  remove_axes()
+orca(p, file = "png/ggseg3d_data.png", format = "png")
+
 p <- ggseg3d(.data = someData, atlas = dkt_3d,
@@
              palette = c("#ff0000", "#00ff00", "#0000ff")) %>%   
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes()
@@
                na.colour = "black") %>%   
-  pan_camera("medial") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes()
@@
 
-include_graphics(c("png/ggseg3d_colour.png","png/ggseg3d_nacolour.png"))
+include_graphics(c(
+  "png/ggseg3d_data.png",
+  "png/ggseg3d_colour.png",
+  "png/ggseg3d_nacolour.png"))
+
 ```
 
-## Glass brain
+### Colours
+
+You can provide custom colour palettes either in hex or R-names, as seen in Figure \@ref(fig:ggseg3d-data1).
+Colours will be evenly spaced when creating the colour-scale.
+A palette may also be supplied as a _named numeric vector_, where the vector names are the colours that users wish to use, and the numeric values are the breakpoint for each colour (e.g.
+`c("red" = 0, "white" = 0.5, "blue" = 1)`).
+This way the users can control the minimum and maximum values of the colour scale, and also how the gradient is applied.
+If another colour than the default gray is wanted for the `NA` regions, supply 'na.colour', either as HEX colour or colour name.
+This option only takes a single colour.
+
+### Glass brain
 Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
 One can control the opacity of the these `NA` structures, to improve visualization.
-Additionally, one can add the `glassbrain` option, which provides a frame of reference for the subcortical structures.
-Glass brains can be added with the function `add_glassbrain()` which takes three extra arguments: hemisphere, colour, and opacity.
+Glassbrains can be added provides a frame of reference for the subcortical structures with the function `add_glassbrain()` (Figure \@ref(fig:glassbrain1)) , which takes three extra arguments: hemisphere, colour, and opacity.
 
 ```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
-# Figure 10
+# Figure 9
 ggseg3d(atlas = aseg_3d,
@@
   add_glassbrain("left") %>% 
-  pan_camera("lateral") %>% 
+  pan_camera("left lateral") %>% 
   remove_axes()
 ```
-```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="When visualising subcortical structures, it might be an idea to add a glassbrain to the plot.
-This will help with understanding structure locations in the brain and make the plot easier to interpret.
-The glassbrain is controlled by three options: opacity:0-1, hemisphere left, right or both, and colour."}
+
+```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="For subcortical structure visualisation, one can add a glassbrain to the plot.
+This will help with understanding structure locations relative to the cortex and make the plot easier to interpret.
+The glassbrain is controlled by three options: opacity, hemisphere, and colour."}
 p <- ggseg3d(atlas = aseg_3d,
@@
   add_glassbrain("left") %>% 
-  pan_camera("lateral") %>% 
+  pan_camera("right lateral") %>% 
   remove_axes()
@@
 
-`ggsed3d()` is based on plotly and thus additional plotly functionalities can be used to modify and improve the 3D atlas representations.
-In addition to Carson Sievert's book on plotly in R (-@plotly), we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with plotly in R (@plotly-trisurf).
-Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line tool to save ggseg3d atlas snapshots.
+`ggsed3d()` is based on _plotly_ and thus additional _plotly_ functionalities can be used to modify and improve the 3D atlas representations.
+In addition to Carson Sievert's book on _plotly_ in R (-@plotly), we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with _plotly_ in R (@plotly-trisurf).
+Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line tool to save _ggseg3d_ atlas snapshots.
+
+## Additional atlases
+The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dkt` (@dkt) and `aseg` (@aseg) atlases, default atlases used by the commonly used imaging software Freesurfer (@@fischl_99, @dale_99, @Fischl2000).
+These are, however, only two among many meaningful ways of segmenting the brain into different regions. 
+Thus, the _ggsegExtra_ package is library with additional datasets for plotting with the _ggseg_ and _ggseg3d_ packages. 
+We hope to expand this library through a combination of our own and community effort, as the need arises for new atlases, as brain new brain atlases are continuously being released.
+A small collection of the atlases currently in the _ggsegExtra_ package may be viewed in Figure \@ref(fig:ggsegExtra)
+
+```{r,  include = FALSE}
+library(ggsegExtra)
+k1 <- ggseg(atlas = yeo7 , mapping=aes(fill=area),
+            show.legend = FALSE, position="stacked") +
+  scale_fill_brain("yeo7") +
+  labs(title="yeo7-dataset") 
+ggsave(filename = "png/ggseg_extra1.png", plot = k1, device = "png")
+
+k2 <- ggseg(atlas = glasser , mapping=aes(fill=area),
+            show.legend = FALSE, position="stacked") +
+  scale_fill_brain("glasser") +
+  labs(title="glasser-dataset")
+kk <- cowplot::plot_grid(k1, k2)
+ggsave(filename = "png/ggseg_extra.png", plot = kk, device = "png")
+
+p <- ggseg3d(atlas = desterieux_3d) %>%   
+  pan_camera("right lateral") %>% 
+  remove_axes() %>% 
+  add_annotations(
+    text="icbm_3d dataset",
+    y = 1,
+    yanchor = 'top',
+    legendtitle=TRUE, showarrow=FALSE,
+    font = list(family = 'sans serif',
+                size = 50)) 
+orca(p, file = "png/ggseg3d_extra1.png", format = "png")
+
+p <- ggseg3d(atlas = icbm_3d) %>%   
+  pan_camera("right lateral") %>%
+  add_glassbrain("left") %>% 
+  remove_axes() %>% 
+  add_annotations(
+    text="icbm_3d dataset",
+    y = 1,
+    yanchor = 'top',
+    legendtitle=TRUE, showarrow=FALSE,
+    font = list(family = 'sans serif',
+                size = 50)) 
+orca(p, file = "png/ggseg3d_extra2.png", format = "png")
+```
+
+```{r "ggsegExtra", out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='Four example datasets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.'}
+include_graphics(c(
+  "png/ggseg_extra.png",
+  "png/ggseg3d_extra1.png",
+  "png/ggseg3d_extra2.png"
+))
+```
 
@@
 
-The main aim of the ggseg, ggseg3d, and ggsegExtra packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
-In this tutorial, we introduce the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
-As a visualization tool, this package adds up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the ggseg-packages as visualization tools.
+The main aim of the _ggseg_, _ggseg3d_, and _ggsegExtra_ packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing customized plotting functions.
+In this tutorial, we introduced the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
+As visualization tools, these packages add up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
+In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the _ggseg_-packages as visualization tools.
 After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
-On the contrary, we believe the ggseg niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
+On the contrary, we believe the _ggseg_ niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
@@
 
-The ggseg, ggseg3d, and ggsegExtra packages contain three main features:
+The _ggseg_, _ggseg3d_, and _ggsegExtra_ packages contain three main features:
 **1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases. 
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
+The atlas data include the necessary coordinates for plotting, but are made more user-friendly.
 **2)** `ggseg()` and `ggseg3d()` functions for visualization.
-`ggseg()` is a wrapper function for `geom_polygon` from `ggplot2` and it can be built upon like any ggplot object.
-`ggseg3d()` is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from ggplot and plotly, respectively.
-**3)** Complimentary features – e.g.
-color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format.
-These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository in ggsegExtra.
+`ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object.
+`ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from _ggplot2_ and _plotly_, respectively.
+**3)** Complimentary features – e.g. color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format. 
+These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in _ggsegExtra_.
 
-The foundations of the ggseg-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
+The foundations of the _ggseg_-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
 (-@vidal_2019)) .
-That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
+That is, _ggseg_ appears with the need to inspect and display brain information over time - i.e.
+including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
 bar plots).
@@
 
-Finally - while the ggseg-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
-We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (@gganatogram).
+Finally - while the _ggseg_-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
+We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the _gganatogram_-package (@gganatogram).
 
 # Planned package improvements
-In the package [github wiki](https://github.com/LCBC-UiO/ggseg/wiki), we offer a pipeline to create and supply atlases for 2D plotting.
-However, the current way to do this is convoluted and difficult, and we are putting efforts into improving the way atlases can be made more easily.
-The aim is to create a set of functions that will call specialized tools like FSL(@fsl), Freesurfer(@fischl_99, @dale_99, @Fischl2000) and Imagemagick(@magick) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid ggseg atlases.
-We encourage users to contribute to the ggsegExtra brain atlas repository by including additional brain atlases.
+In the _ggsegExtra_ [github wiki](https://github.com/LCBC-UiO/ggsegExtra/wiki/Contributing%3A-polygon-atlases-new), we offer a pipeline to create and supply atlases for 2D plotting.
+At the moment, the creation of atlas for ggseg is convoluted and difficult and requires manual intervention. 
+We are in the process of designing a simple, straightforward pipeline to facilitate the creation of new ggseg (2D) atlases.
+The aim is to create a set of functions that will call specialized tools like FSL(@fsl), Freesurfer(@fischl_99, @dale_99, @Fischl2000) and Imagemagick(@magick) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid _ggseg_ and _ggseg3d_ atlases.
+We encourage users to contribute to the _ggsegExtra_ brain atlas repository by including additional brain atlases.
 
 # Conclusion
-Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation and, communicate with colleagues and general audience.
-In this tutorial, we have introduced the ggseg-package, a tool for visualizing brain statistics through brain parcellation atlases in R.
+Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation, and communicate with colleagues and the general audience.
+In this tutorial, we have introduced the _ggseg_-packages, tools for visualizing brain statistics through brain parcellation atlases in R.
 This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines.
@@
 # Author Contributions
-Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts that made the plots.
-He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
+Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts for plot visualization.
+He has also been responsible for converting images from neuroimaging data to ggseg-like data (e.g. polygons and mesh data).
 Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
@@
 # Acknowledgements
-We wish to thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (-@neuroconductor), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
-A.
+We thank John Muschelli for package code comments and its adaptation to neuroconductor ([2018]), and Richard Beare for the first community created atlas, and base code comments. 
+We are indebted to A.
 M.
-Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into `.ply` files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
-Additionally, Inge Amlien has provided helpful pointers to plot Freesurfer images in R, which resulted in us finally managing to plot mesh segmented brains in R.
-He also provided the 'LCBC' surface files, which can be found in the ggsegExtra github repository (@ggsegExt12_lcbc).
-We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
-We are indebted to those users that actively contributed to the package and all the scientists and engineers that worked on the framework on which the package is sustained (e.g.
-R-project, neuroimaging software).
+Winkler scripts for converting neuroimaging data (Freesurfer) into .ply files, a necessary step for neuroimaging to mesh-plots conversion, and to Inge Amlien for the ‘LCBC’ surface files and helpful contributions to the ggseg3d pipeline.
+We are indebted to those users that actively contributed to the package and all the individuals that worked on the framework on which the package is sustained (e.g. R-project, neuroimaging software). 
+Finally, we thank Anders M. Fjell and Kristine B. Walhovd for their encouragement and financial support.
 
@@
 # Prior versions
-Some of the content presented here also appears in the package vignette of `ggseg` and `ggseg3d`, which may be accessed through R or in the package websites (@ggseg, @ggseg3d).
-Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (@ggsegAnim, @ggsegIntro).
+Some of the content presented here also appears in the package vignettes of _ggseg_ and _ggseg3d_, which may be accessed through R or in the package websites (@ggseg, @ggseg3d).
+Athanasia Monika Mowinckel also has several tutorials on her blog regarding _ggseg_ creation and functionality (@ggsegAnim, @ggsegIntro).
 

### 2019-12-11 draft_2.tex
revised as discussed

--- 2019-11-28 draft_2.tex
@@
 \usepackage{xcolor}
+\usepackage[utf8]{inputenc}
+\usepackage{textcomp}
 \usepackage{booktabs}
@@
 
-\title{Visualisation of Brain Statistics with R-package \texttt{ggseg}}
+\title{Visualisation of Brain Statistics with R-packages \emph{ggseg} and \emph{ggseg3d}}
 
@@
 \corrauthor[1]{Athanasia M. Mowinckel}{\href{mailto:a.m.mowinckel@psykologi.uio.no}{\nolinkurl{a.m.mowinckel@psykologi.uio.no}}}
-\author[1]{Didac Vidal Piñeiro}
+\author[1]{Didac Vidal-Piñeiro}
 
@@
 \begin{abstract}
-There is an increasing emphasis on visualizing results from statistical analysis in more intuitive ways.
-Neuroimaging data already have, as a consequence of the source of the data, particular shapes and as such visualization of these data should reflect these same dimensional properties.
-The ggseg package makes it possible to visualize pre-defined brain segmentations of local/structural or connectivity features in both 2D polygons or 3D mesh.
-Presenting both relative position and size of segments in addition to statistical summary information aides interpretation and dissemination of results.
-In this tutorial, we present the data and functions in the ggseg and ggseg3d R-packages for brain atlas visualization.
-The tutorial highlights the main functions to create brain segmentation plots in R, and also introduces the accompanying package data that makes these visualizations possible.
+There is increased emphasis on visualizing results from statistical analysis in more intuitive ways.
+Results from imaging data inherently have a spatial component.
+Commonly used tools to disseminate results (i.e.~bar-charts or similar) ignore this spatial component.
+A tool that easily integrates the spatial aspect of the data could greatly improve dissemination of results, and also increase efficiency of workflows.
+The ggseg-packages make it possible to visualize pre-defined brain segmentations as both 2D polygons or 3D meshes.
+Presenting both the relative position and size of segments in addition to the statistical summary information eases interpretation and dissemination of results.
+Here we present the main data and functions in the \emph{ggseg} and \emph{ggseg3d} R-packages for brain atlas visualization.
+The main highlighted functions create brain segmentation plots in R, and datasets included in the packages, as well as in the \emph{ggsegExtra} companion package, fascilitate these visualizations.
 % Dummy abstract text.
@@
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
-Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on local/structural or on connectivity properties (\citet{eickhoff_2018}).
+Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on structural or connectivity properties (\citet{eickhoff_2018}).
 The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons.
-This facilitates replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (\citet{R}).
+This enables replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (\citet{R}).
 
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, rather than on other types of graphs (e.g.~bar charts) alone.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, thus it is desirable that any visual representation respects this relation.
 The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
-Adopting the grammar of graphics implemented in ggplot2 (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
+Adopting the grammar of graphics implemented in \emph{ggplot2} (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
 Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The ggseg and ggseg3d -- pluss the complimentary ggsegExtra -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
-Two-dimensional functionality in ggseg is based on polygons and ggplot2-based grammar of graphics (\citet{ggplot}), while the 3D functionality in ggseg3d is based on tri-surface mesh plots and plotly (\citet{plotly}).
+The \emph{ggseg} and \emph{ggseg3d} -- pluss the complimentary \emph{ggsegExtra} -- packages include pre-compiled datasets for different brain atlases that allow for 2D and 3D visualization.
+The two-dimensional functionality in \emph{ggseg} is based on polygons and \emph{ggplot2}-based grammar of graphics (\citet{ggplot}), while the 3D functionality in \emph{ggseg3d} is based on tri-surface mesh plots and \emph{plotly} (\citet{plotly}).
 
-Both packages present compiled data sets, tailored functions that allow brain data integration, and plotting, as well as other minor features such as custom colour palettes.
+Both packages present compiled datasets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
 The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
-The ggsegExtra package includes a library of precompiled atlases (currently including 13 additional atlases) and it is updated continuously.
-See Table \ref{tab:atlasTab} for a summary of the available atlases (as of November 2019).
+The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 15 additional atlases) and it is updated frequently.
+A summary of all available atlases compatible with the \emph{ggseg}-packages as of December 2019 can be found in \ref{tab:atlasTab}.
 
@@
 \midrule
+Automated probabilistic reconstruction of white-matter pathways & tracula & None & ggsegExtra & \citet{tracula}\\
 Desikan-Killiany Cortical Atlas & dkt & ggseg3d & ggseg & \citet{dkt}\\
@@
 Genetic topography of brain area morphology & chenAr & None & ggsegExtra & \citet{chen}\\
+\addlinespace
 Genetic topography of brain thickness morphology & chenTh & None & ggsegExtra & \citet{chen}\\
+Harvard-Oxford Cortical atlas & hoCort & None & ggsegExtra & \citet{ho}\\
+HPC - Multi-modal parcellation of human cerebral cortex & glasser & None & ggsegExtra & \citet{glasser}\\
+ICBM white matter parcellation & icbm & ggsegExtra & None & \citet{icbm}\\
+JHU parcellation & jhu & None & ggsegExtra & \citet{jhu}\\
 \addlinespace
-Harvard-Oxford Cortical atlas & hoCort & None & ggsegExtra & \citet{ho}\\
-Parcellation from a midsagittal slice & midsagittal & None & ggsegExtra & \\
-Parcellation from JHU & jhu & ggsegExtra & ggsegExtra & \citet{jhu}\\
-Parcellation from of white matter & icbm & ggsegExtra & None & \citet{icbm}\\
-Parcellation from the Human Connectome Project & glasser & ggsegExtra & ggsegExtra & \citet{glasser}\\
+Local-Global 17 Parcellation of the Human Cerebral Cortex & schaefer17 & ggsegExtra & None & \citet{schaefer}\\
+Local-Global 7 Parcellation of the Human Cerebral Cortex & schaefer7 & ggsegExtra & None & \citet{schaefer}\\
+Parcellation from JHU & jhu & ggsegExtra & None & \citet{jhu}\\
+Parcellation from the Human Connectome Project & glasser & ggsegExtra & None & \citet{glasser}\\
+White matter tract parcellations & tracula & ggsegExtra & None & \citet{tracula}\\
 \addlinespace
-Schaefer 17 Resting-state Cortical Parcellations & schaefer17 & ggsegExtra & None & \citet{schaefer}\\
-Schaefer 7 Resting-state Cortical Parcellations & schaefer7 & ggsegExtra & None & \citet{schaefer}\\
-White matter tract parcellations & tracula & ggsegExtra & ggsegExtra & \citet{tracula}\\
 Yeo 17 Resting-state Cortical Parcellations & yeo17 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
@@
 
-This tutorial will introduce the \texttt{ggseg}, \texttt{ggseg3d}, and \texttt{ggsegExtra} packages and familiarize the reader with the main functions and the general use of the packages.
+This tutorial will introduce the \emph{ggseg}, \emph{ggseg3d}, and \emph{ggsegExtra} packages and familiarize the reader with the main functions and the general use of the packages.
 The tutorial will focus on the two main functions: \texttt{ggseg()} for plotting 2D polygons and \texttt{ggseg3d()} for plotting 3D brains based on tri-surface mesh plots.
@@
 
-\texttt{ggseg} is the main function for plotting 2D data.
+\emph{ggseg} is the main function for plotting 2D data.
 By default, the function automatically plots the DKT atlas (see Figure \ref{fig:init}).
-The \texttt{ggseg} function is a wrapper for \texttt{geom\_polygon} from \texttt{ggplot2}, and it can be built upon and combined like any ggplot object.
+The \texttt{ggseg()}-function is a wrapper for \texttt{geom\_polygon()} from \emph{ggplot2}, and it can be built upon and combined like any \emph{ggplot}-object.
 The image plot consists of a simple brain representation containing no extra information.
-Hence, \texttt{ggseg} plots can be easily complemented with any of the available ggplot features and options.
-We recommend users to get familiarized with ggplot2 grammar (\citet{ggplot}).
+Hence, \emph{ggseg} plots can be easily complemented with any of the available \emph{ggplot2} features and options.
+We recommend users to get familiarized with \emph{ggplot2} (\citet{ggplot}).
 The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (\citet{cran}) some time during 2020.
@@
 \begin{Highlighting}[]
-\CommentTok{# remotes::install_github("LCBC-UiO/ggseg")}
+\CommentTok{\# remotes::install\_github("LCBC{-}UiO/ggseg")}
 \KeywordTok{library}\NormalTok{(ggseg)}
-\KeywordTok{library}\NormalTok{(tidyverse)}
+\KeywordTok{library}\NormalTok{(dplyr)}
+\KeywordTok{library}\NormalTok{(tidyr)}
 
-\CommentTok{# Figure 1}
+\CommentTok{\# Figure 1}
 \KeywordTok{ggseg}\NormalTok{()}
@@
 \includegraphics{draft_2_files/figure-latex/init-1.pdf}
-\caption{\label{fig:init}By default \texttt{ggseg} will plot the dkt atlas in grey shaded polygons.}
+\caption{\label{fig:init}By default \texttt{ggseg()} will plot the dkt atlas in grey shaded polygons.}
 \end{figure}
 
-In addition to the standard options for ggplot polygon geoms, the function also has several options for plotting the main brain representations.
+In addition to the standard options for \emph{ggplot2} polygon geoms, the function also has several options for plotting the main brain representations.
 These options are atlas-specific.
-For cortical atlases, such as the `dkt', one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view.
-See Figure \ref{fig:collection}) for examples.
-For subcortical atlases, such as the `aseg', the options are more limited but one can often choose between axial, sagittal, and coronal views.
+For cortical atlases, such as the \texttt{dkt}, one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \ref{fig:collection} for examples).
+For subcortical atlases, such as the \texttt{aseg} atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# dkt dark theme}
-\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}
+\CommentTok{\# dkt dark theme}
+\NormalTok{p1 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme\_dark}\NormalTok{() }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
 
-\CommentTok{# dkt classic theme}
-\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
+\CommentTok{\# dkt classic theme}
+\NormalTok{p2 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme\_classic}\NormalTok{() }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
 
-\CommentTok{# dkt medial view}
-\NormalTok{med <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
+\CommentTok{\# dkt medial view}
+\NormalTok{med <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
 
-\CommentTok{# dkt left hemisphere}
-\NormalTok{left <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\CommentTok{\# dkt left hemisphere}
+\NormalTok{left <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
 
-\CommentTok{# aseg default theme}
-\NormalTok{p3 <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}
+\CommentTok{\# aseg default theme}
+\NormalTok{p3 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
 
-\CommentTok{# dkt left medial alone}
-\NormalTok{combo <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
+\CommentTok{\# dkt left medial alone}
+\NormalTok{combo <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
                \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
@@
 
-\CommentTok{# Combine plots to Figure 2}
-\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
-                   \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A: dkt - dark"}\NormalTok{, }\StringTok{"B: dkt - medial"}\NormalTok{,}
-                              \StringTok{"C: dkt - combo"}\NormalTok{, }\StringTok{"D: dkt - classic"}\NormalTok{,}
-                              \StringTok{"E: dkt - left"}\NormalTok{, }\StringTok{"F: aseg"}\NormalTok{),}
-                   \DataTypeTok{hjust =} \FloatTok{-.05}\NormalTok{)}
+\CommentTok{\# Combine plots to Figure 2}
+\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot\_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
+                   \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A: dkt {-} dark"}\NormalTok{, }\StringTok{"B: dkt {-} medial"}\NormalTok{,}
+                              \StringTok{"C: dkt {-} combo"}\NormalTok{, }\StringTok{"D: dkt {-} classic"}\NormalTok{,}
+                              \StringTok{"E: dkt {-} left"}\NormalTok{, }\StringTok{"F: aseg"}\NormalTok{),}
+                   \DataTypeTok{hjust =} \FloatTok{{-}.05}\NormalTok{)}
 \end{Highlighting}
@@
 \includegraphics{draft_2_files/figure-latex/collection-1.pdf}
-\caption{\label{fig:collection}ggseg plots can be used with any ggplot feasture such as standard scales and themes.
-For cortical atlases, one can supply special ggseg options to determine hemisphere or view, and control how the plots looks like and is organised.
-\textbf{A:} dkt atlas, stacked with dark theme ; \textbf{B:} dkt with medial view only; \textbf{C:} dkt atlas with only left medial display; \textbf{D:} dkt atlas, stacked, with classic theme; \textbf{E:} dkt atlas with left hemisphere only; \textbf{F:} complete aseg atlas}
+\caption{\label{fig:collection}\emph{ggseg} plots can be used with any \emph{ggplot2} feature such as standard scales and themes.
+For cortical atlases, one can supply special \emph{ggseg} options to determine hemisphere or view, and control how the plots looks like and is organised.
+\textbf{A:} \texttt{dkt} atlas, stacked with dark theme ; \textbf{B:} \texttt{dkt} with medial view only; \textbf{C:} \texttt{dkt} atlas with only left medial display; \textbf{D:} \texttt{dkt} atlas, stacked, with classic theme; \textbf{E:} \texttt{dkt} atlas with left hemisphere only; \textbf{F:} complete \texttt{aseg} atlas}
 \end{figure}
@@
 
-\texttt{ggseg} accepts any argument you can supply to \texttt{geom\_polygon} and therefore is easy to work with for those familiar with ggplot functionality.
-Standard arguments like \texttt{fill} that floods the segments with a colour, or \texttt{colour} that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the ggplot mapping function \texttt{aes}.
-To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales.
-See an example in Figure \ref{fig:fill}.
+\texttt{ggseg()} accepts any argument you can supply to \texttt{geom\_polygon()}, and therefore is easy to work with for those familiar with \emph{ggplot2} functionality.
+Standard arguments like \texttt{fill} that floods the segments with a colour, or \texttt{colour} that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the \emph{ggplot2} mapping function \texttt{aes}.
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales (Figure \ref{fig:fill}).
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure 3}
+\CommentTok{\# Figure 3}
 \KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_fill\_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
         \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
-        \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
+        \DataTypeTok{legend.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
 \includegraphics{draft_2_files/figure-latex/fill-1.pdf}
-\caption{\label{fig:fill}Supplying `area' to the fill option in ggseg, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
-The `dkt' atlas default palette corresponds to the FreeSurferColorLut scheme.}
+\caption{\label{fig:fill}Supplying `area' to the fill option in \texttt{ggseg()}, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
+The \texttt{dkt} atlas default palette corresponds to the FreeSurferColorLut scheme.}
 \end{figure}
 
-Most users will use ggseg to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, we need to explore the structure of the atlas data sets.
-The atlas data set structure will help us understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique data set.
-All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Most users will use \texttt{ggseg()} to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas datasets.
+The atlas dataset structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique dataset.
+All datasets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+
+In any atlas, the column `label' is particularly useful for combining the data of interest with the \emph{ggseg}-polygons.
+The column `label' contains the label (region) names as in the original neuroimaging software.
+For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+Yet, the data in \emph{ggseg} is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
+Often datasets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+See below an example of wide-to-long conversion.
 
@@
 \begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(dkt, }\DecValTok{5}\NormalTok{)}
-\CommentTok{## # A tibble: 5 x 6}
-\CommentTok{##   atlas area               hemi  side    label              ggseg          }
-\CommentTok{##   <chr> <chr>              <chr> <chr>   <chr>              <list>         }
-\CommentTok{## 1 dkt   superior temporal  left  lateral lh_superiortempor~ <tibble [1,494~}
-\CommentTok{## 2 dkt   pre central        left  lateral lh_precentral      <tibble [1,314~}
-\CommentTok{## 3 dkt   post central       left  lateral lh_postcentral     <tibble [1,164~}
-\CommentTok{## 4 dkt   rostral middle fr~ left  lateral lh_rostralmiddlef~ <tibble [1,194~}
-\CommentTok{## 5 dkt   insula             left  lateral lh_insula          <tibble [870 x~}
+\CommentTok{\# Create a mock freesurfer output}
+\NormalTok{freesurfer\_stats <{-}}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
+  \DataTypeTok{id =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\OperatorTok{:}\DecValTok{12}\NormalTok{),}
+  \DataTypeTok{lh\_superiortemporal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.32}\NormalTok{, }\FloatTok{4.1}\NormalTok{, }\FloatTok{3.5}\NormalTok{),}
+  \DataTypeTok{lh\_precentral =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.3}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.1}\NormalTok{),}
+  \DataTypeTok{lh\_rostralmiddlefrontal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.3}\NormalTok{, }\FloatTok{3.2}\NormalTok{, }\FloatTok{3.1}\NormalTok{)}
+\NormalTok{)}
+\NormalTok{freesurfer\_stats}
+
+\CommentTok{\# Wrangle wide format data into long format}
+\NormalTok{freesurfer\_long <{-}}\StringTok{ }\NormalTok{freesurfer\_stats }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{{-}}\NormalTok{id)}
+\NormalTok{freesurfer\_long}
+\CommentTok{\#\#   id lh\_superiortemporal lh\_precentral lh\_rostralmiddlefrontal}
+\CommentTok{\#\# 1 10                3.32           2.3                     3.3}
+\CommentTok{\#\# 2 11                4.10           2.5                     3.2}
+\CommentTok{\#\# 3 12                3.50           2.1                     3.1}
+\CommentTok{\#\#   id                   label thickness}
+\CommentTok{\#\# 1 10     lh\_superiortemporal      3.32}
+\CommentTok{\#\# 2 11     lh\_superiortemporal      4.10}
+\CommentTok{\#\# 3 12     lh\_superiortemporal      3.50}
+\CommentTok{\#\# 4 10           lh\_precentral      2.30}
+\CommentTok{\#\# 5 11           lh\_precentral      2.50}
+\CommentTok{\#\# 6 12           lh\_precentral      2.10}
+\CommentTok{\#\# 7 10 lh\_rostralmiddlefrontal      3.30}
+\CommentTok{\#\# 8 11 lh\_rostralmiddlefrontal      3.20}
+\CommentTok{\#\# 9 12 lh\_rostralmiddlefrontal      3.10}
 \end{Highlighting}
@@
 
-In any atlas, the column `label' is particularly useful for combining the data of interest with the \texttt{ggseg}-polygons.
-The column `label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
-Yet the data in \texttt{ggseg} is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
-Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
-See below an example of wide-to-long conversion.
+Data in long format can then be used directly with the \texttt{ggseg()}-function, as the `label' column corresponds in name and content with the `label' column in the atlas data of \texttt{dkt}.
+The data \textbf{must} include a column that has the same column name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
+In the next example we create some data with 4 rows, with an `area' and `p' column, representing the results of a hypothetical analysis.
+The \texttt{ggseg()}-function will recognise the matching column `area', and merge the supplied data into the atlas using \emph{dplyr}-joins.
+We use the `p' column as the column flooding the segment with colour.
+The appearance of the plot can then be modified similarly to any other \emph{ggplot2} graph using functions such as scales, labs, themes, etc., as seen in Figure \ref{fig:datasupp}
 
@@
 \begin{Highlighting}[]
-\NormalTok{freesurfer_stats <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{id =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\OperatorTok{:}\DecValTok{12}\NormalTok{),}
-  \DataTypeTok{lh_superiortemporal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.32}\NormalTok{, }\FloatTok{4.1}\NormalTok{, }\FloatTok{3.5}\NormalTok{),}
-  \DataTypeTok{lh_precentral =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.3}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.1}\NormalTok{),}
-  \DataTypeTok{lh_rostralmiddlefrontal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.3}\NormalTok{, }\FloatTok{3.2}\NormalTok{, }\FloatTok{3.1}\NormalTok{)}
-\NormalTok{  )}
-\NormalTok{freesurfer_stats}
-\CommentTok{##   id lh_superiortemporal lh_precentral lh_rostralmiddlefrontal}
-\CommentTok{## 1 10                3.32           2.3                     3.3}
-\CommentTok{## 2 11                4.10           2.5                     3.2}
-\CommentTok{## 3 12                3.50           2.1                     3.1}
-
-\NormalTok{freesurfer_long <-}\StringTok{ }\NormalTok{freesurfer_stats }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{-}\NormalTok{id)}
-\NormalTok{freesurfer_long}
-\CommentTok{##   id                   label thickness}
-\CommentTok{## 1 10     lh_superiortemporal      3.32}
-\CommentTok{## 2 11     lh_superiortemporal      4.10}
-\CommentTok{## 3 12     lh_superiortemporal      3.50}
-\CommentTok{## 4 10           lh_precentral      2.30}
-\CommentTok{## 5 11           lh_precentral      2.50}
-\CommentTok{## 6 12           lh_precentral      2.10}
-\CommentTok{## 7 10 lh_rostralmiddlefrontal      3.30}
-\CommentTok{## 8 11 lh_rostralmiddlefrontal      3.20}
-\CommentTok{## 9 12 lh_rostralmiddlefrontal      3.10}
-\end{Highlighting}
-\end{Shaded}
-
-Data in long format can then be used directly with the \texttt{ggseg} function, as the \texttt{label} column corresponds in name and content with the \texttt{label} column in the atlas data of dkt.
-The data \textbf{must} include a column that has the same name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, and an `area' and `p' column, representing the results of a hypothetical analysis.
-The \texttt{ggseg} function will recognise the matching column `area', and merge the supplied data into the atlas using \texttt{dplyr} joins.
-We use the \texttt{p} column as the column flooding the segment with colour.
-The appearance of the plot can then be modified similarly to any other ggplot graph using functions such as scales, labs, themes, etc.
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\CommentTok{\# Make some mock data}
 \NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
@@
 
-\CommentTok{# Figure 4}
+\CommentTok{\# Figure 4}
 \KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))  }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p-value"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p{-}value"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_fill\_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}
 \end{Highlighting}
@@
 \includegraphics{draft_2_files/figure-latex/datasupp-1.pdf}
-\caption{\label{fig:datasupp}Supplying data through the .data option in ggseg will make it possible to use columns in the supplied data to aestheticals (such as fill) in ggseg.
-The ggseg object can be used with any other polygon compatible function from ggplot or ggplot extentions, like here adding title, changing the legend name and the colour scheme with standard ggplot2 functions.}
+\caption{\label{fig:datasupp}Supplying data through the `.data' option in \texttt{ggseg()} will make it possible to use columns in the supplied data to aesthetical arguments (such as `fill').
+The \emph{ggseg} plot can be used with any other polygon compatible function from \emph{ggplot2} or \emph{ggplot2} extentions, for instance adding title, changing the legend name and the colour scheme with standard \emph{ggplot2} functions.}
 \end{figure}
 
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column \texttt{hemi} with either `right' or `left' for this to happen.
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi' with either `right' or `left' for this to happen.
 In this case, data will be merged into the atlas both by `area' and by `hemi'.
@@
 There is often the need to plot a statistic of interest in different groups (e.g.~thickness or brain - cognition relationships in young or older adults).
-This may be obtained also with ggseg, using ggplot's \texttt{facet\_wrap} or \texttt{facet\_grid}, using two guiding rules:
-\textbf{1)} as before, data needs to be in long format (group data should appear in seperate rows, not in separate columns).
-\textbf{2)} The data needs to be grouped using \texttt{dplyr}'s \texttt{group\_by} function \emph{before} providing the data to the \texttt{ggseg} function.
-The ggseg function will detect grouped data, and adapt it to \texttt{facet}'s requirements.
+This may be obtained also with \texttt{ggseg()}, using \emph{ggplot2}'s \texttt{facet\_wrap} or \texttt{facet\_grid}, with three guiding rules:
+\textbf{1)} as before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in seperate rows, not in separate columns).
+\textbf{2)} The data needs to be grouped using \emph{dplyr}'s \texttt{group\_by()} function \emph{before} providing the data to the \texttt{ggseg()}-function.
+The \texttt{ggseg()}-function will detect grouped data, and adapt it to \texttt{facet}`s requirements.
+\textbf{3)} Apply \texttt{facet\_wrap} or \texttt{facet\_grid} to the plot having used the above two rules.
+An example of this can be seen in Figure \ref{fig:datasupp3}, where a mock dataset including summary statistics for two groups ('Young' and `Old') is used when facetting a \emph{ggseg}-plot.
+
+All the concepts described above also work with the \texttt{aseg} atlas for subcortical structures, with the exception of `hemisphere' and `view' arguments that are superfluous in subcortical atlases (Figure \ref{fig:atlases}).
 
@@
 \begin{Highlighting}[]
+\CommentTok{\# Make some mock data}
 \NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
@@
   \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{)),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{group_by}\NormalTok{(AgeG)}
+  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{group\_by}\NormalTok{(AgeG)}
 
-\CommentTok{# Figure 5}
+\CommentTok{\# Figure 5}
 \KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
       \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet\_wrap}\NormalTok{(}\OperatorTok{\textasciitilde{}}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
@@
 \includegraphics{draft_2_files/figure-latex/datasupp3-1.pdf}
-\caption{\label{fig:datasupp3}Data with different groups need the data to be grouped\_by before providing the data to ggseg}
+\caption{\label{fig:datasupp3}Data including different groups need to be grouped using \texttt{group\_by()} \textbf{before} plottinh with \texttt{ggseg()} for facetting to work.}
 \end{figure}
-
-All the concepts described above also work with the `aseg' atlas for subcortical structures, with the exception of `hemisphere' and `view' arguments that are superfluous in subcortical atlases such as aseg.
-Again, inspecting the atlas data directly will be of aid when preparing data to use with the atlas.
-Note the equivalent structures between `aseg' and the `dkt' atlases.
-
-\small
 
@@
 \begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(aseg, }\DecValTok{5}\NormalTok{)}
-\CommentTok{## # A tibble: 5 x 6}
-\CommentTok{##   atlas area             hemi  side  label                ggseg            }
-\CommentTok{##   <chr> <chr>            <chr> <chr> <fct>                <list>           }
-\CommentTok{## 1 aseg  <NA>             right axial <NA>                 <tibble [808 x 5~}
-\CommentTok{## 2 aseg  <NA>             left  axial <NA>                 <tibble [1,014 x~}
-\CommentTok{## 3 aseg  thalamus proper  left  axial Left-Thalamus-Proper <tibble [56 x 5]>}
-\CommentTok{## 4 aseg  thalamus proper  right axial Right-Thalamus-Prop~ <tibble [61 x 5]>}
-\CommentTok{## 5 aseg  lateral ventric~ right axial Right-Lateral-Ventr~ <tibble [66 x 5]>}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Figure 6}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area))}
+\CommentTok{\# Figure 6}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
+        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
+        \DataTypeTok{legend.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{figure}
-\includegraphics[width=0.6\linewidth]{draft_2_files/figure-latex/atlases-1} \caption{The first 10 rows of the `aseg` dataset, that has some specific differences from the dkt, like missing the option to not view both hemispheres, as the plot would be hard to understand that way.}\label{fig:atlases}
+\centering
+\includegraphics{draft_2_files/figure-latex/atlases-1.pdf}
+\caption{\label{fig:atlases}The \texttt{aseg} atlas, showing subcortical structures, has some distinct differences from the \texttt{dkt}.
+For instance, there is no option to show only a single hemisphere, as the resulting plot would be very difficult to understand.
+Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.}
 \end{figure}
@@
 Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
-This problem is exacerbated in atlases that represent subcortical features -- after all, cortical surfaces are 2-dimensional -- such as grey matter structures or white matter tracts.
-Hence, here we also provide the \texttt{ggseg3d} package to plot, view, and print 3D-atlases in R.
-\texttt{ggseg3d} is based on tri-surface mesh plots using plotly (\citet{plotly}).
-The data structure is somewhat more complex than the ggplot polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
-As \texttt{ggseg3d} is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
-We recommend users to familiarize themselves with plotly (\citet{plotly}) when using this function.
+This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical strucctures, such as grey matter structures or white matter tracts, can be flattened to 2-dimensions.
+Hence, here we also provide the \emph{ggseg3d} package to plot, view, and print 3D-atlases in R.
+\emph{ggseg3d} is based on tri-surface mesh plots using \emph{plotly} (\citet{plotly}).
+The data structure is more complex than the \emph{ggplot2} polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
+As \emph{ggseg3d} is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
+We recommend users to familiarize themselves with \emph{plotly} (\citet{plotly}) when using this function.
 
-Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dkt\_3d} atlas in `LCBC' surface.
+Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dkt\_3d} atlas in `LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \ref{fig:ggseg3d-1-out})
 The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage} template subject.
-See \emph{SI SCRIPT} for reproducibility.
-All \texttt{{[}...{]}\_3d} atlases have a built in \texttt{colour} column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
+All \texttt{{[}...{]}\_3d} atlases include a \texttt{colour} column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
 
-\begin{figure}
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in ggseg 3d atlases.
-From left: the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface.}\label{fig:ggseg3d-1-out}
+\begin{figure}[H]
+\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in *ggseg3d* atlases.
+**From left:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface.}\label{fig:ggseg3d-1-out}
 \end{figure}
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
+Each cortical atlas has datasets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
 Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
-The `ggseg\_3d' column includes all necessary information for \texttt{ggseg3d()} to create a 3D mesh-plot, but that the user should not attempt to alter.
-The additional 3D-atlases in ggsegExtra have the same data structure.
-It is important to note that the coordinates in the plot (X, Y, Z) are \textbf{not} any type of radiological coordinate system, but rather Cartesian plot coordinates without further meaning.
+The `ggseg\_3d' column includes all necessary information for \texttt{ggseg3d()} to create a 3D mesh-plot, and should not be modified by the user.
+The additional 3D-atlases in \emph{ggsegExtra} have the same data structure.
+It is important to note that the coordinates in the plot (X, Y, Z) are \textbf{not} any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# remotes::install_github("LCBC-UiO/ggseg")}
+\CommentTok{\# remotes::install\_github("LCBC{-}UiO/ggseg3d")}
 \KeywordTok{library}\NormalTok{(ggseg3d)}
-\NormalTok{dkt_3d}
-\CommentTok{## # A tibble: 6 x 4}
-\CommentTok{##   atlas  surf     hemi  ggseg_3d         }
-\CommentTok{##   <chr>  <chr>    <chr> <list>           }
-\CommentTok{## 1 dkt_3d inflated left  <tibble [36 x 8]>}
-\CommentTok{## 2 dkt_3d inflated right <tibble [36 x 8]>}
-\CommentTok{## 3 dkt_3d LCBC     left  <tibble [36 x 8]>}
-\CommentTok{## 4 dkt_3d LCBC     right <tibble [36 x 8]>}
-\CommentTok{## 5 dkt_3d white    left  <tibble [36 x 8]>}
-\CommentTok{## 6 dkt_3d white    right <tibble [36 x 8]>}
 \end{Highlighting}
@@
 
-Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d} to display through a colour scale some descriptive or inferential statistics.
-If the data is not already in the correct long format, or uses similar naming as the atlas, one needs to grab the data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see what the atlas expects.
+Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d()} to display through a colour scale some descriptive or inferential statistic.
+If the data is not already in the correct long format, or uses similar naming as the atlas, The users should inspect the atlas data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see how the atlas data is organised.
 
@@
 \begin{Highlighting}[]
-\NormalTok{dkt_3d }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{lobe, }\OperatorTok{-}\NormalTok{acronym, }\OperatorTok{-}\NormalTok{roi, }\OperatorTok{-}\NormalTok{annot) }\OperatorTok{%>%}\StringTok{ }
+\CommentTok{\# Select surface and hemisphere, and then unnest to inspect the atlas data}
+\NormalTok{dkt\_3d }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{\&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg\_3d) }\OperatorTok{\%>\%}
 \StringTok{  }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
-\CommentTok{## # A tibble: 5 x 7}
-\CommentTok{## atlas surf hemi area colour mesh label}
-\CommentTok{## <chr> <chr> <chr> <chr> <chr> <list> <chr>}
-\CommentTok{## 1 dkt_3d inflat~ right <NA> <NA> <named li~ rh_medialwall}
-\CommentTok{## 2 dkt_3d inflat~ right banks superior ~ #196428 <named li~}
-\NormalTok{rh_bankssts}
-\CommentTok{## 3 dkt_3d inflat~ right caudal anterior~ #7D64A0 <named li~}
-\NormalTok{rh_caudalanteri}\OperatorTok{~}
-\CommentTok{## 4 dkt_3d inflat~ right caudal middle f~ #641900 <named li~}
-\NormalTok{rh_caudalmiddle}\OperatorTok{~}
-\CommentTok{## 5 dkt_3d inflat~ right corpus callosum <NA> <named li~}
-\NormalTok{rh_corpuscallos}\OperatorTok{~}
 \end{Highlighting}
@@
 It should also be noted that the `label', `annot' and `area' columns could provide matching values for your own data.
-Similarly to the \texttt{ggseg}-function, the `label' column should match the region names used in the original neuroimaging software while `area' and `annot' provide alternative/secondary names.
+Similarly to the \texttt{ggseg()}-function, the `label' column should match the region names used in the original neuroimaging software while `area' and `annot' provide alternative/secondary names.
 It is thus important to match your regional identifiers with those used in the atlas.
-A warning will be issued in case of any mismatch.
-The column you want to use for colour, needs to be supplied to the \texttt{colour} option, and you'll likely want to supply it to the \texttt{text} option, as this will add another line to the plotly hover information.
+To colour the segments using a column from the data, that column name needs to be supplied to the \texttt{colour} option, and providing it to the \texttt{text} option will add another line to the \emph{plotly} hover information.
 
@@
 \begin{Highlighting}[]
-\NormalTok{someData =}\StringTok{ }\NormalTok{dkt_3d }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}
-\StringTok{         }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(area) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{na.omit}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{, }\DataTypeTok{length.out =} \DecValTok{100}\NormalTok{ ), }\KeywordTok{nrow}\NormalTok{(.)) }\OperatorTok{%>%}
-\StringTok{           }\KeywordTok{round}\NormalTok{(}\DecValTok{2}\NormalTok{))}
-  \CommentTok{# Figure 8}
-  \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d, }\DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{)}
+\CommentTok{\# Figure 8 left}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt\_3d, }\DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{)}
+
+\CommentTok{\# Figure 8 middle}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
+        \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
+        \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
+        \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"\#ff0000"}\NormalTok{, }\StringTok{"\#00ff00"}\NormalTok{, }\StringTok{"\#0000ff"}\NormalTok{)) }\OperatorTok{\%>\%}\StringTok{   }
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
+
+\CommentTok{\# Figure 8 right}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
+        \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
+        \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
+        \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
 \end{Highlighting}
@@
 
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{png/ggseg3d_data} \caption{Supplying data so ggseg3d works in much the same way as for ggseg.
-Since ggseg3d is based on plotly rather than ggplot, much adaptation of the plots aesthetics must be done immediately in the main function.
-Here we set the parcellation colours with `colour` and accompanying mouse-hover text with `text`.}\label{fig:ggseg3d-data1}
+\begin{figure}[H]
+\includegraphics[width=0.3\linewidth]{png/ggseg3d_data} \includegraphics[width=0.3\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.3\linewidth]{png/ggseg3d_nacolour} \caption{**Left:** Supplying data to `ggseg3d()` works in much the same way as for `ggseg()`.
+Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function.
+Here we set the parcellation colours with the 'colour' column and accompanying mouse-hover text with the 'text' column.
+**Middle:** Specifying the palette for `ggseg3d()` must be done directly in the main call to the plot.
+The 'palette' option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
+**Right:** The colour of the `NA` values can also be changed through the option 'na.colour', which is here set to be 'black'}\label{fig:ggseg3d-data1}
 \end{figure}
@@
 
-You can provide custom colour palettes either in hex or R-names.
-Colours will be evenly spaced when creating the colour-scale, and you may provide as many colours as you wish.
-If you want to alter the colour of \texttt{NA} regions, supply \texttt{na.colour}, either as HEX colour or colour name.
+You can provide custom colour palettes either in hex or R-names, as seen in Figure \ref{fig:ggseg3d-data1}.
+Colours will be evenly spaced when creating the colour-scale.
+A palette may also be supplied as a \emph{named numeric vector}, where the vector names are the colours that users wish to use, and the numeric values are the breakpoint for each colour (e.g.~\texttt{c("red"\ =\ 0,\ "white"\ =\ 0.5,\ "blue"\ =\ 1)}).
+This way the users can control the minimum and maximum values of the colour scale, and also how the gradient is applied.
+If another colour than the default gray is wanted for the \texttt{NA} regions, supply `na.colour', either as HEX colour or colour name.
 This option only takes a single colour.
+
+\hypertarget{glass-brain}{%
+\subsubsection{Glass brain}\label{glass-brain}}
+
+Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
+One can control the opacity of the these \texttt{NA} structures, to improve visualization.
+Glassbrains can be added provides a frame of reference for the subcortical structures with the function \texttt{add\_glassbrain()} (Figure \ref{fig:glassbrain1}) , which takes three extra arguments: hemisphere, colour, and opacity.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure 9 left}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-          \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"#ff0000"}\NormalTok{, }\StringTok{"#00ff00"}\NormalTok{, }\StringTok{"#0000ff"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
-
-\CommentTok{# Figure 9 right}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt_3d,}
-          \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-          \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{%>%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
+\CommentTok{\# Figure 9}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg\_3d,}
+        \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
+\StringTok{  }\KeywordTok{add\_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"left lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
 \end{Highlighting}
@@
 
-\begin{figure}
-\includegraphics[width=0.4\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.4\linewidth]{png/ggseg3d_nacolour} \caption{Left: Specifying the palette for ggseg3d must be done directly in the main call to the plot, rather than a separate function later like in ggseg.
-The `palette` option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
-Right: The colour of the `NA` values can also be changed through the option `na.colour`.
-Here the medial wall (which is `NA`), has been coloured black using this option.}\label{fig:ggseg3d-nacol}
+\begin{figure}[H]
+\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{For subcortical structure visualisation, one can add a glassbrain to the plot.
+This will help with understanding structure locations relative to the cortex and make the plot easier to interpret.
+The glassbrain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:glassbrain1}
 \end{figure}
 
-\hypertarget{glass-brain}{%
-\subsection{Glass brain}\label{glass-brain}}
+\texttt{ggsed3d()} is based on \emph{plotly} and thus additional \emph{plotly} functionalities can be used to modify and improve the 3D atlas representations.
+In addition to Carson Sievert's book on \emph{plotly} in R (\citeyearpar{plotly}), we recommend resources for modifying axes in 3D plots (\citet{plotly-ax}), the basic introduction to tri-surface plots (\citet{plotly-tri}), and this tutorial on tri-surface plots with \emph{plotly} in R (\citet{plotly-trisurf}).
+Finally, we recommend \href{https://github.com/plotly/orca\#installation}{orca} command line tool to save \emph{ggseg3d} atlas snapshots.
 
-Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these \texttt{NA} structures, to improve visualization.
-Additionally, one can add the \texttt{glassbrain} option, which provides a frame of reference for the subcortical structures.
-Glass brains can be added with the function \texttt{add\_glassbrain()} which takes three extra arguments: hemisphere, colour, and opacity.
+\hypertarget{additional-atlases}{%
+\subsection{Additional atlases}\label{additional-atlases}}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Figure 10}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg_3d,}
-          \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{%>%}\StringTok{   }
-\StringTok{  }\KeywordTok{add_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"lateral"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove_axes}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
+The \emph{ggseg} and \emph{ggseg3d} packages have two atlases each, which are 2D and 3D variations of the same main atlases: the \texttt{dkt} (\citet{dkt}) and \texttt{aseg} (\citet{aseg}) atlases, default atlases used by the commonly used imaging software Freesurfer (@\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}).
+These are, however, only two among many meaningful ways of segmenting the brain into different regions.
+Thus, the \emph{ggsegExtra} package is library with additional datasets for plotting with the \emph{ggseg} and \emph{ggseg3d} packages.
+We hope to expand this library through a combination of our own and community effort, as the need arises for new atlases, as brain new brain atlases are continuously being released.
+A small collection of the atlases currently in the \emph{ggsegExtra} package may be viewed in Figure \ref{fig:ggsegExtra}
 
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{When visualising subcortical structures, it might be an idea to add a glassbrain to the plot.
-This will help with understanding structure locations in the brain and make the plot easier to interpret.
-The glassbrain is controlled by three options: opacity:0-1, hemisphere left, right or both, and colour.}\label{fig:glassbrain1}
+\begin{figure}[H]
+\includegraphics[width=0.8\linewidth]{png/ggseg_extra} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra1} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra2} \caption{Four example datasets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.}\label{fig:ggsegExtra}
 \end{figure}
-
-\texttt{ggsed3d()} is based on plotly and thus additional plotly functionalities can be used to modify and improve the 3D atlas representations.
-In addition to Carson Sievert's book on plotly in R (\citeyearpar{plotly}), we recommend resources for modifying axes in 3D plots (\citet{plotly-ax}), the basic introduction to tri-surface plots (\citet{plotly-tri}), and this tutorial on tri-surface plots with plotly in R (\citet{plotly-trisurf}).
-Finally, we recommend \href{https://github.com/plotly/orca\#installation}{orca} command line tool to save ggseg3d atlas snapshots.
 
@@
 
-The main aim of the ggseg, ggseg3d, and ggsegExtra packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing plotting functions adapted to these atlases.
-In this tutorial, we introduce the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
-As a visualization tool, this package adds up to manifold functionalities such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}) in R, and software-specific image viewers such as FSLeyes (\citet{fsleyes}) and Freeview (\citet{dale_99}).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the ggseg-packages as visualization tools.
+The main aim of the \emph{ggseg}, \emph{ggseg3d}, and \emph{ggsegExtra} packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing customized plotting functions.
+In this tutorial, we introduced the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
+As visualization tools, these packages add up to manifold functionalities such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}) in R, and software-specific image viewers such as FSLeyes (\citet{fsleyes}) and Freeview (\citet{dale_99}).
+In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the \emph{ggseg}-packages as visualization tools.
 After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
-On the contrary, we believe the ggseg niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
+On the contrary, we believe the \emph{ggseg} niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (\citet{binder}) or Shiny (\citet{shiny}), is an added benefit.
@@
 
-The ggseg, ggseg3d, and ggsegExtra packages contain three main features:
+The \emph{ggseg}, \emph{ggseg3d}, and \emph{ggsegExtra} packages contain three main features:
 \textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data are necessary to create the coordinate system for the brain atlas plotting, but are made in such a way to make it easy to use for ones own purposes.
+The atlas data include the necessary coordinates for plotting, but are made more user-friendly.
 \textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
-\texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \texttt{ggplot2} and it can be built upon like any ggplot object.
-\texttt{ggseg3d()} is a plotly wrapper function for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from ggplot and plotly, respectively.
+\texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.
+\texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.
 \textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
-These function should provide users with the possibility of adapting their plots to their wishes, and also to make it possible to create and contribute to the atlas repository in ggsegExtra.
+These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in \emph{ggsegExtra}.
 
-The foundations of the ggseg-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019})) .
-That is, ggseg appears with the need to inspect and display brain information over time - that is, including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
+The foundations of the \emph{ggseg}-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019})) .
+That is, \emph{ggseg} appears with the need to inspect and display brain information over time - i.e.~including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
 The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
@@
 
-Finally - while the ggseg-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
-We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the gganatogram-package (\citet{gganatogram}).
+Finally - while the \emph{ggseg}-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
+We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the \emph{gganatogram}-package (\citet{gganatogram}).
 
@@
 In the package \href{https://github.com/LCBC-UiO/ggseg/wiki}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
-However, the current way to do this is convoluted and difficult, and we are putting efforts into improving the way atlases can be made more easily.
-The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid ggseg atlases.
-We encourage users to contribute to the ggsegExtra brain atlas repository by including additional brain atlases.
+At the moment, the creation of atlas for ggseg is convoluted and difficult and requires manual intervention.
+We are in the process of designing a simple, straightforward pipeline to facilitate the creation of new ggseg (2D) atlases.
+The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid \emph{ggseg} and \emph{ggseg3d} atlases.
+We encourage users to contribute to the \emph{ggsegExtra} brain atlas repository by including additional brain atlases.
 
@@
 
-Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation and, communicate with colleagues and general audience.
-In this tutorial, we have introduced the ggseg-package, a tool for visualizing brain statistics through brain parcellation atlases in R.
+Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation, and communicate with colleagues and the general audience.
+In this tutorial, we have introduced the \emph{ggseg}-packages, tools for visualizing brain statistics through brain parcellation atlases in R.
 This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines.
@@
 
-Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts that made the plots.
-He has also been responsible for converting images from neuroimaging software into the type of data necessary for plotting polygons and mesh plots in R.
+Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts for plot visualization.
+He has also been responsible for converting images from neuroimaging data to ggseg-like data (e.g.~polygons and mesh data).
 Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
@@
 
-We wish to thank John Muschelli for his initial improvements in the package code and preparing it for incorporation in neuroconductor (\citeyearpar{neuroconductor}), and Richard Beare for submitting the first community created atlas, and providing helpful comments to the base code.
-A.
+We thank John Muschelli for package code comments and its adaptation to neuroconductor ({[}2018{]}), and Richard Beare for the first community created atlas, and base code comments.
+We are indebted to A.
 M.
-Winkler, who originally introduced us to the procedure of converting 3D brain images from Freesurfer into \texttt{.ply} files, also needs to be credited with this, we would not have managed to create the mesh-plots without his scripts.
-Additionally, Inge Amlien has provided helpful pointers to plot Freesurfer images in R, which resulted in us finally managing to plot mesh segmented brains in R.
-He also provided the `LCBC' surface files, which can be found in the ggsegExtra github repository (\citet{ggsegExt12_lcbc}).
-We also thank Anders M. Fjell and Kristine B. Walhovd for allowing us to pursue the creation of this tool and encouraging our progress.
-We are indebted to those users that actively contributed to the package and all the scientists and engineers that worked on the framework on which the package is sustained (e.g.~R-project, neuroimaging software).
+Winkler scripts for converting neuroimaging data (Freesurfer) into .ply files, a necessary step for neuroimaging to mesh-plots conversion, and to Inge Amlien for the `LCBC' surface files and helpful contributions to the ggseg3d pipeline.
+We are indebted to those users that actively contributed to the package and all the individuals that worked on the framework on which the package is sustained (e.g.~R-project, neuroimaging software).
+Finally, we thank Anders M. Fjell and Kristine B. Walhovd for their encouragement and financial support.
 
@@
 
-Some of the content presented here also appears in the package vignette of \texttt{ggseg} and \texttt{ggseg3d}, which may be accessed through R or in the package websites (\citet{ggseg}, \citet{ggseg3d}).
-Athanasia Monika Mowinckel also has several tutorials on her blog regarding ggseg creation and functionality (\citet{ggsegAnim}, \citet{ggsegIntro}).
+Some of the content presented here also appears in the package vignettes of \emph{ggseg} and \emph{ggseg3d}, which may be accessed through R or in the package websites (\citet{ggseg}, \citet{ggseg3d}).
+Athanasia Monika Mowinckel also has several tutorials on her blog regarding \emph{ggseg} creation and functionality (\citet{ggsegAnim}, \citet{ggsegIntro}).
 

### 2019-12-17 msc_ggseg.Rmd
preprint submission

--- 2019-12-11 draft_2.Rmd
@@
   - code: 1
-    address: Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway
+    address: Center for Lifespan Changes in Brain and Cognition, University of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway
 abstract: >
-  There is increased emphasis on visualizing results from statistical analysis in more intuitive ways.
-Results from imaging data inherently have a spatial component.
-Commonly used tools to disseminate results (i.e.
-bar-charts or similar) ignore this spatial component.
-A tool that easily integrates the spatial aspect of the data could greatly improve dissemination of results, and also increase efficiency of workflows.
-The ggseg-packages make it possible to visualize pre-defined brain segmentations as both 2D polygons or 3D meshes.
-Presenting both the relative position and size of segments in addition to the statistical summary information eases interpretation and dissemination of results.
-Here we present the main data and functions in the _ggseg_ and _ggseg3d_ R-packages for brain atlas visualization.
-The main highlighted functions create brain segmentation plots in R, and datasets included in the packages, as well as in the _ggsegExtra_ companion package, fascilitate these visualizations.
+  There is an increased emphasis on visualizing neuroimaging results in more intuitive ways.
+Common statistical tools for dissemination, such as bar charts, lack the spatial dimension that is inherent in neuroimaging data.
+Here we present two packages for the statistical software R, _ggseg_ and _ggseg3d_, that integrate this spatial component.
+The _ggseg_ and _ggseg3d_ packages visualize pre-defined brain segmentations as both 2D polygons and 3D meshes, respectively.
+Both packages are integrated with other well-established R-packages, allowing great flexibility.
+In this tutorial, we present the main data and functions in the _ggseg_ and _ggseg3d_ packages for brain atlas visualization.
+The main highlighted functions are able to display brain segmentation plots in R.
+Further, the accompanying _ggsegExtra_-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases to _ggseg_ and _ggseg3d_.
+Overall, the _ggseg_-packages facilitate parcellation-based visualizations in R, improve and ease the dissemination of the results, and increase the efficiency of the workflows.
 bibliography: references.bib
@@
 knitr::opts_chunk$set(eval=TRUE, echo=TRUE, collapse = TRUE, results = 'hold', cache=TRUE, fig.pos = 'H') 
-options(tinytex.verbose = TRUE)
+options(tinytex.verbose = TRUE, tinytex.clean = FALSE)
 
@@
   spread(Type, Package, fill = "None")  %>% 
-  mutate(citation = case_when(
+  mutate(Citation = case_when(
     grepl("dkt", Item) ~ "\\citet{dkt}",
@@
 MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
-Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, thus it is desirable that any visual representation respects this relation.
+Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, thus it is desirable that any visual representation takes this relation into account.
 The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The _ggseg_ and _ggseg3d_ – pluss the complimentary _ggsegExtra_ – packages include pre-compiled datasets for different brain atlases that allow for 2D and 3D visualization.
+The _ggseg_ and _ggseg3d_ – pluss the complimentary _ggsegExtra_ – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in _ggseg_ is based on polygons and _ggplot2_-based grammar of graphics (@ggplot), while the 3D functionality in _ggseg3d_ is based on tri-surface mesh plots and _plotly_ (@plotly).
 
-Both packages present compiled datasets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
+Both packages present compiled data sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
 The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The _ggsegExtra_ package contains a collection of precompiled atlases (currently `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is updated frequently.
-A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in \@ref(tab:atlasTab).
+The _ggsegExtra_ package contains a collection of precompiled atlases (currently `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is frequently updated.
+A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in Table \@ref(tab:atlasTab).
 
@@
 We recommend users to get familiarized with _ggplot2_ (@ggplot).
-The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (@cran) some time during 2020.
+The package is currently only available through github, but we expect to submit the _ggseg_-package to The Comprehensive R Archive Network (@cran) in 2020.
 
@@
 These options are atlas-specific.
-For cortical atlases, such as the `dkt`, one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:collection) for examples).
+For cortical atlases, such as the `dkt`, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:collection) for examples).
 For subcortical atlases, such as the `aseg` atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
@@
 ```{r 'collection', fig.width=5, fig.cap='*ggseg* plots can be used with any *ggplot2* feature such as standard scales and themes.
-For cortical atlases, one can supply special *ggseg* options to determine hemisphere or view, and control how the plots looks like and is organised.
+For cortical atlases, one can supply special *ggseg* options to determine, for example, hemisphere, view, or position.
 **A:** `dkt` atlas, stacked with dark theme ; **B:** `dkt` with medial view only; **C:** `dkt` atlas with only left medial display; **D:** `dkt` atlas, stacked, with classic theme; **E:** `dkt` atlas with left hemisphere only; **F:** complete `aseg` atlas'}
@@
 Most users will use `ggseg()` to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas datasets.
-The atlas dataset structure will help users understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique dataset.
-All datasets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data sets.
+The atlas data set structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data set.
+All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
@@
 For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
-Yet, the data in _ggseg_ is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
-Often datasets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+Yet, the data in _ggseg_ is in a long format - that is each region has a single row - and any data of interest needs to be in this same format.
+Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
 See below an example of wide-to-long conversion.
@@
 
-```{r 'datasupp', fig.width=5, fig.cap="Supplying data through the '.data' option in `ggseg()` will make it possible to use columns in the supplied data to aesthetical arguments (such as 'fill').
+```{r 'datasupp', fig.width=5, fig.cap="Supplying data through the '.data' option in `ggseg()` enables use of columns in the supplied data to aesthetical arguments (such as 'fill').
 The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions."}
@@
 
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column 'hemi' with either 'right' or 'left' for this to happen.
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column 'hemi' with either 'right' or 'left'.
 In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
-For more information about adapting data and vieweing only one hemisphere or side, the [package vignettes](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.
+For more information about adapting data and viewing only one hemisphere or side, the [package vignettes](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.
 
@@
 **3)** Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules. 
-An example of this can be seen in Figure \@ref(fig:datasupp3), where a mock dataset including summary statistics for two groups ('Young' and 'Old') is used when facetting a _ggseg_-plot.
+An example of this can be seen in Figure \@ref(fig:datasupp3), where a mock data set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.
 
-All the concepts described above also work with the `aseg` atlas for subcortical structures, with the exception of 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:atlases)).
+All the concepts described above also work with the `aseg` atlas for subcortical structures, except for 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:atlases)).
 
-```{r 'datasupp3', fig.width=5, fig.cap='Data including different groups need to be grouped using `group_by()` **before** plottinh with `ggseg()` for facetting to work.'}
+```{r 'datasupp3', fig.width=5, fig.cap='To visualize data corresponding to different groups, one needs to use `group_by` or similar **before** plotting with `ggseg()` for facetting to work.'}
 # Make some mock data
@@
 
-```{r 'atlases', fig.width=5, fig.cap='The `aseg` atlas, showing subcortical structures, has some distinct differences from the `dkt`.
-For instance, there is no option to show only a single hemisphere, as the resulting plot would be very difficult to understand.
+```{r 'atlases', fig.width=5, fig.cap='The `aseg` atlas showing subcortical structures has some distinct differences from the `dkt`.
+For instance, there is no option to show only a single hemisphere.
 Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.'}
@@
 Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
-This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical strucctures, such as grey matter structures or white matter tracts, can be flattened to 2-dimensions.
+This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical structures, such as grey matter structures, can be flattened to 2-dimensions.
 Hence, here we also provide the _ggseg3d_ package to plot, view, and print 3D-atlases in R.
@@
 Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \@ref(fig:ggseg3d-1-out))
-The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage_ template subject.
-All `[...]_3d` atlases include a `colour` column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
+The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage5_ template subject.
+All `[...]_3d` atlases include a `colour` column that based on the color scheme used in the source neuroimaging software.
 
-```{r "ggseg3d-1-out", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap="The three surface options provided in *ggseg3d* atlases.
-**From left:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface."}
+(ref:gseg3d-1-out-cap) The three surface options provided in *ggseg3d* atlases.
+**From left to right:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.
+
+```{r "ggseg3d-1-out", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:gseg3d-1-out-cap)'}
 p <- ggseg3d(surface = "white")  %>%
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has datasets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
-Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
+Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
+Only one surface is available for subcortical atlases as inflation procedures are irrelevant.
 The 'ggseg_3d' column includes all necessary information for `ggseg3d()` to create a 3D mesh-plot, and should not be modified by the user.
@@
 Similarly as in the 2D-atlas, the user will use `ggseg3d()` to display through a colour scale some descriptive or inferential statistic.
-If the data is not already in the correct long format, or uses similar naming as the atlas, The users should inspect the atlas data for a specific surface (and hemisphere, if desired) and then `unnest(ggseg_3d)` it to see how the atlas data is organised.
+If the data is not already in the correct long format or uses similar naming as the atlas, the users should inspect the atlas data for a specific surface (and hemisphere, if desired), and then `unnest(ggseg_3d)` it to see how the atlas data is organised.
 
@@
 It is thus important to match your regional identifiers with those used in the atlas.
-To colour the segments using a column from the data, that column name needs to be supplied to the `colour` option, and providing it to the `text` option will add another line to the _plotly_ hover information.
+To colour the segments using a column from the data, a column name from the data needs to be supplied to the `colour` option, and providing it to the `text` option will add another line to the _plotly_ hover information.
 
@@
 
-```{r ggseg3d-data1, out.width='30%', echo=FALSE, fig.cap="**Left:** Supplying data to `ggseg3d()` works in much the same way as for `ggseg()`.
+(ref:ggseg3d-data1-cap) **Left:** Supplying data to `ggseg3d()` works similarly to `ggseg()`.
 Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function.
 Here we set the parcellation colours with the 'colour' column and accompanying mouse-hover text with the 'text' column.
-**Middle:** Specifying the palette for `ggseg3d()` must be done directly in the main call to the plot.
-The 'palette' option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
-**Right:** The colour of the `NA` values can also be changed through the option 'na.colour', which is here set to be 'black'", fig.show='hold', fig.ncol = 3}
+**Middle:** The palette for *ggseg3d* needs to be specified directly in the main plot call.
+The 'palette' option takes vectors of colours either as HEX-codes or R-colour names.
+**Right:** The colour of the `NA` values can also be changed through the option 'na.colour'.
+
+```{r ggseg3d-data1, out.width='30%', echo=FALSE, fig.cap="(ref:ggseg3d-data1-cap)", fig.show='hold', fig.ncol = 3}
 p <- ggseg3d(.data = someData, atlas = dkt_3d,
@@
 
-### Colours
+### Customizing colours and the colour bar
 
@@
 Colours will be evenly spaced when creating the colour-scale.
-A palette may also be supplied as a _named numeric vector_, where the vector names are the colours that users wish to use, and the numeric values are the breakpoint for each colour (e.g.
+A palette may also be supplied as a _named numeric vector_, where the vector names are the colours that users wish to use, and the numeric values are the breakpoints for each colour (e.g.
 `c("red" = 0, "white" = 0.5, "blue" = 1)`).
@@
 
-### Glass brain
+### Adding a glass brain
 Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
 One can control the opacity of the these `NA` structures, to improve visualization.
-Glassbrains can be added provides a frame of reference for the subcortical structures with the function `add_glassbrain()` (Figure \@ref(fig:glassbrain1)) , which takes three extra arguments: hemisphere, colour, and opacity.
+Glass brains can be added to provide a frame of reference for the subcortical structures with the function `add_glassbrain()` (Figure \@ref(fig:glassbrain1)) , which takes three extra arguments: hemisphere, colour, and opacity.
 
@@
 
-```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="For subcortical structure visualisation, one can add a glassbrain to the plot.
-This will help with understanding structure locations relative to the cortex and make the plot easier to interpret.
-The glassbrain is controlled by three options: opacity, hemisphere, and colour."}
+(ref:glassbrain1-cap) For subcortical structure visualization, one can add a glass brain to the plot.
+This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
+The glass brain is controlled by three options: opacity, hemisphere, and colour.
+
+```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="(ref:glassbrain1-cap)"}
 p <- ggseg3d(atlas = aseg_3d,
@@
 ## Additional atlases
-The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dkt` (@dkt) and `aseg` (@aseg) atlases, default atlases used by the commonly used imaging software Freesurfer (@@fischl_99, @dale_99, @Fischl2000).
+The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dkt` (@dkt) and `aseg` (@aseg) atlases. 
 These are, however, only two among many meaningful ways of segmenting the brain into different regions. 
-Thus, the _ggsegExtra_ package is library with additional datasets for plotting with the _ggseg_ and _ggseg3d_ packages. 
-We hope to expand this library through a combination of our own and community effort, as the need arises for new atlases, as brain new brain atlases are continuously being released.
+Thus, the _ggsegExtra_ package is a repository containing additional with additional data sets for plotting with the _ggseg_ and _ggseg3d_ packages. 
+There is an ever-increasing amount of new atlases being created, as research and methods in neuroimaging analysis progresses.
+The _ggsegExtra_-package is intended to be expanded as a community-effort, as new and informative atlases are published.
 A small collection of the atlases currently in the _ggsegExtra_ package may be viewed in Figure \@ref(fig:ggsegExtra)
@@
   scale_fill_brain("yeo7") +
-  labs(title="yeo7-dataset") 
+  labs(title="yeo7-data set") 
 ggsave(filename = "png/ggseg_extra1.png", plot = k1, device = "png")
@@
   scale_fill_brain("glasser") +
-  labs(title="glasser-dataset")
+  labs(title="glasser-data set")
 kk <- cowplot::plot_grid(k1, k2)
@@
   add_annotations(
-    text="icbm_3d dataset",
+    text="icbm_3d data set",
     y = 1,
@@
 
-```{r "ggsegExtra", out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='Four example datasets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.'}
+(ref:ggsegExtra-cap) Four example data sets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.
+
+```{r "ggsegExtra", out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='(ref:ggsegExtra-cap)'}
 include_graphics(c(
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
-An example of this is in the online supplementary information of Vidal-Piñeiro et al. (-@vidal_2019).
+This is exemplified in the online supplementary information of Vidal-Piñeiro et al. (-@vidal_2019).
 
-The _ggseg_, _ggseg3d_, and _ggsegExtra_ packages contain three main features:
+The three _ggseg_-packages contain three main features:
+
 **1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data include the necessary coordinates for plotting, but are made more user-friendly.
+The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.
 **2)** `ggseg()` and `ggseg3d()` functions for visualization.
-`ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object.
-`ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.
 Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from _ggplot2_ and _plotly_, respectively.
+    - `ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object.   
+    - `ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.  
 **3)** Complimentary features – e.g. color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format.
+
 These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in _ggsegExtra_.
 
-The foundations of the _ggseg_-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.
-(-@vidal_2019)) .
+The foundations of the _ggseg_-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supplementary information in Vidal-Piñeiro et al.
+(-@vidal_2019) ) .
 That is, _ggseg_ appears with the need to inspect and display brain information over time - i.e.
@@
 We are indebted to those users that actively contributed to the package and all the individuals that worked on the framework on which the package is sustained (e.g. R-project, neuroimaging software). 
-Finally, we thank Anders M. Fjell and Kristine B. Walhovd for their encouragement and financial support.
+Finally, we thank Anders M. Fjell (AMF) and Kristine B. Walhovd (KBW) for their encouragement and financial support.
 
@@
 
-This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
-The project has also received funding from the European Research Council’s Starting Grant scheme under grant agreements 283634, 725025 (to center PI A.M.
-Fjell), and 313440 (to center PI K.B.
-Walhovd).
+This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement 732592.
+The project has also received funding from the European Research Council’s Starting (grant agreements 283634, to A.M.F.
+and 313440 to K.B.W.) and consolidator Grant Scheme (grant agreement 771355 to KBW and 725025 to AMF).
+The project has received funding through multiple grants from the Norwegian Research Council"
 

### 2019-12-17 msc_ggseg.tex
preprint submission

--- 2019-12-11 draft_2.tex
@@
 
-\affil[1]{Center for Lifespan Changes in Brain and Cognition, Univeristy of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
+\affil[1]{Center for Lifespan Changes in Brain and Cognition, University of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway}
 
@@
 \begin{abstract}
-There is increased emphasis on visualizing results from statistical analysis in more intuitive ways.
-Results from imaging data inherently have a spatial component.
-Commonly used tools to disseminate results (i.e.~bar-charts or similar) ignore this spatial component.
-A tool that easily integrates the spatial aspect of the data could greatly improve dissemination of results, and also increase efficiency of workflows.
-The ggseg-packages make it possible to visualize pre-defined brain segmentations as both 2D polygons or 3D meshes.
-Presenting both the relative position and size of segments in addition to the statistical summary information eases interpretation and dissemination of results.
-Here we present the main data and functions in the \emph{ggseg} and \emph{ggseg3d} R-packages for brain atlas visualization.
-The main highlighted functions create brain segmentation plots in R, and datasets included in the packages, as well as in the \emph{ggsegExtra} companion package, fascilitate these visualizations.
+There is an increased emphasis on visualizing neuroimaging results in more intuitive ways.
+Common statistical tools for dissemination, such as bar charts, lack the spatial dimension that is inherent in neuroimaging data.
+Here we present two packages for the statistical software R, \emph{ggseg} and \emph{ggseg3d}, that integrate this spatial component.
+The \emph{ggseg} and \emph{ggseg3d} packages visualize pre-defined brain segmentations as both 2D polygons and 3D meshes, respectively.
+Both packages are integrated with other well-established R-packages, allowing great flexibility.
+In this tutorial, we present the main data and functions in the \emph{ggseg} and \emph{ggseg3d} packages for brain atlas visualization.
+The main highlighted functions are able to display brain segmentation plots in R.
+Further, the accompanying \emph{ggsegExtra}-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases to \emph{ggseg} and \emph{ggseg3d}.
+Overall, the \emph{ggseg}-packages facilitate parcellation-based visualizations in R, improve and ease the dissemination of the results, and increase the efficiency of the workflows.
 % Dummy abstract text.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The \emph{ggseg} and \emph{ggseg3d} -- pluss the complimentary \emph{ggsegExtra} -- packages include pre-compiled datasets for different brain atlases that allow for 2D and 3D visualization.
+The \emph{ggseg} and \emph{ggseg3d} -- pluss the complimentary \emph{ggsegExtra} -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in \emph{ggseg} is based on polygons and \emph{ggplot2}-based grammar of graphics (\citet{ggplot}), while the 3D functionality in \emph{ggseg3d} is based on tri-surface mesh plots and \emph{plotly} (\citet{plotly}).
 
-Both packages present compiled datasets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
+Both packages present compiled data sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
 The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
-The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 15 additional atlases) and it is updated frequently.
-A summary of all available atlases compatible with the \emph{ggseg}-packages as of December 2019 can be found in \ref{tab:atlasTab}.
+The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 15 additional atlases) and it is frequently updated.
+A summary of all available atlases compatible with the \emph{ggseg}-packages as of December 2019 can be found in Table \ref{tab:atlasTab}.
 
@@
 We recommend users to get familiarized with \emph{ggplot2} (\citet{ggplot}).
-The package is currently only available through github, but we expect to submit to The Comprehensive R Archive Network (\citet{cran}) some time during 2020.
+The package is currently only available through github, but we expect to submit the \emph{ggseg}-package to The Comprehensive R Archive Network (\citet{cran}) in 2020.
 
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/init-1.pdf}
+\includegraphics{msc_ggseg_files/figure-latex/init-1.pdf}
 \caption{\label{fig:init}By default \texttt{ggseg()} will plot the dkt atlas in grey shaded polygons.}
@@
 These options are atlas-specific.
-For cortical atlases, such as the \texttt{dkt}, one can stack the hemispheres, view only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \ref{fig:collection} for examples).
+For cortical atlases, such as the \texttt{dkt}, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \ref{fig:collection} for examples).
 For subcortical atlases, such as the \texttt{aseg} atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/collection-1.pdf}
+\includegraphics{msc_ggseg_files/figure-latex/collection-1.pdf}
 \caption{\label{fig:collection}\emph{ggseg} plots can be used with any \emph{ggplot2} feature such as standard scales and themes.
-For cortical atlases, one can supply special \emph{ggseg} options to determine hemisphere or view, and control how the plots looks like and is organised.
+For cortical atlases, one can supply special \emph{ggseg} options to determine, for example, hemisphere, view, or position.
 \textbf{A:} \texttt{dkt} atlas, stacked with dark theme ; \textbf{B:} \texttt{dkt} with medial view only; \textbf{C:} \texttt{dkt} atlas with only left medial display; \textbf{D:} \texttt{dkt} atlas, stacked, with classic theme; \textbf{E:} \texttt{dkt} atlas with left hemisphere only; \textbf{F:} complete \texttt{aseg} atlas}
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/fill-1.pdf}
+\includegraphics{msc_ggseg_files/figure-latex/fill-1.pdf}
 \caption{\label{fig:fill}Supplying `area' to the fill option in \texttt{ggseg()}, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
@@
 Most users will use \texttt{ggseg()} to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas datasets.
-The atlas dataset structure will help users understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique dataset.
-All datasets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data sets.
+The atlas data set structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data set.
+All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Look at the top 5 rows of the dkt dataset}
+\KeywordTok{head}\NormalTok{(dkt, }\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
 
@@
 For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
-Yet, the data in \emph{ggseg} is in a long format - that is each region has its own row - and any data of interest needs to be in this same format.
-Often datasets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged in order to work with ggseg.
+Yet, the data in \emph{ggseg} is in a long format - that is each region has a single row - and any data of interest needs to be in this same format.
+Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
 See below an example of wide-to-long conversion.
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/datasupp-1.pdf}
-\caption{\label{fig:datasupp}Supplying data through the `.data' option in \texttt{ggseg()} will make it possible to use columns in the supplied data to aesthetical arguments (such as `fill').
+\includegraphics{msc_ggseg_files/figure-latex/datasupp-1.pdf}
+\caption{\label{fig:datasupp}Supplying data through the `.data' option in \texttt{ggseg()} enables use of columns in the supplied data to aesthetical arguments (such as `fill').
 The \emph{ggseg} plot can be used with any other polygon compatible function from \emph{ggplot2} or \emph{ggplot2} extentions, for instance adding title, changing the legend name and the colour scheme with standard \emph{ggplot2} functions.}
@@
 
-If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi' with either `right' or `left' for this to happen.
+If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi' with either `right' or `left'.
 In this case, data will be merged into the atlas both by `area' and by `hemi'.
+For more information about adapting data and viewing only one hemisphere or side, the \href{https://lcbc-uio.github.io/ggseg/articles/ggseg.html\#single-hemisphere-results}{package vignettes} contains more elaborate information.
 
@@
 \textbf{3)} Apply \texttt{facet\_wrap} or \texttt{facet\_grid} to the plot having used the above two rules.
-An example of this can be seen in Figure \ref{fig:datasupp3}, where a mock dataset including summary statistics for two groups ('Young' and `Old') is used when facetting a \emph{ggseg}-plot.
+An example of this can be seen in Figure \ref{fig:datasupp3}, where a mock data set including summary statistics for two groups ('Young' and `Old') is used when faceting a \emph{ggseg}-plot.
 
-All the concepts described above also work with the \texttt{aseg} atlas for subcortical structures, with the exception of `hemisphere' and `view' arguments that are superfluous in subcortical atlases (Figure \ref{fig:atlases}).
+All the concepts described above also work with the \texttt{aseg} atlas for subcortical structures, except for `hemisphere' and `view' arguments that are superfluous in subcortical atlases (Figure \ref{fig:atlases}).
 
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/datasupp3-1.pdf}
-\caption{\label{fig:datasupp3}Data including different groups need to be grouped using \texttt{group\_by()} \textbf{before} plottinh with \texttt{ggseg()} for facetting to work.}
+\includegraphics{msc_ggseg_files/figure-latex/datasupp3-1.pdf}
+\caption{\label{fig:datasupp3}To visualize data corresponding to different groups, one needs to use \texttt{group\_by} or similar \textbf{before} plotting with \texttt{ggseg()} for facetting to work.}
 \end{figure}
@@
 \centering
-\includegraphics{draft_2_files/figure-latex/atlases-1.pdf}
-\caption{\label{fig:atlases}The \texttt{aseg} atlas, showing subcortical structures, has some distinct differences from the \texttt{dkt}.
-For instance, there is no option to show only a single hemisphere, as the resulting plot would be very difficult to understand.
+\includegraphics{msc_ggseg_files/figure-latex/atlases-1.pdf}
+\caption{\label{fig:atlases}The \texttt{aseg} atlas showing subcortical structures has some distinct differences from the \texttt{dkt}.
+For instance, there is no option to show only a single hemisphere.
 Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.}
@@
 Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
-This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical strucctures, such as grey matter structures or white matter tracts, can be flattened to 2-dimensions.
+This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical structures, such as grey matter structures, can be flattened to 2-dimensions.
 Hence, here we also provide the \emph{ggseg3d} package to plot, view, and print 3D-atlases in R.
@@
 Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dkt\_3d} atlas in `LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \ref{fig:ggseg3d-1-out})
-The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage} template subject.
-All \texttt{{[}...{]}\_3d} atlases include a \texttt{colour} column for default colour plotting of the regions, based on the color scheme used in the neuroimaging software.
+The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage5} template subject.
+All \texttt{{[}...{]}\_3d} atlases include a \texttt{colour} column that based on the color scheme used in the source neuroimaging software.
 
 \begin{figure}[H]
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in *ggseg3d* atlases.
-**From left:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a fully inflated grey matter cortical surface.}\label{fig:ggseg3d-1-out}
+\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in \emph{ggseg3d} atlases.
+\textbf{From left to right:} the `white' surface is the white matter surface, `LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the `inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.}\label{fig:ggseg3d-1-out}
 \end{figure}
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has datasets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
-Only one surface is available for subcortical atlases as inflation procedures do not yield better representations.
+Each cortical atlas has data sets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
+Only one surface is available for subcortical atlases as inflation procedures are irrelevant.
 The `ggseg\_3d' column includes all necessary information for \texttt{ggseg3d()} to create a 3D mesh-plot, and should not be modified by the user.
@@
 Similarly as in the 2D-atlas, the user will use \texttt{ggseg3d()} to display through a colour scale some descriptive or inferential statistic.
-If the data is not already in the correct long format, or uses similar naming as the atlas, The users should inspect the atlas data for a specific surface (and hemisphere, if desired) and then \texttt{unnest(ggseg\_3d)} it to see how the atlas data is organised.
+If the data is not already in the correct long format or uses similar naming as the atlas, the users should inspect the atlas data for a specific surface (and hemisphere, if desired), and then \texttt{unnest(ggseg\_3d)} it to see how the atlas data is organised.
 
@@
 It is thus important to match your regional identifiers with those used in the atlas.
-To colour the segments using a column from the data, that column name needs to be supplied to the \texttt{colour} option, and providing it to the \texttt{text} option will add another line to the \emph{plotly} hover information.
+To colour the segments using a column from the data, a column name from the data needs to be supplied to the \texttt{colour} option, and providing it to the \texttt{text} option will add another line to the \emph{plotly} hover information.
 
@@
 \begin{figure}[H]
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_data} \includegraphics[width=0.3\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.3\linewidth]{png/ggseg3d_nacolour} \caption{**Left:** Supplying data to `ggseg3d()` works in much the same way as for `ggseg()`.
-Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function.
-Here we set the parcellation colours with the 'colour' column and accompanying mouse-hover text with the 'text' column.
-**Middle:** Specifying the palette for `ggseg3d()` must be done directly in the main call to the plot.
-The 'palette' option takes vectors of colours either as HEX-codes or R-colour names, but not a combination of these.
-**Right:** The colour of the `NA` values can also be changed through the option 'na.colour', which is here set to be 'black'}\label{fig:ggseg3d-data1}
+\includegraphics[width=0.3\linewidth]{png/ggseg3d_data} \includegraphics[width=0.3\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.3\linewidth]{png/ggseg3d_nacolour} \caption{\textbf{Left:} Supplying data to \texttt{ggseg3d()} works similarly to \texttt{ggseg()}.
+Since \texttt{ggseg3d()} is based on \emph{plotly} rather than \emph{ggplot2}, most aesthetic adaptations must be set in the main function.
+Here we set the parcellation colours with the `colour' column and accompanying mouse-hover text with the `text' column.
+\textbf{Middle:} The palette for \emph{ggseg3d} needs to be specified directly in the main plot call.
+The `palette' option takes vectors of colours either as HEX-codes or R-colour names.
+\textbf{Right:} The colour of the \texttt{NA} values can also be changed through the option `na.colour'.}\label{fig:ggseg3d-data1}
 \end{figure}
 
-\hypertarget{colours}{%
-\subsubsection{Colours}\label{colours}}
+\hypertarget{customizing-colours-and-the-colour-bar}{%
+\subsubsection{Customizing colours and the colour bar}\label{customizing-colours-and-the-colour-bar}}
 
@@
 Colours will be evenly spaced when creating the colour-scale.
-A palette may also be supplied as a \emph{named numeric vector}, where the vector names are the colours that users wish to use, and the numeric values are the breakpoint for each colour (e.g.~\texttt{c("red"\ =\ 0,\ "white"\ =\ 0.5,\ "blue"\ =\ 1)}).
+A palette may also be supplied as a \emph{named numeric vector}, where the vector names are the colours that users wish to use, and the numeric values are the breakpoints for each colour (e.g.~\texttt{c("red"\ =\ 0,\ "white"\ =\ 0.5,\ "blue"\ =\ 1)}).
 This way the users can control the minimum and maximum values of the colour scale, and also how the gradient is applied.
@@
 
-\hypertarget{glass-brain}{%
-\subsubsection{Glass brain}\label{glass-brain}}
+\hypertarget{adding-a-glass-brain}{%
+\subsubsection{Adding a glass brain}\label{adding-a-glass-brain}}
 
@@
 One can control the opacity of the these \texttt{NA} structures, to improve visualization.
-Glassbrains can be added provides a frame of reference for the subcortical structures with the function \texttt{add\_glassbrain()} (Figure \ref{fig:glassbrain1}) , which takes three extra arguments: hemisphere, colour, and opacity.
+Glass brains can be added to provide a frame of reference for the subcortical structures with the function \texttt{add\_glassbrain()} (Figure \ref{fig:glassbrain1}) , which takes three extra arguments: hemisphere, colour, and opacity.
 
@@
 \begin{figure}[H]
-\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{For subcortical structure visualisation, one can add a glassbrain to the plot.
-This will help with understanding structure locations relative to the cortex and make the plot easier to interpret.
-The glassbrain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:glassbrain1}
+\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{For subcortical structure visualization, one can add a glass brain to the plot.
+This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
+The glass brain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:glassbrain1}
 \end{figure}
@@
 
-The \emph{ggseg} and \emph{ggseg3d} packages have two atlases each, which are 2D and 3D variations of the same main atlases: the \texttt{dkt} (\citet{dkt}) and \texttt{aseg} (\citet{aseg}) atlases, default atlases used by the commonly used imaging software Freesurfer (@\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}).
+The \emph{ggseg} and \emph{ggseg3d} packages have two atlases each, which are 2D and 3D variations of the same main atlases: the \texttt{dkt} (\citet{dkt}) and \texttt{aseg} (\citet{aseg}) atlases.
 These are, however, only two among many meaningful ways of segmenting the brain into different regions.
-Thus, the \emph{ggsegExtra} package is library with additional datasets for plotting with the \emph{ggseg} and \emph{ggseg3d} packages.
-We hope to expand this library through a combination of our own and community effort, as the need arises for new atlases, as brain new brain atlases are continuously being released.
+Thus, the \emph{ggsegExtra} package is a repository containing additional with additional data sets for plotting with the \emph{ggseg} and \emph{ggseg3d} packages.
+There is an ever-increasing amount of new atlases being created, as research and methods in neuroimaging analysis progresses.
+The \emph{ggsegExtra}-package is intended to be expanded as a community-effort, as new and informative atlases are published.
 A small collection of the atlases currently in the \emph{ggsegExtra} package may be viewed in Figure \ref{fig:ggsegExtra}
@@
 \begin{figure}[H]
-\includegraphics[width=0.8\linewidth]{png/ggseg_extra} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra1} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra2} \caption{Four example datasets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.}\label{fig:ggsegExtra}
+\includegraphics[width=0.8\linewidth]{png/ggseg_extra} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra1} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra2} \caption{Four example data sets from the \emph{ggsegExtra} library, plotted with the \texttt{ggseg()} and \texttt{ggseg3d()} functions.}\label{fig:ggsegExtra}
 \end{figure}
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (\citet{binder}) or Shiny (\citet{shiny}), is an added benefit.
-An example of this is in the online supplementary information of Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}).
+This is exemplified in the online supplementary information of Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}).
 
-The \emph{ggseg}, \emph{ggseg3d}, and \emph{ggsegExtra} packages contain three main features:
+The three \emph{ggseg}-packages contain three main features:
+
 \textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data include the necessary coordinates for plotting, but are made more user-friendly.
+The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.\\
 \textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
-\texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.
-\texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.\\
+- \texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.\\
+- \texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.\\
 \textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
+
 These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in \emph{ggsegExtra}.
 
-The foundations of the \emph{ggseg}-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supporting information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019})) .
+The foundations of the \emph{ggseg}-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supplementary information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}) ) .
 That is, \emph{ggseg} appears with the need to inspect and display brain information over time - i.e.~including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
@@
 
-In the package \href{https://github.com/LCBC-UiO/ggseg/wiki}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
+In the \emph{ggsegExtra} \href{https://github.com/LCBC-UiO/ggsegExtra/wiki/Contributing\%3A-polygon-atlases-new}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
 At the moment, the creation of atlas for ggseg is convoluted and difficult and requires manual intervention.
@@
 We are indebted to those users that actively contributed to the package and all the individuals that worked on the framework on which the package is sustained (e.g.~R-project, neuroimaging software).
-Finally, we thank Anders M. Fjell and Kristine B. Walhovd for their encouragement and financial support.
+Finally, we thank Anders M. Fjell (AMF) and Kristine B. Walhovd (KBW) for their encouragement and financial support.
 
@@
 
-This work is funded by EU Horizon 2020 Grant `Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement number 732592.
-The project has also received funding from the European Research Council's Starting Grant scheme under grant agreements 283634, 725025 (to center PI A.M.
-Fjell), and 313440 (to center PI K.B.
-Walhovd).
+This work is funded by EU Horizon 2020 Grant `Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement 732592.
+The project has also received funding from the European Research Council's Starting (grant agreements 283634, to A.M.F.
+and 313440 to K.B.W.) and consolidator Grant Scheme (grant agreement 771355 to KBW and 725025 to AMF).
+The project has received funding through multiple grants from the Norwegian Research Council"
 

### 2020-03-17 reviewer_response.Rmd
update with revisions comments

---
title: "Response to reviewers"
subtitle: "for AMPPS manuscript: Visualisation of Brain Statistics with R-packages ggseg and ggseg3d"
author: "Athanasia Monika Mowinckel & Didac Vidal Piñeiro"
date: "2/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reviewer 1

### Major concerns
#### 1. The output of many code chunks are not present to illustrate the concepts described (Pages 5, 9).

The code for Figure 3 on page 5 is printed on page 5, _after_ the figure. 
This is necessary, as there is not really any proper room for the code to come before that plot specifically.
We hoped this was not too confusing, but understand that is it easy to miss.

on the plot for page 9, the missing code is printed in the title of each sub-plot. 
The reason we are omitting the code chunk of this image is that to incorporate static images of 3d plotly images, we need to use the tool `orca`. We mention this tool on page 11, but we do not want to go into orca in this tutorial. This is because it adds a level of complexity which we do not want to introduce, as we wish to focus on the bits of code necessary to make the images in an interactive R session. 
We have added an _unexectuted_ code chunk showing these three plots, but not how you can combine them into the three-part subplot that is actually in the paper, as this is just a little too much information in our opinion.

#### 2. Many users wish to bring in parcellations from individual subjects processing. For example, freesurfer files or PLY files from some other software. It is not clear if ggseg has the ability to read and then render these, but they are fundamental to this area. Similarly, In this vein, the overall functionality of a user creating their own atlas would be helpful.
We agree, this is an area we are working on imporiving. 
Currently, creating one's own custom atlas is cumbersome and not something anyone can do easily. 
While we hope to get this functionality improved in the (near) future, we currently have information on how to make your own `ggseg` atlas in the github wiki'es page. 
**ADD INFORMATION FOR ggseg3d creation!**

#### 3. In this regard, please read in a true freesurfer stats file from the subjects folder installed with freesurfer, and don’t use mock data. This will help those using Freesurfer, which is a likely user of these packages.  
We considerably altered the section introducing how to supply your own data to the function, to include this aspect of reading a freesurfer file directly.
Indeed, when we did this, we realised a bit more data manipulation was necessary to get the data into correct format than we could expect a novice user to handle.
Therefore, we added several new functions to the package, that should make this transition easier.
The section is now changed to also instruct the user on how to use those functions to read in Freesurfer data from two different formats, and these are subsequently plotted in Figure 4.

We do continue to use mock data in the remainder of the tutorial, because the mock data allows us to highlight certain function features more easily.

#### 4. The interactivity of these tools is essential - link out to interactive examples are necessary.
We have added links to online documentation of the tools, which have more features covered, and in case of `ggseg3d` also has interactive versions of plots.

#### 5. Page 9, Line 34 “It is important to note that the coordinates in the plot (X, Y, Z) are not any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.” - it is unclear why this is important - please discuss the signifiance (such as using this to find ROIs on MNI space, etc).
We added a sentence to this section making sure this point was made even more clearly.

> It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in `ggseg3d`, at the moment.
As such, information regarding referecing the regions' location in standard radiological spaces (like MNI or Taliarch) should be aquired directly from imaging software or papers describin the atlases, not `ggseg3d`.

### Minor concerns
#### 1. Why must you group_by and then facet_wrap in line 43 page 7? Is the grouping necessary? Line 23, page 7 seems to indicate that it is, but it’s unclear why
This is a pretty annoying neccessity because the function `ggseg` is just a wrapper for `geom_polygon` and we are in a sense cricumventing the genious of the grammar of graphics. We added a small section in the manuscript explaining this, and also added a subplot to figure 5, now showing when data is grouped before and if not grouped. `ggseg` will merge supplied data with atlas-data, and then feed that merged data to `ggplot` and `geom_polygon`. This makes some unfortunate changes to the data, so that the entire atlas will not be plotted correctly once provided to `ggplot` and facetted. We also added some text explaining this under the **Creating subplots** heading for ggseg.

> In regards to the second rule of needing to group the data before plotting, this is due to the underlying structure of how the provided data and the atlas data are combined before plotting. 
Since the package function, so far, is a wrapper for {ggplot2} rather than a  `geom`, there is certain ggplot2-functionality that is difficult to integrate, like the facets. 
For the complete atlas-data to be merged with _every data group_ correctly, the ggseg function needs to have an indication before plotting what the groups are. 
The result of ungrouped data into `ggseg` is shown in  \@ref(fig:datasupp3)A. 
Notice how there are three panels rather than two, and that the regions _not represented in the data_ have their own panel with all the grey `NA` colour, while the panels with the data only have the regions represented in the data.

#### 2. Can a continuous mapping of triangles be handled? For example, Figure 4 has a parcellation, but can a mapping to each point on the surface be handled? If so, please give example in supplement. 
unfortunately no. We are finding it difficult to figure out how the indexing of each individual triangle actually works, and how to map that in a continuous way. We hope to be able to crack that nut in the future, but at the moment we are not able to do so. If we get there, that would be a great improvement to the package as a whole, as it would greatly minimize the data needed to plot and improve speed of plotting, which now can be quite slow for atlases with many parcellations.

#### 3. Define “glass brains” for the main audience (page 9)
We have added the following centences below the glass brain subheading

> A glass brain is a transluscent representation of the brain that can provide a frame of reference, particularly when looking at subcortical parcellations.
Since individual brain areas or subcortical structures can be very hard to place without the visual reference of the entire brain, glass brains are often used to aid the viewer understand a regions location.
Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.

#### 4. A link to an interactive figure 9 should be provided.

#### 5. Why must title = " " in the code chunks Page (3)
This was something hacky we did to get title alignment and placement to work with cowplot. We switched to patchwork, where this is not needed. We have not made title and subtitle setting in `labs` that describe the plots very shortly.

#### 6. The code chunks in Page 3 should be run in the same order as the figure so that readers flipping back and forth do not have to search around, p1, med, combo, p2, left, p3 is the order they are passed, so they should be the order in which they are created. Also A/B/C/etc should be in the code comments, not simply the name, e.g. F: aseg default theme.
We have corrected the figure and code with your comments, and find it much improved.

#### 7. Why is there an NA in Figure 3?
All cortical atlases have `NA` for the polygon representing the medial wall, and currently subcortical atlases also have `NA` for the polygon(s) representing the cortex.  
This was a design decision we did early on, that areas that are _not_ part of the atlas itself should be represented stored as `NA` , rather than giving them their own fill colour.

### Typos
**• extra “s” in pluss in 3rd paragraph (line 55)**  
Thank you for spotting this, we amended the error.

## Reviewer 2
**I have done many R package tutorials, and I saw no issue with this one. I would rank it as intermediate in difficulty because of some setting up additional dependencies such as the remotes package. I think I would especially utilize this package with Freesurfer so the examples provided in congruence with that workflow was useful to me. Code itself was clean, commented, and easy to follow. While much of the work I do is outside of R completely, I could see this being a useful tool for someone who is basing their analyses in that framework. I hope that the contents of this tutorial are being similarly represented on a Git documentation page, as I am sure most users will default to that mode to learn about this package instead of a manuscript. I look forward to the planned improvements that the authors note elsewhere in the Planned Package Improvements portion of the article and thank them for the open-source distribution of this package.**

**I can provide to the authors minimally commented code I wrote to follow this tutorial, but would note that most of it overlaps with existing content in the manuscript.**

### 2020-03-17 cover.Rmd
update with revisions comments

---
title: "cover_ggseg"
author: "Athanasia Monika Mowinckel & Didac Vidan Piñeiro"
date: "12/18/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

Dear editor,

we would like for you to consider our manuscipt *Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_* for publication in your journal *Advances in Methods and Practices in Psychological Science*. This tutorial walks the reader through the basic principles in our two R-packages for data visualization of brain segmentation data directly through R. As neuroimaging data have innate spatial information, it is crucial to keep this information when disseminating results, but it can be cumbersome for researchers to change between statistical and neuroimaging software. Our packages offer two solutions for visualizing results based on brain segmentations, one using 2D polygons and the other using 3D meshes. We believe these tool conform to the essence of open science, and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs. Furthermore, these packages integrate with other R tools, like Rmarkdown and shiny, creating streamlined, transparent, and reproducible ways to create dissemination reports.

We believe this tutorial is of interest to your subscribers and fits well with the journal scope, and hope you will consider it for review and publication.

All the best,

Athanasia Monika Mowinckel,

_Staff scientist_
_Center for Lifespan Changes in Brain and Cognition_
_Dept. of Psychology_
_University of Oslo_
_Norway_

### 2020-03-17 msc_ggseg.Rmd
update with revisions comments

@@
   rticles::peerj_article: preprint
+keep_tex: true
 ---
@@
 ```{r setup, include=F}
-knitr::opts_chunk$set(eval=TRUE, echo=TRUE, collapse = TRUE, results = 'hold', cache=TRUE, fig.pos = 'H') 
-options(tinytex.verbose = TRUE, tinytex.clean = FALSE)
+knitr::opts_chunk$set(eval=TRUE, echo=TRUE, 
+                      collapse = TRUE, results = 'hold', 
+                      cache=FALSE, fig.pos = 'H') 
+options(tinytex.verbose = TRUE, tinytex.clean = FALSE,
+        tibble.width = 50)
 
-library(kableExtra)
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 library(plotly) 
 library(ggplot2)
-library(knitr)
 library(ggseg)
 library(ggseg3d)
+library(ggsegExtra)
+library(kableExtra)
+library(knitr)
 
@@
   mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
-         Item = gsub("_3d", "", Item)) %>% 
+         Item = gsub("_3d", "", Item),
+         Item = gsub("_", "", Item)) %>% 
   select(Package, Title, Type, Item) %>% 
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The _ggseg_ and _ggseg3d_ – pluss the complimentary _ggsegExtra_ – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+The _ggseg_ and _ggseg3d_ – plus the complimentary _ggsegExtra_ – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in _ggseg_ is based on polygons and _ggplot2_-based grammar of graphics (@ggplot), while the 3D functionality in _ggseg3d_ is based on tri-surface mesh plots and _plotly_ (@plotly).
@@
 library(ggseg)
-library(dplyr)
-library(tidyr)
+
+# package combining plots easily
+library(patchwork)
 
@@
 For cortical atlases, one can supply special *ggseg* options to determine, for example, hemisphere, view, or position.
-**A:** `dkt` atlas, stacked with dark theme ; **B:** `dkt` with medial view only; **C:** `dkt` atlas with only left medial display; **D:** `dkt` atlas, stacked, with classic theme; **E:** `dkt` atlas with left hemisphere only; **F:** complete `aseg` atlas'}
-# dkt dark theme
+**A:** `dkt` atlas, stacked with dark theme ; **B:** `dkt` with medial view only; **C:** `dkt` atlas with only left medial display; **D:** `dkt` atlas, stacked, with classic theme; **E:** `dkt` atlas with left hemisphere only; **F:** axial view of `aseg` atlas'}
+# A: dkt dark theme
 p1 <- ggseg(position = "stacked") +
   theme_dark() +
-  labs(title=" ")
+  labs(title="dkt", subtitle = "dark theme")
 
-# dkt classic theme
-p2 <- ggseg(position = "stacked") +
+# B: dkt medial view
+p2 <- ggseg(view = "medial") +
+  labs(title = "dkt", subtitle = "medial view")
+
+# C: dkt left medial alone
+p3 <- ggseg(view = "medial",
+            hemisphere = "left") +
+  labs(title="dkt", subtitle = "medial left side")
+
+# D: dkt classic theme
+p4 <- ggseg(position = "stacked") +
   theme_classic() +
-  labs(title = " ")
+  labs(title = "dkt", subtitle = "classic theme")
 
-# dkt medial view
-med <- ggseg(view = "medial") +
-  labs(title = " ")
+# E: dkt left hemisphere
+p5 <- ggseg(hemisphere = "left") +
+  labs(title = "dkt", subtitle = "left hemisphere")
 
-# dkt left hemisphere
-left <- ggseg(hemisphere = "left") +
-  labs(title = " ")
+# F: aseg default theme
+p6 <- ggseg(atlas = aseg, view = "axial") +
+  labs(title = "aseg", subtitle = "axial view")
 
-# aseg default theme
-p3 <- ggseg(atlas=aseg) +
-  labs(title = " ")
-
-# dkt left medial alone
-combo <- ggseg(view = "medial",
-               hemisphere = "left") +
-  labs(title=" ")
-
-# Combine plots to Figure 2
-cowplot::plot_grid(p1, med, combo,  p2, left, p3,
-                   labels = c("A: dkt - dark", "B: dkt - medial",
-                              "C: dkt - combo", "D: dkt - classic",
-                              "E: dkt - left", "F: aseg"),
-                   hjust = -.05)
+# Using patchwork to combine all plots into one
+(p1 | p2 | p3) / (p4 | p5 | p6) +
+  plot_annotation(tag_levels = 'A')
 ```
@@
 # Figure 3
-ggseg(mapping=aes(fill = area), colour="black") +
+ggseg(mapping = aes(fill = area), colour="black") +
   scale_fill_brain("dkt") +
-  theme(legend.justification=c(1,0),
-        legend.position="bottom",
+  theme(legend.justification = c(1,0),
+        legend.position = "bottom",
         legend.text = element_text(size = 5)) +
@@
 
-```{r, size='small', eval=FALSE}
+```{r, size='small', eval=FALSE, codecap = "Inspect atlas data"}
 # Look at the top 5 rows of the dkt dataset
@@
 Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
-See below an example of wide-to-long conversion.
+To ease to conversion between Freesurfer files and ggseg compatible files, we have functions to aid import to R. 
+There is an extended vignette in the [online documentation]() of the various options, here we assume that that the data has been compiled into a file using Freesurfer's `aparcstats2table` or `asegstats2table` that combine parcellation metrics for all wanted subjects into a single file.
+This file also contains extra metrics like intracranial volume and total grey matter. 
+These do not map onto ggseg, and as such, one should make sure to remove them from the subsequent data.
 
 ```{r, size='small'}
-# Create a mock freesurfer output
-freesurfer_stats <- data.frame(
-  id = c(10:12),
-  lh_superiortemporal = c(3.32, 4.1, 3.5),
-  lh_precentral = c(2.3, 2.5, 2.1),
-  lh_rostralmiddlefrontal = c(3.3, 3.2, 3.1)
-)
-freesurfer_stats
+file_path <- "data/aparc.volume.table"
 
-# Wrangle wide format data into long format
-freesurfer_long <- freesurfer_stats %>%
-  gather(label, thickness, -id)
-freesurfer_long
+someData <- read_freesurfer_table(file_path, measure = "volume") %>% 
+  # remove rows in the data that do not conain the strings "rh" or "lh"
+  dplyr::filter(grepl("rh|lh", label))
+someData
 ```
 
-Data in long format can then be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dkt`.
+A second option is to read in atlas files directly from the entire subject directory, using regular expression to define the atlas wanted. 
+If we inspect a single subjects stats directory of Freesurfer (here on default installation on macOS), we can see several parcellations to choose from.
+The default aparc has a unique string ending with `aparc.stats` which we will use to read in all subject stats for that particular atlas.
+This data has more columns, as it contains several metrics that can be plotted.
+
+```{bash}
+# in bash terminal
+ls /Applications/freesurfer/subjects/bert/stats
+```
+
+```{r, warning=FALSE}
+atlas_data <- read_atlas_files(
+  subjects_dir = "/Applications/freesurfer/subjects/",
+  atlas = "aparc.stats$"
+)
+atlas_data
+```
+
+This long format data can be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dkt`.
 The data **must** include a column that has the same column name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, with an 'area' and 'p' column, representing the results of a hypothetical analysis.
-The `ggseg()`-function will recognise the matching column 'area', and merge the supplied data into the atlas using _dplyr_-joins.
-We use the 'p' column as the column flooding the segment with colour.
+In the following example, the `ggseg()`-function will recognise the matching column 'label', and merge the supplied data into the atlas using _dplyr_-joins.
 The appearance of the plot can then be modified similarly to any other _ggplot2_ graph using functions such as scales, labs, themes, etc., as seen in Figure \@ref(fig:datasupp)
@@
 ```{r 'datasupp', fig.width=5, fig.cap="Supplying data through the '.data' option in `ggseg()` enables use of columns in the supplied data to aesthetical arguments (such as 'fill').
-The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions."}
-# Make some mock data
-someData = data.frame(
-  area = c("transverse temporal", "insula",
-           "pre central","superior parietal"),
-  p = sample(seq(0,.5,.001), 4),
-  stringsAsFactors = FALSE)
+The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions.", warning = FALSE}
 
-# Figure 4
-ggseg(.data=someData, mapping=aes(fill=p))  +
-  labs(title="A nice plot title", fill="p-value") +
-  scale_fill_gradient(low="firebrick",high="goldenrod")
+# Figure 4A
+p1 <- ggseg(.data=someData, mapping=aes(fill=volume))  +
+  labs(title="from read_freesurfer_table", 
+       fill="volume\n(mm^3)") +
+  scale_fill_gradient(low="firebrick",high="goldenrod") +
+  guides(fill = guide_colourbar(barheight = 3))
+
+# Figure 4B
+p2 <- ggseg(.data=atlas_data, mapping=aes(fill=ThickAvg))  +
+  labs(title="from read_atlas_files", 
+       fill="Thickness\n(mean)") +
+  scale_fill_gradient(low="forestgreen",high="goldenrod")+
+  guides(fill = guide_colourbar(barheight = 3))
+
+(p1 / p2) +
+  plot_annotation(tag_levels = "A")
 ```
@@
 **3)** Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules.
-An example of this can be seen in Figure \@ref(fig:datasupp3), where a mock data set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.
+An example of this can be seen in Figure \@ref(fig:datasupp3) B, where a mock data set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.
+
+In regards to the second rule of needing to group the data before plotting, this is due to the underlying structure of how the provided data and the atlas data are combined before plotting.
+Since the package function, so far, is a wrapper for {ggplot2} rather than a  `geom`, there is certain ggplot2-functionality that is difficult to integrate, like the facets.
+For the complete atlas-data to be merged with _every data group_ correctly, the ggseg function needs to have an indication before plotting what the groups are.
+The result of ungrouped data into `ggseg` is shown in  \@ref(fig:datasupp3) A.
+Notice how there are three panels rather than two, and that the regions _not represented in the data_ have their own panel with all the grey `NA` colour, while the panels with the data only have the regions represented in the data.
 
@@
 
-```{r 'datasupp3', fig.width=5, fig.cap='To visualize data corresponding to different groups, one needs to use `group_by` or similar **before** plotting with `ggseg()` for facetting to work.'}
+(ref:datasupp3-cap) Examples of incorrect and correct faceting with `ggseg`.
+**A**: ungrouped data will facet incorrectly due to the data not being merged correctly with the atlas.
+**B**: grouped data will merge in the expected way and result in correctly facetted plots.
+
+```{r 'datasupp3', fig.width=5,  warning= FALSE, fig.cap='(ref:datasupp3-cap)', out.width="100%"}
 # Make some mock data
-someData = data.frame(
+someData = tibble(
   area = rep(c("transverse temporal", "insula",
@@
   p = sample(seq(0,.5,.001), 8),
-  AgeG = c(rep("Young",4), rep("Old",4)),
-  stringsAsFactors = FALSE) %>%
-  group_by(AgeG)
+  AgeG = c(rep("Young",4), rep("Old",4))
+)
 
-# Figure 5
-ggseg(.data=someData, colour="white", position = "stacked",
-      mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=2) +
-  theme(legend.position = "bottom")
+# Figure 5a
+grouped_plot <- someData %>%
+  ggseg(position = "stacked",
+      mapping=aes(fill=p), show.legend = FALSE) +
+  facet_wrap(~AgeG, ncol=3) +
+  labs(title = "Ungrouped data")
+
+# Figure 5b
+ungrouped_plot <- someData %>%
+  group_by(AgeG) %>%
+  ggseg(position = "stacked",
+      mapping=aes(fill=p), show.legend = FALSE) +
+  facet_wrap(~AgeG, ncol=3) +
+  labs(title = "Grouped data")
+
+# Combining plots with patchwork, applying theme to all
+(grouped_plot / ungrouped_plot) +
+  plot_layout(widths = 1) + 
+  plot_annotation(tag_levels = 'A') &
+  theme(axis.text = element_text(size=8))
 ```
+
+The `ggseg` package also has extensive [online documentation](https://lcbc-uio.github.io/ggseg/), with more functionality than we have covered here.
+The online documentation is also regularly updated with explanations of new features or code improvements.
 
@@
 For instance, there is no option to show only a single hemisphere.
-Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.'}
+Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.', collapse = TRUE}
 # Figure 6
@@
 **From left to right:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.
+The title for each plot is the literal code use to make the subplot.
+
+```{r "ggseg3d-1", message=FALSE, echo=TRUE, eval=FALSE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:gseg3d-1-out-cap)'}
+# Figure 7 left
+ggseg3d(surface = "white")  %>%
+  pan_camera("right lateral") %>%
+  remove_axes() %>%
+  layout(title="\nggseg3d(surface='white')",
+         font =  list(size = 30))
+
+# Figure 7 center
+ggseg3d()  %>%
+  pan_camera("right lateral") %>%
+  remove_axes() %>%
+  layout(title="\nggseg3d(surface='LCBC')",
+         font =  list(size = 30))
+
+# Figure 7 right
+ggseg3d(surface = "inflated") %>%
+  pan_camera("right lateral") %>%
+  remove_axes() %>%
+  layout(title="\nggseg3d(surface='inflated')",
+         font =  list(size = 30))
+```
 
@@
 It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
+While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in `ggseg3d`, at the moment.
+As such, information regarding referencing the regions' location in standard radiological spaces (like MNI or Taliarch) should be acquired directly from imaging software or papers describing the atlases, not `ggseg3d`.
 
@@
 ### Adding a glass brain
+A glass brain is a translucent representation of the brain that can provide a frame of reference, particularly when looking at subcortical parcellations.
+Since individual brain areas or subcortical structures can be very hard to place without the visual reference of the entire brain, glass brains are often used to aid the viewer understand a regions location.
 Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these `NA` structures, to improve visualization.
-Glass brains can be added to provide a frame of reference for the subcortical structures with the function `add_glassbrain()` (Figure \@ref(fig:glassbrain1)) , which takes three extra arguments: hemisphere, colour, and opacity.
+One can control the opacity of the these `NA` structures, to improve visualization, directly in the `ggseg3d` call.
+A cortical glass brain can be added to `ggseg3d` plots through the function `add_glassbrain()` (Figure \@ref(fig:glass brain1)), which takes three extra arguments: hemisphere, colour, and opacity.
+We recommend only using the glass brain function when plotting subcortical atlases.
+An interactive version of this plot can be found at [https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html#colours](https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html#colours)
 
@@
 
-(ref:glassbrain1-cap) For subcortical structure visualization, one can add a glass brain to the plot.
+(ref:glass brain1-cap) For subcortical structure visualization, one can add a glass brain to the plot.
 This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
@@
 
+The package [online documentation](https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html) also includes interactive versions of many of the figures in this tutorial, and several more options not covered here.
+The online documentation will also be updated as new features are required or we improve the current state of user functionality and options.
+
 ## Additional atlases
@@
 The _ggsegExtra_-package is intended to be expanded as a community-effort, as new and informative atlases are published.
-A small collection of the atlases currently in the _ggsegExtra_ package may be viewed in Figure \@ref(fig:ggsegExtra)
+A small collection of the atlases currently in the _ggsegExtra_ package may be viewed in Figure \@ref(fig:ggsegExtra), and explored in our [shinyapps.io demonstration](https://athanasiamo.shinyapps.io/ggsegDemo/).
 
@@
             show.legend = FALSE, position="stacked") +
-  scale_fill_brain("yeo7") +
+  scale_fill_brain("yeo7", package = "ggsegExtra") +
   labs(title="yeo7-data set")
@@
             show.legend = FALSE, position="stacked") +
-  scale_fill_brain("glasser") +
+  scale_fill_brain("glasser", package = "ggsegExtra") +
   labs(title="glasser-data set")
@@
 The aim is to create a set of functions that will call specialized tools like FSL(@fsl), Freesurfer(@fischl_99, @dale_99, @Fischl2000) and Imagemagick(@magick) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid _ggseg_ and _ggseg3d_ atlases.
+Additionally, while `ggseg` works quite nicely as a `ggplot2` wrapper, ideally it should be a proper `geom` or `stat`, making it a `ggplot2` extension rather than a wrapper. 
+This would benefit the user as it would behave more like other `ggplot2` `geoms` and there would be less new functionality to learn.
+Furthermore, a `geom` would make the process of having to group data _before_ faceting become unnecessary, and as such also ease use.   
+We also hope to make it possible to colour each individual mesh triangle in `ggseg3d`, and as such provide the users with more flexibility in the plotting. 
+This should also improve the speed of `ggseg3d`, which currently is quite slow due to each parcellation region being loaded into `plotly` sequentially. 
+If the entire brain mesh is loaded once, and then the mesh triangles coloured, this would greatly improve speed. 
+Additionally, the current procedure makes it necessary to store the 3d-rendered mesh information directly in the package for all atlases, which also takes up a lot of disk space and makes it slow to download.
+When able to colour the mesh triangles directly, rather than each parcellation individually, only text data indexing mesh-locations would be necessary and thus reduce package size considerably. 
 We encourage users to contribute to the _ggsegExtra_ brain atlas repository by including additional brain atlases.

### 2020-03-17 msc_ggseg.tex
update with revisions comments

@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The \emph{ggseg} and \emph{ggseg3d} -- pluss the complimentary \emph{ggsegExtra} -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+The \emph{ggseg} and \emph{ggseg3d} -- plus the complimentary \emph{ggsegExtra} -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in \emph{ggseg} is based on polygons and \emph{ggplot2}-based grammar of graphics (\citet{ggplot}), while the 3D functionality in \emph{ggseg3d} is based on tri-surface mesh plots and \emph{plotly} (\citet{plotly}).
@@
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
-The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 15 additional atlases) and it is frequently updated.
-A summary of all available atlases compatible with the \emph{ggseg}-packages as of December 2019 can be found in Table \ref{tab:atlasTab}.
+The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 16 additional atlases) and it is frequently updated.
+A summary of all available atlases compatible with the \emph{ggseg}-packages as of March 2020 can be found in Table \ref{tab:atlasTab}.
 
@@
 \toprule
-Title & Item & Mesh & Polygon & citation\\
+Title & Item & Mesh & Polygon & Citation\\
 \midrule
@@
 Desterieux cortical parcellations & desterieux & ggsegExtra & None & \citet{desterieux}\\
+Freesurfer ASEG with posterior-anterior hippocampus & hcpa & ggsegExtra & None & \\
 Freesurfer automatic subcortical segmentation of a brain volume & aseg & ggseg3d & ggseg & \citet{aseg}\\
+\addlinespace
 Genetic topography of brain area morphology & chenAr & None & ggsegExtra & \citet{chen}\\
-\addlinespace
 Genetic topography of brain thickness morphology & chenTh & None & ggsegExtra & \citet{chen}\\
@@
 ICBM white matter parcellation & icbm & ggsegExtra & None & \citet{icbm}\\
+\addlinespace
 JHU parcellation & jhu & None & ggsegExtra & \citet{jhu}\\
-\addlinespace
 Local-Global 17 Parcellation of the Human Cerebral Cortex & schaefer17 & ggsegExtra & None & \citet{schaefer}\\
@@
 Parcellation from the Human Connectome Project & glasser & ggsegExtra & None & \citet{glasser}\\
+\addlinespace
 White matter tract parcellations & tracula & ggsegExtra & None & \citet{tracula}\\
-\addlinespace
 Yeo 17 Resting-state Cortical Parcellations & yeo17 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
@@
 \KeywordTok{library}\NormalTok{(ggseg)}
-\KeywordTok{library}\NormalTok{(dplyr)}
-\KeywordTok{library}\NormalTok{(tidyr)}
+
+\CommentTok{\# package combining plots easily}
+\KeywordTok{library}\NormalTok{(patchwork)}
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# dkt dark theme}
+\CommentTok{\# A: dkt dark theme}
 \NormalTok{p1 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme\_dark}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"dark theme"}\NormalTok{)}
 
-\CommentTok{\# dkt classic theme}
-\NormalTok{p2 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\CommentTok{\# B: dkt medial view}
+\NormalTok{p2 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"medial view"}\NormalTok{)}
+
+\CommentTok{\# C: dkt left medial alone}
+\NormalTok{p3 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
+            \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"medial left side"}\NormalTok{)}
+
+\CommentTok{\# D: dkt classic theme}
+\NormalTok{p4 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme\_classic}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"classic theme"}\NormalTok{)}
 
-\CommentTok{\# dkt medial view}
-\NormalTok{med <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{\# E: dkt left hemisphere}
+\NormalTok{p5 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dkt"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"left hemisphere"}\NormalTok{)}
 
-\CommentTok{\# dkt left hemisphere}
-\NormalTok{left <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{\# F: aseg default theme}
+\NormalTok{p6 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg, }\DataTypeTok{view =} \StringTok{"axial"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"aseg"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"axial view"}\NormalTok{)}
 
-\CommentTok{\# aseg default theme}
-\NormalTok{p3 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
-
-\CommentTok{\# dkt left medial alone}
-\NormalTok{combo <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
-               \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
-
-\CommentTok{\# Combine plots to Figure 2}
-\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot\_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
-                   \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A: dkt {-} dark"}\NormalTok{, }\StringTok{"B: dkt {-} medial"}\NormalTok{,}
-                              \StringTok{"C: dkt {-} combo"}\NormalTok{, }\StringTok{"D: dkt {-} classic"}\NormalTok{,}
-                              \StringTok{"E: dkt {-} left"}\NormalTok{, }\StringTok{"F: aseg"}\NormalTok{),}
-                   \DataTypeTok{hjust =} \FloatTok{{-}.05}\NormalTok{)}
+\CommentTok{\# Using patchwork to combine all plots into one}
+\NormalTok{(p1 }\OperatorTok{|}\StringTok{ }\NormalTok{p2 }\OperatorTok{|}\StringTok{ }\NormalTok{p3) }\OperatorTok{/}\StringTok{ }\NormalTok{(p4 }\OperatorTok{|}\StringTok{ }\NormalTok{p5 }\OperatorTok{|}\StringTok{ }\NormalTok{p6) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot\_annotation}\NormalTok{(}\DataTypeTok{tag\_levels =} \StringTok{\textquotesingle{}A\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 For cortical atlases, one can supply special \emph{ggseg} options to determine, for example, hemisphere, view, or position.
-\textbf{A:} \texttt{dkt} atlas, stacked with dark theme ; \textbf{B:} \texttt{dkt} with medial view only; \textbf{C:} \texttt{dkt} atlas with only left medial display; \textbf{D:} \texttt{dkt} atlas, stacked, with classic theme; \textbf{E:} \texttt{dkt} atlas with left hemisphere only; \textbf{F:} complete \texttt{aseg} atlas}
+\textbf{A:} \texttt{dkt} atlas, stacked with dark theme ; \textbf{B:} \texttt{dkt} with medial view only; \textbf{C:} \texttt{dkt} atlas with only left medial display; \textbf{D:} \texttt{dkt} atlas, stacked, with classic theme; \textbf{E:} \texttt{dkt} atlas with left hemisphere only; \textbf{F:} axial view of \texttt{aseg} atlas}
 \end{figure}
@@
 \CommentTok{\# Figure 3}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping =} \KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{scale\_fill\_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
-        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
+        \DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{,}
         \DataTypeTok{legend.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
@@
 Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
-See below an example of wide-to-long conversion.
+To ease to conversion between Freesurfer files and ggseg compatible files, we have functions to aid import to R.
+There is an extended vignette in the \href{}{online documentation} of the various options, here we assume that that the data has been compiled into a file using Freesurfer's \texttt{aparcstats2table} or \texttt{asegstats2table} that combine parcellation metrics for all wanted subjects into a single file.
+This file also contains extra metrics like intracranial volume and total grey matter.
+These do not map onto ggseg, and as such, one should make sure to remove them from the subsequent data.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Create a mock freesurfer output}
-\NormalTok{freesurfer\_stats <{-}}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{id =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\OperatorTok{:}\DecValTok{12}\NormalTok{),}
-  \DataTypeTok{lh\_superiortemporal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.32}\NormalTok{, }\FloatTok{4.1}\NormalTok{, }\FloatTok{3.5}\NormalTok{),}
-  \DataTypeTok{lh\_precentral =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.3}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.1}\NormalTok{),}
-  \DataTypeTok{lh\_rostralmiddlefrontal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.3}\NormalTok{, }\FloatTok{3.2}\NormalTok{, }\FloatTok{3.1}\NormalTok{)}
-\NormalTok{)}
-\NormalTok{freesurfer\_stats}
+\NormalTok{file\_path <{-}}\StringTok{ "data/aparc.volume.table"}
 
-\CommentTok{\# Wrangle wide format data into long format}
-\NormalTok{freesurfer\_long <{-}}\StringTok{ }\NormalTok{freesurfer\_stats }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{{-}}\NormalTok{id)}
-\NormalTok{freesurfer\_long}
-\CommentTok{\#\#   id lh\_superiortemporal lh\_precentral lh\_rostralmiddlefrontal}
-\CommentTok{\#\# 1 10                3.32           2.3                     3.3}
-\CommentTok{\#\# 2 11                4.10           2.5                     3.2}
-\CommentTok{\#\# 3 12                3.50           2.1                     3.1}
-\CommentTok{\#\#   id                   label thickness}
-\CommentTok{\#\# 1 10     lh\_superiortemporal      3.32}
-\CommentTok{\#\# 2 11     lh\_superiortemporal      4.10}
-\CommentTok{\#\# 3 12     lh\_superiortemporal      3.50}
-\CommentTok{\#\# 4 10           lh\_precentral      2.30}
-\CommentTok{\#\# 5 11           lh\_precentral      2.50}
-\CommentTok{\#\# 6 12           lh\_precentral      2.10}
-\CommentTok{\#\# 7 10 lh\_rostralmiddlefrontal      3.30}
-\CommentTok{\#\# 8 11 lh\_rostralmiddlefrontal      3.20}
-\CommentTok{\#\# 9 12 lh\_rostralmiddlefrontal      3.10}
+\NormalTok{someData <{-}}\StringTok{ }\KeywordTok{read\_freesurfer\_table}\NormalTok{(file\_path, }\DataTypeTok{measure =} \StringTok{"volume"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+\StringTok{  }\CommentTok{\# remove rows in the data that do not conain the strings "rh" or "lh"}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(}\KeywordTok{grepl}\NormalTok{(}\StringTok{"rh|lh"}\NormalTok{, label))}
+\NormalTok{someData}
+\CommentTok{\#\# \# A tibble: 34 x 3}
+\CommentTok{\#\#    subject label                      volume}
+\CommentTok{\#\#    <fct>   <chr>                       <dbl>}
+\CommentTok{\#\#  1 bert    rh\_bankssts                  1969}
+\CommentTok{\#\#  2 bert    rh\_caudalanteriorcingulate   2280}
+\CommentTok{\#\#  3 bert    rh\_caudalmiddlefrontal       5390}
+\CommentTok{\#\#  4 bert    rh\_cuneus                    2998}
+\CommentTok{\#\#  5 bert    rh\_entorhinal                1735}
+\CommentTok{\#\#  6 bert    rh\_fusiform                  8144}
+\CommentTok{\#\#  7 bert    rh\_inferiorparietal         14876}
+\CommentTok{\#\#  8 bert    rh\_inferiortemporal         11016}
+\CommentTok{\#\#  9 bert    rh\_isthmuscingulate          1983}
+\CommentTok{\#\# 10 bert    rh\_lateraloccipital         12729}
+\CommentTok{\#\# \# ... with 24 more rows}
 \end{Highlighting}
@@
 
-Data in long format can then be used directly with the \texttt{ggseg()}-function, as the `label' column corresponds in name and content with the `label' column in the atlas data of \texttt{dkt}.
+A second option is to read in atlas files directly from the entire subject directory, using regular expression to define the atlas wanted.
+If we inspect a single subjects stats directory of Freesurfer (here on default installation on macOS), we can see several parcellations to choose from.
+The default aparc has a unique string ending with \texttt{aparc.stats} which we will use to read in all subject stats for that particular atlas.
+This data has more columns, as it contains several metrics that can be plotted.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# in bash terminal}
+\FunctionTok{ls}\NormalTok{ /Applications/freesurfer/subjects/bert/stats}
+\CommentTok{\#\# aseg.stats}
+\CommentTok{\#\# lh.BA\_exvivo.stats}
+\CommentTok{\#\# lh.BA\_exvivo.thresh.stats}
+\CommentTok{\#\# lh.aparc.DKTatlas.stats}
+\CommentTok{\#\# lh.aparc.a2009s.stats}
+\CommentTok{\#\# lh.aparc.pial.stats}
+\CommentTok{\#\# lh.aparc.stats}
+\CommentTok{\#\# lh.curv.stats}
+\CommentTok{\#\# lh.w{-}g.pct.stats}
+\CommentTok{\#\# rh.BA\_exvivo.stats}
+\CommentTok{\#\# rh.BA\_exvivo.thresh.stats}
+\CommentTok{\#\# rh.aparc.DKTatlas.stats}
+\CommentTok{\#\# rh.aparc.a2009s.stats}
+\CommentTok{\#\# rh.aparc.pial.stats}
+\CommentTok{\#\# rh.aparc.stats}
+\CommentTok{\#\# rh.curv.stats}
+\CommentTok{\#\# rh.w{-}g.pct.stats}
+\CommentTok{\#\# wmparc.stats}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{atlas\_data <{-}}\StringTok{ }\KeywordTok{read\_atlas\_files}\NormalTok{(}
+  \DataTypeTok{subjects\_dir =} \StringTok{"/Applications/freesurfer/subjects/"}\NormalTok{,}
+  \DataTypeTok{atlas =} \StringTok{"aparc.stats$"}
+\NormalTok{)}
+\NormalTok{atlas\_data}
+\CommentTok{\#\# \# A tibble: 204 x 11}
+\CommentTok{\#\#    subject label NumVert SurfArea GrayVol ThickAvg}
+\CommentTok{\#\#    <chr>   <chr>   <int>    <int>   <int>    <dbl>}
+\CommentTok{\#\#  1 bert    lh\_b\textasciitilde{}    1181      831    2297     2.77}
+\CommentTok{\#\#  2 bert    lh\_c\textasciitilde{}     843      572    1534     2.72}
+\CommentTok{\#\#  3 bert    lh\_c\textasciitilde{}    2758     1840    5772     2.80}
+\CommentTok{\#\#  4 bert    lh\_c\textasciitilde{}    2683     1654    3074     1.81}
+\CommentTok{\#\#  5 bert    lh\_e\textasciitilde{}     581      416    1840     3.33}
+\CommentTok{\#\#  6 bert    lh\_f\textasciitilde{}    4113     2875    8519     2.72}
+\CommentTok{\#\#  7 bert    lh\_i\textasciitilde{}    4948     3466   10559     2.70}
+\CommentTok{\#\#  8 bert    lh\_i\textasciitilde{}    5056     3542   12358     2.98}
+\CommentTok{\#\#  9 bert    lh\_i\textasciitilde{}    1561      990    2350     2.09}
+\CommentTok{\#\# 10 bert    lh\_l\textasciitilde{}    7961     5077   12743     2.30}
+\CommentTok{\#\# \# ... with 194 more rows, and 5 more variables:}
+\CommentTok{\#\# \#   ThickStd <dbl>, MeanCurv <dbl>,}
+\CommentTok{\#\# \#   GausCurv <dbl>, FoldInd <int>, CurvInd <dbl>}
+\end{Highlighting}
+\end{Shaded}
+
+This long format data can be used directly with the \texttt{ggseg()}-function, as the `label' column corresponds in name and content with the `label' column in the atlas data of \texttt{dkt}.
 The data \textbf{must} include a column that has the same column name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, with an `area' and `p' column, representing the results of a hypothetical analysis.
-The \texttt{ggseg()}-function will recognise the matching column `area', and merge the supplied data into the atlas using \emph{dplyr}-joins.
-We use the `p' column as the column flooding the segment with colour.
+In the following example, the \texttt{ggseg()}-function will recognise the matching column `label', and merge the supplied data into the atlas using \emph{dplyr}-joins.
 The appearance of the plot can then be modified similarly to any other \emph{ggplot2} graph using functions such as scales, labs, themes, etc., as seen in Figure \ref{fig:datasupp}
@@
 \begin{Highlighting}[]
-\CommentTok{\# Make some mock data}
-\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{area =} \KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
-           \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}
-  \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{4}\NormalTok{),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
 
-\CommentTok{\# Figure 4}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))  }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p{-}value"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale\_fill\_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}
+\CommentTok{\# Figure 4A}
+\NormalTok{p1 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{volume))  }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"from read\_freesurfer\_table"}\NormalTok{, }
+       \DataTypeTok{fill=}\StringTok{"volume}\CharTok{\textbackslash{}n}\StringTok{(mm\^{}3)"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_fill\_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_colourbar}\NormalTok{(}\DataTypeTok{barheight =} \DecValTok{3}\NormalTok{))}
+
+\CommentTok{\# Figure 4B}
+\NormalTok{p2 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{atlas\_data, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{ThickAvg))  }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"from read\_atlas\_files"}\NormalTok{, }
+       \DataTypeTok{fill=}\StringTok{"Thickness}\CharTok{\textbackslash{}n}\StringTok{(mean)"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_fill\_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"forestgreen"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_colourbar}\NormalTok{(}\DataTypeTok{barheight =} \DecValTok{3}\NormalTok{))}
+
+\NormalTok{(p1 }\OperatorTok{/}\StringTok{ }\NormalTok{p2) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot\_annotation}\NormalTok{(}\DataTypeTok{tag\_levels =} \StringTok{"A"}\NormalTok{)}
 \end{Highlighting}
@@
 \textbf{3)} Apply \texttt{facet\_wrap} or \texttt{facet\_grid} to the plot having used the above two rules.
-An example of this can be seen in Figure \ref{fig:datasupp3}, where a mock data set including summary statistics for two groups ('Young' and `Old') is used when faceting a \emph{ggseg}-plot.
+An example of this can be seen in Figure \ref{fig:datasupp3} B, where a mock data set including summary statistics for two groups ('Young' and `Old') is used when faceting a \emph{ggseg}-plot.
+
+In regards to the second rule of needing to group the data before plotting, this is due to the underlying structure of how the provided data and the atlas data are combined before plotting.
+Since the package function, so far, is a wrapper for \{ggplot2\} rather than a \texttt{geom}, there is certain ggplot2-functionality that is difficult to integrate, like the facets.
+For the complete atlas-data to be merged with \emph{every data group} correctly, the ggseg function needs to have an indication before plotting what the groups are.
+The result of ungrouped data into \texttt{ggseg} is shown in \ref{fig:datasupp3} A.
+Notice how there are three panels rather than two, and that the regions \emph{not represented in the data} have their own panel with all the grey \texttt{NA} colour, while the panels with the data only have the regions represented in the data.
 
@@
 \CommentTok{\# Make some mock data}
-\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
+\NormalTok{someData =}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
   \DataTypeTok{area =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
@@
   \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{8}\NormalTok{),}
-  \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{)),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{group\_by}\NormalTok{(AgeG)}
+  \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{))}
+\NormalTok{)}
 
-\CommentTok{\# Figure 5}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
-      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet\_wrap}\NormalTok{(}\OperatorTok{\textasciitilde{}}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
+\CommentTok{\# Figure 5a}
+\NormalTok{grouped\_plot <{-}}\StringTok{ }\NormalTok{someData }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
+      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet\_wrap}\NormalTok{(}\OperatorTok{\textasciitilde{}}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Ungrouped data"}\NormalTok{)}
+
+\CommentTok{\# Figure 5b}
+\NormalTok{ungrouped\_plot <{-}}\StringTok{ }\NormalTok{someData }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{group\_by}\NormalTok{(AgeG) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
+      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet\_wrap}\NormalTok{(}\OperatorTok{\textasciitilde{}}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Grouped data"}\NormalTok{)}
+
+\CommentTok{\# Combining plots with patchwork, applying theme to all}
+\NormalTok{(grouped\_plot }\OperatorTok{/}\StringTok{ }\NormalTok{ungrouped\_plot) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot\_layout}\NormalTok{(}\DataTypeTok{widths =} \DecValTok{1}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{plot\_annotation}\NormalTok{(}\DataTypeTok{tag\_levels =} \StringTok{\textquotesingle{}A\textquotesingle{}}\NormalTok{) }\OperatorTok{\&}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{8}\NormalTok{))}
 \end{Highlighting}
@@
 
-\begin{figure}
-\centering
-\includegraphics{msc_ggseg_files/figure-latex/datasupp3-1.pdf}
-\caption{\label{fig:datasupp3}To visualize data corresponding to different groups, one needs to use \texttt{group\_by} or similar \textbf{before} plotting with \texttt{ggseg()} for facetting to work.}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{msc_ggseg_files/figure-latex/datasupp3-1} \caption{Examples of incorrect and correct faceting with \texttt{ggseg}.
+\textbf{A}: ungrouped data will facet incorrectly due to the data not being merged correctly with the atlas.
+\textbf{B}: grouped data will merge in the expected way and result in correctly facetted plots.}\label{fig:datasupp3}
 \end{figure}
+
+The \texttt{ggseg} package also has extensive \href{https://lcbc-uio.github.io/ggseg/}{online documentation}, with more functionality than we have covered here.
+The online documentation is also regularly updated with explanations of new features or code improvements.
 
@@
 \CommentTok{\# Figure 6}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area)) }\OperatorTok{+}\StringTok{ }
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area)) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Figure 7 left}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{surface =} \StringTok{"white"}\NormalTok{)  }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{() }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface=\textquotesingle{}white\textquotesingle{})"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+
+\CommentTok{\# Figure 7 center}
+\KeywordTok{ggseg3d}\NormalTok{()  }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{() }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface=\textquotesingle{}LCBC\textquotesingle{})"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+
+\CommentTok{\# Figure 7 right}
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{surface =} \StringTok{"inflated"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{() }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface=\textquotesingle{}inflated\textquotesingle{})"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}[H]
 \includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in \emph{ggseg3d} atlases.
-\textbf{From left to right:} the `white' surface is the white matter surface, `LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the `inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.}\label{fig:ggseg3d-1-out}
+\textbf{From left to right:} the `white' surface is the white matter surface, `LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the `inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.
+The title for each plot is the literal code use to make the subplot.}\label{fig:ggseg3d-1-out}
 \end{figure}
@@
 It is important to note that the coordinates in the plot (X, Y, Z) are \textbf{not} any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
+While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in \texttt{ggseg3d}, at the moment.
+As such, information regarding referencing the regions' location in standard radiological spaces (like MNI or Taliarch) should be acquired directly from imaging software or papers describing the atlases, not \texttt{ggseg3d}.
 
@@
 \CommentTok{\# Figure 8 middle}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData,}
         \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
         \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-        \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"\#ff0000"}\NormalTok{, }\StringTok{"\#00ff00"}\NormalTok{, }\StringTok{"\#0000ff"}\NormalTok{)) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+        \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"\#ff0000"}\NormalTok{, }\StringTok{"\#00ff00"}\NormalTok{, }\StringTok{"\#0000ff"}\NormalTok{)) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}
 \StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
@@
 \CommentTok{\# Figure 8 right}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
+\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData,}
         \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
         \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-        \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+        \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}
 \StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
@@
 
+A glass brain is a translucent representation of the brain that can provide a frame of reference, particularly when looking at subcortical parcellations.
+Since individual brain areas or subcortical structures can be very hard to place without the visual reference of the entire brain, glass brains are often used to aid the viewer understand a regions location.
 Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these \texttt{NA} structures, to improve visualization.
-Glass brains can be added to provide a frame of reference for the subcortical structures with the function \texttt{add\_glassbrain()} (Figure \ref{fig:glassbrain1}) , which takes three extra arguments: hemisphere, colour, and opacity.
+One can control the opacity of the these \texttt{NA} structures, to improve visualization, directly in the \texttt{ggseg3d} call.
+A cortical glass brain can be added to \texttt{ggseg3d} plots through the function \texttt{add\_glassbrain()} (Figure @ref(fig:glass brain1)), which takes three extra arguments: hemisphere, colour, and opacity.
+We recommend only using the glass brain function when plotting subcortical atlases.
+An interactive version of this plot can be found at \url{https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html\#colours}
 
@@
 \KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg\_3d,}
-        \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{add\_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"left lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
+        \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{add\_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"left lateral"}\NormalTok{) }\OperatorTok{\%>\%}
 \StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
@@
 
+(ref:glass brain1-cap) For subcortical structure visualization, one can add a glass brain to the plot.
+This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
+The glass brain is controlled by three options: opacity, hemisphere, and colour.
+
 \begin{figure}[H]
-\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{For subcortical structure visualization, one can add a glass brain to the plot.
-This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
-The glass brain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:glassbrain1}
+\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{(ref:glassbrain1-cap)}\label{fig:glassbrain1}
 \end{figure}
@@
 Finally, we recommend \href{https://github.com/plotly/orca\#installation}{orca} command line tool to save \emph{ggseg3d} atlas snapshots.
+
+The package \href{https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html}{online documentation} also includes interactive versions of many of the figures in this tutorial, and several more options not covered here.
+The online documentation will also be updated as new features are required or we improve the current state of user functionality and options.
 
@@
 The \emph{ggsegExtra}-package is intended to be expanded as a community-effort, as new and informative atlases are published.
-A small collection of the atlases currently in the \emph{ggsegExtra} package may be viewed in Figure \ref{fig:ggsegExtra}
+A small collection of the atlases currently in the \emph{ggsegExtra} package may be viewed in Figure \ref{fig:ggsegExtra}, and explored in our \href{https://athanasiamo.shinyapps.io/ggsegDemo/}{shinyapps.io demonstration}.
 
@@
 \textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.\\
+The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.
 \textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.\\
-- \texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.\\
-- \texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.\\
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.
+- \texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.
+- \texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.
 \textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
@@
 We are in the process of designing a simple, straightforward pipeline to facilitate the creation of new ggseg (2D) atlases.
-The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid \emph{ggseg} and \emph{ggseg3d} atlases.
+The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid \emph{ggseg} and \emph{ggseg3d} atlases.\\
+Additionally, while \texttt{ggseg} works quite nicely as a \texttt{ggplot2} wrapper, ideally it should be a proper \texttt{geom} or \texttt{stat}, making it a \texttt{ggplot2} extension rather than a wrapper.
+This would benefit the user as it would behave more like other \texttt{ggplot2} \texttt{geoms} and there would be less new functionality to learn.
+Furthermore, a \texttt{geom} would make the process of having to group data \emph{before} faceting become unnecessary, and as such also ease use.\\
+We also hope to make it possible to colour each individual mesh triangle in \texttt{ggseg3d}, and as such provide the users with more flexibility in the plotting.
+This should also improve the speed of \texttt{ggseg3d}, which currently is quite slow due to each parcellation region being loaded into \texttt{plotly} sequentially.
+If the entire brain mesh is loaded once, and then the mesh triangles coloured, this would greatly improve speed.
+Additionally, the current procedure makes it necessary to store the 3d-rendered mesh information directly in the package for all atlases, which also takes up a lot of disk space and makes it slow to download.
+When able to colour the mesh triangles directly, rather than each parcellation individually, only text data indexing mesh-locations would be necessary and thus reduce package size considerably.
 We encourage users to contribute to the \emph{ggsegExtra} brain atlas repository by including additional brain atlases.

### 2020-04-06 msc_ggseg.Rmd
update post-acceptance

--- 2019-12-17 msc_ggseg.Rmd
@@
 The main highlighted functions are able to display brain segmentation plots in R.
-Further, the accompanying _ggsegExtra_-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases to _ggseg_ and _ggseg3d_.
+Further, the accompanying _ggsegExtra_-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases for _ggseg_ and _ggseg3d_.
 Overall, the _ggseg_-packages facilitate parcellation-based visualizations in R, improve and ease the dissemination of the results, and increase the efficiency of the workflows.
@@
 
-dt <- as_tibble(data(package='ggseg')$results)
-dt3 <- as_tibble(data(package='ggseg3d')$results)
-dtE <- as_tibble(data(package='ggsegExtra')$results)
+get_data_list <- function(package){
+  as_tibble(data(package=package)$results)
+}
 
-dd <-  bind_rows(dt, dt3) %>% 
-  bind_rows(dtE) %>% 
+pkglist <- c("ggseg", "ggseg3d",  
+             basename(ggsegExtra::ggseg_atlas_repos()$repo))
+dt <- lapply(pkglist, get_data_list) %>% 
+  bind_rows() %>% 
     filter(!grepl("pal", Item)) %>% 
   mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
-         Item = gsub("_3d", "", Item)) %>% 
+         Item = gsub("_3d", "", Item),
+         Item = gsub("_", "", Item)) %>% 
     select(Package, Title, Type, Item) %>% 
@@
   mutate(Citation = case_when(
-    grepl("dkt", Item) ~ "\\citet{dkt}",
+    grepl("dk", Item) ~ "\\citet{dk}",
     grepl("yeo", Item) ~ "\\citet{yeo2011}",
-    grepl("aseg", Item) ~ "\\citet{aseg}",
+    grepl("aseg|hcpa", Item) ~ "\\citet{aseg}",
     grepl("desterieux", Item) ~ "\\citet{desterieux}",
@@
   ))
-rm(dt, dt3, dtE)
+
 options(warnPartialMatchAttr = FALSE,
@@
 Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on structural or connectivity properties (@eickhoff_2018).
-The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons. 
+The use of brain atlases in neuroscience is widespread as these facilitate interpretation and minimize the amount of data (in comparison to using whole-brain data), hence reducing problems with multiple comparisons.
 This enables replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).
@@
 The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
-Adopting the grammar of graphics implemented in _ggplot2_ (@ggplot), one can plot neuroimaging data directly in R with several tools such as ggBrain (@ggBrain) and ggneuro (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
+Adopting the grammar of graphics implemented in _ggplot2_ (@ggplot), one can plot neuroimaging data directly in R with several tools such as _ggBrain_ (@ggBrain) and _ggneuro_ (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
 Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The _ggseg_ and _ggseg3d_ – pluss the complimentary _ggsegExtra_ – packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+The _ggseg_ and _ggseg3d_ – plus the complimentary _ggsegExtra_ – packages, which include pre-compiled data-sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in _ggseg_ is based on polygons and _ggplot2_-based grammar of graphics (@ggplot), while the 3D functionality in _ggseg3d_ is based on tri-surface mesh plots and _plotly_ (@plotly).
 
-Both packages present compiled data sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
-The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; @dkt), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
+The _ggseg_ and _ggseg3d_ packages present compiled data-sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
+The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DK; @dk), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The _ggsegExtra_ package contains a collection of precompiled atlases (currently `r dd %>% filter(grepl("ggsegExtra", Mesh) | grepl("ggsegExtra", Polygon)) %>% nrow()` additional atlases) and it is frequently updated.
-A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in Table \@ref(tab:atlasTab).
+The _ggsegExtra_ package contains a list of online repositories of pre-compiled atlases (currently `r sum(!grepl("None", dt$Polygon))-1` additional polygon atlases and `r sum(!grepl("None", dt$Mesh))-1` mesh atlases) and it is frequently updated.
+A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in Table \@ref(tab:table1).
 
-```{r 'atlasTab', echo=F, warning=FALSE, message=FALSE}
-dd %>%
+```{r 'table1', echo=F, warning=FALSE, message=FALSE}
+dt %>%
   kable(format="latex", escape = FALSE, booktabs = TRUE,
-        caption = "Table of currently available atlases in the ggseg, ggseg3d, and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlas representations, respectively") %>%
+        caption = "Table of currently available atlases in ggseg and ggseg3d, and listed in the ggsegExtra R-package.
+Polygon and Mesh refer to 2D and 3D brain atlas representations, respectively.
+Cell contents in the Mesh and Polygon columns indicate the source package containing the atlas data-set.") %>%
   kable_styling(latex_options = c("scale_down"))
@@
 The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
+The following tutorial complements the package documentation and vignettes that can be found online for all three packages: [ggseg](https://lcbc-uio.github.io/ggseg), [ggseg3d](https://lcbc-uio.github.io/ggseg3d), [ggsegExtra](https://lcbc-uio.github.io/ggsegExtra).
 
@@
 
-_ggseg_ is the main function for plotting 2D data.
-By default, the function automatically plots the DKT atlas (see Figure \@ref(fig:init)).
+`ggseg()` from the _ggseg_ package is the main function for plotting 2D data.
+By default, the function automatically plots the DK atlas (see Figure \@ref(fig:figure1)).
 The `ggseg()`-function is a wrapper for `geom_polygon()` from _ggplot2_, and it can be built upon and combined like any _ggplot_-object.
@@
 
-```{r 'init', fig.width=3, message=FALSE, fig.cap='By default `ggseg()` will plot the dkt atlas in grey shaded polygons.'}
+```{r 'figure1', fig.width=3, message=FALSE, fig.cap='By default `ggseg()` will plot the dk atlas in grey shaded polygons.'}
 # remotes::install_github("LCBC-UiO/ggseg")
 library(ggseg)
-library(dplyr)
-library(tidyr)
+
+# package combining plots easily
+library(patchwork)
 
@@
 These options are atlas-specific.
-For cortical atlases, such as the `dkt`, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:collection) for examples).
+For cortical atlases, such as the `dk`, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:figure2) for examples).
 For subcortical atlases, such as the `aseg` atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
 
-```{r 'collection', fig.width=5, fig.cap='*ggseg* plots can be used with any *ggplot2* feature such as standard scales and themes.
+(ref:figure2-cap) *ggseg* plots can be used with any *ggplot2* feature such as standard scales and themes.
 For cortical atlases, one can supply special *ggseg* options to determine, for example, hemisphere, view, or position.
-**A:** `dkt` atlas, stacked with dark theme ; **B:** `dkt` with medial view only; **C:** `dkt` atlas with only left medial display; **D:** `dkt` atlas, stacked, with classic theme; **E:** `dkt` atlas with left hemisphere only; **F:** complete `aseg` atlas'}
-# dkt dark theme
-p1 <- ggseg(position = "stacked") +
+**A:** `dk` atlas, stacked with dark theme ; **B:** `dk` with medial view only; **C:** `dk` atlas with only left medial display; **D:** `dk` atlas, stacked, with classic theme; **E:** `dk` atlas with left hemisphere only; **F:** axial view of `aseg` atlas
+
+```{r 'figure2', fig.width=6, fig.cap='(ref:figure2-cap)', out.width = "100%"}
+# A: dk dark theme
+fig2A <- ggseg(position = "stacked") +
   theme_dark() +
-  labs(title=" ")
+  labs(title="dk", subtitle = "dark theme")
 
-# dkt classic theme
-p2 <- ggseg(position = "stacked") +
+# B: dk medial view
+fig2B <- ggseg(view = "medial") +
+  labs(title = "dk", subtitle = "medial view")
+
+# C: dk left medial alone
+fig2C <- ggseg(view = "medial",
+            hemisphere = "left") +
+  labs(title="dk", subtitle = "medial left")
+
+# D: dk classic theme
+fig2D <- ggseg(position = "stacked") +
   theme_classic() +
-  labs(title = " ")
+  labs(title = "dk", subtitle = "classic theme")
 
-# dkt medial view
-med <- ggseg(view = "medial") +
-  labs(title = " ")
+# E: dk left hemisphere
+fig2E <- ggseg(hemisphere = "left") +
+  labs(title = "dk", subtitle = "left hemisphere")
 
-# dkt left hemisphere
-left <- ggseg(hemisphere = "left") +
-  labs(title = " ")
+# F: aseg default theme
+fig2F <- ggseg(atlas = aseg, view = "axial") +
+  labs(title = "aseg", subtitle = "axial view")
 
-# aseg default theme
-p3 <- ggseg(atlas=aseg) +
-  labs(title = " ")
-
-# dkt left medial alone
-combo <- ggseg(view = "medial",
-               hemisphere = "left") +
-  labs(title=" ")
-
-# Combine plots to Figure 2
-cowplot::plot_grid(p1, med, combo,  p2, left, p3,
-                   labels = c("A: dkt - dark", "B: dkt - medial",
-                              "C: dkt - combo", "D: dkt - classic",
-                              "E: dkt - left", "F: aseg"),
-                   hjust = -.05)
+# Using patchwork to combine all plots into one
+(fig2A | fig2B | fig2C) / (fig2D | fig2E | fig2F) +
+  plot_annotation(tag_levels = 'A')
 ```
@@
 Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the _ggplot2_ mapping function `aes`.
-To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales (Figure \@ref(fig:fill)).
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales (Figure \@ref(fig:figure3)).
 
-```{r 'fill', fig.width=5, fig.height=5, fig.cap="Supplying 'area' to the fill option in `ggseg()`, will use the column 'area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
-The `dkt` atlas default palette corresponds to the FreeSurferColorLut scheme."}
+(ref:figure3-cap) Supplying 'region' to the fill option in `ggseg()`, will use the column 'region' from the accompanying data-set to create a discrete colour palette over the segments in the atlas.
+The `dk` atlas default palette corresponds to the FreeSurferColorLut scheme.
+
+```{r 'figure3', fig.width=5, fig.height=5, fig.cap="(ref:figure3-cap)"}
 # Figure 3
-ggseg(mapping=aes(fill = area), colour="black") +
-  scale_fill_brain("dkt") +
-  theme(legend.justification=c(1,0),
-        legend.position="bottom",
+ggseg(mapping = aes(fill = region), colour="black") +
+  scale_fill_brain("dk") +
+  theme(legend.justification = c(1,0),
+        legend.position = "bottom",
         legend.text = element_text(size = 5)) +
@@
 Most users will use `ggseg()` to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data sets.
-The atlas data set structure will help users understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique data set.
-All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data-sets.
+The atlas data-set structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data-set.
+All data-sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
-```{r, size='small', eval=FALSE}
-# Look at the top 5 rows of the dkt dataset
-head(dkt, 5)
+```{r, size='small', eval=FALSE, codecap = "Inspect atlas data"}
+# Look at the top 5 rows of the dk data-set
+head(dk, 5)
 ```
@@
 The column 'label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+For example, the `dk` atlas label column matches the region names from Freesurfer statistics table outputs.
 Yet, the data in _ggseg_ is in a long format - that is each region has a single row - and any data of interest needs to be in this same format.
-Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
-See below an example of wide-to-long conversion.
+Often data-sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
+Summary statistic files from Freesurfer have a file format that can be difficult to work with in R. 
+In a set of functions in _ggseg_, we facilitate conversion between FreeSurfer files and ggseg compatible data. 
+There is an [extended vignette](https://lcbc-uio.github.io/ggseg/articles/freesurfer_files.html) of the various options.
+In the following example, we import summary data that has been compiled using Freesurfer's `aparcstats2table` (for surfaces) or `asegstats2table` (volumes). 
+The summary files can combine parcellation metrics for one or more subjects into a single summary file.
+This file also contains extra metrics like intracranial volume and total grey matter volume. 
+These later measures do not map onto ggseg, and as such, one should filter them out of the data returned from `read_freesurfer_table)`.
 
 ```{r, size='small'}
-# Create a mock freesurfer output
-freesurfer_stats <- data.frame(
-  id = c(10:12),
-  lh_superiortemporal = c(3.32, 4.1, 3.5),
-  lh_precentral = c(2.3, 2.5, 2.1),
-  lh_rostralmiddlefrontal = c(3.3, 3.2, 3.1)
-)
-freesurfer_stats
+file_path <- "data/aparc.volume.table"
 
-# Wrangle wide format data into long format
-freesurfer_long <- freesurfer_stats %>%
-  gather(label, thickness, -id)
-freesurfer_long
+someData <- read_freesurfer_table(file_path, measure = "volume") %>% 
+  # remove rows in the data that do not contain the strings "rh" or "lh"
+  dplyr::filter(grepl("rh|lh", label))
+someData
 ```
 
-Data in long format can then be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dkt`.
+A second option is to read in atlas files directly from the entire subject directory, using regular expression to select the target atlas. 
+The default aparc has a unique string ending with `aparc.stats` which we will use to read in all subject stats for that particular atlas.
+One can filter specific subjects and/or regions from the R-object data.frame.
+
+```{r, warning=FALSE}
+atlas_data <- read_atlas_files(
+  # Path to Freesurfers $SUBJECTS_DIR
+  subjects_dir = "/Applications/freesurfer/subjects/",
+  # Regular expression to find atlas files
+  atlas = "aparc.stats$"
+)
+atlas_data
+```
+
+This long format data can be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dk`.
 The data **must** include a column that has the same column name and at least _some_ data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, with an 'area' and 'p' column, representing the results of a hypothetical analysis.
-The `ggseg()`-function will recognise the matching column 'area', and merge the supplied data into the atlas using _dplyr_-joins.
-We use the 'p' column as the column flooding the segment with colour.
-The appearance of the plot can then be modified similarly to any other _ggplot2_ graph using functions such as scales, labs, themes, etc., as seen in Figure \@ref(fig:datasupp)
+In the following example, the `ggseg()`-function will recognise the matching column 'label', and merge the supplied data into the atlas using _dplyr_-joins.
+The appearance of the plot can then be modified similarly to any other _ggplot2_ graph using functions such as scales, labs, themes, etc., as seen in Figure \@ref(fig:figure4)
 
-```{r 'datasupp', fig.width=5, fig.cap="Supplying data through the '.data' option in `ggseg()` enables use of columns in the supplied data to aesthetical arguments (such as 'fill').
-The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions."}
-# Make some mock data
-someData = data.frame(
-  area = c("transverse temporal", "insula",
-           "pre central","superior parietal"),
-  p = sample(seq(0,.5,.001), 4),
-  stringsAsFactors = FALSE)
+(ref:figure4-cap) Supplying data through the '.data' option in `ggseg()` enables use of columns in the supplied data to aesthetical arguments (such as 'fill').
+The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions.
 
-# Figure 4
-ggseg(.data=someData, mapping=aes(fill=p))  +
-  labs(title="A nice plot title", fill="p-value") +
-  scale_fill_gradient(low="firebrick",high="goldenrod")
+```{r 'figure4', fig.width=5, fig.cap="(ref:figure4-cap)", warning = FALSE}
+
+# Figure 4A
+fig4A <- someData %>% 
+  # plot only subject bert
+  filter(subject == "bert") %>%
+
+  ggseg(mapping=aes(fill=volume))  +
+  labs(title="from read_freesurfer_table", 
+       fill="volume\n(mm^3)") +
+  scale_fill_gradient(low="firebrick",high="goldenrod") +
+  guides(fill = guide_colourbar(barheight = 3))
+
+# Figure 4B
+fig4B <- atlas_data %>% 
+  # plot only subject bert
+  filter(subject == "bert") %>%
+
+  ggseg(mapping=aes(fill=ThickAvg))  +
+  labs(title="from read_atlas_files", 
+       fill="Thickness\n(mean)") +
+  scale_fill_gradient(low="forestgreen",high="goldenrod")+
+  guides(fill = guide_colourbar(barheight = 3))
+
+(fig4A / fig4B) +
+  plot_annotation(tag_levels = "A")
 ```
@@
 If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column 'hemi' with either 'right' or 'left'.
-In this case, data will be merged into the atlas both by 'area' and by 'hemi'.
-For more information about adapting data and viewing only one hemisphere or side, the [package vignettes](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.
+In this case, data will be merged into the atlas both by 'region' and by 'hemi'.
+For more information about adapting data and viewing only one hemisphere or side, the [online documentation](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.
 
@@
 This may be obtained also with `ggseg()`, using _ggplot2_'s `facet_wrap` or `facet_grid`, with three guiding rules:
-**1)** as before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in seperate rows, not in separate columns).
-**2)** The data needs to be grouped using _dplyr_'s `group_by()` function _before_ providing the data to the `ggseg()`-function.
+
+1. As before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in separate rows, not in separate columns).  
+2. The data needs to be grouped using _dplyr_'s `group_by()` function _before_ providing the data to the `ggseg()`-function. 
 The `ggseg()`-function will detect grouped data, and adapt it to `facet`'s requirements.  
-**3)** Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules. 
-An example of this can be seen in Figure \@ref(fig:datasupp3), where a mock data set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.
+3. Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules.
+An example of this can be seen in Figure \@ref(fig:figure5) B, where a mock data-set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.
 
-All the concepts described above also work with the `aseg` atlas for subcortical structures, except for 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:atlases)).
+The second rule is needed as the `ggseg()`-function is a ggplot-wrapper rather than an extension, and thus has a particular structure and pipeline to combine the atlas to supplied data.
+The result of ungrouped data into `ggseg` is shown in  \@ref(fig:figure5) A.
+Note that the data is combined incorrectly, there are three panels rather than two. 
+Regions not represented in the data (`NA`-values) have their own panel, while the remaining two panels only include regions with valid data.
+All the concepts described above also work with the `aseg` atlas for subcortical structures, except for 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:figure6)).
 
-```{r 'datasupp3', fig.width=5, fig.cap='To visualize data corresponding to different groups, one needs to use `group_by` or similar **before** plotting with `ggseg()` for facetting to work.'}
-# Make some mock data
-someData = data.frame(
-  area = rep(c("transverse temporal", "insula",
-               "pre central","superior parietal"),2),
+(ref:figure5-cap) Examples of incorrect and correct faceting with `ggseg`.
+**A**: ungrouped data will facet incorrectly due to the data not being merged correctly with the atlas.
+**B**: grouped data will merge in the expected way and result in correctly facetted plots.
+
+```{r 'figure5', fig.width=5,  warning= FALSE, fig.cap='(ref:figure5-cap)', out.width="100%"}
+# Make some simpler mock data
+someData = tibble(
+  region = rep(c("transverse temporal", "insula",
+               "precentral","superior parietal"),2),
   p = sample(seq(0,.5,.001), 8),
-  AgeG = c(rep("Young",4), rep("Old",4)),
-  stringsAsFactors = FALSE) %>%
-  group_by(AgeG)
+  AgeG = c(rep("Young",4), rep("Old",4))
+)
 
-# Figure 5
-ggseg(.data=someData, colour="white", position = "stacked",
-      mapping=aes(fill=p)) +
-  facet_wrap(~AgeG, ncol=2) +
-  theme(legend.position = "bottom")
+# Figure 5a
+fig5A <- someData %>%
+  ggseg(position = "stacked",
+      mapping=aes(fill=p), show.legend = FALSE) +
+  facet_wrap(~AgeG, ncol=3) +
+  labs(title = "Ungrouped data")
+
+# Figure 5b
+fig5B <- someData %>%
+  group_by(AgeG) %>%
+  ggseg(position = "stacked",
+      mapping=aes(fill=p), show.legend = FALSE) +
+  facet_wrap(~AgeG, ncol=3) +
+  labs(title = "Grouped data")
+
+# Combining plots with patchwork, applying theme to all
+(fig5A / fig5B) +
+  plot_layout(widths = 1) + 
+  plot_annotation(tag_levels = 'A') &
+  theme(axis.text = element_text(size=8))
 ```
 
-```{r 'atlases', fig.width=5, fig.cap='The `aseg` atlas showing subcortical structures has some distinct differences from the `dkt`.
+The `ggseg` package also has extensive [online documentation](https://lcbc-uio.github.io/ggseg/), with more functionality than we have covered here.
+The online documentation is also regularly updated with explanations of new features or code improvements.
+
+(ref:figure6-cap) The `aseg` atlas showing subcortical structures has some distinct differences from the `dk`.
 For instance, there is no option to show only a single hemisphere.
-Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.'}
+Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.
+
+```{r 'figure6', fig.width=5, fig.cap='(ref:figure6-cap)', collapse = TRUE}
 # Figure 6
-ggseg(atlas="aseg", mapping=aes(fill=area)) + 
+ggseg(atlas="aseg", mapping=aes(fill=region)) +
   theme(legend.justification=c(1,0),
@@
 
-Out-of-the-box, `ggseg3d()` plots the `dkt_3d` atlas in 'LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \@ref(fig:ggseg3d-1-out))
-The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage5_ template subject.
-All `[...]_3d` atlases include a `colour` column that based on the color scheme used in the source neuroimaging software.
+Out-of-the-box, `ggseg3d()` plots the `dk_3d` atlas in 'LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \@ref(fig:figure7))
+The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage5_ template subject, a surface template that is issued together with the `ggsegExtra` package.
+All `[...]_3d` atlases include a `colour` column that is based on the color scheme used in the source neuroimaging software.
 
-(ref:gseg3d-1-out-cap) The three surface options provided in *ggseg3d* atlases.
+(ref:figure7-cap) The three surface options provided in *ggseg3d* atlases.
 **From left to right:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.
+The title for each plot is the literal code use to make the subplot.
 
-```{r "ggseg3d-1-out", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:gseg3d-1-out-cap)'}
-p <- ggseg3d(surface = "white")  %>%
+```{r "figure7-show", message=FALSE, echo=TRUE, results = 'hide',  out.width='30%',fig.show='hold',fig.ncol = 3}
+# remotes::install_github("LCBC-UiO/ggseg3d")
+library(ggseg3d)
+
+# Figure 7 left
+fig7l <- ggseg3d(surface = "white")  %>%
   pan_camera("right lateral") %>%
@@
          font =  list(size = 30))
-orca(p, file = "png/ggseg3d_white.png", format = "png")
+fig7l
 
-p <- ggseg3d()  %>%
+# Figure 7 middle
+fig7m <- ggseg3d()  %>%
   pan_camera("right lateral") %>%
@@
          font =  list(size = 30))
-orca(p, file = "png/ggseg3d.png", format = "png")
+fig7m
 
-p <- ggseg3d(surface = "inflated") %>%
+# Figure 7 right
+fig7r <- ggseg3d(surface = "inflated") %>%
   pan_camera("right lateral") %>%
@@
          font =  list(size = 30))
-orca(p, file = "png/ggseg3d_inflated.png", format = "png")
+fig7r
+```
 
-include_graphics(c("png/ggseg3d_white.png","png/ggseg3d.png", "png/ggseg3d_inflated.png"))
+```{r "figure7", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:figure7-cap)'}
+orca(fig7l, file = "msc_ggseg_files/figure-latex/figure7l.png", format = "png")
+orca(fig7m, file = "msc_ggseg_files/figure-latex/figure7m.png", format = "png")
+orca(fig7r, file = "msc_ggseg_files/figure-latex/figure7r.png", format = "png")
+
+include_graphics(c("msc_ggseg_files/figure-latex/figure7l.png",
+                   "msc_ggseg_files/figure-latex/figure7m.png",
+                   "msc_ggseg_files/figure-latex/figure7r.png"))
 ```
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \@ref(fig:ggseg3d-1-out)) and the two hemispheres.
+Each cortical atlas has data-sets for three different surfaces (see Figure \@ref(fig:figure7)) and the two hemispheres.
 Only one surface is available for subcortical atlases as inflation procedures are irrelevant.
@@
 It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
-
-```{r, out.width='90%', size='small'}
-# remotes::install_github("LCBC-UiO/ggseg3d")
-library(ggseg3d)
-```
+While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in `ggseg3d`, at the moment.
+As such, information regarding referencing the regions' location in standard radiological spaces (like MNI or Taliarch) should be acquired directly from imaging software or papers describing the atlases, not `ggseg3d`.
 
@@
 
-```{r out.width='80%', linewidth=100, size='small', eval=FALSE}
+```{r out.width='80%', linewidth=100, size='small', eval=TRUE}
 # Select surface and hemisphere, and then unnest to inspect the atlas data
-dkt_3d %>%
+dk_3d %>%
   filter(surf == "inflated" & hemi == "right") %>%
@@
 Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
-It should also be noted that the 'label', 'annot' and 'area' columns could provide matching values for your own data.
-Similarly to the `ggseg()`-function, the 'label' column should match the region names used in the original neuroimaging software while 'area' and 'annot' provide alternative/secondary names.
+It should also be noted that the 'label', 'annot' and 'region' columns could provide matching values for your own data.
+Similarly to the `ggseg()`-function, the 'label' column should match the region names used in the original neuroimaging software while 'region' and 'annot' provide alternative/secondary names.
 It is thus important to match your regional identifiers with those used in the atlas.
@@
 
-```{r, eval=FALSE, out.width='40%'}
+```{r 'figure8-show', include=FALSE, out.width='40%', results='hide'}
 # Figure 8 left
-ggseg3d(.data = someData, atlas = dkt_3d, colour = "p", text = "p")
+fig8l <- ggseg3d(.data = someData, 
+                 atlas = dk_3d, colour = "p", text = "p") %>% 
+  pan_camera("right lateral")
+fig8l
 
 # Figure 8 middle
-ggseg3d(.data = someData, 
-        atlas = dkt_3d,
+fig8m <- ggseg3d(.data = someData,
+        atlas = dk_3d,
         colour = "p", text = "p",
@@
   remove_axes()
+fig8m
 
 # Figure 8 right
-ggseg3d(.data = someData, 
-        atlas = dkt_3d,
+fig8r <- ggseg3d(.data = someData,
+        atlas = dk_3d,
         colour = "p", text = "p",
@@
   remove_axes()
+fig8r
 ```
 
-(ref:ggseg3d-data1-cap) **Left:** Supplying data to `ggseg3d()` works similarly to `ggseg()`.
+(ref:figure8-cap) **Left:** Supplying data to `ggseg3d()` works similarly to `ggseg()`.
 Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function.
@@
 
-```{r ggseg3d-data1, out.width='30%', echo=FALSE, fig.cap="(ref:ggseg3d-data1-cap)", fig.show='hold', fig.ncol = 3}
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-             colour = "p", text = "p") %>%   
-  pan_camera("right lateral") %>% 
-  remove_axes()
-orca(p, file = "png/ggseg3d_data.png", format = "png")
-
-p <- ggseg3d(.data = someData, atlas = dkt_3d,
-             colour = "p", text = "p",
-             palette = c("#ff0000", "#00ff00", "#0000ff")) %>%   
-  pan_camera("right lateral") %>% 
-  remove_axes()
-orca(p, file = "png/ggseg3d_colour.png", format = "png")
-
-p <-   ggseg3d(.data = someData, atlas = dkt_3d,
-               colour = "p", text = "p",
-               na.colour = "black") %>%   
-  pan_camera("right lateral") %>% 
-  remove_axes()
-orca(p, file = "png/ggseg3d_nacolour.png", format = "png")
+```{r 'figure8', out.width='30%', echo=FALSE, fig.cap="(ref:figure8-cap)", fig.show='hold', fig.ncol = 3}
+orca(fig8l, file = "msc_ggseg_files/figure-latex/figure8l.png", format = "png")
+orca(fig8m, file = "msc_ggseg_files/figure-latex/figure8m.png", format = "png")
+orca(fig8r, file = "msc_ggseg_files/figure-latex/figure8r.png", format = "png")
 
 include_graphics(c(
-  "png/ggseg3d_data.png",
-  "png/ggseg3d_colour.png",
-  "png/ggseg3d_nacolour.png"))
+  "msc_ggseg_files/figure-latex/figure8l.png",
+  "msc_ggseg_files/figure-latex/figure8m.png",
+  "msc_ggseg_files/figure-latex/figure8r.png"))
 
@@
 
-You can provide custom colour palettes either in hex or R-names, as seen in Figure \@ref(fig:ggseg3d-data1).
+You can provide custom colour palettes either in hex or R-names, as seen in Figure \@ref(fig:figure8).
 Colours will be evenly spaced when creating the colour-scale.
@@
 ### Adding a glass brain
-Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these `NA` structures, to improve visualization.
-Glass brains can be added to provide a frame of reference for the subcortical structures with the function `add_glassbrain()` (Figure \@ref(fig:glassbrain1)) , which takes three extra arguments: hemisphere, colour, and opacity.
+A glass brain is a translucent representation of the brain that can provide a frame of reference,
+particularly when looking at subcortical parcellations and it only serves as a visual aid.
+One can control the opacity of the these `NA` structures, to improve visualization, directly in the `ggseg3d` call.
+A cortical glass brain can be added to `ggseg3d` plots through the function `add_glassbrain()` (Figure \@ref(fig:figure9)), which takes three extra arguments: hemisphere, colour, and opacity.
+We recommend only using the glass brain function when plotting subcortical atlases.
+A similar, interactive version of this plot can be found in the [online documentation](https://lcbc-uio.github.io/ggseg3d//articles/ggseg3d.html#colours-1).
 
-```{r, out.width='40%', warning=F, message=FALSE, eval=FALSE}
+```{r 'figure9-show', out.width='40%', warning=F, message=FALSE, results = 'hide'}
 # Figure 9
-ggseg3d(atlas = aseg_3d,
+fig9 <- ggseg3d(atlas = aseg_3d,
         na.alpha= .5) %>%
   add_glassbrain("left") %>%
-  pan_camera("left lateral") %>% 
-  remove_axes()
+  pan_camera("left lateral")
+fig9
 ```
 
-(ref:glassbrain1-cap) For subcortical structure visualization, one can add a glass brain to the plot.
+(ref:figure9-cap) For subcortical structure visualization, one can add a glass brain to the plot.
 This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
@@
 
-```{r glassbrain1, out.width='60%', echo=F, eval=T, fig.cap="(ref:glassbrain1-cap)"}
-p <- ggseg3d(atlas = aseg_3d,
-             na.alpha= .5) %>%   
-  add_glassbrain("left") %>% 
-  pan_camera("right lateral") %>% 
-  remove_axes()
-include_plotly(p, "png/ggseg3d_glass.png")
+```{r 'figure9', out.width='60%', echo=F, eval=T, fig.cap="(ref:figure9-cap)"}
+include_plotly(fig9, "msc_ggseg_files/figure-latex/figure9.png")
 ```
@@
 
+The package [online documentation](https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html) also includes interactive versions of many of the figures in this tutorial, and several more options not covered here.
+The online documentation will also be updated as new features are required or we improve the current state of user functionality and options.
+
 ## Additional atlases
-The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dkt` (@dkt) and `aseg` (@aseg) atlases. 
+The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dk` (@dk) and `aseg` (@aseg) atlases.
 These are, however, only two among many meaningful ways of segmenting the brain into different regions.
-Thus, the _ggsegExtra_ package is a repository containing additional with additional data sets for plotting with the _ggseg_ and _ggseg3d_ packages. 
+Thus, the _ggsegExtra_ package is a library with two main functionalities: 1) to install additional data-sets for plotting with the _ggseg_ and _ggseg3d_ packages from other repositories, and 2) to create custom atlases for _ggseg_ and _ggseg3d_.
+
+### Verified existing atlases
 There is an ever-increasing amount of new atlases being created, as research and methods in neuroimaging analysis progresses.
 The _ggsegExtra_-package is intended to be expanded as a community-effort, as new and informative atlases are published.
-A small collection of the atlases currently in the _ggsegExtra_ package may be viewed in Figure \@ref(fig:ggsegExtra)
+A collection of the atlases currently listed in the _ggsegExtra_ package may be found in Figure \@ref(fig:figure10) and explored in our [shinyapps.io demonstration](https://athanasiamo.shinyapps.io/ggsegDemo/).
 
-```{r,  include = FALSE}
+```{r}
+# remotes::install_github("LCBC-UiO/ggsegExtra")
 library(ggsegExtra)
-k1 <- ggseg(atlas = yeo7 , mapping=aes(fill=area),
+
+# list all known ggseg atlas-repositories
+ggseg_atlas_repos()
+
+# list repositories matching a pattern search
+yeo_repo <- ggseg_atlas_repos("Yeo")
+```
+
+Having found an existing atlas, you will need to install this like an R-library.
+There is a convenience-function in _ggsegExtra_ that will help you do that.
+Alternatively, you can install all the existing atlases. 
+This function will take time to execute and will consume quite some space as each atlas (particularly ggseg3d-atlases) can be large. 
+Once the atlases are installed, the atlas-packages can be loaded like any other R-package using `library`.
+
+```{r, eval = FALSE}
+# Install ggseg-atlas listed by ggseg_atlas_repos("Yeo")
+install_ggseg_atlas(yeo_repo$repo, yeo_repo$source)
+
+# This will take quite some time, there are many atlases with heavy data.
+install_ggseg_atlas_all()
+```
+
+```{r 'figure10-show', results = 'hide', message = FALSE}
+library(ggsegYeo2011)
+library(ggsegGlasser)
+library(ggsegDesterieux)
+library(ggsegICBM)
+
+yeo7_p <- ggseg(atlas = yeo7 , mapping=aes(fill=region),
             show.legend = FALSE, position="stacked") +
-  scale_fill_brain("yeo7") +
-  labs(title="yeo7-data set") 
-ggsave(filename = "png/ggseg_extra1.png", plot = k1, device = "png")
+  scale_fill_brain("yeo7", package = "ggsegYeo2011") +
+  labs(title="yeo7-data-set")
 
-k2 <- ggseg(atlas = glasser , mapping=aes(fill=area),
+glasser_p <- ggseg(atlas = glasser , mapping=aes(fill=region),
             show.legend = FALSE, position="stacked") +
-  scale_fill_brain("glasser") +
-  labs(title="glasser-data set")
-kk <- cowplot::plot_grid(k1, k2)
-ggsave(filename = "png/ggseg_extra.png", plot = kk, device = "png")
+  scale_fill_brain("glasser", package = "ggsegExtra") +
+  labs(title="glasser-data-set")
 
-p <- ggseg3d(atlas = desterieux_3d) %>%   
+ggsegs_p <- yeo7_p + glasser_p + plot_annotation(tag_levels = "A")
+
+desterieux_p <- ggseg3d(atlas = desterieux_3d) %>%
   pan_camera("right lateral") %>%
@@
   add_annotations(
-    text="icbm_3d data set",
+    text="desterieux_3d data-set",
     y = 1,
@@
                 size = 50))
-orca(p, file = "png/ggseg3d_extra1.png", format = "png")
 
-p <- ggseg3d(atlas = icbm_3d) %>%   
+icbm_p <- ggseg3d(atlas = icbm_3d) %>%
   pan_camera("right lateral") %>%
@@
   add_annotations(
-    text="icbm_3d data set",
+    text="icbm_3d data-set",
     y = 1,
@@
                 size = 50))
-orca(p, file = "png/ggseg3d_extra2.png", format = "png")
+
 ```
 
-(ref:ggsegExtra-cap) Four example data sets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.
+(ref:figure10-cap) Four example data-sets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.
 
-```{r "ggsegExtra", out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='(ref:ggsegExtra-cap)'}
+```{r 'figure10', out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='(ref:figure10-cap)'}
+
+invisible(ggsave(filename = "msc_ggseg_files/figure-latex/figure10ab.png", 
+                 plot = ggsegs_p, device = "png"))
+orca(desterieux_p, file = "msc_ggseg_files/figure-latex/figure10c.png", format = "png")
+orca(icbm_p, file = "msc_ggseg_files/figure-latex/figure10d.png", format = "png")
+
 include_graphics(c(
-  "png/ggseg_extra.png",
-  "png/ggseg3d_extra1.png",
-  "png/ggseg3d_extra2.png"
+  "msc_ggseg_files/figure-latex/figure10ab.png",
+  "msc_ggseg_files/figure-latex/figure10c.png",
+  "msc_ggseg_files/figure-latex/figure10d.png"
 ))
 ```
+
+### Creating custom atlases
+In many cases, users would like to create their own atlases based on existing parcellations or ones own imaging results.
+There are several functions in _ggsegExtra_ to aid creating ones own atlas compatible with either _ggseg_ or _ggseg3d_. 
+These processes depend on the origin of the segmentation (surface-based vs. volume-based), and what type of system you are working on. 
+In addition, one will need other software dependencies to create new ggseg-compatible atlases. 
+Therefore, we will not describe the process here, but refer the readers to the "Articles" sections of the [online documentation](https://lcbc-uio.github.io/ggsegExtra/index.html), for further explanations and guides.
+We encourage users who successfully create new atlases, to contribute to the _ggsegExtra_ brain atlas repository by adding the repository containing the brain atlases to the repository-list.
 
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
-This is exemplified in the online supplementary information of Vidal-Piñeiro et al. (-@vidal_2019).
+This is exemplified in the [online supplementary information](https://athanasiamo.shinyapps.io/Virtual_histology_2019/) of Vidal-Piñeiro et al. (-@vidal_2019).
 
+The functions in the _ggseg_-packages provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in _ggsegExtra_.
 The three _ggseg_-packages contain three main features:
 
-**1)** a collection of 2D-polygon and 3D mesh brain parcellation atlases.
+1.
+A collection of 2D-polygon and 3D mesh brain parcellation atlases.
 The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.
-**2)** `ggseg()` and `ggseg3d()` functions for visualization.
+2.
+`ggseg()` and `ggseg3d()` functions for visualization.
 Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from _ggplot2_ and _plotly_, respectively.
-    - `ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object.   
-    - `ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.  
-**3)** Complimentary features – e.g. color scales - and functions such as `as_ggseg_atlas()` and `as_ggseg3d_atlas()` to convert data in the correct atlas format.
-
-These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in _ggsegExtra_.
+`ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object. 
+`ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.  
+3.
+Complimentary features, like color scales, and functions to convert compatible data frames to atlas-classes, read in likely data sources (like Freesurfer analysis output), and help create new atlases for use with _ggseg_ and _ggseg3d_.
 
@@
 
+The ggseg-packages are based on _ggplot2_ and _plotly_, which are only two of several visualisation tools in R, in addition to base-R plotting. 
+While the functionality presented here could also be integrated into packages such as [lattice](https://www.rdocumentation.org/packages/lattice/versions/0.20-40), both _ggplot2_ and _plotly_ have become established tools in the R-community and have extensive online documentation.
+This makes them great packages to base further extensions and wrappers on.
+
 Finally - while the _ggseg_-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
@@
 # Planned package improvements
-In the _ggsegExtra_ [github wiki](https://github.com/LCBC-UiO/ggsegExtra/wiki/Contributing%3A-polygon-atlases-new), we offer a pipeline to create and supply atlases for 2D plotting.
-At the moment, the creation of atlas for ggseg is convoluted and difficult and requires manual intervention. 
-We are in the process of designing a simple, straightforward pipeline to facilitate the creation of new ggseg (2D) atlases.
-The aim is to create a set of functions that will call specialized tools like FSL(@fsl), Freesurfer(@fischl_99, @dale_99, @Fischl2000) and Imagemagick(@magick) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid _ggseg_ and _ggseg3d_ atlases.
-We encourage users to contribute to the _ggsegExtra_ brain atlas repository by including additional brain atlases.
+In terms of planned _ggseg_ improvements, we are working on making a proper `geom` or `stat`.
+While `ggseg()` works quite nicely as a `ggplot` wrapper, it introduces some unexpected work for the user, like having to group data _before_ faceting. 
+A `geom` or `stat` would benefit the user as it would behave more like other _ggplot2_ `geoms` and there would be less new functionalities to learn.
+
+Concerning _ggseg3d_, we also hope to make it possible to colour each individual mesh triangle, and as such provide the users with more flexibility in the plotting. 
+This should also improve the speed of `ggseg3d()`, which currently is quite slow due to each parcellation region being loaded into `plotly` sequentially. 
+Loading the entire brain mesh at once, and then coloring mesh triangles would significiantly improve the speed of `ggseg3d()`. 
+When able to colour the mesh triangles directly, rather than each parcellation individually, only full brain representations of each surface (rather than each surface of each atlas) needs to be stored, in addition to text data indexing mesh-locations.
+This would reduce package size considerably, including package sizes of all atlas-repositories.
 

### 2020-04-06 msc_ggseg.tex
update post-acceptance

--- 2019-12-17 msc_ggseg.tex
@@
 The main highlighted functions are able to display brain segmentation plots in R.
-Further, the accompanying \emph{ggsegExtra}-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases to \emph{ggseg} and \emph{ggseg3d}.
+Further, the accompanying \emph{ggsegExtra}-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases for \emph{ggseg} and \emph{ggseg3d}.
 Overall, the \emph{ggseg}-packages facilitate parcellation-based visualizations in R, improve and ease the dissemination of the results, and increase the efficiency of the workflows.
@@
 Visualization is increasingly important for accurate guidance and interpretation of neuroimaging results, as current research is able to generate a high amount of data and outcomes.
-For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (\textgreater100.000).
+For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (\textgreater{}100.000).
 Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
 Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on structural or connectivity properties (\citet{eickhoff_2018}).
-The use of brain atlases is widespread as these facilitate interpretation and minimize the amount of data, hence reducing problems with multiple comparisons.
+The use of brain atlases in neuroscience is widespread as these facilitate interpretation and minimize the amount of data (in comparison to using whole-brain data), hence reducing problems with multiple comparisons.
 This enables replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (\citet{R}).
@@
 The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
-Adopting the grammar of graphics implemented in \emph{ggplot2} (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as ggBrain (\citet{ggBrain}) and ggneuro (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
+Adopting the grammar of graphics implemented in \emph{ggplot2} (\citet{ggplot}), one can plot neuroimaging data directly in R with several tools such as \emph{ggBrain} (\citet{ggBrain}) and \emph{ggneuro} (\citet{ggneuro}; see neuroconductor \citeyearpar{neuroconductor} for curated neuroimaging packages for R).
 Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.
@@
 In this tutorial, we introduce two packages for visualizing brain atlas data in R.
-The \emph{ggseg} and \emph{ggseg3d} -- pluss the complimentary \emph{ggsegExtra} -- packages include pre-compiled data sets for different brain atlases that allow for 2D and 3D visualization.
+The \emph{ggseg} and \emph{ggseg3d} -- plus the complimentary \emph{ggsegExtra} -- packages, which include pre-compiled data-sets for different brain atlases that allow for 2D and 3D visualization.
 The two-dimensional functionality in \emph{ggseg} is based on polygons and \emph{ggplot2}-based grammar of graphics (\citet{ggplot}), while the 3D functionality in \emph{ggseg3d} is based on tri-surface mesh plots and \emph{plotly} (\citet{plotly}).
 
-Both packages present compiled data sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
-The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DKT; \citet{dkt}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
+The \emph{ggseg} and \emph{ggseg3d} packages present compiled data-sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
+The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DK; \citet{dk}), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; \citet{aseg}), which covers the subcortical structures.
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (\citet{amlien_elaboration_2019}, \citet{WALHOVD20051261}, \citet{Pizzagalli}).
-The \emph{ggsegExtra} package contains a collection of precompiled atlases (currently 15 additional atlases) and it is frequently updated.
-A summary of all available atlases compatible with the \emph{ggseg}-packages as of December 2019 can be found in Table \ref{tab:atlasTab}.
+The \emph{ggsegExtra} package contains a list of online repositories of pre-compiled atlases (currently 10 additional polygon atlases and 11 mesh atlases) and it is frequently updated.
+A summary of all available atlases compatible with the \emph{ggseg}-packages as of April 2020 can be found in Table \ref{tab:table1}.
 
@@
 
-\caption{\label{tab:atlasTab}Table of currently available atlases in the ggseg, ggseg3d, and the ggsegExtra R-packages. Polygon and mesh refer to 2D and 3D brain atlas representations, respectively}
+\caption{\label{tab:table1}Table of currently available atlases in ggseg and ggseg3d, and listed in the ggsegExtra R-package.
+Polygon and Mesh refer to 2D and 3D brain atlas representations, respectively.
+Cell contents in the Mesh and Polygon columns indicate the source package containing the atlas data-set.}
 \centering
@@
 \toprule
-Title & Item & Mesh & Polygon & citation\\
+Title & Item & Mesh & Polygon & Citation\\
 \midrule
-Automated probabilistic reconstruction of white-matter pathways & tracula & None & ggsegExtra & \citet{tracula}\\
-Desikan-Killiany Cortical Atlas & dkt & ggseg3d & ggseg & \citet{dkt}\\
-Desterieux cortical parcellations & desterieux & ggsegExtra & None & \citet{desterieux}\\
+Automated probabilistic reconstruction of white-matter pathways & tracula & None & ggsegTracula & \citet{tracula}\\
+Desikan-Killiany Cortical Atlas & dk & ggseg3d & ggseg & \citet{dk}\\
+Desikan-Killiany Cortical Atlas & dkextra & None & ggsegDefaultExtra & \citet{dk}\\
+Desterieux cortical parcellations & desterieux & ggsegDesterieux & None & \citet{desterieux}\\
+Freesurfer ASEG with posterior-anterior hippocampus & hcpa & ggsegDefaultExtra & None & \citet{aseg}\\
+\addlinespace
 Freesurfer automatic subcortical segmentation of a brain volume & aseg & ggseg3d & ggseg & \citet{aseg}\\
-Genetic topography of brain area morphology & chenAr & None & ggsegExtra & \citet{chen}\\
+Genetic topography of brain area morphology & chenAr & None & ggsegChen & \citet{chen}\\
+Genetic topography of brain thickness morphology & chenTh & None & ggsegChen & \citet{chen}\\
+Harvard-Oxford Cortical atlas & hoCort & None & ggsegHO & \citet{ho}\\
+HPC - Multi-modal parcellation of human cerebral cortex & glasser & None & ggsegGlasser & \citet{glasser}\\
 \addlinespace
-Genetic topography of brain thickness morphology & chenTh & None & ggsegExtra & \citet{chen}\\
-Harvard-Oxford Cortical atlas & hoCort & None & ggsegExtra & \citet{ho}\\
-HPC - Multi-modal parcellation of human cerebral cortex & glasser & None & ggsegExtra & \citet{glasser}\\
-ICBM white matter parcellation & icbm & ggsegExtra & None & \citet{icbm}\\
-JHU parcellation & jhu & None & ggsegExtra & \citet{jhu}\\
+ICBM white matter parcellation & icbm & ggsegICBM & None & \citet{icbm}\\
+JHU parcellation & jhu & None & ggsegJHU & \citet{jhu}\\
+Local-Global 17 Parcellation of the Human Cerebral Cortex & schaefer17 & ggsegSchaefer & None & \citet{schaefer}\\
+Local-Global 7 Parcellation of the Human Cerebral Cortex & schaefer7 & ggsegSchaefer & None & \citet{schaefer}\\
+Parcellation from JHU & jhu & ggsegJHU & None & \citet{jhu}\\
 \addlinespace
-Local-Global 17 Parcellation of the Human Cerebral Cortex & schaefer17 & ggsegExtra & None & \citet{schaefer}\\
-Local-Global 7 Parcellation of the Human Cerebral Cortex & schaefer7 & ggsegExtra & None & \citet{schaefer}\\
-Parcellation from JHU & jhu & ggsegExtra & None & \citet{jhu}\\
-Parcellation from the Human Connectome Project & glasser & ggsegExtra & None & \citet{glasser}\\
-White matter tract parcellations & tracula & ggsegExtra & None & \citet{tracula}\\
-\addlinespace
-Yeo 17 Resting-state Cortical Parcellations & yeo17 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
-Yeo 7 Resting-state Cortical Parcellations & yeo7 & ggsegExtra & ggsegExtra & \citet{yeo2011}\\
+Parcellation from the Human Connectome Project & glasser & ggsegGlasser & None & \citet{glasser}\\
+White matter tract parcellations & tracula & ggsegTracula & None & \citet{tracula}\\
+Yeo 17 Resting-state Cortical Parcellations & yeo17 & ggsegYeo2011 & ggsegYeo2011 & \citet{yeo2011}\\
+Yeo 7 Resting-state Cortical Parcellations & yeo7 & ggsegYeo2011 & ggsegYeo2011 & \citet{yeo2011}\\
 \bottomrule
@@
 The tutorial will focus on the two main functions: \texttt{ggseg()} for plotting 2D polygons and \texttt{ggseg3d()} for plotting 3D brains based on tri-surface mesh plots.
+The following tutorial complements the package documentation and vignettes that can be found online for all three packages: \href{https://lcbc-uio.github.io/ggseg}{ggseg}, \href{https://lcbc-uio.github.io/ggseg3d}{ggseg3d}, \href{https://lcbc-uio.github.io/ggsegExtra}{ggsegExtra}.
 
@@
 
-\emph{ggseg} is the main function for plotting 2D data.
-By default, the function automatically plots the DKT atlas (see Figure \ref{fig:init}).
+\texttt{ggseg()} from the \emph{ggseg} package is the main function for plotting 2D data.
+By default, the function automatically plots the DK atlas (see Figure \ref{fig:figure1}).
 The \texttt{ggseg()}-function is a wrapper for \texttt{geom\_polygon()} from \emph{ggplot2}, and it can be built upon and combined like any \emph{ggplot}-object.
@@
 \begin{Highlighting}[]
-\CommentTok{\# remotes::install\_github("LCBC{-}UiO/ggseg")}
+\CommentTok{# remotes::install_github("LCBC-UiO/ggseg")}
 \KeywordTok{library}\NormalTok{(ggseg)}
-\KeywordTok{library}\NormalTok{(dplyr)}
-\KeywordTok{library}\NormalTok{(tidyr)}
 
-\CommentTok{\# Figure 1}
+\CommentTok{# package combining plots easily}
+\KeywordTok{library}\NormalTok{(patchwork)}
+
+\CommentTok{# Figure 1}
 \KeywordTok{ggseg}\NormalTok{()}
@@
 \centering
-\includegraphics{msc_ggseg_files/figure-latex/init-1.pdf}
-\caption{\label{fig:init}By default \texttt{ggseg()} will plot the dkt atlas in grey shaded polygons.}
+\includegraphics{msc_ggseg_files/figure-latex/figure1-1.pdf}
+\caption{\label{fig:figure1}By default \texttt{ggseg()} will plot the dk atlas in grey shaded polygons.}
 \end{figure}
@@
 These options are atlas-specific.
-For cortical atlases, such as the \texttt{dkt}, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \ref{fig:collection} for examples).
+For cortical atlases, such as the \texttt{dk}, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \ref{fig:figure2} for examples).
 For subcortical atlases, such as the \texttt{aseg} atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.
@@
 \begin{Highlighting}[]
-\CommentTok{\# dkt dark theme}
-\NormalTok{p1 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme\_dark}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
+\CommentTok{# A: dk dark theme}
+\NormalTok{fig2A <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_dark}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dk"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"dark theme"}\NormalTok{)}
 
-\CommentTok{\# dkt classic theme}
-\NormalTok{p2 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme\_classic}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{# B: dk medial view}
+\NormalTok{fig2B <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dk"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"medial view"}\NormalTok{)}
 
-\CommentTok{\# dkt medial view}
-\NormalTok{med <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{# C: dk left medial alone}
+\NormalTok{fig2C <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
+            \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"dk"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"medial left"}\NormalTok{)}
 
-\CommentTok{\# dkt left hemisphere}
-\NormalTok{left <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{# D: dk classic theme}
+\NormalTok{fig2D <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dk"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"classic theme"}\NormalTok{)}
 
-\CommentTok{\# aseg default theme}
-\NormalTok{p3 <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\NormalTok{aseg) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{" "}\NormalTok{)}
+\CommentTok{# E: dk left hemisphere}
+\NormalTok{fig2E <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"dk"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"left hemisphere"}\NormalTok{)}
 
-\CommentTok{\# dkt left medial alone}
-\NormalTok{combo <{-}}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{view =} \StringTok{"medial"}\NormalTok{,}
-               \DataTypeTok{hemisphere =} \StringTok{"left"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{" "}\NormalTok{)}
+\CommentTok{# F: aseg default theme}
+\NormalTok{fig2F <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg, }\DataTypeTok{view =} \StringTok{"axial"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"aseg"}\NormalTok{, }\DataTypeTok{subtitle =} \StringTok{"axial view"}\NormalTok{)}
 
-\CommentTok{\# Combine plots to Figure 2}
-\NormalTok{cowplot}\OperatorTok{::}\KeywordTok{plot\_grid}\NormalTok{(p1, med, combo,  p2, left, p3,}
-                   \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A: dkt {-} dark"}\NormalTok{, }\StringTok{"B: dkt {-} medial"}\NormalTok{,}
-                              \StringTok{"C: dkt {-} combo"}\NormalTok{, }\StringTok{"D: dkt {-} classic"}\NormalTok{,}
-                              \StringTok{"E: dkt {-} left"}\NormalTok{, }\StringTok{"F: aseg"}\NormalTok{),}
-                   \DataTypeTok{hjust =} \FloatTok{{-}.05}\NormalTok{)}
+\CommentTok{# Using patchwork to combine all plots into one}
+\NormalTok{(fig2A }\OperatorTok{|}\StringTok{ }\NormalTok{fig2B }\OperatorTok{|}\StringTok{ }\NormalTok{fig2C) }\OperatorTok{/}\StringTok{ }\NormalTok{(fig2D }\OperatorTok{|}\StringTok{ }\NormalTok{fig2E }\OperatorTok{|}\StringTok{ }\NormalTok{fig2F) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels =} \StringTok{'A'}\NormalTok{)}
 \end{Highlighting}
@@
 
-\begin{figure}
-\centering
-\includegraphics{msc_ggseg_files/figure-latex/collection-1.pdf}
-\caption{\label{fig:collection}\emph{ggseg} plots can be used with any \emph{ggplot2} feature such as standard scales and themes.
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{msc_ggseg_files/figure-latex/figure2-1} \caption{\emph{ggseg} plots can be used with any \emph{ggplot2} feature such as standard scales and themes.
 For cortical atlases, one can supply special \emph{ggseg} options to determine, for example, hemisphere, view, or position.
-\textbf{A:} \texttt{dkt} atlas, stacked with dark theme ; \textbf{B:} \texttt{dkt} with medial view only; \textbf{C:} \texttt{dkt} atlas with only left medial display; \textbf{D:} \texttt{dkt} atlas, stacked, with classic theme; \textbf{E:} \texttt{dkt} atlas with left hemisphere only; \textbf{F:} complete \texttt{aseg} atlas}
+\textbf{A:} \texttt{dk} atlas, stacked with dark theme ; \textbf{B:} \texttt{dk} with medial view only; \textbf{C:} \texttt{dk} atlas with only left medial display; \textbf{D:} \texttt{dk} atlas, stacked, with classic theme; \textbf{E:} \texttt{dk} atlas with left hemisphere only; \textbf{F:} axial view of \texttt{aseg} atlas}\label{fig:figure2}
 \end{figure}
@@
 Standard arguments like \texttt{fill} that floods the segments with a colour, or \texttt{colour} that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the \emph{ggplot2} mapping function \texttt{aes}.
-To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales (Figure \ref{fig:fill}).
+To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific `brain' palette scales (Figure \ref{fig:figure3}).
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Figure 3}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ area), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale\_fill\_brain}\NormalTok{(}\StringTok{"dkt"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
-        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
-        \DataTypeTok{legend.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
+\CommentTok{# Figure 3}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping =} \KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ region), }\DataTypeTok{colour=}\StringTok{"black"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"dk"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
+        \DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{,}
+        \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
 \centering
-\includegraphics{msc_ggseg_files/figure-latex/fill-1.pdf}
-\caption{\label{fig:fill}Supplying `area' to the fill option in \texttt{ggseg()}, will use the column `area' from the accompanying dataset to create a discrete colour palette over the segments in the atlas.
-The \texttt{dkt} atlas default palette corresponds to the FreeSurferColorLut scheme.}
+\includegraphics{msc_ggseg_files/figure-latex/figure3-1.pdf}
+\caption{\label{fig:figure3}Supplying `region' to the fill option in \texttt{ggseg()}, will use the column `region' from the accompanying data-set to create a discrete colour palette over the segments in the atlas.
+The \texttt{dk} atlas default palette corresponds to the FreeSurferColorLut scheme.}
 \end{figure}
@@
 Most users will use \texttt{ggseg()} to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
-Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data sets.
-The atlas data set structure will help users understand what incoming statistical data needs to look like.
-Note that each atlas corresponds to a unique data set.
-All data sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
+Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data-sets.
+The atlas data-set structure will help users understand what incoming statistical data needs to look like.
+Note that each atlas corresponds to a unique data-set.
+All data-sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Look at the top 5 rows of the dkt dataset}
-\KeywordTok{head}\NormalTok{(dkt, }\DecValTok{5}\NormalTok{)}
+\CommentTok{# Look at the top 5 rows of the dk data-set}
+\KeywordTok{head}\NormalTok{(dk, }\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 The column `label' contains the label (region) names as in the original neuroimaging software.
-For example, the DKT atlas label column matches the region names from Freesurfer statistics table outputs.
+For example, the \texttt{dk} atlas label column matches the region names from Freesurfer statistics table outputs.
 Yet, the data in \emph{ggseg} is in a long format - that is each region has a single row - and any data of interest needs to be in this same format.
-Often data sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
-See below an example of wide-to-long conversion.
+Often data-sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
+Summary statistic files from Freesurfer have a file format that can be difficult to work with in R.
+In a set of functions in \emph{ggseg}, we facilitate conversion between FreeSurfer files and ggseg compatible data.
+There is an \href{https://lcbc-uio.github.io/ggseg/articles/freesurfer_files.html}{extended vignette} of the various options.
+In the following example, we import summary data that has been compiled using Freesurfer's \texttt{aparcstats2table} (for surfaces) or \texttt{asegstats2table} (volumes).
+The summary files can combine parcellation metrics for one or more subjects into a single summary file.
+This file also contains extra metrics like intracranial volume and total grey matter volume.
+These later measures do not map onto ggseg, and as such, one should filter them out of the data returned from \texttt{read\_freesurfer\_table)}.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Create a mock freesurfer output}
-\NormalTok{freesurfer\_stats <{-}}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{id =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\OperatorTok{:}\DecValTok{12}\NormalTok{),}
-  \DataTypeTok{lh\_superiortemporal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.32}\NormalTok{, }\FloatTok{4.1}\NormalTok{, }\FloatTok{3.5}\NormalTok{),}
-  \DataTypeTok{lh\_precentral =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.3}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.1}\NormalTok{),}
-  \DataTypeTok{lh\_rostralmiddlefrontal =} \KeywordTok{c}\NormalTok{(}\FloatTok{3.3}\NormalTok{, }\FloatTok{3.2}\NormalTok{, }\FloatTok{3.1}\NormalTok{)}
-\NormalTok{)}
-\NormalTok{freesurfer\_stats}
+\NormalTok{file_path <-}\StringTok{ "data/aparc.volume.table"}
 
-\CommentTok{\# Wrangle wide format data into long format}
-\NormalTok{freesurfer\_long <{-}}\StringTok{ }\NormalTok{freesurfer\_stats }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(label, thickness, }\OperatorTok{{-}}\NormalTok{id)}
-\NormalTok{freesurfer\_long}
-\CommentTok{\#\#   id lh\_superiortemporal lh\_precentral lh\_rostralmiddlefrontal}
-\CommentTok{\#\# 1 10                3.32           2.3                     3.3}
-\CommentTok{\#\# 2 11                4.10           2.5                     3.2}
-\CommentTok{\#\# 3 12                3.50           2.1                     3.1}
-\CommentTok{\#\#   id                   label thickness}
-\CommentTok{\#\# 1 10     lh\_superiortemporal      3.32}
-\CommentTok{\#\# 2 11     lh\_superiortemporal      4.10}
-\CommentTok{\#\# 3 12     lh\_superiortemporal      3.50}
-\CommentTok{\#\# 4 10           lh\_precentral      2.30}
-\CommentTok{\#\# 5 11           lh\_precentral      2.50}
-\CommentTok{\#\# 6 12           lh\_precentral      2.10}
-\CommentTok{\#\# 7 10 lh\_rostralmiddlefrontal      3.30}
-\CommentTok{\#\# 8 11 lh\_rostralmiddlefrontal      3.20}
-\CommentTok{\#\# 9 12 lh\_rostralmiddlefrontal      3.10}
+\NormalTok{someData <-}\StringTok{ }\KeywordTok{read_freesurfer_table}\NormalTok{(file_path, }\DataTypeTok{measure =} \StringTok{"volume"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# remove rows in the data that do not contain the strings "rh" or "lh"}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(}\KeywordTok{grepl}\NormalTok{(}\StringTok{"rh|lh"}\NormalTok{, label))}
+\NormalTok{someData}
+\CommentTok{## # A tibble: 34 x 3}
+\CommentTok{##    subject label                      volume}
+\CommentTok{##    <fct>   <chr>                       <dbl>}
+\CommentTok{##  1 bert    rh_bankssts                  1969}
+\CommentTok{##  2 bert    rh_caudalanteriorcingulate   2280}
+\CommentTok{##  3 bert    rh_caudalmiddlefrontal       5390}
+\CommentTok{##  4 bert    rh_cuneus                    2998}
+\CommentTok{##  5 bert    rh_entorhinal                1735}
+\CommentTok{##  6 bert    rh_fusiform                  8144}
+\CommentTok{##  7 bert    rh_inferiorparietal         14876}
+\CommentTok{##  8 bert    rh_inferiortemporal         11016}
+\CommentTok{##  9 bert    rh_isthmuscingulate          1983}
+\CommentTok{## 10 bert    rh_lateraloccipital         12729}
+\CommentTok{## # ... with 24 more rows}
 \end{Highlighting}
@@
 
-Data in long format can then be used directly with the \texttt{ggseg()}-function, as the `label' column corresponds in name and content with the `label' column in the atlas data of \texttt{dkt}.
-The data \textbf{must} include a column that has the same column name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
-In the next example we create some data with 4 rows, with an `area' and `p' column, representing the results of a hypothetical analysis.
-The \texttt{ggseg()}-function will recognise the matching column `area', and merge the supplied data into the atlas using \emph{dplyr}-joins.
-We use the `p' column as the column flooding the segment with colour.
-The appearance of the plot can then be modified similarly to any other \emph{ggplot2} graph using functions such as scales, labs, themes, etc., as seen in Figure \ref{fig:datasupp}
+A second option is to read in atlas files directly from the entire subject directory, using regular expression to select the target atlas.
+The default aparc has a unique string ending with \texttt{aparc.stats} which we will use to read in all subject stats for that particular atlas.
+One can filter specific subjects and/or regions from the R-object data.frame.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Make some mock data}
-\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{area =} \KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
-           \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}
-  \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{4}\NormalTok{),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{atlas_data <-}\StringTok{ }\KeywordTok{read_atlas_files}\NormalTok{(}
+  \CommentTok{# Path to Freesurfers $SUBJECTS_DIR}
+  \DataTypeTok{subjects_dir =} \StringTok{"/Applications/freesurfer/subjects/"}\NormalTok{,}
+  \CommentTok{# Regular expression to find atlas files}
+  \DataTypeTok{atlas =} \StringTok{"aparc.stats$"}
+\NormalTok{)}
+\NormalTok{atlas_data}
+\CommentTok{## # A tibble: 204 x 11}
+\CommentTok{##    subject label NumVert SurfArea GrayVol ThickAvg}
+\CommentTok{##    <chr>   <chr>   <int>    <int>   <int>    <dbl>}
+\CommentTok{##  1 bert    lh_b~    1181      831    2297     2.77}
+\CommentTok{##  2 bert    lh_c~     843      572    1534     2.72}
+\CommentTok{##  3 bert    lh_c~    2758     1840    5772     2.80}
+\CommentTok{##  4 bert    lh_c~    2683     1654    3074     1.81}
+\CommentTok{##  5 bert    lh_e~     581      416    1840     3.33}
+\CommentTok{##  6 bert    lh_f~    4113     2875    8519     2.72}
+\CommentTok{##  7 bert    lh_i~    4948     3466   10559     2.70}
+\CommentTok{##  8 bert    lh_i~    5056     3542   12358     2.98}
+\CommentTok{##  9 bert    lh_i~    1561      990    2350     2.09}
+\CommentTok{## 10 bert    lh_l~    7961     5077   12743     2.30}
+\CommentTok{## # ... with 194 more rows, and 5 more variables:}
+\CommentTok{## #   ThickStd <dbl>, MeanCurv <dbl>,}
+\CommentTok{## #   GausCurv <dbl>, FoldInd <int>, CurvInd <dbl>}
+\end{Highlighting}
+\end{Shaded}
 
-\CommentTok{\# Figure 4}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p))  }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"A nice plot title"}\NormalTok{, }\DataTypeTok{fill=}\StringTok{"p{-}value"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale\_fill\_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}
+This long format data can be used directly with the \texttt{ggseg()}-function, as the `label' column corresponds in name and content with the `label' column in the atlas data of \texttt{dk}.
+The data \textbf{must} include a column that has the same column name and at least \emph{some} data matching the values in the corresponding column in the atlas data.
+In the following example, the \texttt{ggseg()}-function will recognise the matching column `label', and merge the supplied data into the atlas using \emph{dplyr}-joins.
+The appearance of the plot can then be modified similarly to any other \emph{ggplot2} graph using functions such as scales, labs, themes, etc., as seen in Figure \ref{fig:figure4}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+
+\CommentTok{# Figure 4A}
+\NormalTok{fig4A <-}\StringTok{ }\NormalTok{someData }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# plot only subject bert}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(subject }\OperatorTok{==}\StringTok{ "bert"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{volume))  }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"from read_freesurfer_table"}\NormalTok{, }
+       \DataTypeTok{fill=}\StringTok{"volume}\CharTok{\textbackslash{}n}\StringTok{(mm^3)"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"firebrick"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_colourbar}\NormalTok{(}\DataTypeTok{barheight =} \DecValTok{3}\NormalTok{))}
+
+\CommentTok{# Figure 4B}
+\NormalTok{fig4B <-}\StringTok{ }\NormalTok{atlas_data }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# plot only subject bert}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(subject }\OperatorTok{==}\StringTok{ "bert"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{ThickAvg))  }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"from read_atlas_files"}\NormalTok{, }
+       \DataTypeTok{fill=}\StringTok{"Thickness}\CharTok{\textbackslash{}n}\StringTok{(mean)"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low=}\StringTok{"forestgreen"}\NormalTok{,}\DataTypeTok{high=}\StringTok{"goldenrod"}\NormalTok{)}\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_colourbar}\NormalTok{(}\DataTypeTok{barheight =} \DecValTok{3}\NormalTok{))}
+
+\NormalTok{(fig4A }\OperatorTok{/}\StringTok{ }\NormalTok{fig4B) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels =} \StringTok{"A"}\NormalTok{)}
 \end{Highlighting}
@@
 \centering
-\includegraphics{msc_ggseg_files/figure-latex/datasupp-1.pdf}
-\caption{\label{fig:datasupp}Supplying data through the `.data' option in \texttt{ggseg()} enables use of columns in the supplied data to aesthetical arguments (such as `fill').
+\includegraphics{msc_ggseg_files/figure-latex/figure4-1.pdf}
+\caption{\label{fig:figure4}Supplying data through the `.data' option in \texttt{ggseg()} enables use of columns in the supplied data to aesthetical arguments (such as `fill').
 The \emph{ggseg} plot can be used with any other polygon compatible function from \emph{ggplot2} or \emph{ggplot2} extentions, for instance adding title, changing the legend name and the colour scheme with standard \emph{ggplot2} functions.}
@@
 If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column `hemi' with either `right' or `left'.
-In this case, data will be merged into the atlas both by `area' and by `hemi'.
-For more information about adapting data and viewing only one hemisphere or side, the \href{https://lcbc-uio.github.io/ggseg/articles/ggseg.html\#single-hemisphere-results}{package vignettes} contains more elaborate information.
+In this case, data will be merged into the atlas both by `region' and by `hemi'.
+For more information about adapting data and viewing only one hemisphere or side, the \href{https://lcbc-uio.github.io/ggseg/articles/ggseg.html\#single-hemisphere-results}{online documentation} contains more elaborate information.
 
@@
 This may be obtained also with \texttt{ggseg()}, using \emph{ggplot2}'s \texttt{facet\_wrap} or \texttt{facet\_grid}, with three guiding rules:
-\textbf{1)} as before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in seperate rows, not in separate columns).
-\textbf{2)} The data needs to be grouped using \emph{dplyr}'s \texttt{group\_by()} function \emph{before} providing the data to the \texttt{ggseg()}-function.
-The \texttt{ggseg()}-function will detect grouped data, and adapt it to \texttt{facet}`s requirements.
-\textbf{3)} Apply \texttt{facet\_wrap} or \texttt{facet\_grid} to the plot having used the above two rules.
-An example of this can be seen in Figure \ref{fig:datasupp3}, where a mock data set including summary statistics for two groups ('Young' and `Old') is used when faceting a \emph{ggseg}-plot.
 
-All the concepts described above also work with the \texttt{aseg} atlas for subcortical structures, except for `hemisphere' and `view' arguments that are superfluous in subcortical atlases (Figure \ref{fig:atlases}).
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  As before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in separate rows, not in separate columns).\\
+\item
+  The data needs to be grouped using \emph{dplyr}'s \texttt{group\_by()} function \emph{before} providing the data to the \texttt{ggseg()}-function.
+  The \texttt{ggseg()}-function will detect grouped data, and adapt it to \texttt{facet}'s requirements.\\
+\item
+  Apply \texttt{facet\_wrap} or \texttt{facet\_grid} to the plot having used the above two rules.
+  An example of this can be seen in Figure \ref{fig:figure5} B, where a mock data-set including summary statistics for two groups (`Young' and `Old') is used when faceting a \emph{ggseg}-plot.
+\end{enumerate}
+
+The second rule is needed as the \texttt{ggseg()}-function is a ggplot-wrapper rather than an extension, and thus has a particular structure and pipeline to combine the atlas to supplied data.
+The result of ungrouped data into \texttt{ggseg} is shown in \ref{fig:figure5} A.
+Note that the data is combined incorrectly, there are three panels rather than two.
+Regions not represented in the data (\texttt{NA}-values) have their own panel, while the remaining two panels only include regions with valid data.
+All the concepts described above also work with the \texttt{aseg} atlas for subcortical structures, except for `hemisphere' and `view' arguments that are superfluous in subcortical atlases (Figure \ref{fig:figure6}).
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Make some mock data}
-\NormalTok{someData =}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{area =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
-               \StringTok{"pre central"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}\DecValTok{2}\NormalTok{),}
+\CommentTok{# Make some simpler mock data}
+\NormalTok{someData =}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{region =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"transverse temporal"}\NormalTok{, }\StringTok{"insula"}\NormalTok{,}
+               \StringTok{"precentral"}\NormalTok{,}\StringTok{"superior parietal"}\NormalTok{),}\DecValTok{2}\NormalTok{),}
   \DataTypeTok{p =} \KeywordTok{sample}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{,.}\DecValTok{001}\NormalTok{), }\DecValTok{8}\NormalTok{),}
-  \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{)),}
-  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{group\_by}\NormalTok{(AgeG)}
+  \DataTypeTok{AgeG =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"Young"}\NormalTok{,}\DecValTok{4}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"Old"}\NormalTok{,}\DecValTok{4}\NormalTok{))}
+\NormalTok{)}
 
-\CommentTok{\# Figure 5}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{.data=}\NormalTok{someData, }\DataTypeTok{colour=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
-      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet\_wrap}\NormalTok{(}\OperatorTok{\textasciitilde{}}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
+\CommentTok{# Figure 5a}
+\NormalTok{fig5A <-}\StringTok{ }\NormalTok{someData }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
+      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Ungrouped data"}\NormalTok{)}
+
+\CommentTok{# Figure 5b}
+\NormalTok{fig5B <-}\StringTok{ }\NormalTok{someData }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(AgeG) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{position =} \StringTok{"stacked"}\NormalTok{,}
+      \DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{p), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{AgeG, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Grouped data"}\NormalTok{)}
+
+\CommentTok{# Combining plots with patchwork, applying theme to all}
+\NormalTok{(fig5A }\OperatorTok{/}\StringTok{ }\NormalTok{fig5B) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{plot_layout}\NormalTok{(}\DataTypeTok{widths =} \DecValTok{1}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels =} \StringTok{'A'}\NormalTok{) }\OperatorTok{&}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{8}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{msc_ggseg_files/figure-latex/figure5-1} \caption{Examples of incorrect and correct faceting with \texttt{ggseg}.
+\textbf{A}: ungrouped data will facet incorrectly due to the data not being merged correctly with the atlas.
+\textbf{B}: grouped data will merge in the expected way and result in correctly facetted plots.}\label{fig:figure5}
+\end{figure}
+
+The \texttt{ggseg} package also has extensive \href{https://lcbc-uio.github.io/ggseg/}{online documentation}, with more functionality than we have covered here.
+The online documentation is also regularly updated with explanations of new features or code improvements.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Figure 6}
+\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{region)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
+        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
+        \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
 \centering
-\includegraphics{msc_ggseg_files/figure-latex/datasupp3-1.pdf}
-\caption{\label{fig:datasupp3}To visualize data corresponding to different groups, one needs to use \texttt{group\_by} or similar \textbf{before} plotting with \texttt{ggseg()} for facetting to work.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 6}
-\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas=}\StringTok{"aseg"}\NormalTok{, }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{area)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.justification=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{),}
-        \DataTypeTok{legend.position=}\StringTok{"bottom"}\NormalTok{,}
-        \DataTypeTok{legend.text =} \KeywordTok{element\_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{5}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill =} \KeywordTok{guide\_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\centering
-\includegraphics{msc_ggseg_files/figure-latex/atlases-1.pdf}
-\caption{\label{fig:atlases}The \texttt{aseg} atlas showing subcortical structures has some distinct differences from the \texttt{dkt}.
+\includegraphics{msc_ggseg_files/figure-latex/figure6-1.pdf}
+\caption{\label{fig:figure6}The \texttt{aseg} atlas showing subcortical structures has some distinct differences from the \texttt{dk}.
 For instance, there is no option to show only a single hemisphere.
@@
 
-Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dkt\_3d} atlas in `LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \ref{fig:ggseg3d-1-out})
-The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage5} template subject.
-All \texttt{{[}...{]}\_3d} atlases include a \texttt{colour} column that based on the color scheme used in the source neuroimaging software.
+Out-of-the-box, \texttt{ggseg3d()} plots the \texttt{dk\_3d} atlas in `LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \ref{fig:figure7})
+The `LCBC' surface consists on a semi-inflated white matter surface based on the \emph{fsaverage5} template subject, a surface template that is issued together with the \texttt{ggsegExtra} package.
+All \texttt{{[}...{]}\_3d} atlases include a \texttt{colour} column that is based on the color scheme used in the source neuroimaging software.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# remotes::install_github("LCBC-UiO/ggseg3d")}
+\KeywordTok{library}\NormalTok{(ggseg3d)}
+
+\CommentTok{# Figure 7 left}
+\NormalTok{fig7l <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{surface =} \StringTok{"white"}\NormalTok{)  }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface='white')"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+\NormalTok{fig7l}
+
+\CommentTok{# Figure 7 middle}
+\NormalTok{fig7m <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{()  }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface='LCBC')"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+\NormalTok{fig7m}
+
+\CommentTok{# Figure 7 right}
+\NormalTok{fig7r <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{surface =} \StringTok{"inflated"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{layout}\NormalTok{(}\DataTypeTok{title=}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{ggseg3d(surface='inflated')"}\NormalTok{,}
+         \DataTypeTok{font =}  \KeywordTok{list}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{))}
+\NormalTok{fig7r}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{figure}[H]
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_white} \includegraphics[width=0.3\linewidth]{png/ggseg3d} \includegraphics[width=0.3\linewidth]{png/ggseg3d_inflated} \caption{The three surface options provided in \emph{ggseg3d} atlases.
-\textbf{From left to right:} the `white' surface is the white matter surface, `LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the `inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.}\label{fig:ggseg3d-1-out}
+\includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure7l} \includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure7m} \includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure7r} \caption{The three surface options provided in \emph{ggseg3d} atlases.
+\textbf{From left to right:} the `white' surface is the white matter surface, `LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the `inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software.
+The title for each plot is the literal code use to make the subplot.}\label{fig:figure7}
 \end{figure}
@@
 The 3D-atlas data is stored in nested tibbles.
-Each cortical atlas has data sets for three different surfaces (see Figure \ref{fig:ggseg3d-1-out}) and the two hemispheres.
+Each cortical atlas has data-sets for three different surfaces (see Figure \ref{fig:figure7}) and the two hemispheres.
 Only one surface is available for subcortical atlases as inflation procedures are irrelevant.
@@
 It is important to note that the coordinates in the plot (X, Y, Z) are \textbf{not} any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
-
-\small
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# remotes::install\_github("LCBC{-}UiO/ggseg3d")}
-\KeywordTok{library}\NormalTok{(ggseg3d)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
+While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in \texttt{ggseg3d}, at the moment.
+As such, information regarding referencing the regions' location in standard radiological spaces (like MNI or Taliarch) should be acquired directly from imaging software or papers describing the atlases, not \texttt{ggseg3d}.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Select surface and hemisphere, and then unnest to inspect the atlas data}
-\NormalTok{dkt\_3d }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{\&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{\%>\%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg\_3d) }\OperatorTok{\%>\%}
+\CommentTok{# Select surface and hemisphere, and then unnest to inspect the atlas data}
+\NormalTok{dk_3d }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(surf }\OperatorTok{==}\StringTok{ "inflated"} \OperatorTok{&}\StringTok{ }\NormalTok{hemi }\OperatorTok{==}\StringTok{ "right"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(ggseg_3d) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\CommentTok{## # A tibble: 5 x 9}
+\CommentTok{##   atlas surf  hemi  region colour mesh  label}
+\CommentTok{##   <chr> <chr> <chr> <chr>  <chr>  <lis> <chr>}
+\CommentTok{## 1 dk_3d infl~ right <NA>   <NA>   <nam~ rh_u~}
+\CommentTok{## 2 dk_3d infl~ right banks~ #1964~ <nam~ rh_b~}
+\CommentTok{## 3 dk_3d infl~ right cauda~ #7D64~ <nam~ rh_c~}
+\CommentTok{## 4 dk_3d infl~ right cauda~ #6419~ <nam~ rh_c~}
+\CommentTok{## 5 dk_3d infl~ right corpu~ #7846~ <nam~ rh_c~}
+\CommentTok{## # ... with 2 more variables: roi <chr>,}
+\CommentTok{## #   annot <chr>}
 \end{Highlighting}
@@
 Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
-It should also be noted that the `label', `annot' and `area' columns could provide matching values for your own data.
-Similarly to the \texttt{ggseg()}-function, the `label' column should match the region names used in the original neuroimaging software while `area' and `annot' provide alternative/secondary names.
+It should also be noted that the `label', `annot' and `region' columns could provide matching values for your own data.
+Similarly to the \texttt{ggseg()}-function, the `label' column should match the region names used in the original neuroimaging software while `region' and `annot' provide alternative/secondary names.
 It is thus important to match your regional identifiers with those used in the atlas.
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 8 left}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }\DataTypeTok{atlas =}\NormalTok{ dkt\_3d, }\DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{)}
-
-\CommentTok{\# Figure 8 middle}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
-        \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
-        \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-        \DataTypeTok{palette =} \KeywordTok{c}\NormalTok{(}\StringTok{"\#ff0000"}\NormalTok{, }\StringTok{"\#00ff00"}\NormalTok{, }\StringTok{"\#0000ff"}\NormalTok{)) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
-
-\CommentTok{\# Figure 8 right}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{.data =}\NormalTok{ someData, }
-        \DataTypeTok{atlas =}\NormalTok{ dkt\_3d,}
-        \DataTypeTok{colour =} \StringTok{"p"}\NormalTok{, }\DataTypeTok{text =} \StringTok{"p"}\NormalTok{,}
-        \DataTypeTok{na.colour =} \StringTok{"black"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}[H]
-\includegraphics[width=0.3\linewidth]{png/ggseg3d_data} \includegraphics[width=0.3\linewidth]{png/ggseg3d_colour} \includegraphics[width=0.3\linewidth]{png/ggseg3d_nacolour} \caption{\textbf{Left:} Supplying data to \texttt{ggseg3d()} works similarly to \texttt{ggseg()}.
+\includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure8l} \includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure8m} \includegraphics[width=0.3\linewidth]{msc_ggseg_files/figure-latex/figure8r} \caption{\textbf{Left:} Supplying data to \texttt{ggseg3d()} works similarly to \texttt{ggseg()}.
 Since \texttt{ggseg3d()} is based on \emph{plotly} rather than \emph{ggplot2}, most aesthetic adaptations must be set in the main function.
@@
 The `palette' option takes vectors of colours either as HEX-codes or R-colour names.
-\textbf{Right:} The colour of the \texttt{NA} values can also be changed through the option `na.colour'.}\label{fig:ggseg3d-data1}
+\textbf{Right:} The colour of the \texttt{NA} values can also be changed through the option `na.colour'.}\label{fig:figure8}
 \end{figure}
@@
 
-You can provide custom colour palettes either in hex or R-names, as seen in Figure \ref{fig:ggseg3d-data1}.
+You can provide custom colour palettes either in hex or R-names, as seen in Figure \ref{fig:figure8}.
 Colours will be evenly spaced when creating the colour-scale.
-A palette may also be supplied as a \emph{named numeric vector}, where the vector names are the colours that users wish to use, and the numeric values are the breakpoints for each colour (e.g.~\texttt{c("red"\ =\ 0,\ "white"\ =\ 0.5,\ "blue"\ =\ 1)}).
+A palette may also be supplied as a \emph{named numeric vector}, where the vector names are the colours that users wish to use, and the numeric values are the breakpoints for each colour (e.g.
+\texttt{c("red"\ =\ 0,\ "white"\ =\ 0.5,\ "blue"\ =\ 1)}).
 This way the users can control the minimum and maximum values of the colour scale, and also how the gradient is applied.
@@
 
-Subcortical atlases include cortical surfaces and other landmark structures for visualization purposes only.
-One can control the opacity of the these \texttt{NA} structures, to improve visualization.
-Glass brains can be added to provide a frame of reference for the subcortical structures with the function \texttt{add\_glassbrain()} (Figure \ref{fig:glassbrain1}) , which takes three extra arguments: hemisphere, colour, and opacity.
+A glass brain is a translucent representation of the brain that can provide a frame of reference,
+particularly when looking at subcortical parcellations and it only serves as a visual aid.
+One can control the opacity of the these \texttt{NA} structures, to improve visualization, directly in the \texttt{ggseg3d} call.
+A cortical glass brain can be added to \texttt{ggseg3d} plots through the function \texttt{add\_glassbrain()} (Figure \ref{fig:figure9}), which takes three extra arguments: hemisphere, colour, and opacity.
+We recommend only using the glass brain function when plotting subcortical atlases.
+A similar, interactive version of this plot can be found in the \href{https://lcbc-uio.github.io/ggseg3d//articles/ggseg3d.html\#colours-1}{online documentation}.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Figure 9}
-\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg\_3d,}
-        \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{   }
-\StringTok{  }\KeywordTok{add\_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
-\StringTok{  }\KeywordTok{pan\_camera}\NormalTok{(}\StringTok{"left lateral"}\NormalTok{) }\OperatorTok{\%>\%}\StringTok{ }
-\StringTok{  }\KeywordTok{remove\_axes}\NormalTok{()}
+\CommentTok{# Figure 9}
+\NormalTok{fig9 <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ aseg_3d,}
+        \DataTypeTok{na.alpha=} \FloatTok{.5}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"left lateral"}\NormalTok{)}
+\NormalTok{fig9}
 \end{Highlighting}
@@
 \begin{figure}[H]
-\includegraphics[width=0.6\linewidth]{png/ggseg3d_glass} \caption{For subcortical structure visualization, one can add a glass brain to the plot.
+\includegraphics[width=0.6\linewidth]{msc_ggseg_files/figure-latex/figure9} \caption{For subcortical structure visualization, one can add a glass brain to the plot.
 This will help with locating the structures relative to the cortex, and make the plot easier to interpret.
-The glass brain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:glassbrain1}
+The glass brain is controlled by three options: opacity, hemisphere, and colour.}\label{fig:figure9}
 \end{figure}
@@
 
+The package \href{https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html}{online documentation} also includes interactive versions of many of the figures in this tutorial, and several more options not covered here.
+The online documentation will also be updated as new features are required or we improve the current state of user functionality and options.
+
 \hypertarget{additional-atlases}{%
@@
 
-The \emph{ggseg} and \emph{ggseg3d} packages have two atlases each, which are 2D and 3D variations of the same main atlases: the \texttt{dkt} (\citet{dkt}) and \texttt{aseg} (\citet{aseg}) atlases.
+The \emph{ggseg} and \emph{ggseg3d} packages have two atlases each, which are 2D and 3D variations of the same main atlases: the \texttt{dk} (\citet{dk}) and \texttt{aseg} (\citet{aseg}) atlases.
 These are, however, only two among many meaningful ways of segmenting the brain into different regions.
-Thus, the \emph{ggsegExtra} package is a repository containing additional with additional data sets for plotting with the \emph{ggseg} and \emph{ggseg3d} packages.
+Thus, the \emph{ggsegExtra} package is a library with two main functionalities: 1) to install additional data-sets for plotting with the \emph{ggseg} and \emph{ggseg3d} packages from other repositories, and 2) to create custom atlases for \emph{ggseg} and \emph{ggseg3d}.
+
+\hypertarget{verified-existing-atlases}{%
+\subsubsection{Verified existing atlases}\label{verified-existing-atlases}}
+
 There is an ever-increasing amount of new atlases being created, as research and methods in neuroimaging analysis progresses.
 The \emph{ggsegExtra}-package is intended to be expanded as a community-effort, as new and informative atlases are published.
-A small collection of the atlases currently in the \emph{ggsegExtra} package may be viewed in Figure \ref{fig:ggsegExtra}
+A collection of the atlases currently listed in the \emph{ggsegExtra} package may be found in Figure \ref{fig:figure10} and explored in our \href{https://athanasiamo.shinyapps.io/ggsegDemo/}{shinyapps.io demonstration}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# remotes::install_github("LCBC-UiO/ggsegExtra")}
+\KeywordTok{library}\NormalTok{(ggsegExtra)}
+
+\CommentTok{# list all known ggseg atlas-repositories}
+\KeywordTok{ggseg_atlas_repos}\NormalTok{()}
+
+\CommentTok{# list repositories matching a pattern search}
+\NormalTok{yeo_repo <-}\StringTok{ }\KeywordTok{ggseg_atlas_repos}\NormalTok{(}\StringTok{"Yeo"}\NormalTok{)}
+\CommentTok{## # A tibble: 10 x 5}
+\CommentTok{##    repo       ggseg ggseg3d source comment        }
+\CommentTok{##    <chr>      <lgl> <lgl>   <chr>  <chr>          }
+\CommentTok{##  1 LCBC-UiO/~ TRUE  TRUE    github both 17 and 7 ~}
+\CommentTok{##  2 LCBC-UiO/~ FALSE TRUE    github the 2009 atlas }
+\CommentTok{##  3 LCBC-UiO/~ FALSE TRUE    github both thickness~}
+\CommentTok{##  4 LCBC-UiO/~ FALSE TRUE    github both 17 and 7 ~}
+\CommentTok{##  5 LCBC-UiO/~ TRUE  TRUE    github full atlas     }
+\CommentTok{##  6 LCBC-UiO/~ TRUE  TRUE    github white tract at~}
+\CommentTok{##  7 LCBC-UiO/~ TRUE  TRUE    github white tract at~}
+\CommentTok{##  8 LCBC-UiO/~ FALSE TRUE    github white tract at~}
+\CommentTok{##  9 LCBC-UiO/~ TRUE  FALSE   github Harvard-Oxford~}
+\CommentTok{## 10 LCBC-UiO/~ TRUE  FALSE   github extra 2d view ~}
+\end{Highlighting}
+\end{Shaded}
+
+Having found an existing atlas, you will need to install this like an R-library.
+There is a convenience-function in \emph{ggsegExtra} that will help you do that.
+Alternatively, you can install all the existing atlases.
+This function will take time to execute and will consume quite some space as each atlas (particularly ggseg3d-atlases) can be large.
+Once the atlases are installed, the atlas-packages can be loaded like any other R-package using \texttt{library}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Install ggseg-atlas listed by ggseg_atlas_repos("Yeo")}
+\KeywordTok{install_ggseg_atlas}\NormalTok{(yeo_repo}\OperatorTok{$}\NormalTok{repo, yeo_repo}\OperatorTok{$}\NormalTok{source)}
+
+\CommentTok{# This will take quite some time, there are many atlases with heavy data.}
+\KeywordTok{install_ggseg_atlas_all}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(ggsegYeo2011)}
+\KeywordTok{library}\NormalTok{(ggsegGlasser)}
+\KeywordTok{library}\NormalTok{(ggsegDesterieux)}
+\KeywordTok{library}\NormalTok{(ggsegICBM)}
+
+\NormalTok{yeo7_p <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ yeo7 , }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{region),}
+            \DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{position=}\StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"yeo7"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"ggsegYeo2011"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"yeo7-data-set"}\NormalTok{)}
+
+\NormalTok{glasser_p <-}\StringTok{ }\KeywordTok{ggseg}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ glasser , }\DataTypeTok{mapping=}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{region),}
+            \DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{position=}\StringTok{"stacked"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_brain}\NormalTok{(}\StringTok{"glasser"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"ggsegExtra"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"glasser-data-set"}\NormalTok{)}
+
+\NormalTok{ggsegs_p <-}\StringTok{ }\NormalTok{yeo7_p }\OperatorTok{+}\StringTok{ }\NormalTok{glasser_p }\OperatorTok{+}\StringTok{ }\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels =} \StringTok{"A"}\NormalTok{)}
+
+\NormalTok{desterieux_p <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ desterieux_3d) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_annotations}\NormalTok{(}
+    \DataTypeTok{text=}\StringTok{"desterieux_3d data-set"}\NormalTok{,}
+    \DataTypeTok{y =} \DecValTok{1}\NormalTok{,}
+    \DataTypeTok{yanchor =} \StringTok{'top'}\NormalTok{,}
+    \DataTypeTok{legendtitle=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{showarrow=}\OtherTok{FALSE}\NormalTok{,}
+    \DataTypeTok{font =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{family =} \StringTok{'sans serif'}\NormalTok{,}
+                \DataTypeTok{size =} \DecValTok{50}\NormalTok{))}
+
+\NormalTok{icbm_p <-}\StringTok{ }\KeywordTok{ggseg3d}\NormalTok{(}\DataTypeTok{atlas =}\NormalTok{ icbm_3d) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pan_camera}\NormalTok{(}\StringTok{"right lateral"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_glassbrain}\NormalTok{(}\StringTok{"left"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{remove_axes}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_annotations}\NormalTok{(}
+    \DataTypeTok{text=}\StringTok{"icbm_3d data-set"}\NormalTok{,}
+    \DataTypeTok{y =} \DecValTok{1}\NormalTok{,}
+    \DataTypeTok{yanchor =} \StringTok{'top'}\NormalTok{,}
+    \DataTypeTok{legendtitle=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{showarrow=}\OtherTok{FALSE}\NormalTok{,}
+    \DataTypeTok{font =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{family =} \StringTok{'sans serif'}\NormalTok{,}
+                \DataTypeTok{size =} \DecValTok{50}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Saving 6.5 x 4.5 in image
+\end{verbatim}
 
 \begin{figure}[H]
-\includegraphics[width=0.8\linewidth]{png/ggseg_extra} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra1} \includegraphics[width=0.4\linewidth]{png/ggseg3d_extra2} \caption{Four example data sets from the \emph{ggsegExtra} library, plotted with the \texttt{ggseg()} and \texttt{ggseg3d()} functions.}\label{fig:ggsegExtra}
+\includegraphics[width=0.8\linewidth]{msc_ggseg_files/figure-latex/figure10ab} \includegraphics[width=0.4\linewidth]{msc_ggseg_files/figure-latex/figure10c} \includegraphics[width=0.4\linewidth]{msc_ggseg_files/figure-latex/figure10d} \caption{Four example data-sets from the \emph{ggsegExtra} library, plotted with the \texttt{ggseg()} and \texttt{ggseg3d()} functions.}\label{fig:figure10}
 \end{figure}
+
+\hypertarget{creating-custom-atlases}{%
+\subsubsection{Creating custom atlases}\label{creating-custom-atlases}}
+
+In many cases, users would like to create their own atlases based on existing parcellations or ones own imaging results.
+There are several functions in \emph{ggsegExtra} to aid creating ones own atlas compatible with either \emph{ggseg} or \emph{ggseg3d}.
+These processes depend on the origin of the segmentation (surface-based vs.~volume-based), and what type of system you are working on.
+In addition, one will need other software dependencies to create new ggseg-compatible atlases.
+Therefore, we will not describe the process here, but refer the readers to the ``Articles'' sections of the \href{https://lcbc-uio.github.io/ggsegExtra/index.html}{online documentation}, for further explanations and guides.
+We encourage users who successfully create new atlases, to contribute to the \emph{ggsegExtra} brain atlas repository by adding the repository containing the brain atlases to the repository-list.
 
@@
 The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (\citet{binder}) or Shiny (\citet{shiny}), is an added benefit.
-This is exemplified in the online supplementary information of Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}).
+This is exemplified in the \href{https://athanasiamo.shinyapps.io/Virtual_histology_2019/}{online supplementary information} of Vidal-Piñeiro et al. (\citeyearpar{vidal_2019}).
 
+The functions in the \emph{ggseg}-packages provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in \emph{ggsegExtra}.
 The three \emph{ggseg}-packages contain three main features:
 
-\textbf{1)} a collection of 2D-polygon and 3D mesh brain parcellation atlases.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  A collection of 2D-polygon and 3D mesh brain parcellation atlases.
 The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.\\
-\textbf{2)} \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.\\
-- \texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.\\
-- \texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.\\
-\textbf{3)} Complimentary features -- e.g.~color scales - and functions such as \texttt{as\_ggseg\_atlas()} and \texttt{as\_ggseg3d\_atlas()} to convert data in the correct atlas format.
+\item
+  \texttt{ggseg()} and \texttt{ggseg3d()} functions for visualization.
+Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from \emph{ggplot2} and \emph{plotly}, respectively.
+  \texttt{ggseg()} is a wrapper function for \texttt{geom\_polygon} from \emph{ggplot2} and it can be built upon like any \emph{ggplot2} object.
+  \texttt{ggseg3d()} is a \emph{plotly} wrapper function for tri-surface mesh plots which prints 3D atlases.\\
+\item
+  Complimentary features, like color scales, and functions to convert compatible data frames to atlas-classes, read in likely data sources (like Freesurfer analysis output), and help create new atlases for use with \emph{ggseg} and \emph{ggseg3d}.
+\end{enumerate}
 
-These functions provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in \emph{ggsegExtra}.
-
-The foundations of the \emph{ggseg}-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supplementary information in Vidal-Piñeiro et al.~(\citeyearpar{vidal_2019}) ) .
+The foundations of the \emph{ggseg}-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supplementary information in Vidal-Piñeiro et al.
+(\citeyearpar{vidal_2019}) ) .
 That is, \emph{ggseg} appears with the need to inspect and display brain information over time - i.e.~including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.~bar plots).
@@
 Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public \href{https://github.com/LCBC-UiO/ggseg}{ggseg GitHub repository}, \href{https://github.com/LCBC-UiO/ggseg3d}{ggseg3d GitHub repository}, and \href{https://github.com/LCBC-UiO/ggsegExtra}{ggsegExtra GitHub repository}.
+
+The ggseg-packages are based on \emph{ggplot2} and \emph{plotly}, which are only two of several visualisation tools in R, in addition to base-R plotting.
+While the functionality presented here could also be integrated into packages such as \href{https://www.rdocumentation.org/packages/lattice/versions/0.20-40}{lattice}, both \emph{ggplot2} and \emph{plotly} have become established tools in the R-community and have extensive online documentation.
+This makes them great packages to base further extensions and wrappers on.
 
@@
 
-In the \emph{ggsegExtra} \href{https://github.com/LCBC-UiO/ggsegExtra/wiki/Contributing\%3A-polygon-atlases-new}{github wiki}, we offer a pipeline to create and supply atlases for 2D plotting.
-At the moment, the creation of atlas for ggseg is convoluted and difficult and requires manual intervention.
-We are in the process of designing a simple, straightforward pipeline to facilitate the creation of new ggseg (2D) atlases.
-The aim is to create a set of functions that will call specialized tools like FSL(\citet{fsl}), Freesurfer(\citet{fischl_99}, \citet{dale_99}, \citet{Fischl2000}) and Imagemagick(\citet{magick}) to detect the polygon vectors or the mesh segments given an MRI image containing a parcellation specification, and organize these into valid \emph{ggseg} and \emph{ggseg3d} atlases.
-We encourage users to contribute to the \emph{ggsegExtra} brain atlas repository by including additional brain atlases.
+In terms of planned \emph{ggseg} improvements, we are working on making a proper \texttt{geom} or \texttt{stat}.
+While \texttt{ggseg()} works quite nicely as a \texttt{ggplot} wrapper, it introduces some unexpected work for the user, like having to group data \emph{before} faceting.
+A \texttt{geom} or \texttt{stat} would benefit the user as it would behave more like other \emph{ggplot2} \texttt{geoms} and there would be less new functionalities to learn.
+
+Concerning \emph{ggseg3d}, we also hope to make it possible to colour each individual mesh triangle, and as such provide the users with more flexibility in the plotting.
+This should also improve the speed of \texttt{ggseg3d()}, which currently is quite slow due to each parcellation region being loaded into \texttt{plotly} sequentially.
+Loading the entire brain mesh at once, and then coloring mesh triangles would significiantly improve the speed of \texttt{ggseg3d()}.
+When able to colour the mesh triangles directly, rather than each parcellation individually, only full brain representations of each surface (rather than each surface of each atlas) needs to be stored, in addition to text data indexing mesh-locations.
+This would reduce package size considerably, including package sizes of all atlas-repositories.
 

### 2020-04-06 README.Rmd
fixing binder

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_

<!-- badges: start -->
[![Launch Rstudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/Athanasiamo/ggsegpaper/master?urlpath=rstudio)
<!-- badges: end -->

This is the repository for the tutorial "Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_"

Full [pdf](msc_ggseg.pdf), [Rmarkdown-file](msc_ggseg.Rmd), and latex files can be found directly in this repository.

[arXiv](https://arxiv.org/abs/1912.08200)

### 2020-04-06 install.R

@@ -1,0 +1,10 @@
+install.packages("dplyr")
+install.packages("tidyr")
+install.packages("plotly")
+install.packages("ggplot2")
+install.packages("kableExtra")
+install.packages("knitr")
+install.packages("remotes")
+remotes::install_github("LCBC-UiO/ggseg")
+remotes::install_github("LCBC-UiO/ggseg3d")
+remotes::install_github("LCBC-UiO/ggsegExtra")

### 2020-04-06 README.Rmd
cant get binder to work

@@
 <!-- badges: start -->
-[![Launch Rstudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/Athanasiamo/ggsegpaper/master?urlpath=rstudio)
+<!-- [![Launch Rstudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/Athanasiamo/ggsegpaper/master?urlpath=rstudio) -->
 <!-- badges: end -->

### 2020-04-06 codethrough.Rmd
add code through

--- 2020-04-06 msc_ggseg.Rmd
@@
 ---
-title: Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_
-preprint: true
-author: 
-  - name: Athanasia M. Mowinckel
-    affiliation: 1
-    corresponding: true
-    email: a.m.mowinckel@psykologi.uio.no
-  - name: Didac Vidal-Piñeiro
-    affiliation: 1
-affiliation:
-  - code: 1
-    address: Center for Lifespan Changes in Brain and Cognition, University of Oslo, PO. box 1094 Blindern, 0317 Oslo, Norway
-abstract: >
-  There is an increased emphasis on visualizing neuroimaging results in more intuitive ways.
-Common statistical tools for dissemination, such as bar charts, lack the spatial dimension that is inherent in neuroimaging data.
-Here we present two packages for the statistical software R, _ggseg_ and _ggseg3d_, that integrate this spatial component.
-The _ggseg_ and _ggseg3d_ packages visualize pre-defined brain segmentations as both 2D polygons and 3D meshes, respectively.
-Both packages are integrated with other well-established R-packages, allowing great flexibility.
-In this tutorial, we present the main data and functions in the _ggseg_ and _ggseg3d_ packages for brain atlas visualization.
-The main highlighted functions are able to display brain segmentation plots in R.
-Further, the accompanying _ggsegExtra_-package includes a wider collection of atlases, and is intended for community-based efforts to develop more compatible atlases for _ggseg_ and _ggseg3d_.
-Overall, the _ggseg_-packages facilitate parcellation-based visualizations in R, improve and ease the dissemination of the results, and increase the efficiency of the workflows.
-bibliography: references.bib
-output:
-  bookdown::pdf_book:
-    base_format: rticles::peerj_article # for using bookdown features like \@ref()
-    citation_package: natbib
-    fig_caption: true
-    includes:
-      in_header: preamble.tex
-  rticles::peerj_article: preprint
-keep_tex: true
+title: "Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_"
+subtite: "code through"
+output: html_notebook
 ---
@@
         tibble.width = 50) 
-
-library(dplyr)
-library(tidyr)
-library(plotly) 
-library(ggplot2)
-library(ggseg)
-library(ggseg3d)
-library(ggsegExtra)
-library(kableExtra)
-library(knitr)
-
-include_plotly <- function(plot, file){
-  orca(plot, file = file, format = "png")
-  knitr::include_graphics(file)
-}
-
-# Hook for setting latex text sizes
-def.chunk.hook  <- knitr::knit_hooks$get("chunk") 
-knitr::knit_hooks$set(chunk = function(x, options) {
-  x <- def.chunk.hook(x, options)
-  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
-})
-
-# Hook for setting code chunk wrapping
-hook_output = knit_hooks$get('output')
-knit_hooks$set(output = function(x, options) {
-  # this hook is used only when the linewidth option is not NULL
-  if (!is.null(n <- options$linewidth)) {
-    x = knitr:::split_lines(x)
-    # any lines wider than n should be wrapped
-    if (any(nchar(x) > n)) x = strwrap(x, width = n)
-    x = paste(x, collapse = '\n')
-  }
-  hook_output(x, options)
-})
-
-get_data_list <- function(package){
-  as_tibble(data(package=package)$results)
-}
-
-pkglist <- c("ggseg", "ggseg3d",  
-             basename(ggsegExtra::ggseg_atlas_repos()$repo))
-dt <- lapply(pkglist, get_data_list) %>% 
-  bind_rows() %>% 
-    filter(!grepl("pal", Item)) %>% 
-  mutate(Type = ifelse(grepl("3d", Item), "Mesh", "Polygon"),
-         Item = gsub("_3d", "", Item),
-         Item = gsub("_", "", Item)) %>% 
-    select(Package, Title, Type, Item) %>% 
-  spread(Type, Package, fill = "None")  %>% 
-  mutate(Citation = case_when(
-    grepl("dk", Item) ~ "\\citet{dk}",
-    grepl("yeo", Item) ~ "\\citet{yeo2011}",
-    grepl("aseg|hcpa", Item) ~ "\\citet{aseg}",
-    grepl("desterieux", Item) ~ "\\citet{desterieux}",
-    grepl("^chen", Item) ~ "\\citet{chen}",
-    grepl("glasser", Item) ~ "\\citet{glasser}",
-    grepl("icbm", Item) ~ "\\citet{icbm}",
-    grepl("jhu", Item) ~ "\\citet{jhu}",
-    grepl("schaefer", Item) ~ "\\citet{schaefer}",
-    grepl("tracula", Item) ~ "\\citet{tracula}",
-    grepl("^ho", Item) ~ "\\citet{ho}",
-    TRUE ~ ""
-  ))
-
-options(warnPartialMatchAttr = FALSE,
-        warnPartialMatchDollar = FALSE,
-        warnPartialMatchArgs = FALSE)
 ```
@@
 The _ggsegExtra_ package contains a list of online repositories of pre-compiled atlases (currently `r sum(!grepl("None", dt$Polygon))-1` additional polygon atlases and `r sum(!grepl("None", dt$Mesh))-1` mesh atlases) and it is frequently updated.
-A summary of all available atlases compatible with the _ggseg_-packages as of `r format(Sys.Date(), "%B %Y")` can be found in Table \@ref(tab:table1).
-
-```{r 'table1', echo=F, warning=FALSE, message=FALSE}
-dt %>%
-  kable(format="latex", escape = FALSE, booktabs = TRUE,
-        caption = "Table of currently available atlases in ggseg and ggseg3d, and listed in the ggsegExtra R-package.
-Polygon and Mesh refer to 2D and 3D brain atlas representations, respectively.
-Cell contents in the Mesh and Polygon columns indicate the source package containing the atlas data-set.") %>%
-  kable_styling(latex_options = c("scale_down"))
-```
 
@@
 
-```{r "figure7-show", message=FALSE, echo=TRUE, results = 'hide',  out.width='30%',fig.show='hold',fig.ncol = 3}
+```{r "figure7-show", out.width='30%',fig.show='hold',fig.ncol = 3}
 # remotes::install_github("LCBC-UiO/ggseg3d")
@@
 
-```{r "figure7", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:figure7-cap)'}
-orca(fig7l, file = "msc_ggseg_files/figure-latex/figure7l.png", format = "png")
-orca(fig7m, file = "msc_ggseg_files/figure-latex/figure7m.png", format = "png")
-orca(fig7r, file = "msc_ggseg_files/figure-latex/figure7r.png", format = "png")
-
-include_graphics(c("msc_ggseg_files/figure-latex/figure7l.png",
-                   "msc_ggseg_files/figure-latex/figure7m.png",
-                   "msc_ggseg_files/figure-latex/figure7r.png"))
-```
-
 The 3D-atlas data is stored in nested tibbles.
@@
 
-```{r out.width='80%', linewidth=100, size='small', eval=TRUE}
+```{r out.width='80%'}
 # Select surface and hemisphere, and then unnest to inspect the atlas data
@@
 
-```{r 'figure8-show', include=FALSE, out.width='40%', results='hide'}
+```{r 'figure8-show', include=FALSE, out.width='40%'}
 # Figure 8 left
@@
 **Right:** The colour of the `NA` values can also be changed through the option 'na.colour'.
-
-```{r 'figure8', out.width='30%', echo=FALSE, fig.cap="(ref:figure8-cap)", fig.show='hold', fig.ncol = 3}
-orca(fig8l, file = "msc_ggseg_files/figure-latex/figure8l.png", format = "png")
-orca(fig8m, file = "msc_ggseg_files/figure-latex/figure8m.png", format = "png")
-orca(fig8r, file = "msc_ggseg_files/figure-latex/figure8r.png", format = "png")
-
-include_graphics(c(
-  "msc_ggseg_files/figure-latex/figure8l.png",
-  "msc_ggseg_files/figure-latex/figure8m.png",
-  "msc_ggseg_files/figure-latex/figure8r.png"))
-
-```
 
@@
 
-# Discussion
-
-The main aim of the _ggseg_, _ggseg3d_, and _ggsegExtra_ packages is to ease and streamline visualization of brain atlas data in R, by gathering a collection of atlases from several scientific sources and providing customized plotting functions.
-In this tutorial, we introduced the packages to the readers by presenting some use examples and highlighting the main functions and options that are available.
-As visualization tools, these packages add up to manifold functionalities such as ggBrain (@ggBrain) and ggneuro (@ggneuro) in R, and software-specific image viewers such as FSLeyes (@fsleyes) and Freeview (@dale_99).
-In this regard, we do not aim to compete with software-specific visualizations or advocate for the superiority of the _ggseg_-packages as visualization tools.
-After all, flattened 2D polygons do not rely on a meaningful brain coordinate system and the units of information in 3D meshes are limited to the number of parcellations.
-On the contrary, we believe the _ggseg_ niche among visualization tools resides in its simplicity and its ability to be combined with statistical analysis pipelines.
-The possibility to serve as an interactive tool for dissemination and reproducibility when combined with other technologies, such as Binder (@binder) or Shiny (@shiny), is an added benefit.
-This is exemplified in the [online supplementary information](https://athanasiamo.shinyapps.io/Virtual_histology_2019/) of Vidal-Piñeiro et al. (-@vidal_2019).
-
-The functions in the _ggseg_-packages provide users with the possibility of adapting the plots to their wishes, and also makes it possible to create and contribute to the atlas repository in _ggsegExtra_.
-The three _ggseg_-packages contain three main features:
-
-1.
-A collection of 2D-polygon and 3D mesh brain parcellation atlases.
-The atlas data include the necessary coordinates for plotting, and include other information that should be recognizable for users.
-2.
-`ggseg()` and `ggseg3d()` functions for visualization.
-Both functions are flexible and well-adapted to their environment and can be combined with any additional argument from _ggplot2_ and _plotly_, respectively.
-`ggseg()` is a wrapper function for `geom_polygon` from _ggplot2_ and it can be built upon like any _ggplot2_ object. 
-`ggseg3d()` is a _plotly_ wrapper function for tri-surface mesh plots which prints 3D atlases.  
-3.
-Complimentary features, like color scales, and functions to convert compatible data frames to atlas-classes, read in likely data sources (like Freesurfer analysis output), and help create new atlases for use with _ggseg_ and _ggseg3d_.
-
-The foundations of the _ggseg_-packages trace back to the necessity of visualizing and exploring the lifespan trajectories of cortical thickness across different brain regions (see supplementary information in Vidal-Piñeiro et al.
-(-@vidal_2019) ) .
-That is, _ggseg_ appears with the need to inspect and display brain information over time - i.e.
-including a spatial dimension and a time-varying factor - overcoming the constrains of printed journals and classical 2D plots (e.g.
-bar plots).
-The current state of science requires researches to share the results of studies in both high detail and in an intuitive manner, as it permits communication to wide audiences and facilitates reproducibility.
-Hence, we believe this tool conforms to the essence of open science and invite users to improve the code, provide examples, or tutorials, and contribute to the atlas collection according to their own interest and needs via the public [ggseg GitHub repository](https://github.com/LCBC-UiO/ggseg), [ggseg3d GitHub repository](https://github.com/LCBC-UiO/ggseg3d), and [ggsegExtra GitHub repository](https://github.com/LCBC-UiO/ggsegExtra).
-
-The ggseg-packages are based on _ggplot2_ and _plotly_, which are only two of several visualisation tools in R, in addition to base-R plotting. 
-While the functionality presented here could also be integrated into packages such as [lattice](https://www.rdocumentation.org/packages/lattice/versions/0.20-40), both _ggplot2_ and _plotly_ have become established tools in the R-community and have extensive online documentation.
-This makes them great packages to base further extensions and wrappers on.
-
-Finally - while the _ggseg_-packages are circumscribed to brain parcellations - we believe that the structure and functions of the package can be easily applied to any scientific field that benefits from data being displayed across the spatial dimension.
-We encourage readers to borrow the package functionalities and adapt it to their respective fields and structures of interest, such as has already been done with the _gganatogram_-package (@gganatogram).
-
-# Planned package improvements
-In terms of planned _ggseg_ improvements, we are working on making a proper `geom` or `stat`.
-While `ggseg()` works quite nicely as a `ggplot` wrapper, it introduces some unexpected work for the user, like having to group data _before_ faceting. 
-A `geom` or `stat` would benefit the user as it would behave more like other _ggplot2_ `geoms` and there would be less new functionalities to learn.
-
-Concerning _ggseg3d_, we also hope to make it possible to colour each individual mesh triangle, and as such provide the users with more flexibility in the plotting. 
-This should also improve the speed of `ggseg3d()`, which currently is quite slow due to each parcellation region being loaded into `plotly` sequentially. 
-Loading the entire brain mesh at once, and then coloring mesh triangles would significiantly improve the speed of `ggseg3d()`. 
-When able to colour the mesh triangles directly, rather than each parcellation individually, only full brain representations of each surface (rather than each surface of each atlas) needs to be stored, in addition to text data indexing mesh-locations.
-This would reduce package size considerably, including package sizes of all atlas-repositories.
-
-# Conclusion
-Visualization is a fundamental aspect of neuroimaging to explore and understand data, guide interpretation, and communicate with colleagues and the general audience.
-In this tutorial, we have introduced the _ggseg_-packages, tools for visualizing brain statistics through brain parcellation atlases in R.
-This visualization tool easily combines with interactive routines as well as with diverse statistical analysis pipelines.
-We hope this tool and tutorial proves useful to neuroscientists and inspires others to apply the functions in a wide variety of fields and structures.
-
-# Author Contributions
-Didac Vidal-Piñeiro generated the idea for the tool, and the initial scripts for plot visualization.
-He has also been responsible for converting images from neuroimaging data to ggseg-like data (e.g. polygons and mesh data).
-Athanasia M. Mowinckel adapted the initial scripts and made the functions into package format and has continued developing the functions with the aim of increasing user-friendliness.
-She is also responsible for conceiving and adding the mesh-plot functionality through plotly, and developing the pipeline for making that possible.
-A. M. Mowinckel wrote the first draft of the paper, and both have since critically edited it.
-
-# Conflicts of Interest
-The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
-
-# Acknowledgements
-We thank John Muschelli for package code comments and its adaptation to neuroconductor ([2018]), and Richard Beare for the first community created atlas, and base code comments.
-We are indebted to A.
-M.
-Winkler scripts for converting neuroimaging data (Freesurfer) into .ply files, a necessary step for neuroimaging to mesh-plots conversion, and to Inge Amlien for the ‘LCBC’ surface files and helpful contributions to the ggseg3d pipeline.
-We are indebted to those users that actively contributed to the package and all the individuals that worked on the framework on which the package is sustained (e.g. R-project, neuroimaging software).
-Finally, we thank Anders M. Fjell (AMF) and Kristine B. Walhovd (KBW) for their encouragement and financial support.
-
 # Funding

### 2020-04-06 install.R

@@ -2,9 +2,14 @@ install.packages("dplyr")
 install.packages("tidyr")
 install.packages("plotly")
 install.packages("ggplot2")
+install.packages("patchwork")
 install.packages("kableExtra")
 install.packages("knitr")
 install.packages("remotes")
 remotes::install_github("LCBC-UiO/ggseg")
 remotes::install_github("LCBC-UiO/ggseg3d")
 remotes::install_github("LCBC-UiO/ggsegExtra")
+remotes::install_github("LCBC-UiO/ggsegYeo2011")
+remotes::install_github("LCBC-UiO/ggsegDesterieux")
+remotes::install_github("LCBC-UiO/ggsegGlasser")
+remotes::install_github("LCBC-UiO/ggsegICBM")

### 2020-04-21 codethrough.Rmd
new codethrough

@@
 output: html_notebook
+bibliography: references.bib
 ---
@@
 ```{r, size='small'}
+library(dplyr)
 file_path <- "data/aparc.volume.table"
@@
 library(ggseg3d)
+library(plotly)
 
@@
   filter(surf == "inflated" & hemi == "right") %>%
-  unnest(ggseg_3d) %>%
+  tidyr::unnest(ggseg_3d) %>%
   head(5)
@@
 ```{r 'figure9', out.width='60%', echo=F, eval=T, fig.cap="(ref:figure9-cap)"}
-include_plotly(fig9, "msc_ggseg_files/figure-latex/figure9.png")
+# Convenience function to easily add plotly images to rmarkdown
+include_plotly <- function(plot, file){
+  orca(plot, file = file, format = "pdf")
+  knitr::include_graphics(file)
+}
+
+include_plotly(fig9, "figure9.pdf")
 ```
@@
 
-invisible(ggsave(filename = "msc_ggseg_files/figure-latex/figure10ab.png", 
-                 plot = ggsegs_p, device = "png"))
-orca(desterieux_p, file = "msc_ggseg_files/figure-latex/figure10c.png", format = "png")
-orca(icbm_p, file = "msc_ggseg_files/figure-latex/figure10d.png", format = "png")
+invisible(ggsave(filename = "figure10ab.pdf", 
+                 plot = ggsegs_p, device = "pdf"))
+orca(desterieux_p, file = "figure10c.pdf", format = "pdf")
+orca(icbm_p, file = "figure10d.pdf", format = "pdf")
 
-include_graphics(c(
-  "msc_ggseg_files/figure-latex/figure10ab.png",
-  "msc_ggseg_files/figure-latex/figure10c.png",
-  "msc_ggseg_files/figure-latex/figure10d.png"
+knitr::include_graphics(c(
+  "figure10ab.pdf",
+  "figure10c.pdf",
+  "figure10d.pdf"
 ))

### 2020-04-21 msc_ggseg.Rmd
new codethrough

@@
 include_plotly <- function(plot, file){
-  orca(plot, file = file, format = "png")
+  orca(plot, file = file, format = "pdf")
   knitr::include_graphics(file)
@@
 ```{r "figure7", message=FALSE, echo=F, eval=TRUE, out.width='30%',fig.show='hold',fig.ncol = 3, fig.cap='(ref:figure7-cap)'}
-orca(fig7l, file = "msc_ggseg_files/figure-latex/figure7l.png", format = "png")
-orca(fig7m, file = "msc_ggseg_files/figure-latex/figure7m.png", format = "png")
-orca(fig7r, file = "msc_ggseg_files/figure-latex/figure7r.png", format = "png")
+orca(fig7l, file = "msc_ggseg_files/figure-latex/figure7l.pdf", format = "pdf")
+orca(fig7m, file = "msc_ggseg_files/figure-latex/figure7m.pdf", format = "pdf")
+orca(fig7r, file = "msc_ggseg_files/figure-latex/figure7r.pdf", format = "pdf")
 
-include_graphics(c("msc_ggseg_files/figure-latex/figure7l.png",
-                   "msc_ggseg_files/figure-latex/figure7m.png",
-                   "msc_ggseg_files/figure-latex/figure7r.png"))
+include_graphics(c("msc_ggseg_files/figure-latex/figure7l.pdf",
+                   "msc_ggseg_files/figure-latex/figure7m.pdf",
+                   "msc_ggseg_files/figure-latex/figure7r.pdf"))
 ```
@@
 ```{r 'figure8', out.width='30%', echo=FALSE, fig.cap="(ref:figure8-cap)", fig.show='hold', fig.ncol = 3}
-orca(fig8l, file = "msc_ggseg_files/figure-latex/figure8l.png", format = "png")
-orca(fig8m, file = "msc_ggseg_files/figure-latex/figure8m.png", format = "png")
-orca(fig8r, file = "msc_ggseg_files/figure-latex/figure8r.png", format = "png")
+orca(fig8l, file = "msc_ggseg_files/figure-latex/figure8l.pdf", format = "pdf")
+orca(fig8m, file = "msc_ggseg_files/figure-latex/figure8m.pdf", format = "pdf")
+orca(fig8r, file = "msc_ggseg_files/figure-latex/figure8r.pdf", format = "pdf")
 
 include_graphics(c(
-  "msc_ggseg_files/figure-latex/figure8l.png",
-  "msc_ggseg_files/figure-latex/figure8m.png",
-  "msc_ggseg_files/figure-latex/figure8r.png"))
+  "msc_ggseg_files/figure-latex/figure8l.pdf",
+  "msc_ggseg_files/figure-latex/figure8m.pdf",
+  "msc_ggseg_files/figure-latex/figure8r.pdf"))
 
@@
 ```{r 'figure9', out.width='60%', echo=F, eval=T, fig.cap="(ref:figure9-cap)"}
-include_plotly(fig9, "msc_ggseg_files/figure-latex/figure9.png")
+include_plotly(fig9, "msc_ggseg_files/figure-latex/figure9.pdf")
 ```
@@
 
-invisible(ggsave(filename = "msc_ggseg_files/figure-latex/figure10ab.png", 
-                 plot = ggsegs_p, device = "png"))
-orca(desterieux_p, file = "msc_ggseg_files/figure-latex/figure10c.png", format = "png")
-orca(icbm_p, file = "msc_ggseg_files/figure-latex/figure10d.png", format = "png")
+invisible(ggsave(filename = "msc_ggseg_files/figure-latex/figure10ab.pdf", 
+                 plot = ggsegs_p, device = "pdf"))
+orca(desterieux_p, file = "msc_ggseg_files/figure-latex/figure10c.pdf", format = "pdf")
+orca(icbm_p, file = "msc_ggseg_files/figure-latex/figure10d.pdf", format = "pdf")
 
 include_graphics(c(
-  "msc_ggseg_files/figure-latex/figure10ab.png",
-  "msc_ggseg_files/figure-latex/figure10c.png",
-  "msc_ggseg_files/figure-latex/figure10d.png"
+  "msc_ggseg_files/figure-latex/figure10ab.pdf",
+  "msc_ggseg_files/figure-latex/figure10c.pdf",
+  "msc_ggseg_files/figure-latex/figure10d.pdf"
 ))

### 2020-04-21 codethrough.Rmd
new codethrough

@@
 Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
-The _ggsegExtra_ package contains a list of online repositories of pre-compiled atlases (currently `r sum(!grepl("None", dt$Polygon))-1` additional polygon atlases and `r sum(!grepl("None", dt$Mesh))-1` mesh atlases) and it is frequently updated.
+The _ggsegExtra_ package contains a list of online repositories of pre-compiled atlases and it is frequently updated.
 
